### 2019-08-07 rrd.Rmd
init

---
title: "Reproducible Research with Docker"
author: "Andreas Brandmaier & Aaron Peikert"
date: "8/6/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
if(!require("pacman"))install.packages("pacman")
pacman::p_load("here", "tidyverse")
```

In this tutorial, we describe a principled workflow to ensure reproducible research in R-based environments. The workflow relies on the interplay of various open-source software tools including git, docker, R, Rmarkdown, knitr, and papaja whose interplay ensures a seamless integration of version management, (APA-conformable) dynamic report generation and full cross-platform computational reproducibility. The workflow ensures the primary goals that 1) the reporting of analysis results is perfectly consistent with the actual analysis results (dynamic report generation), the analysis exactly reproduces at later time even if the computing and software platform is updated or changed (computational reproducibility), and 3) intermediate or post-publication changes are tracked, marked, and documented while earlier versions of both data and code remain accessible. While dynamic document generation is increasingly recognized as tool for reproducible analyses, we demonstrate with practical examples that dynamic documents are not sufficient to ensure computational reproducibility.

Reproducible Research has become the gold standard for scientific work in recent years. The following tutorial aims to outline a workflow that does not only garanties a sane standard of reproducibility but also decreases the burden of maintaining software for the researcher and his colleages. The authors beflief that integrating tools from software engineering in the process of scientific reporting to enhance repdoducibility is crucial and give guidelines for the practitioner to use them. These tools include software for literate programming, version control, dependency tracking and containerization.

Literate Programming
: Interwoven text and computer code that together form a human readable document.
: Example: Including text that descibes a statistical test, code that is uesed to exucte it and the results from it in the same document.

Version Control
: The management of change of documents/code.
: Example: Changes from the manuscript text upon request by the reviewer are clearly identifiable.

Dependency Tracking
: How different parts of a manuscript/analysis do relate to each other and how these dependences are resolved.
: Example: The manuscript depends on the statistical analysis which depends on the data.

Containerization
: A virtual computational environment that does encapsulate all needed software. More formally called operating-system-level virtualization.
: Example: A statistical analysis depends on the statistical software that is used and they are bundled together.

Within the community of users of the R Programming Language for each purpose a specific software solutions is clearly the most popular. For literate programming, `rmarkdown`, for version control `git`, for dependency management `make` and for container `docker`. Each of them serves a meta-scientific valuable goal (reproducibility) and increases the producitivity of the researcher. Each of these software solutions are extremly mighty, so in order to master them fully often years of practice are required. Luckily the minimal valuable subset needed can be learned with a little practice.

### Random Number Generation with R

```{bash}
docker run --rm rocker/r-ver:3.3.0 Rscript -e "set.seed(1234);sample(1:10, 5)"
```

```{bash}
docker run --rm rocker/r-ver:3.4.0 Rscript -e "set.seed(1234);sample(1:10, 5)"
```

```{bash}
docker run --rm rocker/r-ver:3.5.0 Rscript -e "set.seed(1234);sample(1:10, 5)"
```

```{bash}
docker run --rm rocker/r-ver:3.6.0 Rscript -e "set.seed(1234);sample(1:10, 5)"
```

```{bash}
docker run --rm rocker/r-ver:latest Rscript -e "set.seed(1234);sample(1:10, 5)"
```

### 2019-08-07 rrd.Rmd
add first steps

@@
 : Interwoven text and computer code that together form a human readable document.
-: Example: Including text that descibes a statistical test, code that is uesed to exucte it and the results from it in the same document.
+: Example: One document that includes text that descibes a statistical test, code that is used to execute the statistical and the output of the code in the same document.
 
@@
 Each of these software solutions are extremly mighty, so in order to master them fully often years of practice are required.
-Luckily the minimal valuable subset needed can be learned with a little practice.
+Luckily the minimal valuable subset needed to enusre reproducibility can be learned with a little practice.
+
+The authors believe that for sufficient chances of reproducibility, good documentation is crucial, since it anwers the question "What steps have to be taken to reproduce the results?".
+The most precise documentation is computer code and the most productive computer code is code that runs automaticly without human interaction.
+These principles are what guides the the use of `rmarkdown`, `git`, `make`, `docker` and there interaction.
+
+The first step towards reproducibility is to have a R script, that can be executed on the own computer without error, after everything within R and the files that are the output of the script deleted.
+We assume that the reader is familar with R and has the ability to archive that for a simple analysis.
+A next step is to make sure that all files relevant, can be moved to another computer.
+To archive that it is important that all files are within one folder (and subfolders within it) and all paths are relative to that folder.
+A nice solution to that problem are [Rstudio Projects](https://r4ds.had.co.nz/workflow-projects.html) and/or the `here` package.
+
+Reproducibility at a basic level assumes that results are the same, if neither script nor data has changed.
+It is often not trivial to find out wether anything has changed and if so to "go back in time".
+`git` enables you to do both.
+A good mental model for `git` is that it takes a sequenze of snapshot of all files it is supposed to track.
+These snapshots are reffered to as "commits".
+Each commit has a unique identifier (a hash) and each commits represents a complete copy of the state of the files at the point when they where added to a commit.
 

### 2019-08-12 rrd.Rmd
add draft for git

@@
 In this tutorial, we describe a principled workflow to ensure reproducible research in R-based environments.
-The workflow relies on the interplay of various open-source software tools including git, docker, R, Rmarkdown, knitr, and papaja whose interplay ensures a seamless integration of version management, (APA-conformable) dynamic report generation and full cross-platform computational reproducibility.
+The workflow relies on the interplay of various open-source software tools including Rmarkdown, git, make and docker whose interplay ensures a seamless integration of version management, (APA-conformable) dynamic report generation and full cross-platform computational reproducibility.
 The workflow ensures the primary goals that 1) the reporting of analysis results is perfectly consistent with the actual analysis results (dynamic report generation), the analysis exactly reproduces at later time even if the computing and software platform is updated or changed (computational reproducibility), and 3) intermediate or post-publication changes are tracked, marked, and documented while earlier versions of both data and code remain accessible.
@@
 Each of these software solutions are extremly mighty, so in order to master them fully often years of practice are required.
-Luckily the minimal valuable subset needed to enusre reproducibility can be learned with a little practice.
+Luckily the minimal valuable subset needed to ensure reproducibility can be learned with little practice.
 
@@
 
+```{r, eval=FALSE}
+# BAD
+iris <- read.csv("/home/aaron/Documents/reproducible-research/data/iris.csv")
+# GOOD
+iris <- read.csv("data/iris.csv")
+# GOOD
+iris <- read.csv(here("data", "iris.csv"))
+```
+
+This folder, where everything resides that you need for an analysis, will be refered to as a Projekt.
+Working with Projects is especially easy with RStudio an integrated development environment (IDE) for R.
+
 Reproducibility at a basic level assumes that results are the same, if neither script nor data has changed.
@@
 Each commit has a unique identifier (a hash) and each commits represents a complete copy of the state of the files at the point when they where added to a commit.
+Going back to one state is as easy has finding out the hash of the commit.
+This can be made even easier by tagging specific commits e.g.
+as "submission", "preprint", "publication".
+This collection of commits is called "repository", which is almost the same as your Project.
+
+A typicall workflow looks like that:
+
+```{bash, eval=FALSE}
+git init # to initialize git
+git add /data/iris.csv /R/analysis.R # track specific files/changes
+git commit -m "add data & analysis" # take snapshot
+# new data comes in
+git commit -a -m "complete data colections" # add (-a) and comit all changes
+```
+
+So to just keep track of all changes you only need to use `git add` & `git commit` or `git commit -a` to conviently do both.
+These commands need to be executed in the `terminal`, which you can access from within Rstudio (`Shift + Alt + R`), but if you use Git with Rstudio, you also get a graphical user interface.
+
+![Git Pane in RStudio](Images/git-pane.png)
+
+Now you can check everything that has happend (`git log`) and look at the exact state by suppling the hash or the first few digits of it to `git checkout`:
+
+```{bash, eval = FALSE}
+git log
+git checkout 77db06f78e
+```
+
+Git also makes it especially easy to share and collaborate on a project with another researcher.
+A popular service for sharing via git is [Github](https://github.com).
+Just sharing Git repositories with the public is allways free, privat repositories (only visible to persons you invite) are [free for researcher](https://help.github.com/en/articles/applying-for-an-educator-or-researcher-discount).
+After creating a user account, you can create a new repository where Github advises you how to upload you repository from the terminal:
+
+```{bash, eval=FALSE}
+git remote add origin git@github.com:aaronpeikert/reproducible-research.git
+git push -u origin master
+```
+
+Each update can then be uploaded with `git push` or the green up arrow in the git pane. To download this git repo to another computer just type:
+
+```{bash, eval=FALSE}
+git clone git@github.com:aaronpeikert/reproducible-research.git
+```
+
+Git & Github can do even more to help you to collaborate with your fellow researcher, but this is beyond the already broad scope of this tutorial.
 

### 2019-08-12 rr-flow.Rmd
change title to reflect the broad scope

--- 2019-08-12 rrd.Rmd
@@
 ---
-title: "Reproducible Research with Docker"
+title: "A Reproducible Workflow with RMarkdown, Git, Make & Docker"
+shorttitle: "Reproducible Workflow"
 author: "Andreas Brandmaier & Aaron Peikert"

### 2019-08-12 rr-flow.Rmd
add make section

@@
 iris <- read.csv("data/iris.csv")
-# GOOD
+# BETTER
 iris <- read.csv(here("data", "iris.csv"))
@@
 So to just keep track of all changes you only need to use `git add` & `git commit` or `git commit -a` to conviently do both.
-These commands need to be executed in the `terminal`, which you can access from within Rstudio (`Shift + Alt + R`), but if you use Git with Rstudio, you also get a graphical user interface.
+These commands need to be executed in the `terminal`, which you can access from within Rstudio (`Shift + Alt + R`), but if you use Git with Rstudio, you also get a graphical user interface and quite likly you won't need the terminal more then once per project.
 
@@
 
-Now you can check everything that has happend (`git log`) and look at the exact state by suppling the hash or the first few digits of it to `git checkout`:
+Now you can check everything that has happend (`git log`) and look at the exact state by suppling the id of the commit or the first few digits of it to `git checkout`:
 
@@
 Git & Github can do even more to help you to collaborate with your fellow researcher, but this is beyond the already broad scope of this tutorial.
+Another benefit of using Git & Github is that experimentation is highly encoured, since you can go back to any state easily and everything is backed up on the Github servers.
+
+Even when you have obtained a version of the projekt and you can ensure that this version is unchanged, you may not know how to reproduce the results, because it is unclear what has to be run in which order.
+This job gets really easy with `make`, since it allows you to create (computational) recipes to (re-)create files.
+
+A `Makefile` obayes the following scheme:
+
+```
+target: dependency1 dependency2
+  command-to-create-target
+```
+
+A typical `Makefile` might look like:
+
+```
+analysis.pdf: data/clean.csv
+  Rscript -e 'rmarkdown::render("analysis.Rmd")'
+
+data/clean.csv: R/clean.R data/iris.csv
+  Rscript -e 'source("R/clean.R")'
+```
+
+The first line reads: to create `analysis.pdf` the `analysis.Rmd` needs to be rendered, which depends on `data/clean.csv`.
+This dependency is itself a target.
+To create `data/clean.csv`, `R/clean.R` and `data/iris.csv` are needed.
+If you type `make analysis.pdf` `make` will first check wether the dependencys do exist and if not weather or it knows how to create them.
+So if `data/clean.csv` does not exist, `make` will creates it.
+The same thing happens if one of the dependencies of a target is newer then the target itself, then make updates everything that directly or indirectly depends on it.
+So if `data\iris.csv` is newer then `data\clean.csv`, make atempts to recreate `data\clean.csv` and `analysis.pdf`.
+If there is a dependency which is missing, and there is no recipe to make it, `make` stops with an error.
+It is a convention to have a target `all` which depends on everything and make it the first target in a `Makefile`, then the command `make` without any argument will automaticly create everything.
+This can also be trigered from within RStudio via `Build All`.
+
+```
+all: analysis.pdf
+```
+
+![Build Pane in RStudio](Images/build-pane.png)
+
+If you have followed the above workflow a fellow researcher two commands away from fully reproducing your analysis:
+
+```{bash, eval=FALSE}
+git clone https://github.com/sirfisher/iris.git
+make all
+```
+
+But this relies on the important assumption that your computational environments are sufficiently compatible e.g. everything needed is installed (R + Packages) at a compatible version of the software.
 
 ### Random Number Generation with R
-
-```{bash}
-docker run --rm rocker/r-ver:3.3.0 Rscript -e "set.seed(1234);sample(1:10, 5)"
-```
-
-```{bash}
-docker run --rm rocker/r-ver:3.4.0 Rscript -e "set.seed(1234);sample(1:10, 5)"
-```
 
@@
 ```
-
-```{bash}
-docker run --rm rocker/r-ver:latest Rscript -e "set.seed(1234);sample(1:10, 5)"
-```

### 2019-08-14 rr-flow.Rmd
check spelling & grammar

--- 2019-08-12 rr-flow.Rmd
@@
 Reproducible Research has become the gold standard for scientific work in recent years.
-The following tutorial aims to outline a workflow that does not only garanties a sane standard of reproducibility but also decreases the burden of maintaining software for the researcher and his colleages.
-The authors beflief that integrating tools from software engineering in the process of scientific reporting to enhance repdoducibility is crucial and give guidelines for the practitioner to use them.
-These tools include software for literate programming, version control, dependency tracking and containerization.
+The following tutorial aims to outline a workflow that does guarantees not only a reasonable standard of reproducibility but also decreases the burden of maintaining software for the researcher and his colleagues.
+The authors' believe that integrating tools from software engineering in the process of scientific reporting to enhance reproducibility is crucial and give guidelines for the practitioner to use them.
+These tools include software for literate programming, version control, dependency tracking, and containerization.
 
 Literate Programming
-: Interwoven text and computer code that together form a human readable document.
-: Example: One document that includes text that descibes a statistical test, code that is used to execute the statistical and the output of the code in the same document.
-
+: Interwoven text and computer code that together forms a human-readable document.
+: Example: One document that includes text that describes a statistical test, code that is used to execute the statistical and the output of the code in the same document.
 Version Control
 : The management of change of documents/code.
-: Example: Changes from the manuscript text upon request by the reviewer are clearly identifiable.
+: Example: Changes from the manuscript text upon request by the reviewer are identifiable.
 
 Dependency Tracking
-: How different parts of a manuscript/analysis do relate to each other and how these dependences are resolved.
-: Example: The manuscript depends on the statistical analysis which depends on the data.
+: How different parts of a manuscript/analysis relate and how these dependencies resolve via execution of code.
+: Example: The manuscript depends on the statistical analysis, which depends on the data.
 
@@
 : A virtual computational environment that does encapsulate all needed software. More formally called operating-system-level virtualization.
-: Example: A statistical analysis depends on the statistical software that is used and they are bundled together.
+: Example: A statistical analysis depends on the used statistical software and a virtual machine bundles them together.
 
@@
 
-This folder, where everything resides that you need for an analysis, will be refered to as a Projekt.
-Working with Projects is especially easy with RStudio an integrated development environment (IDE) for R.
+This folder, where everything resides that you need for analysis, is referred to as a Projekt.
+Working with Projects is exceptionally comfortable with RStudio an integrated development environment (IDE) for R.
 
-Reproducibility at a basic level assumes that results are the same, if neither script nor data has changed.
-It is often not trivial to find out wether anything has changed and if so to "go back in time".
-`git` enables you to do both.
-A good mental model for `git` is that it takes a sequenze of snapshot of all files it is supposed to track.
-These snapshots are reffered to as "commits".
-Each commit has a unique identifier (a hash) and each commits represents a complete copy of the state of the files at the point when they where added to a commit.
-Going back to one state is as easy has finding out the hash of the commit.
-This can be made even easier by tagging specific commits e.g.
-as "submission", "preprint", "publication".
-This collection of commits is called "repository", which is almost the same as your Project.
+Reproducibility at a basic level assumes that results are the same if neither script nor data have changed.
+It is often not trivial to find out whether anything has changed and if so to "go back in time." `git` enables you to do both.
+A good mental model for `git` is that it takes a sequence of snapshots of all files it is supposed to track.
+In the language of git, these snapshots are "commits".
+A commit represents a complete copy of the state of the files when added to a commit.
+Each commit has a unique identifier (a hash).
+Going back to one state is as easy as finding the hash of the commit.
+This can be made even easier by tagging specific commits e.g., as "submission," "preprint," "publication." This collection of commits is called "repository," which is almost the same as your Project.
 
-A typicall workflow looks like that:
+A typical git workflow looks like that:
 
@@
 
-So to just keep track of all changes you only need to use `git add` & `git commit` or `git commit -a` to conviently do both.
-These commands need to be executed in the `terminal`, which you can access from within Rstudio (`Shift + Alt + R`), but if you use Git with Rstudio, you also get a graphical user interface.
+So to keep track of all changes, you only need to use `git add` & `git commit` or `git commit -a` to do both.
+These commands need to be executed in the `terminal`, which you can access from within Rstudio (`Shift + Alt + R`), but if you use Git with Rstudio, you also get a graphical user interface and quite likely you won't need the terminal more then once per project.
 
@@
 
-Now you can check everything that has happend (`git log`) and look at the exact state by suppling the hash or the first few digits of it to `git checkout`:
+Now you can check everything that has happened (`git log`) and look at the exact state by supplying the id of the commit or the first few digits of it to `git checkout`:
 
@@
 A popular service for sharing via git is [Github](https://github.com).
-Just sharing Git repositories with the public is allways free, privat repositories (only visible to persons you invite) are [free for researcher](https://help.github.com/en/articles/applying-for-an-educator-or-researcher-discount).
-After creating a user account, you can create a new repository where Github advises you how to upload you repository from the terminal:
+Just sharing Git repositories with the public is always free, private repositories (only visible to persons you invite) are [free for researcher](https://help.github.com/en/articles/applying-for-an-educator-or-researcher-discount).
+After creating a user account, you can create a new repository where Github advises you how to upload your repository from the terminal e.g., for this repository:
 
@@
 
-Each update can then be uploaded with `git push` or the green up arrow in the git pane. To download this git repo to another computer just type:
+`git push` or the green up arrow in the git pane uploads local updates. To download this git repo to another computer type into the terminal:
 
@@
 Git & Github can do even more to help you to collaborate with your fellow researcher, but this is beyond the already broad scope of this tutorial.
+Another benefit of using Git & Github is that experimentation is highly encouraged since you can go back to any state quickly and even when you lose access to the file on your computer, everything is backed up on the Github servers.
+
+Even when you have obtained a version of the project, and you can ensure that this version is unchanged, you may not know how to reproduce the results, because it is unclear what has to be run in which order.
+This job gets comfortable with `make`, since it allows you to create (computational) recipes too (re-)create files.
+
+A `Makefile` obeys the following scheme:
+
+```
+target: dependency1 dependency2
+  command-to-create-target
+```
+
+A typical `Makefile` might look like:
+
+```
+analysis.pdf: data/clean.csv
+  Rscript -e 'rmarkdown::render("analysis.Rmd")'
+
+data/clean.csv: R/clean.R data/iris.csv
+  Rscript -e 'source("R/clean.R")'
+```
+
+The first line reads: to create `analysis.pdf` the `analysis.Rmd` needs to be rendered, which depends on `data/clean.csv`.
+This dependency is itself a target.
+To create `data/clean.csv`, `R/clean.R` and `data/iris.csv` are needed.
+If you type `make analysis.pdf` `make` first checks whether the dependencies do exist and if not whether or it knows how to create them.
+So if `data/clean.csv` does not exist, `make` creates it.
+The same thing happens if one of the dependencies of a target is newer then the target itself, then updates everything that directly or indirectly depends on it.
+So if `data\iris.csv` is newer then `data\clean.csv`, make attempts to recreate `data\clean.csv` and `analysis.pdf`.
+If there is a dependency which is missing, and there is no recipe to make it, `make` stops with an error.
+It is a convention to have a target `all` which depends on everything and make it the first target in a `Makefile`, then the command `make` without any argument automatically creates everything.
+`Build All` from within RStudio also triggers this process.
+
+```
+all: analysis.pdf
+```
+
+![Build Pane in RStudio](Images/build-pane.png)
+
+If you have followed the above workflow a fellow researcher is three commands away from fully reproducing your analysis:
+
+```{bash, eval=FALSE}
+git clone https://github.com/sirfisher/iris.git
+cd iris
+make all
+```
+
+However, this relies on the crucial assumption that your computational environments are sufficiently compatible e.g., everything needed is installed (R + Packages) at a compatible version of the software.
 
 ### Random Number Generation with R
-
-```{bash}
-docker run --rm rocker/r-ver:3.3.0 Rscript -e "set.seed(1234);sample(1:10, 5)"
-```
-
-```{bash}
-docker run --rm rocker/r-ver:3.4.0 Rscript -e "set.seed(1234);sample(1:10, 5)"
-```
 

### 2019-08-16 rr-flow.Rmd
add section on docker

@@
 date: "8/6/2019"
+bibliography: Reproducible-Research.bib
 output: html_document
@@
 Reproducible Research has become the gold standard for scientific work in recent years.
-The following tutorial aims to outline a workflow that does guarantees not only a reasonable standard of reproducibility but also decreases the burden of maintaining software for the researcher and his colleagues.
+The following tutorial aims to outline a workflow that guarantees not only a reasonable standard of reproducibility but also decreases the burden of maintaining software for the researcher and her colleagues.
 The authors' believe that integrating tools from software engineering in the process of scientific reporting to enhance reproducibility is crucial and give guidelines for the practitioner to use them.
 These tools include software for literate programming, version control, dependency tracking, and containerization.
+If implemented, such workflow enables the researcher to transparently create, dynamicly render and puplicly share reports, which she and her colleages can trust to reproduce later, regardless of operating system or software version.
 
@@
 Luckily the minimal valuable subset needed to ensure reproducibility can be learned with little practice.
+However even if only a part of this workflow is implemented, the
 
@@
 The most precise documentation is computer code and the most productive computer code is code that runs automaticly without human interaction.
-These principles are what guides the the use of `rmarkdown`, `git`, `make`, `docker` and there interaction.
+These principles are what guides the use of `rmarkdown`, `git`, `make`, `docker` and there interaction.
 
-The first step towards reproducibility is to have a R script, that can be executed on the own computer without error, after everything within R and the files that are the output of the script deleted.
+The first step towards reproducibility is to have a R script or R markdown, that can be executed on the own computer without error.
 We assume that the reader is familar with R and has the ability to archive that for a simple analysis.
@@
 
-Now you can check everything that has happened (`git log`) and look at the exact state by supplying the id of the commit or the first few digits of it to `git checkout`:
+Now one can check everything that has happened (`git log`) and look at the exact state by supplying the id of the commit or the first few digits of it to `git checkout`:
 
@@
 This job gets comfortable with `make`, since it allows you to create (computational) recipes too (re-)create files.
+A downside of `make` is, that the installation on Windows is not straigtfowrward.
+However if `docker` (see later sections) is availible this is not an issue, since `make` ships with (almost) every `docker` image.
 
@@
 It is a convention to have a target `all` which depends on everything and make it the first target in a `Makefile`, then the command `make` without any argument automatically creates everything.
-`Build All` from within RStudio also triggers this process.
+The button `Build All` from within RStudio also triggers this process.
 
@@
 
-### Random Number Generation with R
+`docker` is a tool that allows to precisely state what computational environment an analysis relies on, in a way that can automatically be recreated on most operating systems (Windows, OS X & Linux).
+It does that, by creating a virtual computer on which a series of commands (e.g., installing software) is executed.
+The resulting state of the virtual computer is saved in what is called an "image." This image can be started and execute commands, e.g., running an R-script or `make`.
+A running instance of an image is called a container.
+An image can be transferred and executed on any machine that has docker installed.
+Regardless of the machine the container is executed on, from the perspective of a programm running inside a container, it always looks identical.
+There are essential differences to a traditional virtual machine, but thinking about a docker container as a computer running inside your computer is still a plausible mental model.
+The most important advantage over traditional virtual machines is that container are lightweight, meaning they start fast and don't need much storage.
+Docker achieves that by reusing large parts of the hosts operating system (how much is varying widely between Linux, OS X & Windows).
 
-```{bash}
+The following example show why researcher might wan't to invest time into docker.
+While the R programming language is considered largely stable and much effort is put into backwarts comptibility, sometimes even basic functions like `sample()` change their behavior from one version to another.
+
+Consider the following R-code:
+
+```{r,eval=FALSE}
+set.seed(1234)
+sample(1:10, 5)
+```
+
+The usual expectation is that this code will deliver the same pseudo random five numbers regardless of operating system or R-Version (due to `set.seed()`).
+Using docker we can start an image which contains the R-Version 3.5.0 and execute the code there.
+
+```{bash, eval=FALSE}
 docker run --rm rocker/r-ver:3.5.0 Rscript -e "set.seed(1234);sample(1:10, 5)"
@@
 
-```{bash}
+However in an image with a more recent version of R (3.6.0) this results in another sample.
+
+```{bash, eval=FALSE}
 docker run --rm rocker/r-ver:3.6.0 Rscript -e "set.seed(1234);sample(1:10, 5)"
 ```
+
+Note that this is intended behavior und the result of a [bugfix](https://bugs.r-project.org/bugzilla/show_bug.cgi?id=17494) implemented beginning of R 3.6.0.
+
+Such change will prevent many analysis from beeing replicated, if they contain e.g., multiple imputation, bootstrapping, simulations studies, graphics with random jitter, bayesian estimations, or similar technics which involve random sampling.
+However seldom does a analysis depend only on R, often a considerable number of packages is required, and they will most likly more often introduce breaking changes with updates.
+The whole endouver of reproducibility is therefore at stake, everytime a update is rolled out.
+In order to prevent that `docker` replicates the  computational environment of an analysis exactly.
+However we do not intend to advocate that software should not be updated, it should just be documented and easily replicated which version was used.
+Quite to the contrary with solutions like docker it gets easier then ever to savely update to new versions, by just bumping the R Version number of the docker image the analysis is relying on.
+
+This convinience is possible to the efforts of the [rocker project](https://github.com/rocker-org/rocker) [@boettigerIntroductionRockerDocker2017a] which provides docker images preconfigured with an R-installation of a fixed R-version.
+These rely on MRAN [@revolutionanalyticsReproducibilityUsingFixed2019], a mirror of CRAN, the repository from which R-packages are installed from, which is fixed to the last date the R-version of the image was the most recent R-version.
+Building upon these rocker-images researcher can build there own docker-images with the required R-packages.
+The rocker project provides also images which do include RStudio (`rocker/rstudio`), the `tidyverse` package (`rocker/tidyverse`) and the `rmarkdown` package with Latex (`rocker/verse`).
+Since the describes workflow relies on `rmarkdown` we suggest using the `rocker/verse` image (which also contains `rstudio` & `tidyverse`).
+These images are stored on [dockerhub](https://hub.docker.com) [@dockerinc.DockerHub2019].
+
+To extent the rocker images e.g. with other packages a Dockerfile (a File actually named `Dockerfile`) is needed, the docker image for this article is based on the following Dockerfile:
+
+```
+FROM rocker/verse:3.6.1
+RUN install2.r --error --skipinstalled\
+  pacman here pander
+WORKDIR /home/rstudio
+```
+
+The `FROM` statement specifies which docker image to use, in this case, the `rocker/verse` image which is tagged with 3.6.1 (referring to the R version 3.6.1).
+The `RUN` statement describes a command to execute, in this case, to run an R-script `install2.r` which is available on all rocker images, to install the packages `pacman`,  `here` & `pander`.
+A Dockerfile allows more than one `RUN` statement, executing any available command.
+Those `RUN` statements can install dependencies that are not an R package e.g., other programming languages like python or Matlab.
+The `WORKDIR` statement is not strictly necessary, but saves writing the working directory out every time a command is issued to docker.
+The command `docker build -t image-name` creates an image named `image-name` from the Dockerfile in the project.
+Following there are two principle ways to share a docker image, either by sharing the Dockerfile which created the image or by sharing the image itself e.g.
+through [dockerhub](https://hub.docker.com).
+While both ways garuantee a replicable computational environment, sharing the Dockerfile is more transparent and changes to it are tracked by git.
+
+There are two ways to use a rocker image. Either one starts with the `docker run` command. The first way is to run a command inside the container. These commands take the form:
+
+```{bash, eval=FALSE}
+docker run --rm -it <IMAGENAME> <COMMAND>
+```
+
+The `--rm` flag means to not save the state of the docker after the command finishes in an image.
+The `-it` flag means to accept inputs and return outputs to the terminal.
+This is especially usefull when using `R` as a command, because then you have acces to the standard R-terminal.
+
+```{bash, eval=FALSE}
+docker run --rm -it reproducible-research R
+```
+
+![R-terminal running inside docker](Images/docker-r-terminal.png)
+
+The second way is to supply no commands and to interact with the container via the webbrowser and the rstudio server instance that runs in it.
+For that you need to supply a password to log into rstudio server (`-e PASSWORD=<YOUR_PASS>`) and open a port (`-p 8787:8787`).
+
+```{bash, eval=FALSE}
+docker run -e PASSWORD=<YOUR_PASS> -p 8787:8787 image-name
+```
+
+The adress to connect to the rstudio server is your ip adress (or on linux also `localhost`) in this scheme: `<IPADRESS>:8787`.
+There is a fully functioning RStudio instance availible, that can be used exactly like one installed locally on the maschine.
+
+None of the above containers can access files on your computer, instead a folder needs to be linked explizitly:
+
+```{bash, eval=FALSE}
+docker run -v /folder/on/your/computer:/folder/in/docker 
+```
+
+The main directory for RStudio inside docker is `/home/rstudio` so an apropriate command is:
+
+```{bash, eval=FALSE}
+docker run --rm -it -e PASSWORD=<YOUR_PASS> -p 8787:8787 -v /path/to/project:/home/rstudio reproducible-research
+```
+
+![RStudio running inside Docker](Images/docker-rstudio.png)
+
+Since `docker`'s commands grow long and are tedius to type, we advocate to use some automatic way to generate them.
+Unfortunatly is the most comfortable way only availible on linux, which uses `make DOCKER=TRUE` together with a Makefile like the one used for that article.
+
+```
+projekt := $(shell basename `git rev-parse --show-toplevel`)
+current_dir := $(shell pwd)
+uid = $(shell id -u)
+
+ifeq ($(DOCKER),TRUE)
+	run:=docker run --rm --user $(uid) -v $(current_dir):/home/rstudio $(projekt)
+	current_dir=/home/rstudio
+endif
+
+all: rr-flow.html
+
+build: Dockerfile
+	sudo docker build -t $(projekt) .
+
+clean:
+	Ruby/clean.rb
+	rm -rf kitematic
+.PHONY: clean
+
+rr-flow.html: rr-flow.Rmd
+	$(run) Rscript -e 'rmarkdown::render("$(current_dir)/$<")'
+```
+
+# References

### 2019-08-20 rr-flow.Rmd
add references

@@
 
-Reproducible Research has become the gold standard for scientific work in recent years.
-The following tutorial aims to outline a workflow that guarantees not only a reasonable standard of reproducibility but also decreases the burden of maintaining software for the researcher and her colleagues.
+Reproducible Research has become the minimal standard for scientific work [@hardwicketome.DataAvailabilityReusability] in recent years.
+The following tutorial aims to outline a workflow that guarantees not only a reasonable standard of reproducibility but also decreases the burden of maintaining software and code for the researcher and her colleagues.
 The authors' believe that integrating tools from software engineering in the process of scientific reporting to enhance reproducibility is crucial and give guidelines for the practitioner to use them.
@@
 : Example: One document that includes text that describes a statistical test, code that is used to execute the statistical and the output of the code in the same document.
+
 Version Control
@@
 
-Within the community of users of the R Programming Language for each purpose a specific software solutions is clearly the most popular.
+While it is widely recognized that these tools enhance reproducibility and there are several guidelines for using them (Literate Programming: @ruleTenSimpleRules2019; Version Control: @barbaHardRoadReproducibility2016; Dependency Management: @askrenUsingMakeReproducible2016; Containerization: @clyburne-sherinComputationalReproducibilityContainers2018) Within the community of users of the R Programming Language for each purpose a specific software solutions is clearly the most popular.
 For literate programming, `rmarkdown`, for version control `git`, for dependency management `make` and for container `docker`.
@@
 Luckily the minimal valuable subset needed to ensure reproducibility can be learned with little practice.
-However even if only a part of this workflow is implemented, the
 

### 2019-08-20 rr-flow.Rmd
correct grammar in abstract

@@
 In this tutorial, we describe a principled workflow to ensure reproducible research in R-based environments.
-The workflow relies on the interplay of various open-source software tools including Rmarkdown, git, make and docker whose interplay ensures a seamless integration of version management, (APA-conformable) dynamic report generation and full cross-platform computational reproducibility.
-The workflow ensures the primary goals that 1) the reporting of analysis results is perfectly consistent with the actual analysis results (dynamic report generation), the analysis exactly reproduces at later time even if the computing and software platform is updated or changed (computational reproducibility), and 3) intermediate or post-publication changes are tracked, marked, and documented while earlier versions of both data and code remain accessible.
-While dynamic document generation is increasingly recognized as tool for reproducible analyses, we demonstrate with practical examples that dynamic documents are not sufficient to ensure computational reproducibility.
+The workflow relies on the interplay of various open-source software tools including Rmarkdown, git, make and docker whose interplay ensures seamless integration of version management, (APA-conformable) dynamic report generation and full cross-platform computational reproducibility.
+The workflow ensures the primary goals that 1) the reporting of analysis results is perfectly consistent with the actual analysis results (dynamic report generation), the analysis exactly reproduces at a later time even if the computing and software platform is updated or changed (computational reproducibility), and 3) intermediate or post-publication changes are tracked, marked, and documented while earlier versions of both data and code remain accessible.
+While dynamic document generation is increasingly recognized as a tool for reproducible analyses, we demonstrate with practical examples that dynamic documents are not sufficient to ensure computational reproducibility.
 

### 2019-08-20 rr-flow.Rmd
correct grammary and spelling

@@
 These tools include software for literate programming, version control, dependency tracking, and containerization.
-If implemented, such workflow enables the researcher to transparently create, dynamicly render and puplicly share reports, which she and her colleages can trust to reproduce later, regardless of operating system or software version.
+If implemented, such workflow enables the researcher to transparently create, dynamically render, and publicly share reports, which she and her colleagues can trust to reproduce later, regardless of operating system or software version.
 
@@
 
-While it is widely recognized that these tools enhance reproducibility and there are several guidelines for using them (Literate Programming: @ruleTenSimpleRules2019; Version Control: @barbaHardRoadReproducibility2016; Dependency Management: @askrenUsingMakeReproducible2016; Containerization: @clyburne-sherinComputationalReproducibilityContainers2018) Within the community of users of the R Programming Language for each purpose a specific software solutions is clearly the most popular.
+The scientific community widely recognizes that these tools enhance reproducibility, and provides guidelines for their use (Literate Programming: @ruleTenSimpleRules2019; Version Control: @barbaHardRoadReproducibility2016; Dependency Management: @askrenUsingMakeReproducible2016; Containerization: @clyburne-sherinComputationalReproducibilityContainers2018).
+However, only a workflow using them in unison allows confidence in reproducing a scientific report.
+Within the community of users of the R Programming Language for each purpose, a specific software solution is the most popular.
 For literate programming, `rmarkdown`, for version control `git`, for dependency management `make` and for container `docker`.
-Each of them serves a meta-scientific valuable goal (reproducibility) and increases the producitivity of the researcher.
-Each of these software solutions are extremly mighty, so in order to master them fully often years of practice are required.
-Luckily the minimal valuable subset needed to ensure reproducibility can be learned with little practice.
+Each of them serves a meta-scientific valuable goal (reproducibility) and increases the productivity of the researcher.
+Each of these software solutions is exceptionally mighty, so to master them thoroughly often years of practice are required.
+Luckily the researcher can learn the minimal valuable subset needed to ensure reproducibility with little practice.
 
-The authors believe that for sufficient chances of reproducibility, good documentation is crucial, since it anwers the question "What steps have to be taken to reproduce the results?".
-The most precise documentation is computer code and the most productive computer code is code that runs automaticly without human interaction.
-These principles are what guides the use of `rmarkdown`, `git`, `make`, `docker` and there interaction.
+The authors believe that for sufficient chances of reproducibility, proper documentation is crucial since it answers the question "What steps have to be taken to reproduce the results?".
+The most precise documentation is computer code, and the most productive computer code is code that runs automatically without human interaction.
+These principles are what guides the use of `rmarkdown`, `git`, `make`, `docker` and their interaction.
 
-The first step towards reproducibility is to have a R script or R markdown, that can be executed on the own computer without error.
-We assume that the reader is familar with R and has the ability to archive that for a simple analysis.
-A next step is to make sure that all files relevant, can be moved to another computer.
-To archive that it is important that all files are within one folder (and subfolders within it) and all paths are relative to that folder.
-A nice solution to that problem are [Rstudio Projects](https://r4ds.had.co.nz/workflow-projects.html) and/or the `here` package.
+The first step towards reproducibility is to have an R-script or R-markdown, that runs on the own computer without error.
+A process, we assume, that is familiar to the reader.
+A next step is to make sure that all files relevant for the analysis can be moved to another computer.
+To archive that it is essential that all files are within one folder (and subfolders within it) and all paths are relative to that folder.
+A friendly solution to that problem are [Rstudio Projects](https://r4ds.had.co.nz/workflow-projects.html) and the `here` package.
 
@@
 Going back to one state is as easy as finding the hash of the commit.
-This can be made even easier by tagging specific commits e.g., as "submission," "preprint," "publication." This collection of commits is called "repository," which is almost the same as your Project.
+Finding a specific hash can be made even easier by tagging specific commits, e.g., as "submission," "preprint," "publication." This collection of commits is called "repository," which is almost the same as your Project.
 
@@
 So to keep track of all changes, you only need to use `git add` & `git commit` or `git commit -a` to do both.
-These commands need to be executed in the `terminal`, which you can access from within Rstudio (`Shift + Alt + R`), but if you use Git with Rstudio, you also get a graphical user interface and quite likely you won't need the terminal more then once per project.
+These commands need to be executed in the `terminal`, which you can access from within Rstudio (`Shift + Alt + R`).
+Rstudio also offers a graphical user interface for Git and quite likely this suffice for almost every interaction with Git.
 
@@
 
-Even when you have obtained a version of the project, and you can ensure that this version is unchanged, you may not know how to reproduce the results, because it is unclear what has to be run in which order.
+Even when you have obtained a version of the Project, and you can ensure that this version is unchanged, you may not know how to reproduce the results, because it is unclear what has to be run in which order.
 This job gets comfortable with `make`, since it allows you to create (computational) recipes too (re-)create files.
-A downside of `make` is, that the installation on Windows is not straigtfowrward.
-However if `docker` (see later sections) is availible this is not an issue, since `make` ships with (almost) every `docker` image.
+A downside of `make` is, that the installation on Windows is not straightforward.
+However, if `docker` (see later sections) is available this is not an issue, since `make` ships with (almost) every `docker` image.
 
@@
 
-If you have followed the above workflow a fellow researcher is three commands away from fully reproducing your analysis:
+If you have followed the above workflow, a fellow researcher is three commands away from fully reproducing your analysis:
 
@@
 
-However, this relies on the crucial assumption that your computational environments are sufficiently compatible e.g., everything needed is installed (R + Packages) at a compatible version of the software.
+However, this relies on the crucial assumption that your computational environments are sufficiently compatible, e.g., everything needed is installed (R + Packages) at a compatible version of the software.
 
-`docker` is a tool that allows to precisely state what computational environment an analysis relies on, in a way that can automatically be recreated on most operating systems (Windows, OS X & Linux).
-It does that, by creating a virtual computer on which a series of commands (e.g., installing software) is executed.
-The resulting state of the virtual computer is saved in what is called an "image." This image can be started and execute commands, e.g., running an R-script or `make`.
+`docker` is a tool that allows to precisely state what computational environment an analysis relies on and recreates this environment automatically on most operating systems (Windows, OS X & Linux).
+`docker` does that, by setting up a virtual computer on which it executes a series of commands (e.g., installing software).
+It then saves the resulting state of the virtual computer in what is called an "image." This image can be started and execute commands, e.g., running an R-script or `make`.
 A running instance of an image is called a container.
 An image can be transferred and executed on any machine that has docker installed.
-Regardless of the machine the container is executed on, from the perspective of a programm running inside a container, it always looks identical.
+Unregarding the machine executing the container, from the perspective of a program running inside the container, it always looks identical.
 There are essential differences to a traditional virtual machine, but thinking about a docker container as a computer running inside your computer is still a plausible mental model.
@@
 
-The following example show why researcher might wan't to invest time into docker.
-While the R programming language is considered largely stable and much effort is put into backwarts comptibility, sometimes even basic functions like `sample()` change their behavior from one version to another.
+While the R programming language is considered stable and much effort is put into backward compatibility, sometimes even basic functions like `sample()` change their behavior from one version to another.
 
@@
 
-The usual expectation is that this code will deliver the same pseudo random five numbers regardless of operating system or R-Version (due to `set.seed()`).
-Using docker we can start an image which contains the R-Version 3.5.0 and execute the code there.
+The usual expectation is that this code delivers the same pseudo-random five numbers regardless of the operating system or R-Version (due to `set.seed()`).
+Using docker, we can start an image which contains the R-Version 3.5.0, and execute the code there.
 
@@
 
-However in an image with a more recent version of R (3.6.0) this results in another sample.
+Surprisingly in an image with a more recent version of R (3.6.0), this results in another sample:
 
@@
 
-Note that this is intended behavior und the result of a [bugfix](https://bugs.r-project.org/bugzilla/show_bug.cgi?id=17494) implemented beginning of R 3.6.0.
+Note that this is the intended behavior and the result of a [bugfix](https://bugs.r-project.org/bugzilla/show_bug.cgi?id=17494) implemented from R 3.6.0 upwards.
 
-Such change will prevent many analysis from beeing replicated, if they contain e.g., multiple imputation, bootstrapping, simulations studies, graphics with random jitter, bayesian estimations, or similar technics which involve random sampling.
-However seldom does a analysis depend only on R, often a considerable number of packages is required, and they will most likly more often introduce breaking changes with updates.
-The whole endouver of reproducibility is therefore at stake, everytime a update is rolled out.
-In order to prevent that `docker` replicates the  computational environment of an analysis exactly.
-However we do not intend to advocate that software should not be updated, it should just be documented and easily replicated which version was used.
-Quite to the contrary with solutions like docker it gets easier then ever to savely update to new versions, by just bumping the R Version number of the docker image the analysis is relying on.
+Such change prevents many reports from beeing replicated, if they contain, e.g., multiple imputations, bootstrapping, simulations studies, graphics with random jitter, bayesian estimations, or similar technics which involve random sampling.
+However seldom does an analysis depend only on R, often a considerable number of packages is required, and they will most likely more often introduce breaking changes with updates.
+The whole endeavor of reproducibility is therefore at stake, every time an update is rolled out.
+To prevent that `docker` replicates the computational environment of analysis exactly.
+We do not intend to advocate that software should not be updated; the used software version should just be documented and easily replicated.
+Quite to the contrary with solutions like docker, it gets more comfortable than ever to safely update to new versions, by just bumping the R Version number of the docker image.
 
-This convinience is possible to the efforts of the [rocker project](https://github.com/rocker-org/rocker) [@boettigerIntroductionRockerDocker2017a] which provides docker images preconfigured with an R-installation of a fixed R-version.
-These rely on MRAN [@revolutionanalyticsReproducibilityUsingFixed2019], a mirror of CRAN, the repository from which R-packages are installed from, which is fixed to the last date the R-version of the image was the most recent R-version.
-Building upon these rocker-images researcher can build there own docker-images with the required R-packages.
-The rocker project provides also images which do include RStudio (`rocker/rstudio`), the `tidyverse` package (`rocker/tidyverse`) and the `rmarkdown` package with Latex (`rocker/verse`).
+This convenience is possible to the efforts of the [rocker project](https://github.com/rocker-org/rocker) [@boettigerIntroductionRockerDocker2017a] which provides docker images preconfigured with an R-installation of a fixed R-version.
+These rely on MRAN [@revolutionanalyticsReproducibilityUsingFixed2019], a repository for R-Packages, which is fixed to the last date the R-version of the image was the most recent.
+Building upon these rocker-images researchers can build there own docker-images with the required R-packages.
+The rocker project also provides images which do include RStudio (`rocker/rstudio`), the `tidyverse` package (`rocker/tidyverse`) and the `rmarkdown` package with Latex (`rocker/verse`).
 Since the describes workflow relies on `rmarkdown` we suggest using the `rocker/verse` image (which also contains `rstudio` & `tidyverse`).
@@
 
-To extent the rocker images e.g. with other packages a Dockerfile (a File actually named `Dockerfile`) is needed, the docker image for this article is based on the following Dockerfile:
+The basic rocker images are extendable by a Dockerfile, e.g., the basis for docker image for this article is the following Dockerfile:
 
@@
 
-The `FROM` statement specifies which docker image to use, in this case, the `rocker/verse` image which is tagged with 3.6.1 (referring to the R version 3.6.1).
+The `FROM` statement specifies which docker image to use, in this case, the `rocker/verse` image with the tag 3.6.1 (referring to the R version 3.6.1).
 The `RUN` statement describes a command to execute, in this case, to run an R-script `install2.r` which is available on all rocker images, to install the packages `pacman`,  `here` & `pander`.
 A Dockerfile allows more than one `RUN` statement, executing any available command.
-Those `RUN` statements can install dependencies that are not an R package e.g., other programming languages like python or Matlab.
-The `WORKDIR` statement is not strictly necessary, but saves writing the working directory out every time a command is issued to docker.
+Those `RUN` statements can install dependencies that are not an R package, e.g., other programming languages like python or Matlab.
+The `WORKDIR` statement is not strictly necessary but saves time spent writing the working directory.
 The command `docker build -t image-name` creates an image named `image-name` from the Dockerfile in the project.
-Following there are two principle ways to share a docker image, either by sharing the Dockerfile which created the image or by sharing the image itself e.g.
-through [dockerhub](https://hub.docker.com).
-While both ways garuantee a replicable computational environment, sharing the Dockerfile is more transparent and changes to it are tracked by git.
+Following there are two principal ways to share a docker image, either by sharing the Dockerfile which created the image or by sharing the image itself e.g., through [dockerhub](https://hub.docker.com).
+While both ways guarantee a replicable computational environment, sharing the Dockerfile is more transparent, and git tracks changes to it.
 
@@
 The `-it` flag means to accept inputs and return outputs to the terminal.
-This is especially usefull when using `R` as a command, because then you have acces to the standard R-terminal.
+Interacting with docker this way is especially useful when using `R` as a command because then you have access to the standard R-terminal.
 

### 2019-08-20 rr-flow.Rmd
resolve #6

@@
 
-This folder, where everything resides that you need for analysis, is referred to as a Projekt.
+This folder, where everything resides that you need for analysis (code & data), is referred to as a Projekt.
 Working with Projects is exceptionally comfortable with RStudio an integrated development environment (IDE) for R.
+It is usefull to organize this Project in way that strictly segregates (raw) data and code by placing them in directories called `data` and `R` [see Section 4 @marwickPackagingDataAnalytical2018].
 

### 2019-08-20 abstract.Rmd
abstract in extra file closes #9

--- 2019-08-20 rr-flow.Rmd
@@
----
-title: "A Reproducible Workflow with RMarkdown, Git, Make & Docker"
-shorttitle: "Reproducible Workflow"
-author: "Andreas Brandmaier & Aaron Peikert"
-date: "8/6/2019"
-bibliography: Reproducible-Research.bib
-output: html_document
----
-
-```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
-if(!require("pacman"))install.packages("pacman")
-pacman::p_load("here", "tidyverse")
-```
-
 In this tutorial, we describe a principled workflow to ensure reproducible research in R-based environments.
@@
 If implemented, such workflow enables the researcher to transparently create, dynamically render, and publicly share reports, which she and her colleagues can trust to reproduce later, regardless of operating system or software version.
-
-Literate Programming
-: Interwoven text and computer code that together forms a human-readable document.
-: Example: One document that includes text that describes a statistical test, code that is used to execute the statistical and the output of the code in the same document.
-
-Version Control
-: The management of change of documents/code.
-: Example: Changes from the manuscript text upon request by the reviewer are identifiable.
-
-Dependency Tracking
-: How different parts of a manuscript/analysis relate and how these dependencies resolve via execution of code.
-: Example: The manuscript depends on the statistical analysis, which depends on the data.
-
-Containerization
-: A virtual computational environment that does encapsulate all needed software. More formally called operating-system-level virtualization.
-: Example: A statistical analysis depends on the used statistical software and a virtual machine bundles them together.
-
-The scientific community widely recognizes that these tools enhance reproducibility, and provides guidelines for their use (Literate Programming: @ruleTenSimpleRules2019; Version Control: @barbaHardRoadReproducibility2016; Dependency Management: @askrenUsingMakeReproducible2016; Containerization: @clyburne-sherinComputationalReproducibilityContainers2018).
-However, only a workflow using them in unison allows confidence in reproducing a scientific report.
-Within the community of users of the R Programming Language for each purpose, a specific software solution is the most popular.
-For literate programming, `rmarkdown`, for version control `git`, for dependency management `make` and for container `docker`.
-Each of them serves a meta-scientific valuable goal (reproducibility) and increases the productivity of the researcher.
-Each of these software solutions is exceptionally mighty, so to master them thoroughly often years of practice are required.
-Luckily the researcher can learn the minimal valuable subset needed to ensure reproducibility with little practice.
-
-The authors believe that for sufficient chances of reproducibility, proper documentation is crucial since it answers the question "What steps have to be taken to reproduce the results?".
-The most precise documentation is computer code, and the most productive computer code is code that runs automatically without human interaction.
-These principles are what guides the use of `rmarkdown`, `git`, `make`, `docker` and their interaction.
-
-The first step towards reproducibility is to have an R-script or R-markdown, that runs on the own computer without error.
-A process, we assume, that is familiar to the reader.
-A next step is to make sure that all files relevant for the analysis can be moved to another computer.
-To archive that it is essential that all files are within one folder (and subfolders within it) and all paths are relative to that folder.
-A friendly solution to that problem are [Rstudio Projects](https://r4ds.had.co.nz/workflow-projects.html) and the `here` package.
-
-```{r, eval=FALSE}
-# BAD
-iris <- read.csv("/home/aaron/Documents/reproducible-research/data/iris.csv")
-# GOOD
-iris <- read.csv("data/iris.csv")
-# BETTER
-iris <- read.csv(here("data", "iris.csv"))
-```
-
-This folder, where everything resides that you need for analysis, is referred to as a Projekt.
-Working with Projects is exceptionally comfortable with RStudio an integrated development environment (IDE) for R.
-
-Reproducibility at a basic level assumes that results are the same if neither script nor data have changed.
-It is often not trivial to find out whether anything has changed and if so to "go back in time." `git` enables you to do both.
-A good mental model for `git` is that it takes a sequence of snapshots of all files it is supposed to track.
-In the language of git, these snapshots are "commits".
-A commit represents a complete copy of the state of the files when added to a commit.
-Each commit has a unique identifier (a hash).
-Going back to one state is as easy as finding the hash of the commit.
-Finding a specific hash can be made even easier by tagging specific commits, e.g., as "submission," "preprint," "publication." This collection of commits is called "repository," which is almost the same as your Project.
-
-A typical git workflow looks like that:
-
-```{bash, eval=FALSE}
-git init # to initialize git
-git add /data/iris.csv /R/analysis.R # track specific files/changes
-git commit -m "add data & analysis" # take snapshot
-# new data comes in
-git commit -a -m "complete data colections" # add (-a) and comit all changes
-```
-
-So to keep track of all changes, you only need to use `git add` & `git commit` or `git commit -a` to do both.
-These commands need to be executed in the `terminal`, which you can access from within Rstudio (`Shift + Alt + R`).
-Rstudio also offers a graphical user interface for Git and quite likely this suffice for almost every interaction with Git.
-
-![Git Pane in RStudio](Images/git-pane.png)
-
-Now one can check everything that has happened (`git log`) and look at the exact state by supplying the id of the commit or the first few digits of it to `git checkout`:
-
-```{bash, eval = FALSE}
-git log
-git checkout 77db06f78e
-```
-
-Git also makes it especially easy to share and collaborate on a project with another researcher.
-A popular service for sharing via git is [Github](https://github.com).
-Just sharing Git repositories with the public is always free, private repositories (only visible to persons you invite) are [free for researcher](https://help.github.com/en/articles/applying-for-an-educator-or-researcher-discount).
-After creating a user account, you can create a new repository where Github advises you how to upload your repository from the terminal e.g., for this repository:
-
-```{bash, eval=FALSE}
-git remote add origin git@github.com:aaronpeikert/reproducible-research.git
-git push -u origin master
-```
-
-`git push` or the green up arrow in the git pane uploads local updates. To download this git repo to another computer type into the terminal:
-
-```{bash, eval=FALSE}
-git clone git@github.com:aaronpeikert/reproducible-research.git
-```
-
-Git & Github can do even more to help you to collaborate with your fellow researcher, but this is beyond the already broad scope of this tutorial.
-Another benefit of using Git & Github is that experimentation is highly encouraged since you can go back to any state quickly and even when you lose access to the file on your computer, everything is backed up on the Github servers.
-
-Even when you have obtained a version of the Project, and you can ensure that this version is unchanged, you may not know how to reproduce the results, because it is unclear what has to be run in which order.
-This job gets comfortable with `make`, since it allows you to create (computational) recipes too (re-)create files.
-A downside of `make` is, that the installation on Windows is not straightforward.
-However, if `docker` (see later sections) is available this is not an issue, since `make` ships with (almost) every `docker` image.
-
-A `Makefile` obeys the following scheme:
-
-```
-target: dependency1 dependency2
-  command-to-create-target
-```
-
-A typical `Makefile` might look like:
-
-```
-analysis.pdf: data/clean.csv
-  Rscript -e 'rmarkdown::render("analysis.Rmd")'
-
-data/clean.csv: R/clean.R data/iris.csv
-  Rscript -e 'source("R/clean.R")'
-```
-
-The first line reads: to create `analysis.pdf` the `analysis.Rmd` needs to be rendered, which depends on `data/clean.csv`.
-This dependency is itself a target.
-To create `data/clean.csv`, `R/clean.R` and `data/iris.csv` are needed.
-If you type `make analysis.pdf` `make` first checks whether the dependencies do exist and if not whether or it knows how to create them.
-So if `data/clean.csv` does not exist, `make` creates it.
-The same thing happens if one of the dependencies of a target is newer then the target itself, then updates everything that directly or indirectly depends on it.
-So if `data\iris.csv` is newer then `data\clean.csv`, make attempts to recreate `data\clean.csv` and `analysis.pdf`.
-If there is a dependency which is missing, and there is no recipe to make it, `make` stops with an error.
-It is a convention to have a target `all` which depends on everything and make it the first target in a `Makefile`, then the command `make` without any argument automatically creates everything.
-The button `Build All` from within RStudio also triggers this process.
-
-```
-all: analysis.pdf
-```
-
-![Build Pane in RStudio](Images/build-pane.png)
-
-If you have followed the above workflow, a fellow researcher is three commands away from fully reproducing your analysis:
-
-```{bash, eval=FALSE}
-git clone https://github.com/sirfisher/iris.git
-cd iris
-make all
-```
-
-However, this relies on the crucial assumption that your computational environments are sufficiently compatible, e.g., everything needed is installed (R + Packages) at a compatible version of the software.
-
-`docker` is a tool that allows to precisely state what computational environment an analysis relies on and recreates this environment automatically on most operating systems (Windows, OS X & Linux).
-`docker` does that, by setting up a virtual computer on which it executes a series of commands (e.g., installing software).
-It then saves the resulting state of the virtual computer in what is called an "image." This image can be started and execute commands, e.g., running an R-script or `make`.
-A running instance of an image is called a container.
-An image can be transferred and executed on any machine that has docker installed.
-Unregarding the machine executing the container, from the perspective of a program running inside the container, it always looks identical.
-There are essential differences to a traditional virtual machine, but thinking about a docker container as a computer running inside your computer is still a plausible mental model.
-The most important advantage over traditional virtual machines is that container are lightweight, meaning they start fast and don't need much storage.
-Docker achieves that by reusing large parts of the hosts operating system (how much is varying widely between Linux, OS X & Windows).
-
-While the R programming language is considered stable and much effort is put into backward compatibility, sometimes even basic functions like `sample()` change their behavior from one version to another.
-
-Consider the following R-code:
-
-```{r,eval=FALSE}
-set.seed(1234)
-sample(1:10, 5)
-```
-
-The usual expectation is that this code delivers the same pseudo-random five numbers regardless of the operating system or R-Version (due to `set.seed()`).
-Using docker, we can start an image which contains the R-Version 3.5.0, and execute the code there.
-
-```{bash, eval=FALSE}
-docker run --rm rocker/r-ver:3.5.0 Rscript -e "set.seed(1234);sample(1:10, 5)"
-```
-
-Surprisingly in an image with a more recent version of R (3.6.0), this results in another sample:
-
-```{bash, eval=FALSE}
-docker run --rm rocker/r-ver:3.6.0 Rscript -e "set.seed(1234);sample(1:10, 5)"
-```
-
-Note that this is the intended behavior and the result of a [bugfix](https://bugs.r-project.org/bugzilla/show_bug.cgi?id=17494) implemented from R 3.6.0 upwards.
-
-Such change prevents many reports from beeing replicated, if they contain, e.g., multiple imputations, bootstrapping, simulations studies, graphics with random jitter, bayesian estimations, or similar technics which involve random sampling.
-However seldom does an analysis depend only on R, often a considerable number of packages is required, and they will most likely more often introduce breaking changes with updates.
-The whole endeavor of reproducibility is therefore at stake, every time an update is rolled out.
-To prevent that `docker` replicates the computational environment of analysis exactly.
-We do not intend to advocate that software should not be updated; the used software version should just be documented and easily replicated.
-Quite to the contrary with solutions like docker, it gets more comfortable than ever to safely update to new versions, by just bumping the R Version number of the docker image.
-
-This convenience is possible to the efforts of the [rocker project](https://github.com/rocker-org/rocker) [@boettigerIntroductionRockerDocker2017a] which provides docker images preconfigured with an R-installation of a fixed R-version.
-These rely on MRAN [@revolutionanalyticsReproducibilityUsingFixed2019], a repository for R-Packages, which is fixed to the last date the R-version of the image was the most recent.
-Building upon these rocker-images researchers can build there own docker-images with the required R-packages.
-The rocker project also provides images which do include RStudio (`rocker/rstudio`), the `tidyverse` package (`rocker/tidyverse`) and the `rmarkdown` package with Latex (`rocker/verse`).
-Since the describes workflow relies on `rmarkdown` we suggest using the `rocker/verse` image (which also contains `rstudio` & `tidyverse`).
-These images are stored on [dockerhub](https://hub.docker.com) [@dockerinc.DockerHub2019].
-
-The basic rocker images are extendable by a Dockerfile, e.g., the basis for docker image for this article is the following Dockerfile:
-
-```
-FROM rocker/verse:3.6.1
-RUN install2.r --error --skipinstalled\
-  pacman here pander
-WORKDIR /home/rstudio
-```
-
-The `FROM` statement specifies which docker image to use, in this case, the `rocker/verse` image with the tag 3.6.1 (referring to the R version 3.6.1).
-The `RUN` statement describes a command to execute, in this case, to run an R-script `install2.r` which is available on all rocker images, to install the packages `pacman`,  `here` & `pander`.
-A Dockerfile allows more than one `RUN` statement, executing any available command.
-Those `RUN` statements can install dependencies that are not an R package, e.g., other programming languages like python or Matlab.
-The `WORKDIR` statement is not strictly necessary but saves time spent writing the working directory.
-The command `docker build -t image-name` creates an image named `image-name` from the Dockerfile in the project.
-Following there are two principal ways to share a docker image, either by sharing the Dockerfile which created the image or by sharing the image itself e.g., through [dockerhub](https://hub.docker.com).
-While both ways guarantee a replicable computational environment, sharing the Dockerfile is more transparent, and git tracks changes to it.
-
-There are two ways to use a rocker image. Either one starts with the `docker run` command. The first way is to run a command inside the container. These commands take the form:
-
-```{bash, eval=FALSE}
-docker run --rm -it <IMAGENAME> <COMMAND>
-```
-
-The `--rm` flag means to not save the state of the docker after the command finishes in an image.
-The `-it` flag means to accept inputs and return outputs to the terminal.
-Interacting with docker this way is especially useful when using `R` as a command because then you have access to the standard R-terminal.
-
-```{bash, eval=FALSE}
-docker run --rm -it reproducible-research R
-```
-
-![R-terminal running inside docker](Images/docker-r-terminal.png)
-
-The second way is to supply no commands and to interact with the container via the webbrowser and the rstudio server instance that runs in it.
-For that you need to supply a password to log into rstudio server (`-e PASSWORD=<YOUR_PASS>`) and open a port (`-p 8787:8787`).
-
-```{bash, eval=FALSE}
-docker run -e PASSWORD=<YOUR_PASS> -p 8787:8787 image-name
-```
-
-The adress to connect to the rstudio server is your ip adress (or on linux also `localhost`) in this scheme: `<IPADRESS>:8787`.
-There is a fully functioning RStudio instance availible, that can be used exactly like one installed locally on the maschine.
-
-None of the above containers can access files on your computer, instead a folder needs to be linked explizitly:
-
-```{bash, eval=FALSE}
-docker run -v /folder/on/your/computer:/folder/in/docker 
-```
-
-The main directory for RStudio inside docker is `/home/rstudio` so an apropriate command is:
-
-```{bash, eval=FALSE}
-docker run --rm -it -e PASSWORD=<YOUR_PASS> -p 8787:8787 -v /path/to/project:/home/rstudio reproducible-research
-```
-
-![RStudio running inside Docker](Images/docker-rstudio.png)
-
-Since `docker`'s commands grow long and are tedius to type, we advocate to use some automatic way to generate them.
-Unfortunatly is the most comfortable way only availible on linux, which uses `make DOCKER=TRUE` together with a Makefile like the one used for that article.
-
-```
-projekt := $(shell basename `git rev-parse --show-toplevel`)
-current_dir := $(shell pwd)
-uid = $(shell id -u)
-
-ifeq ($(DOCKER),TRUE)
-	run:=docker run --rm --user $(uid) -v $(current_dir):/home/rstudio $(projekt)
-	current_dir=/home/rstudio
-endif
-
-all: rr-flow.html
-
-build: Dockerfile
-	sudo docker build -t $(projekt) .
-
-clean:
-	Ruby/clean.rb
-	rm -rf kitematic
-.PHONY: clean
-
-rr-flow.html: rr-flow.Rmd
-	$(run) Rscript -e 'rmarkdown::render("$(current_dir)/$<")'
-```
-
-# References

### 2019-08-20 rr-flow.Rmd
abstract in extra file closes #9

@@
 
-In this tutorial, we describe a principled workflow to ensure reproducible research in R-based environments.
-The workflow relies on the interplay of various open-source software tools including Rmarkdown, git, make and docker whose interplay ensures seamless integration of version management, (APA-conformable) dynamic report generation and full cross-platform computational reproducibility.
-The workflow ensures the primary goals that 1) the reporting of analysis results is perfectly consistent with the actual analysis results (dynamic report generation), the analysis exactly reproduces at a later time even if the computing and software platform is updated or changed (computational reproducibility), and 3) intermediate or post-publication changes are tracked, marked, and documented while earlier versions of both data and code remain accessible.
-While dynamic document generation is increasingly recognized as a tool for reproducible analyses, we demonstrate with practical examples that dynamic documents are not sufficient to ensure computational reproducibility.
-
-Reproducible Research has become the minimal standard for scientific work [@hardwicketome.DataAvailabilityReusability] in recent years.
-The following tutorial aims to outline a workflow that guarantees not only a reasonable standard of reproducibility but also decreases the burden of maintaining software and code for the researcher and her colleagues.
-The authors' believe that integrating tools from software engineering in the process of scientific reporting to enhance reproducibility is crucial and give guidelines for the practitioner to use them.
-These tools include software for literate programming, version control, dependency tracking, and containerization.
-If implemented, such workflow enables the researcher to transparently create, dynamically render, and publicly share reports, which she and her colleagues can trust to reproduce later, regardless of operating system or software version.
+```{r abstract, child = 'abstract.Rmd'}
+```
 

### 2019-08-20 rr-flow.Rmd
added ideas on checksums added ideas on licensing in summary

@@
 ---
-title: "A Reproducible Workflow with RMarkdown, Git, Make & Docker"
+title: "A Reproducible Data Analysis Workflow with RMarkdown, Git, Make & Docker"
 shorttitle: "Reproducible Workflow"
-author: "Andreas Brandmaier & Aaron Peikert"
+author: "Aaron Peikert & Andreas M. Brandmaier"
 date: "8/6/2019"
@@
 ```
+
+! TODO - Unterschied zwischen computational, empirical, and statistical reproducibility (cf. Stodden et al., 2014)
 
@@
 Each of them serves a meta-scientific valuable goal (reproducibility) and increases the productivity of the researcher.
-Each of these software solutions is exceptionally mighty, so to master them thoroughly often years of practice are required.
+Each of these software solutions is highly flexible and powerful, so to master them thoroughly often years of practice are required.
 Luckily the researcher can learn the minimal valuable subset needed to ensure reproducibility with little practice.
 
-The authors believe that for sufficient chances of reproducibility, proper documentation is crucial since it answers the question "What steps have to be taken to reproduce the results?".
+We believe that for sufficient chances of reproducibility, proper documentation is crucial since it answers the question "What steps have to be taken to reproduce the results?".
 The most precise documentation is computer code, and the most productive computer code is code that runs automatically without human interaction.
@@
 
-This folder, where everything resides that you need for analysis (code & data), is referred to as a Projekt.
+Sometimes external requirements mean that the data cannot be stored and shared with the scripts.
+In most of the cases we have seen, these are either space constraints or privacy considerations.
+In these cases, of course, unrestricted reproducibility is not guaranteed.
+If this way of dividing data and scripts into different databases is unavoidable, we recommend that you crosscheck using checksums before loading data.
+To do this, a checksum must be created and filed at the time of the original analysis, and at the time of reproduction the current checksum must be compared with the filed checksum.
+
+! TODO - Beispiel einbauen?
+
+This folder, where everything resides that you need for analysis (code and data), is referred to as a Project.
 Working with Projects is exceptionally comfortable with RStudio an integrated development environment (IDE) for R.
@@
 
-Git & Github can do even more to help you to collaborate with your fellow researcher, but this is beyond the already broad scope of this tutorial.
-Another benefit of using Git & Github is that experimentation is highly encouraged since you can go back to any state quickly and even when you lose access to the file on your computer, everything is backed up on the Github servers.
+Git and Github can do even more to help you to collaborate with your fellow researcher, such as formal means to track issues and their status (open/closed/resolved) and further means to managing and merging multiple, parallel versions of code (such as branches, pull requests, merges) but this is beyond the already broad scope of this tutorial.
+Another benefit of using Git and Github is that experimentation is highly encouraged since you can go back to any state quickly and even when you lose access to the file on your computer, everything is backed up on the Github servers.
 
@@
 
-Note that this is the intended behavior and the result of a [bugfix](https://bugs.r-project.org/bugzilla/show_bug.cgi?id=17494) implemented from R 3.6.0 upwards.
+Note that this is the intended behavior and the result of a [bugfix](https://bugs.r-project.org/bugzilla/show_bug.cgi?id=17494) in the random number generator implemented from R 3.6.0 upwards.
 
@@
 
+# Summary
+
+! TODO - Summary of all this and all benefits
+
+## Related approaches
+
+! TODO
+
+## Sharing reproducible workflows
+
+How to best share a reproducible workflow?
+We believe that, ideally, a service provider should be found that guarantuees permanent and reliable hosting of the reproducible workflow, such as the Open Science Framework (citation needed) or PsychData (citation needed).
+Second, to ensure that other users are legally able to benefit from the shared materials, authors must choose an appropriate license format.
+Typically, there is currently no single license that works for code, data, and media (such as text or figures).
+We encourage authors to choose appropriate license forms that do not hinder others from freely downloading, using, and moifying the shared workflows and materials while, at the same time, ensuring recognition for the time and effort invested in creating the workflow in the first place.
+In our experience, the Creative Commons license (CC-BY) is often appropriate for sharing texts, rmarkdown files, generated figures, and other media whereas scripts and any other computer code is often best shared according to the MIT license.
+Both licenses assure maximal freedom for future users while requiring the attribution of the original authors in derivative work.
+
 # References

### 2019-08-20 README.Rmd

---
output:
  md_document:
    variant: gfm
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(here)
```

This is the accompanying GitHub repository to a work in progress paper, its abstract reads:

```{r abstract, child = 'abstract.Rmd'}
```

# Compile

## Usual Way

Requires: `Git`, `RStudio`, `pandoc`, `pandoc-citeproc` & `rmarkdown`.

Open RStudio -> File -> New Project -> Version Control -> Git

Insert:

```
https://github.com/aaronpeikert/reproducible-research.git
```

Open `rr-flow.Rmd` click on `Knit`.

## Using Workflow

Does not require R or RStudio, but `make` & `docker`.

Execute in Terminal:

```{bash, eval=FALSE}
git clone https://github.com/aaronpeikert/reproducible-research.git
cd reproducible-research
make build
make all DOCKER=TRUE
```

# Session Info

```{r}
sessioninfo::session_info()
```

### 2019-08-21 rr-flow.Rmd
supply metadata to papaja

@@
 ---
-title: "A Reproducible Data Analysis Workflow with RMarkdown, Git, Make & Docker"
-shorttitle: "Reproducible Workflow"
-author: "Aaron Peikert & Andreas M. Brandmaier"
-date: "8/6/2019"
-bibliography: Reproducible-Research.bib
-output: html_document
+title             : "A Reproducible Data Analysis Workflow with RMarkdown, Git, Make & Docker"
+shorttitle        : "Reproducible Data Analysis Workflow"
+
+author: 
+  - name          : "Andreas Brandmaier"
+    affiliation   : "1"
+    corresponding : yes    # Define only one corresponding author
+    address       : "Max Planck Institute for Human Development, Lentzeallee 94, 14195 Berlin, Germany"
+    email         : "brandmaier@mpib-berlin.mpg.de"
+  - name          : "Aaron Peikert"
+    affiliation   : "1,2"
+
+affiliation:
+  - id            : "1"
+    institution   : "Center for Lifespan Psychology, Max Planck Institute for Human Development, Berlin, Germany"
+  - id            : "2"
+    institution   : "Humboldt-Universität zu Berlin"
+
+authornote: |
+
+abstract: "`r readr::read_file('abstract.Rmd')`"
+
+keywords          : "keywords"
+wordcount         : "X"
+
+bibliography      : ["Reproducible-Research.bib"]
+
+floatsintext      : yes
+figurelist        : yes
+tablelist         : no
+footnotelist      : no
+linenumbers       : no
+mask              : no
+draft             : no
+
+documentclass     : "apa6"
+classoption       : "man"
+output            : papaja::apa6_pdf
 ---
@@
 pacman::p_load("here", "tidyverse")
-```
-
-```{r abstract, child = 'abstract.Rmd'}
 ```
@@
 
+\newpage
+
 # References
+
+\begingroup
+\setlength{\parindent}{-0.5in}
+\setlength{\leftskip}{0.5in}
+
+<div id = "refs"></div>
+\endgroup

### 2019-08-23 abstract.Rmd
revised and merged abstract

@@
-In this tutorial, we describe a principled workflow to ensure reproducible research in R-based environments.
-The workflow relies on the interplay of various open-source software tools including Rmarkdown, git, make and docker whose interplay ensures seamless integration of version management, (APA-conformable) dynamic report generation and full cross-platform computational reproducibility.
-The workflow ensures the primary goals that 1) the reporting of analysis results is perfectly consistent with the actual analysis results (dynamic report generation), the analysis exactly reproduces at a later time even if the computing and software platform is updated or changed (computational reproducibility), and 3) intermediate or post-publication changes are tracked, marked, and documented while earlier versions of both data and code remain accessible.
-While dynamic document generation is increasingly recognized as a tool for reproducible analyses, we demonstrate with practical examples that dynamic documents are not sufficient to ensure computational reproducibility.
-
-Reproducible Research has become the minimal standard for scientific work [@hardwicketome.DataAvailabilityReusability] in recent years.
-The following tutorial aims to outline a workflow that guarantees not only a reasonable standard of reproducibility but also decreases the burden of maintaining software and code for the researcher and her colleagues.
-The authors' believe that integrating tools from software engineering in the process of scientific reporting to enhance reproducibility is crucial and give guidelines for the practitioner to use them.
-These tools include software for literate programming, version control, dependency tracking, and containerization.
-If implemented, such workflow enables the researcher to transparently create, dynamically render, and publicly share reports, which she and her colleagues can trust to reproduce later, regardless of operating system or software version.
+In this tutorial, we describe a principled workflow to ensure long-term reproducibility of R-based data analyses.
+The workflow is based on established tools and practices from software engineering.
+It combines the benefits of various open-source software tools including Rmarkdown, git, make and docker whose interplay ensures seamless integration of version management, (APA-conformable) dynamic report generation and full cross-platform and long-term computational reproducibility.
+The workflow ensures meeting the primary goals that 1) the reporting of analysis results is perfectly consistent with the actual analysis results (dynamic report generation), the analysis exactly reproduces at a later time even if the computing and software platform is updated or changed (computational reproducibility), and 3) changes at any time (during development and post-publication) are tracked, tagged, and documented while earlier versions of both data and code remain accessible.
+While dynamic document generation is increasingly recognized as a tool to ensure reproducibility, we demonstrate with practical examples that dynamic documents alone are not sufficient to ensure long-term computational reproducibility.

### 2019-08-23 abstract.Rmd
remove passive voice #26

@@
 In this tutorial, we describe a principled workflow to ensure long-term reproducibility of R-based data analyses.
-The workflow is based on established tools and practices from software engineering.
+The workflow utilizes established tools and practices from software engineering.
 It combines the benefits of various open-source software tools including Rmarkdown, git, make and docker whose interplay ensures seamless integration of version management, (APA-conformable) dynamic report generation and full cross-platform and long-term computational reproducibility.
 The workflow ensures meeting the primary goals that 1) the reporting of analysis results is perfectly consistent with the actual analysis results (dynamic report generation), the analysis exactly reproduces at a later time even if the computing and software platform is updated or changed (computational reproducibility), and 3) changes at any time (during development and post-publication) are tracked, tagged, and documented while earlier versions of both data and code remain accessible.
-While dynamic document generation is increasingly recognized as a tool to ensure reproducibility, we demonstrate with practical examples that dynamic documents alone are not sufficient to ensure long-term computational reproducibility.
+While the research community recognizes dynamic document generation as a tool to ensure reproducibility, we demonstrate with practical examples that dynamic documents alone are not sufficient to ensure long-term computational reproducibility.

### 2019-08-23 rr-flow.Rmd
swapped order of authors added hash example

@@
 author: 
+  - name          : "Aaron Peikert"
+    affiliation   : "1,2"
   - name          : "Andreas Brandmaier"
-    affiliation   : "1"
+    affiliation   : "1,3"
     corresponding : yes    # Define only one corresponding author
@@
     email         : "brandmaier@mpib-berlin.mpg.de"
-  - name          : "Aaron Peikert"
-    affiliation   : "1,2"
 
@@
     institution   : "Humboldt-Universität zu Berlin"
+  - id			  : "3"
+  	institution   : "Max Planck UCL Centre for Computational Psychiatry and Ageing Research, Berlin, Germany & London, U.K."
 
@@
 The most precise documentation is computer code, and the most productive computer code is code that runs automatically without human interaction.
-These principles are what guides the use of `rmarkdown`, `git`, `make`, `docker` and their interaction.
+These principles are what guides the use of `rmarkdown`, `git`, `make`, `docker`.
 
-The first step towards reproducibility is to have an R-script or R-markdown, that runs on the own computer without error.
+The first step towards long-term reproducibility is to have an R-script or R-markdown, that runs on a local computer without error.
 A process, we assume, that is familiar to the reader.
 A next step is to make sure that all files relevant for the analysis can be moved to another computer.
-To archive that it is essential that all files are within one folder (and subfolders within it) and all paths are relative to that folder.
-A friendly solution to that problem are [Rstudio Projects](https://r4ds.had.co.nz/workflow-projects.html) and the `here` package.
+To this end, it is essential that all files are within one folder (or, subfolders within it) and all paths are relative to that folder.
+A robust solution to the problem of making sure that file access does not break across computing platforms are [Rstudio Projects](https://r4ds.had.co.nz/workflow-projects.html) and the `here` package.
 
@@
 
-Sometimes external requirements mean that the data cannot be stored and shared with the scripts.
+This folder, where everything resides that you need for analysis (code and data), is referred to as a Project.
+Working with Projects is exceptionally comfortable with RStudio an integrated development environment (IDE) for R.
+It is usefull to organize this Project in way that strictly segregates (raw) data and code by placing them in directories called `data` and `R` [see Section 4 @marwickPackagingDataAnalytical2018].
+
+Sometimes external requirements mean that the data cannot be stored and shared together with the scripts.
 In most of the cases we have seen, these are either space constraints or privacy considerations.
@@
 If this way of dividing data and scripts into different databases is unavoidable, we recommend that you crosscheck using checksums before loading data.
-To do this, a checksum must be created and filed at the time of the original analysis, and at the time of reproduction the current checksum must be compared with the filed checksum.
+To do this, a checksum must be created and filed at the time of the original analysis, and at the time of reproduction the current checksum must be compared with the stored checksum (also called a "hash", e.g., using the functions provided in package `digest`).
 
-! TODO - Beispiel einbauen?
-
-This folder, where everything resides that you need for analysis (code and data), is referred to as a Project.
-Working with Projects is exceptionally comfortable with RStudio an integrated development environment (IDE) for R.
-It is usefull to organize this Project in way that strictly segregates (raw) data and code by placing them in directories called `data` and `R` [see Section 4 @marwickPackagingDataAnalytical2018].
+```{r}
+x <- data.frame(VAR1=c(1,2,3,4),VAR2=c(0,4,6,9) )
+digest::sha1(x)
+```
 

### 2019-08-23 README.Rmd
add license to readme

@@
 ```
+
+[![licensebuttons by](https://licensebuttons.net/l/by/3.0/88x31.png)](https://creativecommons.org/licenses/by/4.0)
+
+The repo is maintained by Aaron Peikert[![ORCID iD](https://orcid.org/sites/default/files/images/orcid_16x16.png)](https://orcid.org/0000-0001-7813-818X) and Andreas Brandmaier [![ORCID iD](https://orcid.org/sites/default/files/images/orcid_16x16.png)](http://orcid.org/0000-0001-8765-6982)
 

### 2019-08-23 README.Rmd
reorder readme

--- 2019-08-20 README.Rmd
@@
 
-This is the accompanying GitHub repository to a work in progress paper, its abstract reads:
+The repo is maintained by [Aaron Peikert ![ORCID iD](https://orcid.org/sites/default/files/images/orcid_16x16.png)](https://orcid.org/0000-0001-7813-818X) and [Andreas Brandmaier ![ORCID iD](https://orcid.org/sites/default/files/images/orcid_16x16.png)](http://orcid.org/0000-0001-8765-6982) [![licensebuttons by](https://licensebuttons.net/l/by/3.0/88x31.png)](https://creativecommons.org/licenses/by/4.0)
+
+This is the accompanying GitHub repository to a work in progress paperby Aaron Peikert[![ORCID iD](https://orcid.org/sites/default/files/images/orcid_16x16.png)](https://orcid.org/0000-0001-7813-818X) and Andreas Brandmaier [![ORCID iD](https://orcid.org/sites/default/files/images/orcid_16x16.png)](http://orcid.org/0000-0001-8765-6982).
+
+# Abstract
 

### 2019-08-23 README.Rmd
forgot to delete

@@
 ```
-
-The repo is maintained by [Aaron Peikert ![ORCID iD](https://orcid.org/sites/default/files/images/orcid_16x16.png)](https://orcid.org/0000-0001-7813-818X) and [Andreas Brandmaier ![ORCID iD](https://orcid.org/sites/default/files/images/orcid_16x16.png)](http://orcid.org/0000-0001-8765-6982) [![licensebuttons by](https://licensebuttons.net/l/by/3.0/88x31.png)](https://creativecommons.org/licenses/by/4.0)
 

### 2019-08-23 README.Rmd
readd license
add space

--- 2019-08-23 README.Rmd
@@
 
+This is the accompanying GitHub repository to a work in progress paper by Aaron Peikert[![ORCID iD](https://orcid.org/sites/default/files/images/orcid_16x16.png)](https://orcid.org/0000-0001-7813-818X) and Andreas Brandmaier [![ORCID iD](https://orcid.org/sites/default/files/images/orcid_16x16.png)](http://orcid.org/0000-0001-8765-6982).
 [![licensebuttons by](https://licensebuttons.net/l/by/3.0/88x31.png)](https://creativecommons.org/licenses/by/4.0)
 
-The repo is maintained by Aaron Peikert[![ORCID iD](https://orcid.org/sites/default/files/images/orcid_16x16.png)](https://orcid.org/0000-0001-7813-818X) and Andreas Brandmaier [![ORCID iD](https://orcid.org/sites/default/files/images/orcid_16x16.png)](http://orcid.org/0000-0001-8765-6982)
-
-This is the accompanying GitHub repository to a work in progress paper, its abstract reads:
+# Abstract
 

### 2019-08-23 rr-flow.Rmd
fix compile issue introduced by tabs instead of spaces
added officer package description and some limitations

@@
 
-Such change prevents many reports from beeing replicated, if they contain, e.g., multiple imputations, bootstrapping, simulations studies, graphics with random jitter, bayesian estimations, or similar technics which involve random sampling.
-However seldom does an analysis depend only on R, often a considerable number of packages is required, and they will most likely more often introduce breaking changes with updates.
+Such change prevents many reports from beeing replicated, if they contain, e.g., multiple imputations, bootstrapping, simulations studies, graphics with random jitter, Bayesian estimations using sampling algorithms (such as Markov Chain Monte Carlo), or similar technics which involve random sampling.
+However only rarelydoes an analysis depend only on base R, often a considerable number of packages is required, and they will most likely more often introduce breaking changes with updates.
 The whole endeavor of reproducibility is therefore at stake, every time an update is rolled out.
-To prevent that `docker` replicates the computational environment of analysis exactly.
-We do not intend to advocate that software should not be updated; the used software version should just be documented and easily replicated.
+To ensure long-term reproducibility, `docker` replicates the original computational environment of an analysis exactly.
+However, we do not intend to advocate that software should not be updated; updates typically promote bugfixes and provide new functionality; but full computational reproducibility is only achieved if the originally used software version(s) are documented.
+Among other things, this allows to trace back recent updates as to what change in what software caused a non-replication of the original results.
 Quite to the contrary with solutions like docker, it gets more comfortable than ever to safely update to new versions, by just bumping the R Version number of the docker image.
@@
 
+## Limitations
+
+We are aware that implementing this workflow is not straightforward and the difficulty of its implementation may vary by platform.
+For example, already the installation of all tools is easier on POSIX-compatible platforms such as Unix, Linux, or macOS (but not Windows).
+However, once such a workflow is established as a default, it can be used with minimal changes to the familiar use of the R environment.
+It is even possible to move the entire development process to a docker environment.
+
+In our own experience, it is often not possible to convince all co-authors to switch to a different document processing environment, such as RMarkdown.
+That is, we have experienced that after writing up the first draft in Rmarkdown, one must eventually switch to a dynamically generated Word-file that is from then on used as static file that is used as basis for multiple iterations among the co-authors.
+This means, that ideally, one would have to manually merge the final Word file back to the RMarkdown souce, which generally is only possible manually.
+But we believe that even if this final step does not happen that - as long as dynamically generated elements are not manipulated - ensuring the reproducibility of the analysis and the initial draft is a big step forward.
+A less elegant but potentially more practical workflow may involve the officer package (CITATION NEEDED).
+The officer package allows the manipulation of  Word (.docx) and PowerPoint (*.pptx) documents, that is, given a formatted template file, one can dynamically paste images, tables and text into documents from R.
+
 ## Sharing reproducible workflows
@@
 Both licenses assure maximal freedom for future users while requiring the attribution of the original authors in derivative work.
+These licenses are also in line with the recommendations by the Reproducibile Research Standard (see stodden2016enhancing).
 

### 2019-08-23 R/cfcs-example-reproducible-analysis.Rmd
added minimal reproducible example (CFCS)

---
title: "CFCS-Reproducible-Analysis-Example"
author: "Peikert and Brandmaier"
date: "8/23/2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading packages

```{r}
library(lavaan)
library(digest)
```

## Loading data

```{r}
precomputed.digest <- "98c7e634dd7bb2ddf9d8f3972fff303365c55ac6"
temp <- tempfile()
download.file("http://openpsychometrics.org/_rawdata/CFCS.zip",destfile = temp)
dat <- read.table(unz(temp, "CFCS/data.csv"), header=TRUE)
unlink(temp)

if (precomputed.digest != digest::sha1(dat)) {
  error("Mismatch between original and current data file!")
}
```

## Fitting the model

```{r}
model <- paste0(" G =~ ", paste(paste0("Q",1:5),collapse="+"))

model.fit <- cfa(model, data=dat)

fitMeasures(model.fit)[c("cfi","rmsea")]

knitr::kable(lavaan::parameterEstimates(model.fit))
```

### 2019-08-23 rr-flow.Rmd
fix starttoc
execute code in different R Versions
resolves #4

@@
 floatsintext      : yes
-figurelist        : yes
+figurelist        : no
 tablelist         : no
@@
 pacman::p_load("here", "tidyverse")
+
+knitr::read_chunk(here("R", "random.R"))
 ```
@@
 
-```{bash, eval=FALSE}
-docker run --rm rocker/r-ver:3.5.0 Rscript -e "set.seed(1234);sample(1:10, 5)"
+```{r random, eval=FALSE}
+```
+
+```{r, child="R/random35.md"}
 ```
@@
 
-```{bash, eval=FALSE}
-docker run --rm rocker/r-ver:3.6.0 Rscript -e "set.seed(1234);sample(1:10, 5)"
+```{r random, eval=FALSE}
+```
+
+```{r, child="R/random36.md"}
 ```

### 2019-08-23 rr-flow.Rmd
add citations
closes #5

@@
 
-bibliography      : ["Reproducible-Research.bib"]
+bibliography      : ["reproducible-research.bib"]
 
@@
 But we believe that even if this final step does not happen that - as long as dynamically generated elements are not manipulated - ensuring the reproducibility of the analysis and the initial draft is a big step forward.
-A less elegant but potentially more practical workflow may involve the officer package (CITATION NEEDED).
+A less elegant but potentially more practical workflow may involve the officer package (@gohel2019officer).
 The officer package allows the manipulation of  Word (.docx) and PowerPoint (*.pptx) documents, that is, given a formatted template file, one can dynamically paste images, tables and text into documents from R.
@@
 Both licenses assure maximal freedom for future users while requiring the attribution of the original authors in derivative work.
-These licenses are also in line with the recommendations by the Reproducibile Research Standard (see stodden2016enhancing).
+These licenses are also in line with the recommendations by the Reproducibile Research Standard (see @stoddenEnhancingReproducibilityComputational2016).
 

### 2019-08-23 README.Rmd
add funny badges

@@
 This is the accompanying GitHub repository to a work in progress paper by Aaron Peikert[![ORCID iD](https://orcid.org/sites/default/files/images/orcid_16x16.png)](https://orcid.org/0000-0001-7813-818X) and Andreas Brandmaier [![ORCID iD](https://orcid.org/sites/default/files/images/orcid_16x16.png)](http://orcid.org/0000-0001-8765-6982).
-[![licensebuttons by](https://licensebuttons.net/l/by/3.0/88x31.png)](https://creativecommons.org/licenses/by/4.0)
+
+[![licensebuttons by](https://licensebuttons.net/l/by/3.0/88x31.png)](https://creativecommons.org/licenses/by/4.0) [![Ask Me Anything !](https://img.shields.io/badge/Ask%20me-anything-1abc9c.svg)](https://GitHub.com/Naereen/ama) [![Open Source Love](https://badges.frapsoft.com/os/v2/open-source.svg?v=103)](https://github.com/ellerbrock/open-source-badges/)
 

### 2019-08-26 README.Rmd
change links for badges

--- 2019-08-23 README.Rmd
@@
 This is the accompanying GitHub repository to a work in progress paper by Aaron Peikert[![ORCID iD](https://orcid.org/sites/default/files/images/orcid_16x16.png)](https://orcid.org/0000-0001-7813-818X) and Andreas Brandmaier [![ORCID iD](https://orcid.org/sites/default/files/images/orcid_16x16.png)](http://orcid.org/0000-0001-8765-6982).
-[![licensebuttons by](https://licensebuttons.net/l/by/3.0/88x31.png)](https://creativecommons.org/licenses/by/4.0)
+
+[![licensebuttons by](https://licensebuttons.net/l/by/3.0/88x31.png)](https://creativecommons.org/licenses/by/4.0) [![Ask Me Anything !](https://img.shields.io/badge/Ask%20me-anything-1abc9c.svg)](https://github.com/aaronpeikert/reproducible-research/issues/new) ![Open Source Love](https://badges.frapsoft.com/os/v2/open-source.svg?v=103)
 

### 2019-08-26 R/random.R
add outputs from random so that docker is not needed

@@ -1,0 +1,4 @@
+#----random----
+R.version$version.string
+set.seed(1234)
+sample(1:10, 5)

### 2019-08-26 rr-flow.Rmd
add summary

@@
   pacman here pander
+RUN installGithub.r\
+  crsh/papaja benmarwick/wordcountaddin
 WORKDIR /home/rstudio
@@
 
-! TODO - Summary of all this and all benefits
+The overarching goal of this paper was to provide an easy to use workflow, that allows confidence in reproducibility.
+This workflow can be almost unfailingly trusted to be replicable with the following commands:
+
+```{bash, eval=FALSE}
+git clone https://github.com/aaronpeikert/reproducible-research.git
+cd reproducible-research
+make build
+make all DOCKER=TRUE
+```
+
+That is, it replicates a scientific report exactly without regard to the user, the operating system, the software, the timepoint or the version of the files.
+To that end, it relies on tools that have been the foundation of reliable software for years or in most cases, decades.
+On the way, it makes transparent how statistical results relate to the software that created them and enables, through that transparency, reuse by other researchers.
+
+Each tool reduces the number of ways things can go wrong.
+Dynamic reporting with `RMarkdown` fixes the relation between statistical report and statistical results; Version control with `git` fixes the version of the used files; Dependency management with `make` fixes relation of files and their processing; Containerization with `docker` fixes the software environment.
+This tightening of loose ends, we believe, does not restrict the researcher but enables her to operate on a solid basis to deliver sound research.
 
@@
 How to best share a reproducible workflow?
-We believe that, ideally, a service provider should be found that guarantuees permanent and reliable hosting of the reproducible workflow, such as the Open Science Framework (citation needed) or PsychData (citation needed).
+We believe that, ideally, a service provider should be found that guarantuees permanent and reliable hosting of the reproducible workflow, such as the Open Science Framework [@fosterOpenScienceFramework2017] or PsychData (citation needed).
 Second, to ensure that other users are legally able to benefit from the shared materials, authors must choose an appropriate license format.

### 2019-08-26 rr-flow.Rmd
add automated wordcount closes #16
take the edge of the officer part

@@
 keywords          : "keywords"
-wordcount         : "X"
+wordcount         : "`r wordcountaddin::word_count(here::here('rr-flow.Rmd'))`"
 
@@
 But we believe that even if this final step does not happen that - as long as dynamically generated elements are not manipulated - ensuring the reproducibility of the analysis and the initial draft is a big step forward.
-A less elegant but potentially more practical workflow may involve the officer package (@gohel2019officer).
-The officer package allows the manipulation of  Word (.docx) and PowerPoint (*.pptx) documents, that is, given a formatted template file, one can dynamically paste images, tables and text into documents from R.
+Making such static Word-file more dynamic is is possible with the officer package [@gohel2019officer], which allows the manipulation of  Word (.docx) and PowerPoint (.pptx) documents.
+That is, given a formatted static template file, one can dynamically paste images, tables and text into documents from R.
+We consider such static file with dynamic content, less elegant but potentially more practical, when not all authors can adhere to the described process.
 

### 2019-08-26 rr-flow.Rmd
closes #8

@@
 How to best share a reproducible workflow?
-We believe that, ideally, a service provider should be found that guarantuees permanent and reliable hosting of the reproducible workflow, such as the Open Science Framework [@fosterOpenScienceFramework2017] or PsychData (citation needed).
+We believe that, ideally, a service provider should be found that guarantuees permanent and reliable hosting of the reproducible workflow, such as the Open Science Framework [@fosterOpenScienceFramework2017] or PsychData (citation needed), that is independ of Gihub, Docker Hub & MRAN.
 Second, to ensure that other users are legally able to benefit from the shared materials, authors must choose an appropriate license format.

### 2019-08-26 rr-flow.Rmd
add (unfinished) setup

@@
 
+# Set up Workflow
+
+This workflow directly relies on three software packages: `git`, `make` and `docker`.
+Of course it depends also indirectly on R, RStudio & RMarkdown, but they do not need to be installed.
+The installation process depends on the operating system, e.g.
+on OS X, make is allways availible or on Linux git & make are often allready installed.
+In the following section we share what have experienced as the easiest options.
+
+## Windows
+
+Windows has the bigest hurdle to install what is needed.
+However there is a great package manager for Windows called chocolately which you can install from: [https://chocolatey.org](https://chocolatey.org).
+After having installed chocolately, all dependencies can be installed in an **admin terminal** via:
+
+```{bash, eval=FALSE}
+choc install git make docker-desktop
+```
+
 # Summary

### 2019-08-26 rr-flow.Rmd
fix typo

@@
 However there is a great package manager for Windows called chocolately which you can install from: [https://chocolatey.org](https://chocolatey.org).
-After having installed chocolately, all dependencies can be installed in an **admin terminal** via:
+After having installed chocolately (and restarted the computer), all dependencies can be installed in an **admin terminal** via:
 
 ```{bash, eval=FALSE}
-choc install git make docker-desktop
+choco install git make docker-desktop
 ```

### 2019-08-27 R/bootstrap-01-data-generation.R
added reproducible non-reproducibility of bootstrap using lm()

@@ -1,0 +1,20 @@
+#
+# generate dataset for pathological non-replication
+# of bootstrap CI-based decisions
+#
+# run on R 3.6.1
+
+# generate data and make permanent
+set.seed(1234)  
+x <- rnorm(100)
+y <- rnorm(100)
+
+y <- y + .225*x
+
+x <- round(x,2)
+y <- round(y,2)
+
+simdata <- data.frame(x,y)
+
+write.csv(file = "../data/pathological-example2.csv", 
+          x=simdata, quote=FALSE, row.names = FALSE)

### 2019-08-27 R/bootstrap-02-case-generator.R

@@ -1,0 +1,30 @@
+
+# load library
+library("boot")
+
+# load data
+simdata <- read.csv("/home/rstudio/data/pathological-example2.csv")
+
+# function to obtain regression weights 
+bs <- function(formula, data, indices) {
+  d <- data[indices,] # allows boot to select sample 
+  fit <- lm(formula, data=d)
+  return(coef(fit)) 
+} 
+
+bcis <- c()
+seeds <- c(1:10)
+
+for (seed in seeds) {
+  set.seed(seed)
+  # bootstrapping with many replications 
+  results <- boot(data=simdata, statistic=bs, 
+                  R=1000, formula=y~1+x)
+  
+  bci <- boot.ci(results,type="bca", index=2) # b estimate
+  print(bci)
+  
+  bcis <- rbind(bcis,bci$bca)
+}
+
+#save(bci,file = paste0("results",".Rda"))

### 2019-08-27 R/bootstrap-03-fit.R

@@ -1,0 +1,28 @@
+
+# load library
+library("boot")
+
+# load data
+simdata <- read.csv("/home/rstudio/data/pathological-example2.csv")
+
+# function to obtain regression weights 
+bs <- function(formula, data, indices) {
+  d <- data[indices,] # allows boot to select sample 
+  fit <- lm(formula, data=d)
+  return(coef(fit)) 
+} 
+
+set.seed(2)
+
+# bootstrapping with many replications 
+results <- boot(data=simdata, statistic=bs, 
+                  R=1000, formula=y~1+x)
+  
+# get beta estimates' confidence intervals
+bci <- boot.ci(results,type="bca", index=2) # b estimate
+print(bci)
+  
+
+
+
+

### 2019-08-27 rr-flow.Rmd
add bootstrap

@@
 keywords          : "keywords"
-wordcount         : "`r wordcountaddin::word_count(here::here('rr-flow.Rmd'))`"
+wordcount         : "`r tryCatch(wordcountaddin::word_count(here::here('rr-flow.Rmd')), error=function(err) 'NA')`"
 
@@
 knitr::read_chunk(here("R", "random.R"))
+knitr::read_chunk(here("R", "bootstrap.R"))
 ```
@@
 Such change prevents many reports from beeing replicated, if they contain, e.g., multiple imputations, bootstrapping, simulations studies, graphics with random jitter, Bayesian estimations using sampling algorithms (such as Markov Chain Monte Carlo), or similar technics which involve random sampling.
+A result may be that a in one version the result of a bootstrap-test is significant and in another version is not anymore:
+
+```{r bootstrap, echo=TRUE, eval=FALSE}
+```
+
+```{r, child="R/bootstrap35.md"}
+```
+
+```{r bootstrap, echo=TRUE, eval=FALSE}
+```
+
+```{r, child="R/bootstrap36.md"}
+```
+
 However only rarelydoes an analysis depend only on base R, often a considerable number of packages is required, and they will most likely more often introduce breaking changes with updates.

### 2019-08-27 R/bootstrap-03-fit.R

@@ -1,28 +1,0 @@
-
-# load library
-library("boot")
-
-# load data
-simdata <- read.csv("/home/rstudio/data/pathological-example2.csv")
-
-# function to obtain regression weights 
-bs <- function(formula, data, indices) {
-  d <- data[indices,] # allows boot to select sample 
-  fit <- lm(formula, data=d)
-  return(coef(fit)) 
-} 
-
-set.seed(2)
-
-# bootstrapping with many replications 
-results <- boot(data=simdata, statistic=bs, 
-                  R=1000, formula=y~1+x)
-  
-# get beta estimates' confidence intervals
-bci <- boot.ci(results,type="bca", index=2) # b estimate
-print(bci)
-  
-
-
-
-

### 2019-08-28 rr-flow.Rmd
finish install section closes #30 #29

@@
 on OS X, make is allways availible or on Linux git & make are often allready installed.
-In the following section we share what have experienced as the easiest options.
+In the following section we share what have experienced as the easiest options, however in the past how docker should be installed has changed several times, so when in doubt search the internet with: `how to install docker on [your os]`.
 
@@
 
-Windows has the bigest hurdle to install what is needed.
-However there is a great package manager for Windows called chocolately which you can install from: [https://chocolatey.org](https://chocolatey.org).
-After having installed chocolately (and restarted the computer), all dependencies can be installed in an **admin terminal** via:
+Windows has the biggest hurdle to install what is needed.
+An essential requirement is that you have Windows Pro, Education or Server installed, as Microsoft prevents the use of Docker on Windows Home.
+However, having met this requirement, there is a great package manager for Windows called chocolately, which you can install from: [https://chocolatey.org/install](https://chocolatey.org/install).
+After having installed chocolately (and restarted the computer), all dependencies can be installed in an **admin terminal** (Windows-Key, then type `cmd`, right-click *Run as administrator*) via:
 
@@
 ```
+
+To use `docker` you need to start Docker Desktop.
+
+## OS X
+
+As make already ships with OS X, you only need `git` and `docker-desktop`.
+We suggest using the package manager `homebrew`, which you can install from [https://docs.brew.sh/Installation](https://docs.brew.sh/Installation), to install `docker` (git will be installed during the installation of `homebrew`):
+
+```{bash, eval=FALSE}
+brew cask docker
+```
+
+To use `docker` you need to start Docker Desktop.
+
+## Linux
+
+There is a pletora of linux distribuation and almost as many package managers.
+However to our knowledge there is no (recent) Linux where `git`, `make` and `docker` are not availible.
+Example for Ubuntu:
+
+```{bash, eval=FALSE}
+apt install git make docker 
+```
+
+Replace `apt` with your package manager. Possiblie you need elevated rights in this case add `sudo` before that.
+
+An alternative to `docker` on Linux is `singularity`.
+Just replace any docker calls with `singularity docker`, because singularity fully supports docker-images.
+A possible advantage is that singularity works well in  HPC environments and on old Linux versions, the downside is that `singularity` is only availibe on Linux.
 
@@
 That is, it replicates a scientific report exactly without regard to the user, the operating system, the software, the timepoint or the version of the files.
-To that end, it relies on tools that have been the foundation of reliable software for years or in most cases, decades.
+To that end, it relies on tools that have been the foundation of reliable software development for years or in most cases, decades.
 On the way, it makes transparent how statistical results relate to the software that created them and enables, through that transparency, reuse by other researchers.

### 2019-08-28 rr-flow.Rmd
addsection on related approaches closes #20

@@
 
-! TODO
+While this approach here was designed to scale well with the complexity of a computing-intense Project, we realise that this flexibility comes with a steep learning curve.
+A few packages are easing this burden for different parts of the workflow when this flexibility is not needed.
+The use of `rmarkdown` within a project, tracked with `git` can be simplified with the [workflowr-package](https://github.com/jdblischak/workflowr) [@workflowr].
+The [drake-package](https://github.com/ropensci/drake) [@drake] is directly inspired by `make` and does only require knowledge in R, but can only handle dependencies within R.
+The [holepunch-package](https://karthik.github.io/holepunch/) [@holepunch] automizes the use of docker and uses [binder](https://mybinder.org) to move the analysis to the cloud so that no local installation of docker is required.
+It is well suited for simple analysis with low computational demands, because of memory and computing time limitations of binder.
+There are several alternatives to docker which fix the dependancies on R-packages.
+[packrat](https://rstudio.github.io/packrat/) [@packrat] is a way to fix package version via installing locally in the project, however it does not garuantees more then a fixed set of packages (Not R-version or system dependencies).
+Similar approaches are taken by [jetpack](https://github.com/ankane/jetpack) [@jetpack], [miniCRAN](https://cran.r-project.org/web/packages/miniCRAN/vignettes/miniCRAN-introduction.html) [@miniCRAN] and [checkpoint](https://cran.r-project.org/web/packages/checkpoint/vignettes/checkpoint.html) [@checkpoint].
+Notable is also the package [`reprex`](https://github.com/tidyverse/reprex) [**repr**oducible **ex**ample, @reprex], but its scope is limited to small examples.
+
+Another way to increase the chance of reproducibility is to develop an R-package for each analysis.
+The reasoning is that by abiding the strict rules of package development and specifieng all dependencies a high degree of stability can be archieved.
+This is approach is extensivly developed by @marwickPackagingDataAnalytical2018.
 

### 2019-08-28 rr-flow.Rmd
added some intro prose

--- 2019-08-28 rr-flow.Rmd
@@
     affiliation   : "1,2"
-  - name          : "Andreas Brandmaier"
+  - name          : "Andreas M. Brandmaier"
     affiliation   : "1,3"
@@
 
-! TODO - Unterschied zwischen computational, empirical, and statistical reproducibility (cf. Stodden et al., 2014)
+In this tutorial, we describe a principled workflow to ensure long-term reproducibility of
+R-based data analyses.
+In this context, reproducibility means to us that scientific results together with their computational scientific workflows are reported fully, transparently and sustainably available, such that the workflow can be re-run by a different person and/or later in time and the results will be identical to the ones intially reported.
+Here, we combine four tools from the computer sciences, whose interplay can guarantuee full and sustainable computational reproducibility of data analyses.
+The major four elements are:
 

### 2019-08-29 rr-flow.Rmd
revised the manuscript

@@
 
+# Introduction
+
 In this tutorial, we describe a principled workflow to ensure long-term reproducibility of
@@
 
-We believe that for sufficient chances of reproducibility, proper documentation is crucial since it answers the question "What steps have to be taken to reproduce the results?".
+We believe that for sufficient chances of long-term reproducibility, proper documentation is crucial since it answers the question "What steps have to be taken to reproduce the results?".
 The most precise documentation is computer code, and the most productive computer code is code that runs automatically without human interaction.
@@
 
-The first step towards long-term reproducibility is to have an R-script or R-markdown, that runs on a local computer without error.
+# Project organization
+
+The first step towards reproducibility is to have an R-script or R-markdown as primary entry point for the analysis, that runs on a local computer without error.
 A process, we assume, that is familiar to the reader.
@@
 # BAD
-iris <- read.csv("/home/aaron/Documents/reproducible-research/data/iris.csv")
+iris <- read.csv("/home/aaron/reproducible-research/data/iris.csv")
 # GOOD
@@
 
-This folder, where everything resides that you need for analysis (code and data), is referred to as a Project.
-Working with Projects is exceptionally comfortable with RStudio an integrated development environment (IDE) for R.
-It is usefull to organize this Project in way that strictly segregates (raw) data and code by placing them in directories called `data` and `R` [see Section 4 @marwickPackagingDataAnalytical2018].
+This folder, where everything resides that you need for analysis (code and data), is referred to as a 'project'.
+Working with projects is exceptionally comfortable with RStudio, an integrated development environment (IDE) for R.
+It is useful to organize a data-analysis project in way that strictly segregates (raw) data and code by placing them in directories called `data` and `R` [see Section 4 @marwickPackagingDataAnalytical2018]; there are also tools that automatize the standardized creation of folder structures such as `workflowr` [@workflowr].
 
-Sometimes external requirements mean that the data cannot be stored and shared together with the scripts.
+Sometimes external requirements make it impossible that the data cannot be stored and shared together with the scripts.
 In most of the cases we have seen, these are either space constraints or privacy considerations.
@@
 x <- data.frame(VAR1=c(1,2,3,4),VAR2=c(0,4,6,9) )
-digest::sha1(x)
+checksum <- digest::sha1(x)
+if (checksum != "740e88404c04af1d67ffc87b008231c093539cc1") 
+	error("Wrong data file loaded or data has been tempered with!")
 ```
 
-Reproducibility at a basic level assumes that results are the same if neither script nor data have changed.
-It is often not trivial to find out whether anything has changed and if so to "go back in time." `git` enables you to do both.
+# Version Management
+
+Reproducibility at a basic level assumes that results remain identical if neither script nor data have changed.
+It is often not trivial to find out whether anything has changed over time and if so to "go back in time." `git` enables you to do both.
 A good mental model for `git` is that it takes a sequence of snapshots of all files it is supposed to track.
 In the language of git, these snapshots are "commits".
-A commit represents a complete copy of the state of the files when added to a commit.
-Each commit has a unique identifier (a hash).
-Going back to one state is as easy as finding the hash of the commit.
-Finding a specific hash can be made even easier by tagging specific commits, e.g., as "submission," "preprint," "publication." This collection of commits is called "repository," which is almost the same as your Project.
+A commit represents a complete copy of the state of the tracked files.
+Each commit has a unique identifier (a hash) and a human-readable description.
+Going back to one state is as easy as traversing the history of all commits; it is possible to visually track changes between different versions.
+Finding a specific version can be made even easier by tagging snapshots, e.g., as "submission," "preprint," or "publishged." This collection of snapshots is called a "repository," which ideally tracks your entire R project.
 
-A typical git workflow looks like that:
+A typical git workflow in the terminal looks like that:
 
 ```{bash, eval=FALSE}
-git init # to initialize git
-git add /data/iris.csv /R/analysis.R # track specific files/changes
-git commit -m "add data & analysis" # take snapshot
-# new data comes in
-git commit -a -m "complete data colections" # add (-a) and comit all changes
+git init # to initialize git in the current directory
+git add ./data/iris.csv ./R/analysis.R # track specific files/changes
+git commit -m "add data & analysis" # take snapshot with comment
+# once script or data were changed, take a new snapshot
+git commit -a -m "complete data colections" # add (-a) and commit all changes
 ```
 
-So to keep track of all changes, you only need to use `git add` & `git commit` or `git commit -a` to do both.
+So to keep track of all changes on your local computer, you only need to use `git add` & `git commit` or `git commit -a` to do both.
 These commands need to be executed in the `terminal`, which you can access from within Rstudio (`Shift + Alt + R`).
-Rstudio also offers a graphical user interface for Git and quite likely this suffice for almost every interaction with Git.
+Rstudio also offers a graphical user interface for Git and, quite likely, this interface is sufficient for almost every interaction with Git.
 
@@
 
-Now one can check everything that has happened (`git log`) and look at the exact state by supplying the id of the commit or the first few digits of it to `git checkout`:
+Now one can check everything that has happened (`git log`) and look at the exact state by supplying the id of the commit or the first few digits of its hash to `git checkout`:
 
@@
 
-Git also makes it especially easy to share and collaborate on a project with another researcher.
+Git also makes it especially easy to share and collaborate on a project with other researchers.
 A popular service for sharing via git is [Github](https://github.com).
 Just sharing Git repositories with the public is always free, private repositories (only visible to persons you invite) are [free for researcher](https://help.github.com/en/articles/applying-for-an-educator-or-researcher-discount).
-After creating a user account, you can create a new repository where Github advises you how to upload your repository from the terminal e.g., for this repository:
+After creating a user account, you can create a new repository where Github advises you how to upload your repository from the terminal e.g., for this repository (here with user name 'aaronpeikert' and repository name 'reproducible-research'):
 
@@
 
-`git push` or the green up arrow in the git pane uploads local updates. To download this git repo to another computer type into the terminal:
+`git push` or the green up-arrow in the git pane uploads local updates. To download this git repo to another computer type into the terminal:
 
@@
 
-Git and Github can do even more to help you to collaborate with your fellow researcher, such as formal means to track issues and their status (open/closed/resolved) and further means to managing and merging multiple, parallel versions of code (such as branches, pull requests, merges) but this is beyond the already broad scope of this tutorial.
+Git and Github can do even more to help you to collaborate with your fellow researchers, such as formal means to track issues and their status (open/closed/resolved) and further means to managing and merging multiple, parallel versions of code (such as branches, pull requests, merges) but this is beyond the already broad scope of this tutorial.
 Another benefit of using Git and Github is that experimentation is highly encouraged since you can go back to any state quickly and even when you lose access to the file on your computer, everything is backed up on the Github servers.
 
-Even when you have obtained a version of the Project, and you can ensure that this version is unchanged, you may not know how to reproduce the results, because it is unclear what has to be run in which order.
+# Dependency Tracking and Management
+
+Even when you have obtained a version of the project, and you can ensure that this version is unchanged, you may not know how to reproduce the results, because it is unclear what has to be run in which order.
 This job gets comfortable with `make`, since it allows you to create (computational) recipes too (re-)create files.
@@
 
-However, this relies on the crucial assumption that your computational environments are sufficiently compatible, e.g., everything needed is installed (R + Packages) at a compatible version of the software.
+However, this relies on the crucial assumption that your computational environments are identically or sufficiently compatible, e.g., all software dependencies are installed (R and all additional R Packages) and no updates or other changes to the computational environment break or alter your analysis.
+As we will see shortly, this is ensuring full computational reproducibility is necessary at a deeper level than just providing the workflow itself.
+
+# Containerization
 
@@
 
-While the R programming language is considered stable and much effort is put into backward compatibility, sometimes even basic functions like `sample()` change their behavior from one version to another.
+While the R programming language is considered stable and much effort is put into backward compatibility, sometimes even basic functions like `sample()`, to randomly sample from a set, change their behavior from one version to another.
+To ensure reproducibility in analyses based on a computers pseudo random number generator (PRNG), it is good practice to rely on fixed PRNG seeds, which are numeric values that set the PRNG into a deterministic state, i.e., the sequence of random numbers reproduces exactly.
 
@@
 
-Surprisingly in an image with a more recent version of R (3.6.0), this results in another sample:
+Surprisingly in an image with a more recent version of R (3.6.0), this results in another sample despite the identical seed:
 
@@
 
-Such change prevents many reports from beeing replicated, if they contain, e.g., multiple imputations, bootstrapping, simulations studies, graphics with random jitter, Bayesian estimations using sampling algorithms (such as Markov Chain Monte Carlo), or similar technics which involve random sampling.
-A result may be that a in one version the result of a bootstrap-test is significant and in another version is not anymore:
+Such change may render analyses run on previous R versions strictly not reproducible if they contain, e.g., multiple imputations, bootstrapping, simulations studies, graphics with random jitter, Bayesian estimations using sampling algorithms (such as Markov Chain Monte Carlo), or similar techniques that involve random sampling.
+We would like to illustrate this with a more concrete example (the full R code to reproduce this non-reproducibility is provided in the github repository of this manuscript).
+Using a simulated dataset with two variables ´x´ and ´y´, we ran a linear regression model with R's ´lm()´ function regressing ´x´ on ´y´.
+Using the ´boot´ package [@boot], we bootstrapped the 95% confidence intervals around the regression estimate with 1000 bootstrap samples to evaluate whether zero was inside the confindence interval.
+To make the analysis reproducible, we set a random seed before we bootstrapped.
+We ran this once in R 3.5.0:
 
@@
 
+And then, we ran the identical script with identical seed in R 3.6.1:
+
 ```{r bootstrap, echo=TRUE, eval=FALSE}
@@
 
-However only rarelydoes an analysis depend only on base R, often a considerable number of packages is required, and they will most likely more often introduce breaking changes with updates.
+As we see from these R outputs, the confidence intervals once include zero and once do not include zero.
+Please note that one could discuss deeper issues about null hypothesis significance testing here but we would simply like to stress the point that computational reproducibility in the strict sense needs to capture the full computational environment.
+
+However only rarely does an analysis depend only on base R, often a considerable number of packages is required, and they will most likely more often introduce breaking changes with updates.
 The whole endeavor of reproducibility is therefore at stake, every time an update is rolled out.
@@
 
-# Set up Workflow
+# Setting Up the Workflow
 
@@
 
+## Outlook
+
+The presented workflow leverages various existing tools that are partly integrated into Rstudio already.
+Parts of the proposed workflow have been integrated into stand-alone packages (such as @workflowr or @holepunch) but they are either incomplete or rely on propietary service providers.
+We hope that with increasing awareness of the problems associated with computational reproducibility, the increased demand for unified and open solutions, will lead to an even better integration of existing tools to allow reproducible workflows to become a standard in the psychological research.
+
 \newpage

### 2019-08-29 rr-flow.Rmd
add further references close #15

@@
 R-based data analyses.
-In this context, reproducibility means to us that scientific results together with their computational scientific workflows are reported fully, transparently and sustainably available, such that the workflow can be re-run by a different person and/or later in time and the results will be identical to the ones intially reported.
+In this context, reproducibility means to us that scientific results together with their computational scientific workflows are reported fully, transparently and sustainably available, such that the workflow can be re-run by a different person and/or later in time and the results will be identical to the ones intially reported [@herouxCompatibleReproducibilityTaxonomy2018; @claerboutElectronicDocumentsGive1992].
 Here, we combine four tools from the computer sciences, whose interplay can guarantuee full and sustainable computational reproducibility of data analyses.
@@
 
-The scientific community widely recognizes that these tools enhance reproducibility, and provides guidelines for their use (Literate Programming: @ruleTenSimpleRules2019; Version Control: @barbaHardRoadReproducibility2016; Dependency Management: @askrenUsingMakeReproducible2016; Containerization: @clyburne-sherinComputationalReproducibilityContainers2018).
+The scientific community widely recognizes that these tools enhance reproducibility [@piccoloToolsTechniquesComputational2016], and provides guidelines for their use (Literate Programming: @ruleTenSimpleRules2019; Version Control: @barbaHardRoadReproducibility2016; Dependency Management: @askrenUsingMakeReproducible2016; Containerization: @clyburne-sherinComputationalReproducibilityContainers2018).
 However, only a workflow using them in unison allows confidence in reproducing a scientific report.
@@
 
+However, beside these tools that ease the process of creating a workflow on the researchers side, we see increased interest in changing how research is published and used[@perkelToolkitDataTransparency2018], with *life code* [@perkelPioneeringLivecodeArticle2019] and *continues integration* [@beaulieu-jonesReproducibilityComputationalWorkflows2017, @yenniDevelopingModernData2018].
+For such further technical advancements reproducibility is key and researcher using this workflow should be well prepared for such era of interactive publication.
+
 ## Limitations
@@
 Both licenses assure maximal freedom for future users while requiring the attribution of the original authors in derivative work.
-These licenses are also in line with the recommendations by the Reproducibile Research Standard (see @stoddenEnhancingReproducibilityComputational2016).
+These licenses are also in line with the recommendations by the Reproducibile Research Standard [@stoddenEnhancingReproducibilityComputational2016; @stoddenEnablingReproducibleResearch2009].
 

### 2019-08-29 rr-flow.Rmd
revised manuscript

--- 2019-08-29 rr-flow.Rmd
@@
 
+## Reproducing an Analysis
+
+We provide a reproducible analysis as working example via github as a testbed for users.
+The example shows a minimalistic analysis of the 'Considerations of Future Consequences (CFC) Scale'.
+The analysis demonstrates a succesful implementation of our workflow including downloading external data, comparing their state using a checksum, and running a confirmatory factor analysis on the first few items using lavaan [@lavaan2012].
+Once all mentioned tools are installed on a computer, the following few steps are necessary to reproduce our analysis:
+
+```{bash}
+
+TODO: Aaron
+
+```
+
 # Summary

### 2019-08-29 rr-flow.Rmd
revised examples

@@
 
+The overarching goal of this paper was to provide an easy to use workflow, that allows confidence in reproducibility. T
 We provide a reproducible analysis as working example via github as a testbed for users.
@@
 
-```{bash}
+```{bash, eval=FALSE}
 
 TODO: Aaron
-
+git clone https://github.com/aaronpeikert/reproducible-research-example.git
+cd reproducible-research-example
+make build
+make all DOCKER=TRUE
 ```
 
-# Summary
-
-The overarching goal of this paper was to provide an easy to use workflow, that allows confidence in reproducibility.
-This workflow can be almost unfailingly trusted to be replicable with the following commands:
+Similarly, this manuscript and the differing results across R versions were created using the proposed workflow and can be reproduced directly from out github repository:
 
@@
 ```
+
+# Summary
 

### 2019-08-29 rr-flow.Rmd
add bash to reproduce example #2
dont eval bash inside markdown

--- 2019-08-29 rr-flow.Rmd
@@
 
-```{bash}
-
-TODO: Aaron
-
+```{bash, eval=FALSE}
+git clone https://github.com/aaronpeikert/workflow-showcase.git
+cd reproducible-research
+make build
+make all DOCKER=TRUE
 ```
@@
 To that end, it relies on tools that have been the foundation of reliable software development for years or in most cases, decades.
-On the way, it makes transparent how statistical results relate to the software that created them and enables, through that transparency, reuse by other researchers.
+On the way, it makes transparent how statistical results depend upon the software that created them and enables, through that transparency, reuse by other researchers.
 
 Each tool reduces the number of ways things can go wrong.
-Dynamic reporting with `RMarkdown` fixes the relation between statistical report and statistical results; Version control with `git` fixes the version of the used files; Dependency management with `make` fixes relation of files and their processing; Containerization with `docker` fixes the software environment.
-This tightening of loose ends, we believe, does not restrict the researcher but enables her to operate on a solid basis to deliver sound research.
+Dynamic reporting with `RMarkdown` guarantuees consistency between computational results and their reporting; Version control with `git` guarantuees permanence and consistency across multiple versions of data and code; Dependency management with `make` guarantuees defined entry-points and dependency resolition; Containerization with `docker` guarantuees reproducibility of the full computational environment.
+This tightening of loose ends, we believe, does not restrict researchers but enables them to operate on a solid basis to deliver sound research.
 

### 2019-08-29 rr-flow.Rmd
revised ms
add singularity citation closes #37
added call for issues

@@
 
-authornote: |
+authornote: This paper is fully reproducible using the workflow described in this paper.
+All materials can be found on the accompaniying Github repository:  To comment on our paper, please add an issue to the Github repository of our paper: <https://github.com/aaronpeikert/reproducible-research/>.
+To provide feedback or any other type of comment or questions regarding our workflow, please add an issue to the Github repository of our paper: <https://github.com/aaronpeikert/reproducible-research/issues>.
 
@@
 Git and Github can do even more to help you to collaborate with your fellow researchers, such as formal means to track issues and their status (open/closed/resolved) and further means to managing and merging multiple, parallel versions of code (such as branches, pull requests, merges) but this is beyond the already broad scope of this tutorial.
+In particular, github's issue management can be leveraged as post-publication platform to discuss manuscripts and their results [^1].
 Another benefit of using Git and Github is that experimentation is highly encouraged since you can go back to any state quickly and even when you lose access to the file on your computer, everything is backed up on the Github servers.
+
+[^1]: To comment on our paper, please add an issue to the Github repository of our paper: <https://github.com/aaronpeikert/reproducible-research/issues>.
 
@@
 
-```{bash, eval=FALSE}
+```{bash}
 git clone https://github.com/aaronpeikert/workflow-showcase.git

### 2019-08-29 rr-flow.Rmd

@@
 
-An alternative to `docker` on Linux is `singularity`.
+An alternative to `docker` on Linux is `singularity` [@kurtzerSingularityScientificContainers2017].
 Just replace any docker calls with `singularity docker`, because singularity fully supports docker-images.
@@
 
-```{bash}
+```{bash, eval=FALSE}
 git clone https://github.com/aaronpeikert/workflow-showcase.git

### 2019-08-29 rr-flow.Rmd
fixed broken author note

@@
 authornote: This paper is fully reproducible using the workflow described in this paper.
-All materials can be found on the accompaniying Github repository:  To comment on our paper, please add an issue to the Github repository of our paper: <https://github.com/aaronpeikert/reproducible-research/>.
-To provide feedback or any other type of comment or questions regarding our workflow, please add an issue to the Github repository of our paper: <https://github.com/aaronpeikert/reproducible-research/issues>.
+All materials can be found on the accompaniying Github repository <https://github.com/aaronpeikert/reproducible-research/>.
+To provide feedback or any other type of comment or questions regarding our workflow, please add an issue to the Github repository of our paper [https://github.com/aaronpeikert/reproducible-research/issues](https://github.com/aaronpeikert/reproducible-research/issues).
 

### 2019-08-29 rr-flow.Rmd
first stub on literate programming

@@
 ```
+
+# Literate Programming
+
+Typically, the translation of the human-readable summary of ones computational results, be it an internal report, a presentation, or a journal manuscript, is time-consuming and error-prone.
+To not only create fully reproducible results but also fully reproducible reports, one needs to resort to the Literate Programming paradigm [@knuth1984literate], in which human-readable language and computer-code are interspersed to create dynamic documents whose order follow the logic of thought rather then the order of the computer.
+R-Markdown is a simple markup language to create dynamic documents with embedded chunks of R code that can be exported to standard formats such as documents (docx, rtf, epub), presentations (ppt, html) or websites (html).
 

### 2019-08-30 rr-flow.Rmd
revised manuscript

@@
 ---
-title             : "A Reproducible Data Analysis Workflow with RMarkdown, Git, Make & Docker"
+title             : "A Reproducible Data Analysis Workflow with RMarkdown, Git, Make, & Docker"
 shorttitle        : "Reproducible Data Analysis Workflow"
@@
 R-based data analyses.
-In this context, reproducibility means to us that scientific results together with their computational scientific workflows are reported fully, transparently and sustainably available, such that the workflow can be re-run by a different person and/or later in time and the results will be identical to the ones intially reported [@herouxCompatibleReproducibilityTaxonomy2018; @claerboutElectronicDocumentsGive1992].
-Here, we combine four tools from the computer sciences, whose interplay can guarantuee full and sustainable computational reproducibility of data analyses.
-The major four elements are:
+In this context, reproducibility means to us that scientific results together with their computational scientific workflows are reported fully, transparently and remain sustainably available, such that the workflow can be re-run by a different person and/or later in time and the results will be identical to the ones intially reported [@herouxCompatibleReproducibilityTaxonomy2018; @claerboutElectronicDocumentsGive1992].
+The need to ensure reproducibility directly follows from commonly accepted rules of good scientific practice (such as the guidelines of the German Research Foundation; @dfg2019) and ensuring reproducibility is a prerequisite for replicability and a means to increase trustworthiness of empirical results [@epskamp2019rep].
+
+Here, we combine four sofrware tools, whose interplay can guarantuee full and sustainable computational reproducibility of data analyses and their reporting. The major four elements are:
 
 Literate Programming
-: Interwoven text and computer code that together forms a human-readable document.
+: Interwoven text and computer code that together forms a dynamically generated, human-readable document.
 : Example: One document that includes text that describes a statistical test, code that is used to execute the statistical and the output of the code in the same document.
@@
 : The management of change of documents/code.
-: Example: Changes from the manuscript text upon request by the reviewer are identifiable.
+: Example: Changes of the analysis code or even single parameters are documented both at development and post-publication.
 
@@
 : How different parts of a manuscript/analysis relate and how these dependencies resolve via execution of code.
-: Example: The manuscript depends on the statistical analysis, which depends on the data.
+: Example: The manuscript depends on the statistical analysis, which depends on a preprocessing pipeline, which depends on a certain subset of the data.
 
@@
 
-Another way to increase the chance of reproducibility is to develop an R-package for each analysis.
-The reasoning is that by abiding the strict rules of package development and specifieng all dependencies a high degree of stability can be archieved.
-This is approach is extensivly developed by @marwickPackagingDataAnalytical2018.
+Another way to increase the chance of reproducibility is to develop an R-package, which is a self-contained set of files with well-defined meta data, for each analysis.
+The reasoning is that by abiding the strict rules of package development and specifieng all dependencies a high degree of stability can be archieved (e.g., @marwickPackagingDataAnalytical2018).
 
-However, beside these tools that ease the process of creating a workflow on the researchers side, we see increased interest in changing how research is published and used[@perkelToolkitDataTransparency2018], with *life code* [@perkelPioneeringLivecodeArticle2019] and *continues integration* [@beaulieu-jonesReproducibilityComputationalWorkflows2017, @yenniDevelopingModernData2018].
-For such further technical advancements reproducibility is key and researcher using this workflow should be well prepared for such era of interactive publication.
+However, beside these tools that ease the process of creating a workflow on the researchers side, we see increased interest in changing how research is published and used[@perkelToolkitDataTransparency2018], with *life code* [@perkelPioneeringLivecodeArticle2019] and *continuous integration* [@beaulieu-jonesReproducibilityComputationalWorkflows2017, @yenniDevelopingModernData2018].
+These techniques give us a glimpse of a paradigm shift from static to dynamic, interactive and living publications.
 
@@
 
-We are aware that implementing this workflow is not straightforward and the difficulty of its implementation may vary by platform.
+We are aware that implementing the proposed workflow is not straightforward and the difficulty of its implementation may vary by platform.
 For example, already the installation of all tools is easier on POSIX-compatible platforms such as Unix, Linux, or macOS (but not Windows).

### 2019-08-30 rr-flow.Rmd
added to literate programming added refs to R packages

@@
 To not only create fully reproducible results but also fully reproducible reports, one needs to resort to the Literate Programming paradigm [@knuth1984literate], in which human-readable language and computer-code are interspersed to create dynamic documents whose order follow the logic of thought rather then the order of the computer.
-R-Markdown is a simple markup language to create dynamic documents with embedded chunks of R code that can be exported to standard formats such as documents (docx, rtf, epub), presentations (ppt, html) or websites (html).
+R-Markdown is a simple markup language to create dynamic documents with embedded chunks of R code that can be exported to standard formats such as documents (docx, rtf, epub), presentations (ppt, html) or websites (html) using the knitr package [@xie2015, xie2019].
+Figure XXX illustrates RMarkdown syntax and the resulting output.
+Several packages extend the scope of knitr.
+Of particular note are the papaja package [@papaja2018], which additional functions to enable APA-conformable document formatting, including a journal-style final typeset format, and the stargazer package [@stargazer2018], which provides journal-ready tables and reports of statistical models.
+
+<FIGURE 1 HERE>
 

### 2019-09-03 R/rmarkdown.Rmd
add image of rmarkdown closes #38

---
title: "An Example"
author: "Aaron Peikert & Andreas Brandmaier"
date: "9/3/2019"
output: pdf_document
papersize: "a6"
geometry:
  - "landscape"
  - margin=0.1cm
---

```{r setup}
library("knitr")
library("papaja")
```

# Silly Heading

```{r t-test}
data("sleep")
result <- t.test(extra ~ group, data = sleep)
```

The data was never before anylized in this way. Except in `help(t.test)`.

`r apa_print.htest(result)$full_result`

I can now assert that what I *believe* to be true
--- that there is a difference in mean between the groups ---
is `r ifelse(result$p.value > .025, "**not**", "")` supported by data.

### 2019-09-03 rr-flow.Rmd
correct spelling
add image of rmarkdown closes #38
fix little error

@@
 
-Sometimes external requirements make it impossible that the data cannot be stored and shared together with the scripts.
+Sometimes external requirements make it impossible that the data can be stored and shared together with the scripts.
 In most of the cases we have seen, these are either space constraints or privacy considerations.
@@
 checksum <- digest::sha1(x)
-if (checksum != "740e88404c04af1d67ffc87b008231c093539cc1") 
+if (checksum != "740e88404c04af1d67ffc87b008231c093539cc1"){
   error("Wrong data file loaded or data has been tempered with!")
+}
 ```
@@
 
-<FIGURE 1 HERE>
+![Source of an RMarkdown](Images/rmarkdown.png)
+
+![Rendered result](R/rmarkdown.pdf)
 
@@
 
-Replace `apt` with your package manager. Possiblie you need elevated rights in this case add `sudo` before that.
+Replace `apt` with your package manager. Possiblie you need elevated rights; in this case add `sudo` before that.
 

### 2019-09-03 rr-flow.Rmd
remove make on windows section
break line so it fits on page
fix make docker relationship

@@
 This job gets comfortable with `make`, since it allows you to create (computational) recipes too (re-)create files.
-A downside of `make` is, that the installation on Windows is not straightforward.
-However, if `docker` (see later sections) is available this is not an issue, since `make` ships with (almost) every `docker` image.
 
@@
 ```{bash, eval=FALSE}
-docker run --rm -it -e PASSWORD=<YOUR_PASS> -p 8787:8787 -v /path/to/project:/home/rstudio reproducible-research
+docker run --rm -it -e PASSWORD=<YOUR_PASS> -p 8787:8787 -v
+/path/to/project:/home/rstudio reproducible-research
 ```
@@
 Since `docker`'s commands grow long and are tedius to type, we advocate to use some automatic way to generate them.
-Unfortunatly is the most comfortable way only availible on linux, which uses `make DOCKER=TRUE` together with a Makefile like the one used for that article.
+Fortunatly one can use `make` to automaticly generate the `docker` commands, e.g.
+the (simplified) Makefile for this paper allows command after `$(run)` to be condtionally passed through docker if one calls `make DOCKER=TRUE`:
 
@@
 build: Dockerfile
-	sudo docker build -t $(projekt) .
-
-clean:
-	Ruby/clean.rb
-	rm -rf kitematic
-.PHONY: clean
+	docker build -t $(projekt) .
 

### 2019-09-03 rr-flow.Rmd
add link to showcase

@@
 
-There is a pletora of linux distribuation and almost as many package managers.
+There is a pletora of linux distributions and almost as many package managers.
 However to our knowledge there is no (recent) Linux where `git`, `make` and `docker` are not availible.
@@
 
-We provide a reproducible analysis as working example via github as a testbed for users.
+We provide a reproducible analysis as working example via [github](https://github.com/aaronpeikert/workflow-showcase) as a testbed for users.
 The example shows a minimalistic analysis of the 'Considerations of Future Consequences (CFC) Scale'.

### 2019-09-03 rr-flow.Rmd
add minor corrections

@@
 
-Here, we combine four sofrware tools, whose interplay can guarantuee full and sustainable computational reproducibility of data analyses and their reporting. The major four elements are:
+Here, we combine four software tools, whose interplay can guarantuee full and sustainable computational reproducibility of data analyses and their reporting. The major four elements are:
 
@@
 We would like to illustrate this with a more concrete example (the full R code to reproduce this non-reproducibility is provided in the github repository of this manuscript).
-Using a simulated dataset with two variables ´x´ and ´y´, we ran a linear regression model with R's ´lm()´ function regressing ´x´ on ´y´.
-Using the ´boot´ package [@boot], we bootstrapped the 95% confidence intervals around the regression estimate with 1000 bootstrap samples to evaluate whether zero was inside the confindence interval.
+Using a simulated dataset with two variables `x` and `y`, we ran a linear regression model with R's `lm()` function regressing `x` on `y`.
+Using the `boot` package [@boot], we bootstrapped the 95% confidence intervals around the regression estimate with 1000 bootstrap samples to evaluate whether zero was inside the confindence interval.
 To make the analysis reproducible, we set a random seed before we bootstrapped.
@@
 The use of `rmarkdown` within a project, tracked with `git` can be simplified with the [workflowr-package](https://github.com/jdblischak/workflowr) [@workflowr].
-The [drake-package](https://github.com/ropensci/drake) [@drake] is directly inspired by `make` and does only require knowledge in R, but can only handle dependencies within R.
+The [drake-package](https://github.com/ropensci/drake) [@drake] is directly inspired by `make` but does only require knowledge in R, but can only handle dependencies within R.
 The [holepunch-package](https://karthik.github.io/holepunch/) [@holepunch] automizes the use of docker and uses [binder](https://mybinder.org) to move the analysis to the cloud so that no local installation of docker is required.
 It is well suited for simple analysis with low computational demands, because of memory and computing time limitations of binder.
-There are several alternatives to docker which fix the dependancies on R-packages.
-[packrat](https://rstudio.github.io/packrat/) [@packrat] is a way to fix package version via installing locally in the project, however it does not garuantees more then a fixed set of packages (Not R-version or system dependencies).
+There are several alternatives to docker which fix the dependencies on R-packages.
+[packrat](https://rstudio.github.io/packrat/) [@packrat] is a way to fix package version via local installations of packages in the project, however it does not garuantees more then a fixed set of packages (Not R-version or system dependencies).
 Similar approaches are taken by [jetpack](https://github.com/ankane/jetpack) [@jetpack], [miniCRAN](https://cran.r-project.org/web/packages/miniCRAN/vignettes/miniCRAN-introduction.html) [@miniCRAN] and [checkpoint](https://cran.r-project.org/web/packages/checkpoint/vignettes/checkpoint.html) [@checkpoint].
@@
 How to best share a reproducible workflow?
-We believe that, ideally, a service provider should be found that guarantuees permanent and reliable hosting of the reproducible workflow, such as the Open Science Framework [@fosterOpenScienceFramework2017] or PsychData (citation needed), that is independ of Gihub, Docker Hub & MRAN.
+We believe that, ideally, a service provider should be found that guarantuees permanent and reliable hosting of the reproducible workflow, such as the Open Science Framework [@fosterOpenScienceFramework2017] or PsychData (citation needed), that is independ of Github, Docker Hub & MRAN.
 Second, to ensure that other users are legally able to benefit from the shared materials, authors must choose an appropriate license format.

### 2019-09-03 rr-flow.Rmd
remove psych-data as there is no citation availible

@@
 How to best share a reproducible workflow?
-We believe that, ideally, a service provider should be found that guarantuees permanent and reliable hosting of the reproducible workflow, such as the Open Science Framework [@fosterOpenScienceFramework2017] or PsychData (citation needed), that is independ of Github, Docker Hub & MRAN.
+We believe that, ideally, a service provider should be found that guarantuees permanent and reliable hosting of the reproducible workflow, such as the Open Science Framework [@fosterOpenScienceFramework2017], that is independ of Github, Docker Hub & MRAN.
 Second, to ensure that other users are legally able to benefit from the shared materials, authors must choose an appropriate license format.

### 2019-09-03 abstract.Rmd
add productivity #26
minor revision

--- 2019-08-23 abstract.Rmd
@@
 In this tutorial, we describe a principled workflow to ensure long-term reproducibility of R-based data analyses.
-The workflow is based on established tools and practices from software engineering.
-It combines the benefits of various open-source software tools including Rmarkdown, git, make and docker whose interplay ensures seamless integration of version management, (APA-conformable) dynamic report generation and full cross-platform and long-term computational reproducibility.
+The workflow leverages established tools and practices from software engineering.
+It combines the benefits of various open-source software tools including Rmarkdown, git, make, and docker, whose interplay ensures seamless integration of version management, APA-conformable dynamic report generation and full cross-platform and long-term computational reproducibility.
 The workflow ensures meeting the primary goals that 1) the reporting of analysis results is perfectly consistent with the actual analysis results (dynamic report generation), the analysis exactly reproduces at a later time even if the computing and software platform is updated or changed (computational reproducibility), and 3) changes at any time (during development and post-publication) are tracked, tagged, and documented while earlier versions of both data and code remain accessible.
-While dynamic document generation is increasingly recognized as a tool to ensure reproducibility, we demonstrate with practical examples that dynamic documents alone are not sufficient to ensure long-term computational reproducibility.
+While the research community increasingly recognizes dynamic document generation as a tool to ensure reproducibility, we demonstrate with practical examples that dynamic documents alone are not sufficient to ensure long-term computational reproducibility.
+By ensuring full computational reproducibility together with version management, dependence management and literate programming,  the workflow increases scientific productivity by facilitating later reuse.

### 2019-09-03 rr-flow.Rmd
minor revision

--- 2019-09-03 rr-flow.Rmd
@@
 
-In this tutorial, we describe a principled workflow to ensure long-term reproducibility of
+In this tutorial, we describe a principled workflow to ensure long-term and cross-platform reproducibility of
 R-based data analyses.
@@
 
-Sometimes external requirements make it impossible that the data can be stored and shared together with the scripts.
+Sometimes external requirements make it impossible that the data cannot be stored and shared together with the scripts.
 In most of the cases we have seen, these are either space constraints or privacy considerations.
@@
 checksum <- digest::sha1(x)
-if (checksum != "740e88404c04af1d67ffc87b008231c093539cc1"){
+if (checksum != "740e88404c04af1d67ffc87b008231c093539cc1") 
 	error("Wrong data file loaded or data has been tempered with!")
-}
 ```
@@
 
-![Source of an RMarkdown](Images/rmarkdown.png)
-
-![Rendered result](R/rmarkdown.pdf)
+<FIGURE 1 HERE>
 
@@
 This job gets comfortable with `make`, since it allows you to create (computational) recipes too (re-)create files.
+A downside of `make` is, that the installation on Windows is not straightforward.
+However, if `docker` (see later sections) is available this is not an issue, since `make` ships with (almost) every `docker` image.
 
@@
 ```{bash, eval=FALSE}
-docker run --rm -it -e PASSWORD=<YOUR_PASS> -p 8787:8787 -v
-/path/to/project:/home/rstudio reproducible-research
+docker run --rm -it -e PASSWORD=<YOUR_PASS> -p 8787:8787 -v /path/to/project:/home/rstudio reproducible-research
 ```
@@
 Since `docker`'s commands grow long and are tedius to type, we advocate to use some automatic way to generate them.
-Fortunatly one can use `make` to automaticly generate the `docker` commands, e.g.
-the (simplified) Makefile for this paper allows command after `$(run)` to be condtionally passed through docker if one calls `make DOCKER=TRUE`:
+Unfortunatly is the most comfortable way only availible on linux, which uses `make DOCKER=TRUE` together with a Makefile like the one used for that article.
 
@@
 build: Dockerfile
-	docker build -t $(projekt) .
+	sudo docker build -t $(projekt) .
+
+clean:
+	Ruby/clean.rb
+	rm -rf kitematic
+.PHONY: clean
 
@@
 
-There is a pletora of linux distributions and almost as many package managers.
+There is a pletora of linux distribuation and almost as many package managers.
 However to our knowledge there is no (recent) Linux where `git`, `make` and `docker` are not availible.
@@
 
-Replace `apt` with your package manager. Possiblie you need elevated rights; in this case add `sudo` before that.
+Replace `apt` with your package manager. Possiblie you need elevated rights in this case add `sudo` before that.
 
@@
 
-We provide a reproducible analysis as working example via [github](https://github.com/aaronpeikert/workflow-showcase) as a testbed for users.
+We provide a reproducible analysis as working example via github as a testbed for users.
 The example shows a minimalistic analysis of the 'Considerations of Future Consequences (CFC) Scale'.

### 2019-09-03 R/gen_jou_version.R
attempt to make a journal version #40

@@ -1,0 +1,6 @@
+library(readr)
+library(here)
+doc <- read_lines(here("rr-flow.Rmd"))
+pos_classoption <- grep("^classoption", doc)
+doc[pos_classoption] <- 'classoption: "jou"'
+write_lines(doc, here("rr-flow-jou.Rmd"))

### 2019-09-03 R/rmarkdown.Rmd
add border to pdf to distingish it from backround closes #39

@@
   - "landscape"
-  - margin=0.1cm
+  - margin=0.5cm
+header-includes:
+   - \usepackage[center,dvips,frame]{crop}
 ---

### 2019-09-04 rr-flow.Rmd
wrap code for jou version #40

--- 2019-09-03 rr-flow.Rmd
@@
 knitr::opts_chunk$set(echo = TRUE)
+knitr::opts_chunk$set(linewidth = Inf) # see wrap_code.R
+
 if(!require("pacman"))install.packages("pacman")
@@
 knitr::read_chunk(here("R", "bootstrap.R"))
+
+source(here("R", "wrap_code.R"))
 ```

### 2019-09-04 R/gen_jou_version.R

@@ -1,6 +1,12 @@
 library(readr)
 library(here)
 doc <- read_lines(here("rr-flow.Rmd"))
+
 pos_classoption <- grep("^classoption", doc)
 doc[pos_classoption] <- 'classoption: "jou"'
+
+pos_linewidth <- grep("knitr::opts_chunk$set(linewidth = Inf)",
+                      doc,
+                      fixed = TRUE)
+doc[pos_linewidth] <- "knitr::opts_chunk$set(linewidth = 40)"
 write_lines(doc, here("rr-flow-jou.Rmd"))

### 2019-09-04 rr-flow.Rmd
declare files as bash, so they get wrapped

@@
 
-```
+```{bash, eval=FALSE}
 target: dependency1 dependency2
@@
 
-```
+```{bash, eval=FALSE}
 analysis.pdf: data/clean.csv
@@
 
-```
+```{bash, eval=FALSE}
 all: analysis.pdf
@@
 
-```
+```{bash, eval=FALSE}
 FROM rocker/verse:3.6.1
@@
 
-```
+```{bash, eval=FALSE}
 projekt := $(shell basename `git rev-parse --show-toplevel`)

### 2019-09-04 R/gen_jou_version.R
add note about wrapping closes #40

@@ -9,4 +9,8 @@ pos_linewidth <- grep("knitr::opts_chunk$set(linewidth = Inf)",
                       doc,
                       fixed = TRUE)
 doc[pos_linewidth] <- "knitr::opts_chunk$set(linewidth = 40)"
+pos_authornote <- grep("^authornote", doc)
+doc[pos_authornote] <- paste(doc[pos_authornote],
+                             "Please note that code longer then 40 characters is continued on the next line and indended by four spaces.")
+
 write_lines(doc, here("rr-flow-jou.Rmd"))

### 2019-09-04 rr-flow.Rmd
add windows enterprise to list of supported versions
automatically confirm install for choco

@@
 Windows has the biggest hurdle to install what is needed.
-An essential requirement is that you have Windows Pro, Education or Server installed, as Microsoft prevents the use of Docker on Windows Home.
+An essential requirement is that you have Windows Pro, Enterprise, Education or Server installed, as Microsoft prevents the use of Docker on Windows Home.
 However, having met this requirement, there is a great package manager for Windows called chocolately, which you can install from: [https://chocolatey.org/install](https://chocolatey.org/install).
@@
 ```{bash, eval=FALSE}
-choco install git make docker-desktop
+choco install -y git make docker-desktop
 ```

### 2019-09-04 rr-flow.Rmd
add liftr package closes #41

@@
 The [drake-package](https://github.com/ropensci/drake) [@drake] is directly inspired by `make` but does only require knowledge in R, but can only handle dependencies within R.
-The [holepunch-package](https://karthik.github.io/holepunch/) [@holepunch] automizes the use of docker and uses [binder](https://mybinder.org) to move the analysis to the cloud so that no local installation of docker is required.
-It is well suited for simple analysis with low computational demands, because of memory and computing time limitations of binder.
+The [liftr-package](https://cran.r-project.org/web/packages/liftr/vignettes/liftr-intro.html) [@liftr] and the [holepunch-package](https://karthik.github.io/holepunch/) [@holepunch] automize the use of docker.
+The former is perfectly compatible with the described workflow and we recommend it to user uncomftable with docker.
+The later uses [binder](https://mybinder.org) to move the analysis to the cloud so that no local installation of docker is required.
+holepunch is well suited for simple analysis with low computational demands, because of memory and computing time limitations of binder.
 There are several alternatives to docker which fix the dependencies on R-packages.

### 2019-09-04 R/bootstrap.R
smaller output for bootstrap closes #42

@@ -11,6 +11,17 @@ bs <- function(formula, data, indices) {
   fit <- lm(formula, data=d)
   return(coef(fit)) 
 }
+confint.boot <- function (object, parm, level = 0.95, type, ...) {
+  if (!requireNamespace("boot")) stop("boot package is missing")
+  if(length(parm) != 1)stop("Only one parameter")
+  bci <- boot::boot.ci(object, conf = level, type = type, index = parm, ...)
+  which <- 4:5 #lower and upper
+  int <- as.vector(t(bci[[type]][, which]))
+  lev <- (1 - .95)/2
+  levs <- c(lev, 1-lev)
+  names(int) <- paste0(round(100 * levs, 1), " %")
+  return(int)
+}
 
 #-----find-seed----
 bcis35 <- read.csv("../data/bcis-3.5.0.csv")
@@ -27,5 +38,4 @@ results <- boot(data=simdata, statistic=bs,
                   R=1000, formula=y~1+x)
   
 # get beta estimates' confidence intervals
-bci <- boot.ci(results,type="bca", index=2) # b estimate
-print(bci)
+round(confint(results, type = "bca", parm = 2), 4) # parm = 2 -> b

### 2019-09-15 R/rmarkdown.Rmd

--- 2019-09-03 R/rmarkdown.Rmd
@@
 author: "Aaron Peikert & Andreas Brandmaier"
-date: "9/3/2019"
+date: '`r format(Sys.time(), "%d %B, %Y")`'
 output: pdf_document

### 2019-09-15 rr-flow.Rmd
recomend post instalation steps on linux #19

@@
 To not only create fully reproducible results but also fully reproducible reports, one needs to resort to the Literate Programming paradigm [@knuth1984literate], in which human-readable language and computer-code are interspersed to create dynamic documents whose order follow the logic of thought rather then the order of the computer.
-R-Markdown is a simple markup language to create dynamic documents with embedded chunks of R code that can be exported to standard formats such as documents (docx, rtf, epub), presentations (ppt, html) or websites (html) using the knitr package [@xie2015, xie2019].
-Figure XXX illustrates RMarkdown syntax and the resulting output.
+R-Markdown is a simple markup language to create dynamic documents with embedded chunks of R code that can be exported to standard formats such as documents (docx, rtf, epub), presentations (ppt, html) or websites (html) using the knitr package [@xie2015; @xie2019].
+Figure 1 illustrates RMarkdown syntax and Figure 2 the resulting output.
 Several packages extend the scope of knitr.
@@
 A commit represents a complete copy of the state of the tracked files.
-Each commit has a unique identifier (a hash) and a human-readable description.
+Each commit has a unique identifier (a hash) and a human-readable description (commit message).
 Going back to one state is as easy as traversing the history of all commits; it is possible to visually track changes between different versions.
-Finding a specific version can be made even easier by tagging snapshots, e.g., as "submission," "preprint," or "publishged." This collection of snapshots is called a "repository," which ideally tracks your entire R project.
+Finding a specific version can be made even easier by tagging snapshots, e.g., as "submission," "preprint," or "published." This collection of snapshots is called a "repository," which ideally tracks your entire R project.
 
@@
 
-The basic rocker images are extendable by a Dockerfile, e.g., the basis for docker image for this article is the following Dockerfile:
+The basic rocker images are extendable by a Dockerfile, e.g., the basis for this articles docker image is the following Dockerfile:
 
@@
 The command `docker build -t image-name` creates an image named `image-name` from the Dockerfile in the project.
-Following there are two principal ways to share a docker image, either by sharing the Dockerfile which created the image or by sharing the image itself e.g., through [dockerhub](https://hub.docker.com).
+A way to automatically find out the dependencies and genereate a docker image from them is provided in the [liftr-package](https://cran.r-project.org/web/packages/liftr/vignettes/liftr-intro.html) [@liftr].
+
+There are two principal ways to share a docker image, either by sharing the Dockerfile which created the image or by sharing the image itself e.g., through [dockerhub](https://hub.docker.com).
 While both ways guarantee a replicable computational environment, sharing the Dockerfile is more transparent, and git tracks changes to it.
@@
 
-The `--rm` flag means to not save the state of the docker after the command finishes in an image.
+The `--rm` flag means to not save the state of the docker after the command finishes in an container.
 The `-it` flag means to accept inputs and return outputs to the terminal.
@@
 ```{bash, eval=FALSE}
-projekt := $(shell basename `git rev-parse --show-toplevel`)
-current_dir := $(shell pwd)
+projekt := reproducible-research
+current_dir := $(shell git rev-parse --show-toplevel)
+home_dir := $(shell git rev-parse --show-toplevel)
 uid = $(shell id -u)
@@
 ifeq ($(DOCKER),TRUE)
-	run:=docker run --rm --user $(uid) -v $(current_dir):/home/rstudio $(projekt)
+	run:=docker run --rm --user $(uid) -v $(home_dir):/home/rstudio $(projekt)
 	current_dir=/home/rstudio
@@
 
-all: rr-flow.html
+all: rr-flow.pdf
 
@@
 
-rr-flow.html: rr-flow.Rmd
+rr-flow.pdf: rr-flow.Rmd reproducible-research.bib
 	$(run) Rscript -e 'rmarkdown::render("$(current_dir)/$<")'
@@
 
-Replace `apt` with your package manager. Possiblie you need elevated rights; in this case add `sudo` before that.
+Replace `apt` with your package manager.
+Possiblie you need elevated rights; in this case add `sudo` before that.
+`docker` also needs elevated rights, we recomend therefore to add your user to the `docker` group, following the [documentation of docker](https://docs.docker.com/install/linux/linux-postinstall/#manage-docker-as-a-non-root-user).
 
@@
 Each tool reduces the number of ways things can go wrong.
-Dynamic reporting with `RMarkdown` guarantuees consistency between computational results and their reporting; Version control with `git` guarantuees permanence and consistency across multiple versions of data and code; Dependency management with `make` guarantuees defined entry-points and dependency resolition; Containerization with `docker` guarantuees reproducibility of the full computational environment.
+Dynamic reporting with `RMarkdown` guarantuees consistency between computational results and their reporting; Version control with `git` guarantuees permanence and consistency across multiple versions of data and code; Dependency management with `make` guarantuees defined entry-points and dependency resolution; containerization with `docker` guarantuees reproducibility of the full computational environment.
 This tightening of loose ends, we believe, does not restrict researchers but enables them to operate on a solid basis to deliver sound research.
@@
 The [liftr-package](https://cran.r-project.org/web/packages/liftr/vignettes/liftr-intro.html) [@liftr] and the [holepunch-package](https://karthik.github.io/holepunch/) [@holepunch] automize the use of docker.
-The former is perfectly compatible with the described workflow and we recommend it to user uncomftable with docker.
+The former is perfectly compatible with the described workflow and we recommend it to users uncomftable with docker.
 The later uses [binder](https://mybinder.org) to move the analysis to the cloud so that no local installation of docker is required.
@@
 Typically, there is currently no single license that works for code, data, and media (such as text or figures).
-We encourage authors to choose appropriate license forms that do not hinder others from freely downloading, using, and moifying the shared workflows and materials while, at the same time, ensuring recognition for the time and effort invested in creating the workflow in the first place.
+We encourage authors to choose appropriate license forms that do not hinder others from freely downloading, using, and modifying the shared workflows and materials while, at the same time, ensuring recognition for the time and effort invested in creating the workflow in the first place.
 In our experience, the Creative Commons license (CC-BY) is often appropriate for sharing texts, rmarkdown files, generated figures, and other media whereas scripts and any other computer code is often best shared according to the MIT license.

### 2019-10-02 rr-flow.Rmd
add redoc, solve #44

@@
 In our own experience, it is often not possible to convince all co-authors to switch to a different document processing environment, such as RMarkdown.
-That is, we have experienced that after writing up the first draft in Rmarkdown, one must eventually switch to a dynamically generated Word-file that is from then on used as static file that is used as basis for multiple iterations among the co-authors.
-This means, that ideally, one would have to manually merge the final Word file back to the RMarkdown souce, which generally is only possible manually.
-But we believe that even if this final step does not happen that - as long as dynamically generated elements are not manipulated - ensuring the reproducibility of the analysis and the initial draft is a big step forward.
-Making such static Word-file more dynamic is is possible with the officer package [@gohel2019officer], which allows the manipulation of  Word (.docx) and PowerPoint (.pptx) documents.
-That is, given a formatted static template file, one can dynamically paste images, tables and text into documents from R.
-We consider such static file with dynamic content, less elegant but potentially more practical, when not all authors can adhere to the described process.
+That is, we have experienced that after writing up the first draft in Rmarkdown, one must eventually to genereate a Word-file that is from then on used as static file serving as a basis for multiple iterations among the co-authors.
+Reteining reproduciblity in such situation requires tediues manuel synchronization.
+This labor can be saved with the [rodoc package](https://noamross.github.io/redoc/articles/mixed-workflows-with-redoc.html), which enables a two-way workflow between Word and Rmarkdown.
+The conversion between Rmarkdown and Word retains all changes and even supports tracked changes.
+We consider such workflow less elegant since changes can be hard to track in git, but potentially more practical, when not all authors can adhere to the described process.
 

### 2019-10-02 rr-flow.Rmd
add compendium and related files solves #45

--- 2019-09-15 rr-flow.Rmd
@@
 
-This folder, where everything resides that you need for analysis (code and data), is referred to as a 'project'.
+This folder, where everything resides that you need for analysis (code and data), is referred to as a 'project' (or sometimes 'research compendium').
 Working with projects is exceptionally comfortable with RStudio, an integrated development environment (IDE) for R.
-It is useful to organize a data-analysis project in way that strictly segregates (raw) data and code by placing them in directories called `data` and `R` [see Section 4 @marwickPackagingDataAnalytical2018]; there are also tools that automatize the standardized creation of folder structures such as `workflowr` [@workflowr].
+It is useful to organize a data-analysis project in way that strictly segregates (raw) data and code by placing them in directories called `data` and `R` [see Section 4 @marwickPackagingDataAnalytical2018]; there are tools that automatize the standardized creation of more sophisticated folder structures such as `workflowr` [@workflowr] or `rrtools` [@rrtools].
 
@@
 We believe that, ideally, a service provider should be found that guarantuees permanent and reliable hosting of the reproducible workflow, such as the Open Science Framework [@fosterOpenScienceFramework2017], that is independ of Github, Docker Hub & MRAN.
-Second, to ensure that other users are legally able to benefit from the shared materials, authors must choose an appropriate license format.
+Whereever the project is published a Readme file should be included that describes what it is about and how to reproduce the presented results.
+To ensure that other users are legally able to benefit from the shared materials, authors must choose an appropriate license format.
 Typically, there is currently no single license that works for code, data, and media (such as text or figures).
@@
 These licenses are also in line with the recommendations by the Reproducibile Research Standard [@stoddenEnhancingReproducibilityComputational2016; @stoddenEnablingReproducibleResearch2009].
+To facilitate a inclusive environment we recommend to name all contributer and include a [Code of Conduct](https://opensource.guide/code-of-conduct/).
 

### 2019-10-02 rr-flow.Rmd
add missing citation

--- 2019-10-02 rr-flow.Rmd
@@
 Reteining reproduciblity in such situation requires tediues manuel synchronization.
-This labor can be saved with the [rodoc package](https://noamross.github.io/redoc/articles/mixed-workflows-with-redoc.html), which enables a two-way workflow between Word and Rmarkdown.
+This labor can be saved with the [rodoc package](https://noamross.github.io/redoc/articles/mixed-workflows-with-redoc.html) [@redoc], which enables a two-way workflow between Word and Rmarkdown.
 The conversion between Rmarkdown and Word retains all changes and even supports tracked changes.

### 2019-10-24 rr-flow.Rmd
some fixes

--- 2019-10-02 rr-flow.Rmd
@@
 authornote: This paper is fully reproducible using the workflow described in this paper.
-All materials can be found on the accompaniying Github repository <https://github.com/aaronpeikert/reproducible-research/>.
+All materials can be found on the accompanying GitHub repository <https://github.com/aaronpeikert/reproducible-research/>.
 To provide feedback or any other type of comment or questions regarding our workflow, please add an issue to the Github repository of our paper [https://github.com/aaronpeikert/reproducible-research/issues](https://github.com/aaronpeikert/reproducible-research/issues).
@@
 R-based data analyses.
-In this context, reproducibility means to us that scientific results together with their computational scientific workflows are reported fully, transparently and remain sustainably available, such that the workflow can be re-run by a different person and/or later in time and the results will be identical to the ones intially reported [@herouxCompatibleReproducibilityTaxonomy2018; @claerboutElectronicDocumentsGive1992].
-The need to ensure reproducibility directly follows from commonly accepted rules of good scientific practice (such as the guidelines of the German Research Foundation; @dfg2019) and ensuring reproducibility is a prerequisite for replicability and a means to increase trustworthiness of empirical results [@epskamp2019rep].
+Reproducibility means to us that scientific results together with their computational scientific workflows are reported fully, transparently and sustainably remain available, such that the workflow can be re-run by a different person and/or later in time and the results will be identical to the ones intially reported [@herouxCompatibleReproducibilityTaxonomy2018; @claerboutElectronicDocumentsGive1992].
+The need to ensure reproducibility directly follows from commonly accepted rules of good scientific practice (such as the guidelines of the German Research Foundation; @dfg2019).
+Ensuring reproducibility is a prerequisite for replicability and a means to increase trustworthiness of empirical results [@epskamp2019rep].
 
-Here, we combine four software tools, whose interplay can guarantuee full and sustainable computational reproducibility of data analyses and their reporting. The major four elements are:
-
-Literate Programming
-: Interwoven text and computer code that together forms a dynamically generated, human-readable document.
-: Example: One document that includes text that describes a statistical test, code that is used to execute the statistical and the output of the code in the same document.
-
-Version Control
-: The management of change of documents/code.
-: Example: Changes of the analysis code or even single parameters are documented both at development and post-publication.
-
-Dependency Tracking
-: How different parts of a manuscript/analysis relate and how these dependencies resolve via execution of code.
-: Example: The manuscript depends on the statistical analysis, which depends on a preprocessing pipeline, which depends on a certain subset of the data.
-
-Containerization
-: A virtual computational environment that does encapsulate all needed software. More formally called operating-system-level virtualization.
-: Example: A statistical analysis depends on the used statistical software and a virtual machine bundles them together.
-
-The scientific community widely recognizes that these tools enhance reproducibility [@piccoloToolsTechniquesComputational2016], and provides guidelines for their use (Literate Programming: @ruleTenSimpleRules2019; Version Control: @barbaHardRoadReproducibility2016; Dependency Management: @askrenUsingMakeReproducible2016; Containerization: @clyburne-sherinComputationalReproducibilityContainers2018).
-However, only a workflow using them in unison allows confidence in reproducing a scientific report.
-Within the community of users of the R Programming Language for each purpose, a specific software solution is the most popular.
+Here, we combine four software tools, whose interplay can guarantuee full and sustainable computational reproducibility of data analyses and their reporting.
+There are various tools to enhance reproducibility [@piccoloToolsTechniquesComputational2016], of which we believe four of them are particularly important: literate programming: @ruleTenSimpleRules2019, version control @barbaHardRoadReproducibility2016, dependency management @askrenUsingMakeReproducible2016 and containerization @clyburne-sherinComputationalReproducibilityContainers2018.
+Importantly, only a workflow using all four in unison allows confidence in reproducing a scientific report.
+Even though there are various implementations of these concepts, we believe that for the R environment, the following four implementations are best suited.
 For literate programming, `rmarkdown`, for version control `git`, for dependency management `make` and for container `docker`.
@@
 
-This folder, where everything resides that you need for analysis (code and data), is referred to as a 'project' (or sometimes 'research compendium').
+This folder, where everything resides that you need for analysis (code and data), is referred to as a 'project'.
 Working with projects is exceptionally comfortable with RStudio, an integrated development environment (IDE) for R.
-It is useful to organize a data-analysis project in way that strictly segregates (raw) data and code by placing them in directories called `data` and `R` [see Section 4 @marwickPackagingDataAnalytical2018]; there are tools that automatize the standardized creation of more sophisticated folder structures such as `workflowr` [@workflowr] or `rrtools` [@rrtools].
+It is useful to organize a data-analysis project in way that strictly segregates (raw) data and code by placing them in directories called `data` and `R` [see Section 4 @marwickPackagingDataAnalytical2018]; there are also tools that automatize the standardized creation of folder structures such as `workflowr` [@workflowr].
 
@@
 We believe that, ideally, a service provider should be found that guarantuees permanent and reliable hosting of the reproducible workflow, such as the Open Science Framework [@fosterOpenScienceFramework2017], that is independ of Github, Docker Hub & MRAN.
-Whereever the project is published a Readme file should be included that describes what it is about and how to reproduce the presented results.
-To ensure that other users are legally able to benefit from the shared materials, authors must choose an appropriate license format.
+Second, to ensure that other users are legally able to benefit from the shared materials, authors must choose an appropriate license format.
 Typically, there is currently no single license that works for code, data, and media (such as text or figures).
@@
 These licenses are also in line with the recommendations by the Reproducibile Research Standard [@stoddenEnhancingReproducibilityComputational2016; @stoddenEnablingReproducibleResearch2009].
-To facilitate a inclusive environment we recommend to name all contributer and include a [Code of Conduct](https://opensource.guide/code-of-conduct/).
 

### 2019-10-24 rr-flow.Rmd
added disclosures according to AMPPS submission guidelines
added author contributions and conflict of interests

@@
 These principles are what guides the use of `rmarkdown`, `git`, `make`, `docker`.
+
+# Disclosures
+
+This paper is fully reproducible using the workflow described in this paper.
+All materials, including the manuscript source and all R codes and files used to generate this manuscript, can be found on the accompanying GitHub repository <https://github.com/aaronpeikert/reproducible-research/>.
 
@@
 
+## Author Contributions
+
+Aaron Peikert and Andres Brandmaier jointly generated the idea for the manuscript.
+Aaron Peikert implemented the workflow.
+Aaron Peikert and Andreas Brandmaier jointly generated the R code examples.
+Aaron Peikert wrote the first draft of the manuscript, and both authors critically edited it.
+Both authors approved the final submitted version of the manuscript.
+
+## Conflicts of Interest
+
+The authors declare that there were no conflicts of interest with respect to the authorship or the publication of this article.
+
+\newpage
+
 # References

### 2019-10-28 rr-flow.Rmd
revision of the manuscript

--- 2019-10-24 rr-flow.Rmd
@@
 ---
-title             : "A Reproducible Data Analysis Workflow with RMarkdown, Git, Make, & Docker"
+title             : "A Reproducible Data Analysis Workflow with R Markdown, Git, Make, & Docker"
 shorttitle        : "Reproducible Data Analysis Workflow"
@@
 authornote: This paper is fully reproducible using the workflow described in this paper.
-All materials can be found on the accompanying GitHub repository <https://github.com/aaronpeikert/reproducible-research/>.
-To provide feedback or any other type of comment or questions regarding our workflow, please add an issue to the Github repository of our paper [https://github.com/aaronpeikert/reproducible-research/issues](https://github.com/aaronpeikert/reproducible-research/issues).
+All materials can be found on the accompanying GitHub repository <https://GitHub.com/aaronpeikert/reproducible-research/>.
+To provide feedback or any other type of comment or questions regarding our workflow, please add an issue to the GitHub repository of our paper [https://GitHub.com/aaronpeikert/reproducible-research/issues](https://GitHub.com/aaronpeikert/reproducible-research/issues).
 
@@
 R-based data analyses.
-Reproducibility means to us that scientific results together with their computational scientific workflows are reported fully, transparently and sustainably remain available, such that the workflow can be re-run by a different person and/or later in time and the results will be identical to the ones intially reported [@herouxCompatibleReproducibilityTaxonomy2018; @claerboutElectronicDocumentsGive1992].
+Reproducibility means to us that statistica results together with their computational workflows are reported fully, transparently and sustainably remain available, such that the workflow can be re-run by a different person and/or later in time and the results will be identical to the ones intially reported [@herouxCompatibleReproducibilityTaxonomy2018; @claerboutElectronicDocumentsGive1992].
 The need to ensure reproducibility directly follows from commonly accepted rules of good scientific practice (such as the guidelines of the German Research Foundation; @dfg2019).
@@
 Here, we combine four software tools, whose interplay can guarantuee full and sustainable computational reproducibility of data analyses and their reporting.
-There are various tools to enhance reproducibility [@piccoloToolsTechniquesComputational2016], of which we believe four of them are particularly important: literate programming: @ruleTenSimpleRules2019, version control @barbaHardRoadReproducibility2016, dependency management @askrenUsingMakeReproducible2016 and containerization @clyburne-sherinComputationalReproducibilityContainers2018.
-Importantly, only a workflow using all four in unison allows confidence in reproducing a scientific report.
-Even though there are various implementations of these concepts, we believe that for the R environment, the following four implementations are best suited.
-For literate programming, `rmarkdown`, for version control `git`, for dependency management `make` and for container `docker`.
+There are various tools to enhance reproducibility [@piccoloToolsTechniquesComputational2016], of which we believe four of them are particularly important: literate programming: [@ruleTenSimpleRules2019], version control [@barbaHardRoadReproducibility2016], dependency management @[askrenUsingMakeReproducible2016] and containerization [@clyburne-sherinComputationalReproducibilityContainers2018].
+Importantly, only a workflow using all four concepts in unison guarantuee confidence in reproducing a scientific report.
+Even though there are various implementations of these concepts, we believe that for the R environment <CITATION NEEDED>, the following four implementations are best suited.
+For literate programming, R Markdown <CITATION NEEDED>, for version control Git <CITATION NEEDED>, for dependency management Make <CITATION NEEDED>, and for container Docker <CITATION NEEDED>.
 Each of them serves a meta-scientific valuable goal (reproducibility) and increases the productivity of the researcher.
-Each of these software solutions is highly flexible and powerful, so to master them thoroughly often years of practice are required.
-Luckily the researcher can learn the minimal valuable subset needed to ensure reproducibility with little practice.
+Each of these software solutions is highly flexible and powerful, so to master them fully, a considerable amount of practice is necessary.
+However, for our purposes, it is sufficient to master a minimal valuable subset of functions to ensure reproducibility.
 
-We believe that for sufficient chances of long-term reproducibility, proper documentation is crucial since it answers the question "What steps have to be taken to reproduce the results?".
+To maximize chances of long-term reproducibility, proper documentation of all steps taken and their context is crucial since it answers the question "What steps have to be taken to reproduce the results?".
 The most precise documentation is computer code, and the most productive computer code is code that runs automatically without human interaction.
-These principles are what guides the use of `rmarkdown`, `git`, `make`, `docker`.
+This principle is what guided our choice to use of R Markdown, Git, Make, and Docker.
 
@@
 This paper is fully reproducible using the workflow described in this paper.
-All materials, including the manuscript source and all R codes and files used to generate this manuscript, can be found on the accompanying GitHub repository <https://github.com/aaronpeikert/reproducible-research/>.
+All materials, including the manuscript source and all R codes and files used to generate this manuscript, can be found on the accompanying GitHub repository <https://GitHub.com/aaronpeikert/reproducible-research/>.
 
@@
 
-The first step towards reproducibility is to have an R-script or R-markdown as primary entry point for the analysis, that runs on a local computer without error.
-A process, we assume, that is familiar to the reader.
-A next step is to make sure that all files relevant for the analysis can be moved to another computer.
+The first step towards reproducibility is to have an R script or R Markdown file as primary entry point for the analysis, which runs on a local computer without error.
+Next, one has to make sure that all files relevant for the analysis can be moved to another computer.
 To this end, it is essential that all files are within one folder (or, subfolders within it) and all paths are relative to that folder.
-A robust solution to the problem of making sure that file access does not break across computing platforms are [Rstudio Projects](https://r4ds.had.co.nz/workflow-projects.html) and the `here` package.
+A robust solution to the problem of making sure that file access does not break across computing platforms are [Rstudio projects](https://r4ds.had.co.nz/workflow-projects.html) and the `here` <CITATION NEEDED> package.
 
@@
 
-This folder, where everything resides that you need for analysis (code and data), is referred to as a 'project'.
-Working with projects is exceptionally comfortable with RStudio, an integrated development environment (IDE) for R.
-It is useful to organize a data-analysis project in way that strictly segregates (raw) data and code by placing them in directories called `data` and `R` [see Section 4 @marwickPackagingDataAnalytical2018]; there are also tools that automatize the standardized creation of folder structures such as `workflowr` [@workflowr].
+The folder, in which all fies reside that you need for analysis (code and data), is referred to as a 'project'.
+Working with projects is particularly comfortable with RStudio, an integrated development environment (IDE) for R.
+It is useful to organize a data-analysis project in way that strictly segregates (raw) data and code by placing them in directories called `data` and `R` [see Section 4 in @marwickPackagingDataAnalytical2018]; there are also tools that automatize the standardized creation of folder structures such as `workflowr` [@workflowr].
 
@@
 In these cases, of course, unrestricted reproducibility is not guaranteed.
-If this way of dividing data and scripts into different databases is unavoidable, we recommend that you crosscheck using checksums before loading data.
-To do this, a checksum must be created and filed at the time of the original analysis, and at the time of reproduction the current checksum must be compared with the stored checksum (also called a "hash", e.g., using the functions provided in package `digest`).
+If this way of dividing data and scripts into different databases is unavoidable, we recommend to validate all data files using checksums (also called a "hash", e.g., using the functions provided in package `digest` <CITATION NEEDED>) before analyzing them.
+To do this, a checksum must be created and stored at the time of the original analysis.
+At the time of reproduction, the current checksum must be compared with the stored checksum.
 
@@
 
-Typically, the translation of the human-readable summary of ones computational results, be it an internal report, a presentation, or a journal manuscript, is time-consuming and error-prone.
-To not only create fully reproducible results but also fully reproducible reports, one needs to resort to the Literate Programming paradigm [@knuth1984literate], in which human-readable language and computer-code are interspersed to create dynamic documents whose order follow the logic of thought rather then the order of the computer.
-R-Markdown is a simple markup language to create dynamic documents with embedded chunks of R code that can be exported to standard formats such as documents (docx, rtf, epub), presentations (ppt, html) or websites (html) using the knitr package [@xie2015; @xie2019].
-Figure 1 illustrates RMarkdown syntax and Figure 2 the resulting output.
-Several packages extend the scope of knitr.
-Of particular note are the papaja package [@papaja2018], which additional functions to enable APA-conformable document formatting, including a journal-style final typeset format, and the stargazer package [@stargazer2018], which provides journal-ready tables and reports of statistical models.
+The translation of computational results into a human-readable summary, for example into a technical report, a presentation, or a journal manuscript, is time-consuming and error-prone.
+Typical errors result from copy\&paste mistakes, erroneous rounding, or missed updates of the manuscript when the computer coder has changed.
+To not only create fully reproducible results but also fully reproducible reports, one needs to resort to the Literate Programming paradigm [@knuth1984literate], in which human-readable language and computer-code are interspersed to create dynamic documents whose order follows the logic of thought rather then the order of the computer.
+R Markdown is a simple markup language to create dynamic documents with embedded chunks of R code that can be exported to standard formats such as documents (docx, rtf, epub), presentations (ppt, html) or websites (html) using the knitr package [@xie2015; @xie2019].
+Figure 1 illustrates R Markdown syntax and Figure 2 the resulting output.
+Several packages extend the functionality of knitr.
+Of particular note are the `papaja` package [@papaja2018], which additional functions to enable APA-conformable document formatting, including a journal-style final typeset format, and the `stargazer` package [@stargazer2018], which provides journal-ready tables and reports of statistical models.
+Here is a simple example of how `papaja` can be used to create dynamic reports in APA style:
 
-![Source of an RMarkdown](Images/rmarkdown.png)
+![Source of an R Markdown](Images/R Markdown.png)
 
-![Rendered result](R/rmarkdown.pdf)
+![Rendered result](R/R Markdown.pdf)
 
@@
 In the language of git, these snapshots are "commits".
-A commit represents a complete copy of the state of the tracked files.
+A commit represents a complete copy of the state of all tracked files.
 Each commit has a unique identifier (a hash) and a human-readable description (commit message).
@@
 
-So to keep track of all changes on your local computer, you only need to use `git add` & `git commit` or `git commit -a` to do both.
-These commands need to be executed in the `terminal`, which you can access from within Rstudio (`Shift + Alt + R`).
+To keep track of all changes on your local computer, one only needs to use `git add` & `git commit` or `git commit -a` to do both.
+These commands need to be executed in the `terminal`, which one can access from within Rstudio (`Shift + Alt + R`).
 Rstudio also offers a graphical user interface for Git and, quite likely, this interface is sufficient for almost every interaction with Git.
@@
 
-Now one can check everything that has happened (`git log`) and look at the exact state by supplying the id of the commit or the first few digits of its hash to `git checkout`:
+Now one can inspect all changes (`git log`) and examine any previous state by stating the identifier of the commit to `git checkout`:
 
@@
 Git also makes it especially easy to share and collaborate on a project with other researchers.
-A popular service for sharing via git is [Github](https://github.com).
-Just sharing Git repositories with the public is always free, private repositories (only visible to persons you invite) are [free for researcher](https://help.github.com/en/articles/applying-for-an-educator-or-researcher-discount).
-After creating a user account, you can create a new repository where Github advises you how to upload your repository from the terminal e.g., for this repository (here with user name 'aaronpeikert' and repository name 'reproducible-research'):
+A popular service for sharing materials via Git is [GitHub](https://GitHub.com).
+Just sharing Git repositories with the public is always free, private repositories (only visible to persons you invite) are [free for researchers](https://help.GitHub.com/en/articles/applying-for-an-educator-or-researcher-discount).
+After creating a user account, one can create a new repository where GitHub advises you how to upload your repository from the terminal, e.g., for our repository (here with user name 'aaronpeikert' and repository name 'reproducible-research'):
 
 ```{bash, eval=FALSE}
-git remote add origin git@github.com:aaronpeikert/reproducible-research.git
+git remote add origin git@GitHub.com:aaronpeikert/reproducible-research.git
 git push -u origin master
@@
 ```{bash, eval=FALSE}
-git clone git@github.com:aaronpeikert/reproducible-research.git
+git clone git@GitHub.com:aaronpeikert/reproducible-research.git
 ```
 
-Git and Github can do even more to help you to collaborate with your fellow researchers, such as formal means to track issues and their status (open/closed/resolved) and further means to managing and merging multiple, parallel versions of code (such as branches, pull requests, merges) but this is beyond the already broad scope of this tutorial.
-In particular, github's issue management can be leveraged as post-publication platform to discuss manuscripts and their results [^1].
-Another benefit of using Git and Github is that experimentation is highly encouraged since you can go back to any state quickly and even when you lose access to the file on your computer, everything is backed up on the Github servers.
+Git and GitHub can do even more to help you to collaborate with your fellow researchers, such as formal means to track issues and their status (open/closed/resolved) and further means to managing and merging multiple, parallel versions of code (such as branches, pull requests, merges) but this is beyond the scope of this tutorial.
+In particular, GitHub's issue management can be leveraged as post-publication platform to discuss manuscripts and their results [^1].
+Another benefit of using Git and GitHub is that experimentation is highly encouraged since you can go back to any state quickly and even when you lose access to the file on your computer, everything is backed up on the GitHub servers.
 
-[^1]: To comment on our paper, please add an issue to the Github repository of our paper: <https://github.com/aaronpeikert/reproducible-research/issues>.
+[^1]: To comment on our paper, please add an issue to the GitHub repository of our paper: <https://GitHub.com/aaronpeikert/reproducible-research/issues>.
 
@@
 
-Even when you have obtained a version of the project, and you can ensure that this version is unchanged, you may not know how to reproduce the results, because it is unclear what has to be run in which order.
-This job gets comfortable with `make`, since it allows you to create (computational) recipes too (re-)create files.
+Even when you have obtained a given version of a project to reproduce, and you can ensure that this version is unchanged, you may not know how to reproduce the results, because it may be unclear what files have to be run in which order.
+This is particularly the case when complex preprocessing pipelines are part of the computation and/or external dependencies are many.
+Handling such dependencies is comfortable with Make because it allows you to create (computational) recipes to (re-)create files.
 
@@
 analysis.pdf: data/clean.csv
-  Rscript -e 'rmarkdown::render("analysis.Rmd")'
+  Rscript -e 'R Markdown::render("analysis.Rmd")'
 
@@
 To create `data/clean.csv`, `R/clean.R` and `data/iris.csv` are needed.
-If you type `make analysis.pdf` `make` first checks whether the dependencies do exist and if not whether or it knows how to create them.
-So if `data/clean.csv` does not exist, `make` creates it.
-The same thing happens if one of the dependencies of a target is newer then the target itself, then updates everything that directly or indirectly depends on it.
-So if `data\iris.csv` is newer then `data\clean.csv`, make attempts to recreate `data\clean.csv` and `analysis.pdf`.
+If you type `make analysis.pdf`, `make` first checks whether the dependencies do exist and, if not, creates them.
+Here, if `data/clean.csv` does not exist, `make` creates it.
+Also, if one of the dependencies of a target is newer then the target itself, then Make updates everything that directly or indirectly depends on the target.
+Here, if `data\iris.csv` is newer then `data\clean.csv`, Make attempts to recreate `data\clean.csv` and `analysis.pdf`.
 If there is a dependency which is missing, and there is no recipe to make it, `make` stops with an error.
 It is a convention to have a target `all` which depends on everything and make it the first target in a `Makefile`, then the command `make` without any argument automatically creates everything.
-The button `Build All` from within RStudio also triggers this process.
-
-```{bash, eval=FALSE}
-all: analysis.pdf
-```
+The button `Build All` from within RStudio triggers this process.
 
@@
 
-If you have followed the above workflow, a fellow researcher is three commands away from fully reproducing your analysis:
+If you have followed our workflow as presented thus far, a fellow researcher is only three commands away from fully reproducing your analysis:
 
 ```{bash, eval=FALSE}
-git clone https://github.com/sirfisher/iris.git
+git clone https://GitHub.com/sirfisher/iris.git
 cd iris
@@
 
-However, this relies on the crucial assumption that your computational environments are identically or sufficiently compatible, e.g., all software dependencies are installed (R and all additional R Packages) and no updates or other changes to the computational environment break or alter your analysis.
-As we will see shortly, this is ensuring full computational reproducibility is necessary at a deeper level than just providing the workflow itself.
+However, succesful reproducibility relies on the crucial assumption that your computational environments are identically or sufficiently compatible, e.g., all software dependencies are installed (R and all additional R Packages) and no updates or other changes to the computational environment break or alter your analysis.
+As we will shortly see, ensuring full computational reproducibility requires one further level of documentation, that is, documentation of the computational context.
 
@@
 
-`docker` is a tool that allows to precisely state what computational environment an analysis relies on and recreates this environment automatically on most operating systems (Windows, OS X & Linux).
-`docker` does that, by setting up a virtual computer on which it executes a series of commands (e.g., installing software).
+Docker is a tool that allows to encapsulate, share, and re-create a computational environment on most operating systems (Windows, macOS & Linux).
+Docker achieves these goals by setting up a virtual computer on which it can execute well-defined series of commands (e.g., installing software).
 It then saves the resulting state of the virtual computer in what is called an "image." This image can be started and execute commands, e.g., running an R-script or `make`.
@@
 An image can be transferred and executed on any machine that has docker installed.
-Unregarding the machine executing the container, from the perspective of a program running inside the container, it always looks identical.
-There are essential differences to a traditional virtual machine, but thinking about a docker container as a computer running inside your computer is still a plausible mental model.
-The most important advantage over traditional virtual machines is that container are lightweight, meaning they start fast and don't need much storage.
-Docker achieves that by reusing large parts of the hosts operating system (how much is varying widely between Linux, OS X & Windows).
+Irrespetictive of the machine that is executing the container, the computational environment is the same for the programs running inside the container.
+The most important advantage over traditional virtual machines is that containers are lightweight, that is, they start fast and don't need much storage space.
+Docker achieves this by reusing large parts of the hosts operating system (how much is varying widely between Linux, macOS & Windows).
 
+With the following example, we demonstrate why it is important to store the computational environment.
+Generally, with containers we would like to safe-guard against changes to the computational environment with unexpected changes, e.g., changes in the functionality or default options in packages or even in the R environment itself.
 While the R programming language is considered stable and much effort is put into backward compatibility, sometimes even basic functions like `sample()`, to randomly sample from a set, change their behavior from one version to another.
@@
 
-Consider the following R-code:
+Consider the following R-code to randomly draw five numbers between 1 and 10:
 
@@
 The usual expectation is that this code delivers the same pseudo-random five numbers regardless of the operating system or R-Version (due to `set.seed()`).
-Using docker, we can start an image which contains the R-Version 3.5.0, and execute the code there.
+Using Docker, we can start an image which contains the R-Version 3.5.0, and execute the code there.
 
@@
 ```
+This outputs:
 
@@
 ```
-
+This outputs:
 ```{r, child="R/random36.md"}
@@
 Note that this is the intended behavior and the result of a [bugfix](https://bugs.r-project.org/bugzilla/show_bug.cgi?id=17494) in the random number generator implemented from R 3.6.0 upwards.
-
-Such change may render analyses run on previous R versions strictly not reproducible if they contain, e.g., multiple imputations, bootstrapping, simulations studies, graphics with random jitter, Bayesian estimations using sampling algorithms (such as Markov Chain Monte Carlo), or similar techniques that involve random sampling.
-We would like to illustrate this with a more concrete example (the full R code to reproduce this non-reproducibility is provided in the github repository of this manuscript).
+Such changes may render analyses run on previous R versions strictly not reproducible if they contain, e.g., multiple imputations, bootstrapping, simulations studies, graphics with random jitter, Bayesian estimations using sampling algorithms (such as Markov Chain Monte Carlo), or similar techniques that involve random sampling.
+We would like to illustrate this with a more concrete example (the full R code to reproduce this non-reproducibility is provided in the GitHub repository of this manuscript).
 Using a simulated dataset with two variables `x` and `y`, we ran a linear regression model with R's `lm()` function regressing `x` on `y`.
 Using the `boot` package [@boot], we bootstrapped the 95% confidence intervals around the regression estimate with 1000 bootstrap samples to evaluate whether zero was inside the confindence interval.
-To make the analysis reproducible, we set a random seed before we bootstrapped.
+To make the analysis reproducible, we set a random seed.
 We ran this once in R 3.5.0:
@@
 
-And then, we ran the identical script with identical seed in R 3.6.1:
+Then, we ran the identical script with identical seed in R 3.6.1:
 
@@
 As we see from these R outputs, the confidence intervals once include zero and once do not include zero.
-Please note that one could discuss deeper issues about null hypothesis significance testing here but we would simply like to stress the point that computational reproducibility in the strict sense needs to capture the full computational environment.
+Please note that one could discuss deeper issues about null hypothesis significance testing here but we would simply like to stress the point that computational reproducibility in the strict sense requires capturing the full computational environment.
 
-However only rarely does an analysis depend only on base R, often a considerable number of packages is required, and they will most likely more often introduce breaking changes with updates.
+Only rarely does an analysis depend only on base R.
+Often, a considerable number of packages is required, and they will increase the likelihood of breaking reproducibiity with every update to any of the packages of base R.
 The whole endeavor of reproducibility is therefore at stake, every time an update is rolled out.
 To ensure long-term reproducibility, `docker` replicates the original computational environment of an analysis exactly.
-However, we do not intend to advocate that software should not be updated; updates typically promote bugfixes and provide new functionality; but full computational reproducibility is only achieved if the originally used software version(s) are documented.
-Among other things, this allows to trace back recent updates as to what change in what software caused a non-replication of the original results.
-Quite to the contrary with solutions like docker, it gets more comfortable than ever to safely update to new versions, by just bumping the R Version number of the docker image.
+Note that we do not intend to advocate that software should not be updated; updates typically promote bugfixes and provide new functionality; but full computational reproducibility is only achieved if the originally used software versions are documented.
+Among other things, this allows to trace back recent updates as to what change in what package caused a non-reproduction of the original results.
+Quite to the contrary, with solutions like docker, it gets more comfortable than ever to safely update to new versions, by just bumping the R Version number of the docker image.
 
-This convenience is possible to the efforts of the [rocker project](https://github.com/rocker-org/rocker) [@boettigerIntroductionRockerDocker2017a] which provides docker images preconfigured with an R-installation of a fixed R-version.
+This convenience is possible to the efforts of the [rocker project](https://GitHub.com/rocker-org/rocker) [@boettigerIntroductionRockerDocker2017a] which provides docker images preconfigured with an R-installation of a fixed R-version.
 These rely on MRAN [@revolutionanalyticsReproducibilityUsingFixed2019], a repository for R-Packages, which is fixed to the last date the R-version of the image was the most recent.
 Building upon these rocker-images researchers can build there own docker-images with the required R-packages.
-The rocker project also provides images which do include RStudio (`rocker/rstudio`), the `tidyverse` package (`rocker/tidyverse`) and the `rmarkdown` package with Latex (`rocker/verse`).
-Since the describes workflow relies on `rmarkdown` we suggest using the `rocker/verse` image (which also contains `rstudio` & `tidyverse`).
+The rocker project also provides images which do include RStudio (`rocker/rstudio`), the `tidyverse` package (`rocker/tidyverse`) and the `R Markdown` package with Latex (`rocker/verse`).
+Since the describes workflow relies on `R Markdown` we suggest using the `rocker/verse` image (which also contains `rstudio` & `tidyverse`).
 These images are stored on [dockerhub](https://hub.docker.com) [@dockerinc.DockerHub2019].
@@
   pacman here pander
-RUN installGithub.r\
+RUN installGitHub.r\
   crsh/papaja benmarwick/wordcountaddin
@@
 rr-flow.pdf: rr-flow.Rmd reproducible-research.bib
-	$(run) Rscript -e 'rmarkdown::render("$(current_dir)/$<")'
+	$(run) Rscript -e 'R Markdown::render("$(current_dir)/$<")'
 ```
@@
 This workflow directly relies on three software packages: `git`, `make` and `docker`.
-Of course it depends also indirectly on R, RStudio & RMarkdown, but they do not need to be installed.
+Of course it depends also indirectly on R, RStudio & R Markdown, but they do not need to be installed.
 The installation process depends on the operating system, e.g.
-on OS X, make is allways availible or on Linux git & make are often allready installed.
+on macOS, make is allways availible or on Linux git & make are often allready installed.
 In the following section we share what have experienced as the easiest options, however in the past how docker should be installed has changed several times, so when in doubt search the internet with: `how to install docker on [your os]`.
@@
 
-## OS X
+## macOS
 
-As make already ships with OS X, you only need `git` and `docker-desktop`.
+As make already ships with macOS, you only need `git` and `docker-desktop`.
 We suggest using the package manager `homebrew`, which you can install from [https://docs.brew.sh/Installation](https://docs.brew.sh/Installation), to install `docker` (git will be installed during the installation of `homebrew`):
@@
 
-We provide a reproducible analysis as working example via [github](https://github.com/aaronpeikert/workflow-showcase) as a testbed for users.
+We provide a reproducible analysis as working example via [GitHub](https://GitHub.com/aaronpeikert/workflow-showcase) as a testbed for users.
 The example shows a minimalistic analysis of the 'Considerations of Future Consequences (CFC) Scale'.
@@
 ```{bash, eval=FALSE}
-git clone https://github.com/aaronpeikert/workflow-showcase.git
+git clone https://GitHub.com/aaronpeikert/workflow-showcase.git
 cd reproducible-research
@@
 ```{bash, eval=FALSE}
-git clone https://github.com/aaronpeikert/reproducible-research.git
+git clone https://GitHub.com/aaronpeikert/reproducible-research.git
 cd reproducible-research
@@
 Each tool reduces the number of ways things can go wrong.
-Dynamic reporting with `RMarkdown` guarantuees consistency between computational results and their reporting; Version control with `git` guarantuees permanence and consistency across multiple versions of data and code; Dependency management with `make` guarantuees defined entry-points and dependency resolution; containerization with `docker` guarantuees reproducibility of the full computational environment.
+Dynamic reporting with `R Markdown` guarantuees consistency between computational results and their reporting; Version control with `git` guarantuees permanence and consistency across multiple versions of data and code; Dependency management with `make` guarantuees defined entry-points and dependency resolution; containerization with `docker` guarantuees reproducibility of the full computational environment.
 This tightening of loose ends, we believe, does not restrict researchers but enables them to operate on a solid basis to deliver sound research.
@@
 A few packages are easing this burden for different parts of the workflow when this flexibility is not needed.
-The use of `rmarkdown` within a project, tracked with `git` can be simplified with the [workflowr-package](https://github.com/jdblischak/workflowr) [@workflowr].
-The [drake-package](https://github.com/ropensci/drake) [@drake] is directly inspired by `make` but does only require knowledge in R, but can only handle dependencies within R.
-The [liftr-package](https://cran.r-project.org/web/packages/liftr/vignettes/liftr-intro.html) [@liftr] and the [holepunch-package](https://karthik.github.io/holepunch/) [@holepunch] automize the use of docker.
+The use of `R Markdown` within a project, tracked with `git` can be simplified with the [workflowr-package](https://GitHub.com/jdblischak/workflowr) [@workflowr].
+The [drake-package](https://GitHub.com/ropensci/drake) [@drake] is directly inspired by `make` but does only require knowledge in R, but can only handle dependencies within R.
+The [liftr-package](https://cran.r-project.org/web/packages/liftr/vignettes/liftr-intro.html) [@liftr] and the [holepunch-package](https://karthik.GitHub.io/holepunch/) [@holepunch] automize the use of docker.
 The former is perfectly compatible with the described workflow and we recommend it to users uncomftable with docker.
@@
 There are several alternatives to docker which fix the dependencies on R-packages.
-[packrat](https://rstudio.github.io/packrat/) [@packrat] is a way to fix package version via local installations of packages in the project, however it does not garuantees more then a fixed set of packages (Not R-version or system dependencies).
-Similar approaches are taken by [jetpack](https://github.com/ankane/jetpack) [@jetpack], [miniCRAN](https://cran.r-project.org/web/packages/miniCRAN/vignettes/miniCRAN-introduction.html) [@miniCRAN] and [checkpoint](https://cran.r-project.org/web/packages/checkpoint/vignettes/checkpoint.html) [@checkpoint].
-Notable is also the package [`reprex`](https://github.com/tidyverse/reprex) [**repr**oducible **ex**ample, @reprex], but its scope is limited to small examples.
+[packrat](https://rstudio.GitHub.io/packrat/) [@packrat] is a way to fix package version via local installations of packages in the project, however it does not garuantees more then a fixed set of packages (Not R-version or system dependencies).
+Similar approaches are taken by [jetpack](https://GitHub.com/ankane/jetpack) [@jetpack], [miniCRAN](https://cran.r-project.org/web/packages/miniCRAN/vignettes/miniCRAN-introduction.html) [@miniCRAN] and [checkpoint](https://cran.r-project.org/web/packages/checkpoint/vignettes/checkpoint.html) [@checkpoint].
+Notable is also the package [`reprex`](https://GitHub.com/tidyverse/reprex) [**repr**oducible **ex**ample, @reprex], but its scope is limited to small examples.
 
@@
 
-In our own experience, it is often not possible to convince all co-authors to switch to a different document processing environment, such as RMarkdown.
-That is, we have experienced that after writing up the first draft in Rmarkdown, one must eventually switch to a dynamically generated Word-file that is from then on used as static file that is used as basis for multiple iterations among the co-authors.
-This means, that ideally, one would have to manually merge the final Word file back to the RMarkdown souce, which generally is only possible manually.
+In our own experience, it is often not possible to convince all co-authors to switch to a different document processing environment, such as R Markdown.
+That is, we have experienced that after writing up the first draft in R Markdown, one must eventually switch to a dynamically generated Word-file that is from then on used as static file that is used as basis for multiple iterations among the co-authors.
+This means, that ideally, one would have to manually merge the final Word file back to the R Markdown souce, which generally is only possible manually.
 But we believe that even if this final step does not happen that - as long as dynamically generated elements are not manipulated - ensuring the reproducibility of the analysis and the initial draft is a big step forward.
@@
 How to best share a reproducible workflow?
-We believe that, ideally, a service provider should be found that guarantuees permanent and reliable hosting of the reproducible workflow, such as the Open Science Framework [@fosterOpenScienceFramework2017], that is independ of Github, Docker Hub & MRAN.
+We believe that, ideally, a service provider should be found that guarantuees permanent and reliable hosting of the reproducible workflow, such as the Open Science Framework [@fosterOpenScienceFramework2017], that is independ of GitHub, Docker Hub & MRAN.
 Second, to ensure that other users are legally able to benefit from the shared materials, authors must choose an appropriate license format.
@@
 We encourage authors to choose appropriate license forms that do not hinder others from freely downloading, using, and modifying the shared workflows and materials while, at the same time, ensuring recognition for the time and effort invested in creating the workflow in the first place.
-In our experience, the Creative Commons license (CC-BY) is often appropriate for sharing texts, rmarkdown files, generated figures, and other media whereas scripts and any other computer code is often best shared according to the MIT license.
+In our experience, the Creative Commons license (CC-BY) is often appropriate for sharing texts, R Markdown files, generated figures, and other media whereas scripts and any other computer code is often best shared according to the MIT license.
 Both licenses assure maximal freedom for future users while requiring the attribution of the original authors in derivative work.

### 2019-10-28 rr-flow.Rmd
manuscript revision

@@
 Quite to the contrary, with solutions like docker, it gets more comfortable than ever to safely update to new versions, by just bumping the R Version number of the docker image.
-
-This convenience is possible to the efforts of the [rocker project](https://GitHub.com/rocker-org/rocker) [@boettigerIntroductionRockerDocker2017a] which provides docker images preconfigured with an R-installation of a fixed R-version.
-These rely on MRAN [@revolutionanalyticsReproducibilityUsingFixed2019], a repository for R-Packages, which is fixed to the last date the R-version of the image was the most recent.
-Building upon these rocker-images researchers can build there own docker-images with the required R-packages.
-The rocker project also provides images which do include RStudio (`rocker/rstudio`), the `tidyverse` package (`rocker/tidyverse`) and the `R Markdown` package with Latex (`rocker/verse`).
-Since the describes workflow relies on `R Markdown` we suggest using the `rocker/verse` image (which also contains `rstudio` & `tidyverse`).
+This convenience is possible because of the efforts of the [rocker project](https://GitHub.com/rocker-org/rocker) [@boettigerIntroductionRockerDocker2017a], which provides docker images preconfigured with an R-installation of given R versions.
+These packages are taken from MRAN [@revolutionanalyticsReproducibilityUsingFixed2019], a repository for R packages fixed to the last date the R-version of the image was the most recent.
+Building upon these rocker-images, researchers can easily build there own Docker images with all required R packages.
+The rocker project also provides images that include RStudio (`rocker/rstudio`), the `tidyverse` package (`rocker/tidyverse`) and the R Markdown package with LaTeX (`rocker/verse`).
+Because our workflow relies on R Markdown, we suggest using the `rocker/verse` image (which also contains `rstudio` & `tidyverse`).
 These images are stored on [dockerhub](https://hub.docker.com) [@dockerinc.DockerHub2019].
@@
 The `FROM` statement specifies which docker image to use, in this case, the `rocker/verse` image with the tag 3.6.1 (referring to the R version 3.6.1).
-The `RUN` statement describes a command to execute, in this case, to run an R-script `install2.r` which is available on all rocker images, to install the packages `pacman`,  `here` & `pander`.
-A Dockerfile allows more than one `RUN` statement, executing any available command.
+The `RUN` statement describes a command to execute, in this case, to run an R-script `install2.r` which is available on all rocker images, to install the specified packages (here, `pacman`,  `here` & `pander`).
+A Dockerfile allows more than one `RUN` statement, executing arbitrary system commands.
 Those `RUN` statements can install dependencies that are not an R package, e.g., other programming languages like python or Matlab.
@@
 The command `docker build -t image-name` creates an image named `image-name` from the Dockerfile in the project.
-A way to automatically find out the dependencies and genereate a docker image from them is provided in the [liftr-package](https://cran.r-project.org/web/packages/liftr/vignettes/liftr-intro.html) [@liftr].
+A way to automatically find out the dependencies and genereate a docker image from them is provided in the [liftR](https://cran.r-project.org/web/packages/liftr/vignettes/liftr-intro.html) package [@liftr].
 
-There are two principal ways to share a docker image, either by sharing the Dockerfile which created the image or by sharing the image itself e.g., through [dockerhub](https://hub.docker.com).
-While both ways guarantee a replicable computational environment, sharing the Dockerfile is more transparent, and git tracks changes to it.
+There are two principal ways to share a docker image, either by sharing the Dockerfile that creates the image, or by sharing the image itself, e.g., through a service like [dockerhub](https://hub.docker.com).
+While both ways guarantee a replicable computational environment, sharing the Dockerfile is more transparent; in our workflow, we can use Git to track changes in the Dockerfile (such as updates to dependencies).
 
-There are two ways to use a rocker image. Either one starts with the `docker run` command. The first way is to run a command inside the container. These commands take the form:
+There are two ways to use a rocker image. Both are based on the `docker run` command. The first way is to run a command inside the container. The call takes the form:
 
@@
 The `--rm` flag means to not save the state of the docker after the command finishes in an container.
-The `-it` flag means to accept inputs and return outputs to the terminal.
-Interacting with docker this way is especially useful when using `R` as a command because then you have access to the standard R-terminal.
+The `-it` flag tells Docker to accept inputs and return outputs to the terminal.
+Interacting with docker this way is especially useful when using `R` as a command because then you have access to the standard R terminal.
+This is the command to start an interactive R session inside a Docker image called `reproducible-research`:
 
@@
 The second way is to supply no commands and to interact with the container via the webbrowser and the rstudio server instance that runs in it.
-For that you need to supply a password to log into rstudio server (`-e PASSWORD=<YOUR_PASS>`) and open a port (`-p 8787:8787`).
+In order to do so, one needs to supply a password to log into Rstudio server (`-e PASSWORD=<YOUR_PASS>`) and open a port (`-p 8787:8787`).
 
@@
 
-The adress to connect to the rstudio server is your ip adress (or on linux also `localhost`) in this scheme: `<IPADRESS>:8787`.
-There is a fully functioning RStudio instance availible, that can be used exactly like one installed locally on the maschine.
+The adress to connect to the Rstudio server is your ip adress (or, on Linux also `localhost`) in this scheme: `<IPADRESS>:8787`.
+This offers a fully functioning Rstudio instance that runs in the image but is accessible through a local web browser.
 
-None of the above containers can access files on your computer, instead a folder needs to be linked explizitly:
+By default, programs inside the container cannot access files on the local computer; to enable access, a local folder needs to be linked explicitly:
 
@@
 
-The main directory for RStudio inside docker is `/home/rstudio` so an apropriate command is:
+The main directory for RStudio inside docker is `/home/rstudio`, so the final call to start Rstudio inside a docker image may look like this:
 
@@
 
-Since `docker`'s commands grow long and are tedius to type, we advocate to use some automatic way to generate them.
-Fortunatly one can use `make` to automaticly generate the `docker` commands, e.g.
-the (simplified) Makefile for this paper allows command after `$(run)` to be condtionally passed through docker if one calls `make DOCKER=TRUE`:
+Since Docker commands grow long and are tedious to type manually, we advocate to use some automatic way to generate them.
+Fortunately, one can use `make` to automatically generate the `docker` commands, e.g.
+the (simplified) Makefile for this paper allows command after `$(run)` to be condtionally passed through docker if one calls `make DOCKER=TRUE` (otherwise, they are run locally):
 
@@
 
-# Setting Up the Workflow
+# Installing and Setting Up the Workflow
 
-This workflow directly relies on three software packages: `git`, `make` and `docker`.
-Of course it depends also indirectly on R, RStudio & R Markdown, but they do not need to be installed.
-The installation process depends on the operating system, e.g.
-on macOS, make is allways availible or on Linux git & make are often allready installed.
-In the following section we share what have experienced as the easiest options, however in the past how docker should be installed has changed several times, so when in doubt search the internet with: `how to install docker on [your os]`.
+Other than on R, RStudio, and R Markdown, our workflow relies on three software external pieces of software: `git`, `make`, and `docker`.
+The installation process of these software packages depends on the operating system, e.g., on macOS, Make is allways availible, or on Linux, Git and Make are often allready installed.
+In the following section, we share what have experienced as the easiest way to install those packages.
+However, installation processes may be subject to change and we advise readers to also consult the documentations of the packages.
 
@@
 
-Windows has the biggest hurdle to install what is needed.
-An essential requirement is that you have Windows Pro, Enterprise, Education or Server installed, as Microsoft prevents the use of Docker on Windows Home.
-However, having met this requirement, there is a great package manager for Windows called chocolately, which you can install from: [https://chocolatey.org/install](https://chocolatey.org/install).
+Windows systems typically require the biggest efforts to install all necessary software.
+Note that you have to have Windows Pro, Enterprise, Education or Server installed, as Microsoft prevents the use of Docker on Windows Home.
+There is a package manager for Windows called chocolately, which you can install from: [https://chocolatey.org/install](https://chocolatey.org/install), providing all dependencies.
 After having installed chocolately (and restarted the computer), all dependencies can be installed in an **admin terminal** (Windows-Key, then type `cmd`, right-click *Run as administrator*) via:
@@
 
-There is a pletora of linux distributions and almost as many package managers.
-However to our knowledge there is no (recent) Linux where `git`, `make` and `docker` are not availible.
-Example for Ubuntu:
+There is a host of diffferent Linux distributions and almost as many package managers.
+Still, to our knowledge there is no (recent) Linux edition, where `git`, `make` and `docker` are not availible.
+For example, in Ubuntu Linux, installation is straightforward using the shipped package manager:
 
@@
 
-Replace `apt` with your package manager.
-Possiblie you need elevated rights; in this case add `sudo` before that.
-`docker` also needs elevated rights, we recomend therefore to add your user to the `docker` group, following the [documentation of docker](https://docs.docker.com/install/linux/linux-postinstall/#manage-docker-as-a-non-root-user).
+For other distributions, replace `apt` with the native package manager.
+One may need elevated rights for the installation; in this case add `sudo` before the installation command.
+`docker` also needs elevated rights to run; therefore, we recomend to add the local user to the `docker` group, following the [documentation of docker](https://docs.docker.com/install/linux/linux-postinstall/#manage-docker-as-a-non-root-user).
 
@@
 Just replace any docker calls with `singularity docker`, because singularity fully supports docker-images.
-A possible advantage is that singularity works well in  HPC environments and on old Linux versions, the downside is that `singularity` is only availibe on Linux.
+A possible advantage is that singularity works well in high-performance computing environments and on old Linux versions, the downside is that `singularity` is only availibe on Linux.
 
@@
 The overarching goal of this paper was to provide an easy to use workflow, that allows confidence in reproducibility.
-This workflow can be almost unfailingly trusted to be replicable with the following commands:
+This workflow can be almost unfailingly trusted to be reproduced with the following commands:
 
@@
 
-That is, it replicates a scientific report exactly without regard to the user, the operating system, the software, the timepoint or the version of the files.
+That is, it reproduces a scientific report exactly without regard to the user, the operating system, the software, the timepoint or the recent changes to the involved files.
 To that end, it relies on tools that have been the foundation of reliable software development for years or in most cases, decades.
-On the way, it makes transparent how statistical results depend upon the software that created them and enables, through that transparency, reuse by other researchers.
+As a by-product, it makes transparent how statistical results depend on the software that created them and enables, through that transparency, reuse by other researchers.
 
-Each tool reduces the number of ways things can go wrong.
-Dynamic reporting with `R Markdown` guarantuees consistency between computational results and their reporting; Version control with `git` guarantuees permanence and consistency across multiple versions of data and code; Dependency management with `make` guarantuees defined entry-points and dependency resolution; containerization with `docker` guarantuees reproducibility of the full computational environment.
-This tightening of loose ends, we believe, does not restrict researchers but enables them to operate on a solid basis to deliver sound research.
+Each tool reduces the likelihood of non-reproduction.
+Dynamic reporting with R Markdown guarantuees consistency between computational results and their reporting; Version control with `git` guarantuees permanence and consistency across multiple versions of data and code; Dependency management with `make` guarantuees defined entry-points and dependency resolution; containerization with `docker` guarantuees reproducibility of the full computational environment.
+We believe that this tightening of loose ends does not restrict researchers but enables them to operate on a solid basis to deliver sound and sustainable research.
 
@@
 
-While this approach here was designed to scale well with the complexity of a computing-intense Project, we realise that this flexibility comes with a steep learning curve.
-A few packages are easing this burden for different parts of the workflow when this flexibility is not needed.
-The use of `R Markdown` within a project, tracked with `git` can be simplified with the [workflowr-package](https://GitHub.com/jdblischak/workflowr) [@workflowr].
-The [drake-package](https://GitHub.com/ropensci/drake) [@drake] is directly inspired by `make` but does only require knowledge in R, but can only handle dependencies within R.
-The [liftr-package](https://cran.r-project.org/web/packages/liftr/vignettes/liftr-intro.html) [@liftr] and the [holepunch-package](https://karthik.GitHub.io/holepunch/) [@holepunch] automize the use of docker.
-The former is perfectly compatible with the described workflow and we recommend it to users uncomftable with docker.
-The later uses [binder](https://mybinder.org) to move the analysis to the cloud so that no local installation of docker is required.
-holepunch is well suited for simple analysis with low computational demands, because of memory and computing time limitations of binder.
-There are several alternatives to docker which fix the dependencies on R-packages.
-[packrat](https://rstudio.GitHub.io/packrat/) [@packrat] is a way to fix package version via local installations of packages in the project, however it does not garuantees more then a fixed set of packages (Not R-version or system dependencies).
-Similar approaches are taken by [jetpack](https://GitHub.com/ankane/jetpack) [@jetpack], [miniCRAN](https://cran.r-project.org/web/packages/miniCRAN/vignettes/miniCRAN-introduction.html) [@miniCRAN] and [checkpoint](https://cran.r-project.org/web/packages/checkpoint/vignettes/checkpoint.html) [@checkpoint].
-Notable is also the package [`reprex`](https://GitHub.com/tidyverse/reprex) [**repr**oducible **ex**ample, @reprex], but its scope is limited to small examples.
+While our approach was designed to scale well with the complexity of a computing-intense project, we realise that this flexibility may not be straightforward to integrate into a researcher's everyday workflow.
+There are R packages that implement parts of our workflow and, thus, lower the threshold to adoption when the full flexibility is not needed.
+The use of R Markdown within a project, tracked with `git` can be simplified with the [workflowR](https://GitHub.com/jdblischak/workflowr) package [@workflowr].
+The [drake](https://GitHub.com/ropensci/drake) package [@drake] is directly inspired by `make` but can only handle dependencies within R.
+The [liftR ](https://cran.r-project.org/web/packages/liftr/vignettes/liftr-intro.html) package [@liftr] and the [holepunch](https://karthik.GitHub.io/holepunch/) package [@holepunch] automatize the use of docker.
+The former is perfectly compatible with the described workflow and we recommend it to users uncomfortable with command-line use of docker.
+The later uses [binder](https://mybinder.org) <CITATION NEEDED> to move the analysis to the cloud, so that no local installation of docker is required.
+`holepunch` is well suited for simple analysis with low computational demands, because of memory and computing time limitations of binder.
+There are several alternatives to docker that manage dependencies on R packages.
+[`packrat`](https://rstudio.GitHub.io/packrat/) [@packrat] is a way to freeze package version via local copies of packages in the project, however it does not garuantee a given base R version or system dependencies.
+Similar approaches are taken by [`jetpack`](https://GitHub.com/ankane/jetpack) [@jetpack], [`miniCRAN`](https://cran.r-project.org/web/packages/miniCRAN/vignettes/miniCRAN-introduction.html) [@miniCRAN] and [`checkpoint`](https://cran.r-project.org/web/packages/checkpoint/vignettes/checkpoint.html) [@checkpoint].
+Notable is also the package [`reprex`](https://GitHub.com/tidyverse/reprex) [**repr**oducible **ex**ample, @reprex], but its scope is limited.
 
-Another way to increase the chance of reproducibility is to develop an R-package, which is a self-contained set of files with well-defined meta data, for each analysis.
+Another way to increase the chance of reproducibility is to develop an R package, which is a self-contained set of files with well-defined meta data, for each analysis.
 The reasoning is that by abiding the strict rules of package development and specifieng all dependencies a high degree of stability can be archieved (e.g., @marwickPackagingDataAnalytical2018).
 
-However, beside these tools that ease the process of creating a workflow on the researchers side, we see increased interest in changing how research is published and used[@perkelToolkitDataTransparency2018], with *life code* [@perkelPioneeringLivecodeArticle2019] and *continuous integration* [@beaulieu-jonesReproducibilityComputationalWorkflows2017, @yenniDevelopingModernData2018].
+However, beside these tools that ease the process of creating a workflow on the researchers side, we see increased interest in changing how research is published and used [@perkelToolkitDataTransparency2018], with *life code* [@perkelPioneeringLivecodeArticle2019] and *continuous integration* [@beaulieu-jonesReproducibilityComputationalWorkflows2017, @yenniDevelopingModernData2018].
 These techniques give us a glimpse of a paradigm shift from static to dynamic, interactive and living publications.

### 2019-10-28 abstract.Rmd

@@
 The workflow leverages established tools and practices from software engineering.
-It combines the benefits of various open-source software tools including Rmarkdown, git, make, and docker, whose interplay ensures seamless integration of version management, APA-conformable dynamic report generation and full cross-platform and long-term computational reproducibility.
-The workflow ensures meeting the primary goals that 1) the reporting of analysis results is perfectly consistent with the actual analysis results (dynamic report generation), the analysis exactly reproduces at a later time even if the computing and software platform is updated or changed (computational reproducibility), and 3) changes at any time (during development and post-publication) are tracked, tagged, and documented while earlier versions of both data and code remain accessible.
+It combines the benefits of various open-source software tools including R Markdown, Git, Make, and Docker, whose interplay ensures seamless integration of version management, APA-conformable dynamic report generation and full cross-platform and long-term computational reproducibility.
+The workflow ensures meeting the primary goals that 1) the reporting of statistical results is consistent with the actual statistical results (dynamic report generation), the analysis exactly reproduces at a later time even if the computing and software platform is updated or changed (computational reproducibility), and 3) changes at any time (during development and post-publication) are tracked, tagged, and documented while earlier versions of both data and code remain accessible.
 While the research community increasingly recognizes dynamic document generation as a tool to ensure reproducibility, we demonstrate with practical examples that dynamic documents alone are not sufficient to ensure long-term computational reproducibility.
-By ensuring full computational reproducibility together with version management, dependence management and literate programming,  the workflow increases scientific productivity by facilitating later reuse.
+By ensuring full computational reproducibility together with version management, dependence management and literate programming,  the workflow increases scientific productivity by facilitating later reuse of code and dat.

### 2019-10-28 rr-flow.Rmd

@@
 
-The folder, in which all fies reside that you need for analysis (code and data), is referred to as a 'project'.
+The folder, in which all fies reside that you need for analysis (code and data), is referred to as a 'project' (or sometimes 'research compendium').
 Working with projects is particularly comfortable with RStudio, an integrated development environment (IDE) for R.
@@
 These licenses are also in line with the recommendations by the Reproducibile Research Standard [@stoddenEnhancingReproducibilityComputational2016; @stoddenEnablingReproducibleResearch2009].
+To facilitate an inclusive environment we recommend to name all contributer and include a [Code of Conduct](https://opensource.guide/code-of-conduct/).
 

### 2019-10-28 R/rmarkdown.Rmd

--- 2019-09-03 R/rmarkdown.Rmd
@@
 ---
-title: "An Example"
-author: "Aaron Peikert & Andreas Brandmaier"
-date: "9/3/2019"
+title: "A simple RMarkdown Example"
+author: "Aaron Peikert & Andreas M. Brandmaier"
+date: '`r format(Sys.time(), "%d %B, %Y")`'
 output: pdf_document
@@
 
-The data was never before anylized in this way. Except in `help(t.test)`.
+This is an example on students' sleep data taken from `help(t.test)`.
 
@@
 I can now assert that what I *believe* to be true
---- that there is a difference in mean between the groups ---
+--- that there is a difference in means between the groups ---
 is `r ifelse(result$p.value > .025, "**not**", "")` supported by data.

### 2019-10-28 rr-flow.Rmd

@@
 ---
-title             : "A Reproducible Data Analysis Workflow with R Markdown, Git, Make, & Docker"
+title             : "A Reproducible Data Analysis Workflow with RMarkdown, Git, Make, & Docker"
 shorttitle        : "Reproducible Data Analysis Workflow"
@@
 authornote: This paper is fully reproducible using the workflow described in this paper.
-All materials can be found on the accompanying GitHub repository <https://GitHub.com/aaronpeikert/reproducible-research/>.
-To provide feedback or any other type of comment or questions regarding our workflow, please add an issue to the GitHub repository of our paper [https://GitHub.com/aaronpeikert/reproducible-research/issues](https://GitHub.com/aaronpeikert/reproducible-research/issues).
+All materials can be found on the accompanying GitHub repository <https://github.com/aaronpeikert/reproducible-research/>.
+To provide feedback or any other type of comment or questions regarding our workflow, please add an issue to the GitHub repository of our paper [https://github.com/aaronpeikert/reproducible-research/issues](https://github.com/aaronpeikert/reproducible-research/issues).
 
@@
 
-The first step towards reproducibility is to have an R script or R Markdown file as primary entry point for the analysis, which runs on a local computer without error.
+The first step towards reproducibility is to have an R-script or RMarkdown file as primary entry point for the analysis, that runs on a local computer without error.
 Next, one has to make sure that all files relevant for the analysis can be moved to another computer.
@@
 
-The folder, in which all fies reside that you need for analysis (code and data), is referred to as a 'project' (or sometimes 'research compendium').
+The folder, where all files reside that you need for analysis (code and data), is referred to as a 'project' (or sometimes 'research compendium').
 Working with projects is particularly comfortable with RStudio, an integrated development environment (IDE) for R.
@@
 
-![Source of an R Markdown](Images/R Markdown.png)
+![Source of an RMarkdown](Images/rmarkdown.png)
 
-![Rendered result](R/R Markdown.pdf)
+![Rendered result](R/rmarkdown.pdf)
 
@@
 
-To keep track of all changes on your local computer, one only needs to use `git add` & `git commit` or `git commit -a` to do both.
-These commands need to be executed in the `terminal`, which one can access from within Rstudio (`Shift + Alt + R`).
+So to keep track of all changes on your local computer, you only need to use `git add` & `git commit` or `git commit -a` to do both.
+These commands need to be executed in the `terminal`, which you can access from within Rstudio (`Shift + Alt + R`).
 Rstudio also offers a graphical user interface for Git and, quite likely, this interface is sufficient for almost every interaction with Git.
@@
 
-Now one can inspect all changes (`git log`) and examine any previous state by stating the identifier of the commit to `git checkout`:
+Now you can inspect all changes (`git log`) and examine any previous state by stating the identifier of the commit to `git checkout`:
 
@@
 Git also makes it especially easy to share and collaborate on a project with other researchers.
-A popular service for sharing materials via Git is [GitHub](https://GitHub.com).
+A popular service for sharing materials via Git is [GitHub](https://github.com).
 Just sharing Git repositories with the public is always free, private repositories (only visible to persons you invite) are [free for researchers](https://help.GitHub.com/en/articles/applying-for-an-educator-or-researcher-discount).
@@
 ```{bash, eval=FALSE}
-git remote add origin git@GitHub.com:aaronpeikert/reproducible-research.git
+git remote add origin git@github.com:aaronpeikert/reproducible-research.git
 git push -u origin master
@@
 ```{bash, eval=FALSE}
-git clone git@GitHub.com:aaronpeikert/reproducible-research.git
+https://github.com/aaronpeikert/reproducible-research.git
 ```
 
-Git and GitHub can do even more to help you to collaborate with your fellow researchers, such as formal means to track issues and their status (open/closed/resolved) and further means to managing and merging multiple, parallel versions of code (such as branches, pull requests, merges) but this is beyond the scope of this tutorial.
+Git and GitHub can do even more to help you to collaborate with your fellow researchers, such as formal means to track issues and their status (open/closed/resolved) and further means to managing and merging multiple, parallel versions of code (such as branches, pull requests, merges) but this is beyond the already broad scope of this tutorial.
 In particular, GitHub's issue management can be leveraged as post-publication platform to discuss manuscripts and their results [^1].
@@
 
-[^1]: To comment on our paper, please add an issue to the GitHub repository of our paper: <https://GitHub.com/aaronpeikert/reproducible-research/issues>.
+[^1]: To comment on our paper, please add an issue to the GitHub repository of our paper: <https://github.com/aaronpeikert/reproducible-research/issues>.
 
@@
 analysis.pdf: data/clean.csv
-  Rscript -e 'R Markdown::render("analysis.Rmd")'
+  Rscript -e 'rmarkdown::render("analysis.Rmd")'
 
@@
 ```{bash, eval=FALSE}
-git clone https://GitHub.com/sirfisher/iris.git
+git clone https://github.com/sirfisher/iris.git
 cd iris
@@
 With the following example, we demonstrate why it is important to store the computational environment.
-Generally, with containers we would like to safe-guard against changes to the computational environment with unexpected changes, e.g., changes in the functionality or default options in packages or even in the R environment itself.
+Generally, with containers we would like to safe-guard against changes to the computational environment with unexpected consequences, e.g., changes in the functionality or default options in packages or even in the R environment itself.
 While the R programming language is considered stable and much effort is put into backward compatibility, sometimes even basic functions like `sample()`, to randomly sample from a set, change their behavior from one version to another.
@@
   pacman here pander
-RUN installGitHub.r\
+RUN installGithub.r\
   crsh/papaja benmarwick/wordcountaddin
@@
 The second way is to supply no commands and to interact with the container via the webbrowser and the rstudio server instance that runs in it.
-In order to do so, one needs to supply a password to log into Rstudio server (`-e PASSWORD=<YOUR_PASS>`) and open a port (`-p 8787:8787`).
+In order to do so, you need to supply a password to log into Rstudio server (`-e PASSWORD=<YOUR_PASS>`) and open a port (`-p 8787:8787`).
 
@@
 rr-flow.pdf: rr-flow.Rmd reproducible-research.bib
-	$(run) Rscript -e 'R Markdown::render("$(current_dir)/$<")'
+	$(run) Rscript -e 'rmarkdown::render("$(current_dir)/$<")'
 ```
@@
 
-We provide a reproducible analysis as working example via [GitHub](https://GitHub.com/aaronpeikert/workflow-showcase) as a testbed for users.
+We provide a reproducible analysis as working example via [GitHub](https://github.com/aaronpeikert/workflow-showcase) as a testbed for users.
 The example shows a minimalistic analysis of the 'Considerations of Future Consequences (CFC) Scale'.
@@
 ```{bash, eval=FALSE}
-git clone https://GitHub.com/aaronpeikert/workflow-showcase.git
+git clone https://github.com/aaronpeikert/workflow-showcase.git
 cd reproducible-research
@@
 ```{bash, eval=FALSE}
-git clone https://GitHub.com/aaronpeikert/reproducible-research.git
+git clone https://github.com/aaronpeikert/reproducible-research.git
 cd reproducible-research
@@
 
-In our own experience, it is often not possible to convince all co-authors to switch to a different document processing environment, such as R Markdown.
-That is, we have experienced that after writing up the first draft in R Markdown, one must eventually switch to a dynamically generated Word-file that is from then on used as static file that is used as basis for multiple iterations among the co-authors.
-This means, that ideally, one would have to manually merge the final Word file back to the R Markdown souce, which generally is only possible manually.
-But we believe that even if this final step does not happen that - as long as dynamically generated elements are not manipulated - ensuring the reproducibility of the analysis and the initial draft is a big step forward.
-Making such static Word-file more dynamic is is possible with the officer package [@gohel2019officer], which allows the manipulation of  Word (.docx) and PowerPoint (.pptx) documents.
-That is, given a formatted static template file, one can dynamically paste images, tables and text into documents from R.
-We consider such static file with dynamic content, less elegant but potentially more practical, when not all authors can adhere to the described process.
+In our own experience, it is often not possible to convince all co-authors to switch to a different document processing environment, such as RMarkdown.
+That is, we have experienced that after writing up the first draft in Rmarkdown, one must eventually to genereate a Word-file that is from then on used as static file serving as a basis for multiple iterations among the co-authors.
+Reteining reproduciblity in such situation requires tediues manuel synchronization.
+This labor can be saved with the [rodoc package](https://noamross.github.io/redoc/articles/mixed-workflows-with-redoc.html) [@redoc], which enables a two-way workflow between Word and Rmarkdown.
+The conversion between Rmarkdown and Word retains all changes and even supports words tracked changes feature.
 

### 2019-10-28 rr-flow.Rmd

@@
 Here, we combine four software tools, whose interplay can guarantuee full and sustainable computational reproducibility of data analyses and their reporting.
-There are various tools to enhance reproducibility [@piccoloToolsTechniquesComputational2016], of which we believe four of them are particularly important: literate programming: [@ruleTenSimpleRules2019], version control [@barbaHardRoadReproducibility2016], dependency management @[askrenUsingMakeReproducible2016] and containerization [@clyburne-sherinComputationalReproducibilityContainers2018].
+There are various tools to enhance reproducibility [@piccoloToolsTechniquesComputational2016], of which we believe four of them are particularly important: literate programming: [@ruleTenSimpleRules2019], version control [@barbaHardRoadReproducibility2016], dependency management [@askrenUsingMakeReproducible2016] and containerization [@clyburne-sherinComputationalReproducibilityContainers2018].
 Importantly, only a workflow using all four concepts in unison guarantuee confidence in reproducing a scientific report.
-Even though there are various implementations of these concepts, we believe that for the R environment <CITATION NEEDED>, the following four implementations are best suited.
-For literate programming, R Markdown <CITATION NEEDED>, for version control Git <CITATION NEEDED>, for dependency management Make <CITATION NEEDED>, and for container Docker <CITATION NEEDED>.
+Even though there are various implementations of these concepts, we believe that for the R environment [@r2019], the following four implementations are best suited.
+For literate programming, R Markdown [@rmarkdown2019], for version control Git [@chaconProGit2014], for dependency management Make [@feldmanMakeProgramMaintaining1979], and for container Docker [@merkelDockerLightweightLinux2014].
 Each of them serves a meta-scientific valuable goal (reproducibility) and increases the productivity of the researcher.
@@
 To this end, it is essential that all files are within one folder (or, subfolders within it) and all paths are relative to that folder.
-A robust solution to the problem of making sure that file access does not break across computing platforms are [Rstudio projects](https://r4ds.had.co.nz/workflow-projects.html) and the `here` <CITATION NEEDED> package.
+A robust solution to the problem of making sure that file access does not break across computing platforms are [Rstudio projects](https://r4ds.had.co.nz/workflow-projects.html) and the `here` [@here2017] package.
 
@@
 In these cases, of course, unrestricted reproducibility is not guaranteed.
-If this way of dividing data and scripts into different databases is unavoidable, we recommend to validate all data files using checksums (also called a "hash", e.g., using the functions provided in package `digest` <CITATION NEEDED>) before analyzing them.
+If this way of dividing data and scripts into different databases is unavoidable, we recommend to validate all data files using checksums (also called a "hash", e.g., using the functions provided in package `digest` [@digest2019]) before analyzing them.
 To do this, a checksum must be created and stored at the time of the original analysis.
@@
 The former is perfectly compatible with the described workflow and we recommend it to users uncomfortable with command-line use of docker.
-The later uses [binder](https://mybinder.org) <CITATION NEEDED> to move the analysis to the cloud, so that no local installation of docker is required.
+The later uses [binder](https://mybinder.org) [@binder2018] to move the analysis to the cloud, so that no local installation of docker is required.
 `holepunch` is well suited for simple analysis with low computational demands, because of memory and computing time limitations of binder.
@@
 
-However, beside these tools that ease the process of creating a workflow on the researchers side, we see increased interest in changing how research is published and used [@perkelToolkitDataTransparency2018], with *life code* [@perkelPioneeringLivecodeArticle2019] and *continuous integration* [@beaulieu-jonesReproducibilityComputationalWorkflows2017, @yenniDevelopingModernData2018].
+However, beside these tools that ease the process of creating a workflow on the researchers side, we see increased interest in changing how research is published and used [@perkelToolkitDataTransparency2018], with *life code* [@perkelPioneeringLivecodeArticle2019] and *continuous integration* [@beaulieu-jonesReproducibilityComputationalWorkflows2017; @yenniDevelopingModernData2018].
 These techniques give us a glimpse of a paradigm shift from static to dynamic, interactive and living publications.

### 2019-10-29 rr-flow.Rmd
solves #54

@@
 For other distributions, replace `apt` with the native package manager.
-One may need elevated rights for the installation; in this case add `sudo` before the installation command.
+You may need elevated rights for the installation; in this case add `sudo` before the installation command.
 `docker` also needs elevated rights to run; therefore, we recomend to add the local user to the `docker` group, following the [documentation of docker](https://docs.docker.com/install/linux/linux-postinstall/#manage-docker-as-a-non-root-user).
@@
 In our own experience, it is often not possible to convince all co-authors to switch to a different document processing environment, such as RMarkdown.
-That is, we have experienced that after writing up the first draft in Rmarkdown, one must eventually to genereate a Word-file that is from then on used as static file serving as a basis for multiple iterations among the co-authors.
+That is, we have experienced that after writing up the first draft in Rmarkdown, one must eventually genereate a Word-file that is from then on used as static file serving as a basis for multiple iterations among the co-authors.
 Reteining reproduciblity in such situation requires tediues manuel synchronization.

### 2019-10-29 rr-flow.Rmd
unify R script and R Markdown spelling closes #53

@@
 
-The first step towards reproducibility is to have an R-script or RMarkdown file as primary entry point for the analysis, that runs on a local computer without error.
+The first step towards reproducibility is to have an R script or R Markdown file as primary entry point for the analysis, that runs on a local computer without error.
 Next, one has to make sure that all files relevant for the analysis can be moved to another computer.
@@
 
-![Source of an RMarkdown](Images/rmarkdown.png)
+![Source of an R Markdown](Images/rmarkdown.png)
 
@@
 Docker achieves these goals by setting up a virtual computer on which it can execute well-defined series of commands (e.g., installing software).
-It then saves the resulting state of the virtual computer in what is called an "image." This image can be started and execute commands, e.g., running an R-script or `make`.
+It then saves the resulting state of the virtual computer in what is called an "image." This image can be started and execute commands, e.g., running `Rscript` or `make`.
 A running instance of an image is called a container.
@@
 The `FROM` statement specifies which docker image to use, in this case, the `rocker/verse` image with the tag 3.6.1 (referring to the R version 3.6.1).
-The `RUN` statement describes a command to execute, in this case, to run an R-script `install2.r` which is available on all rocker images, to install the specified packages (here, `pacman`,  `here` & `pander`).
+The `RUN` statement describes a command to execute, in this case, to run an R script `install2.r` which is available on all rocker images, to install the specified packages (here, `pacman`,  `here` & `pander`).
 A Dockerfile allows more than one `RUN` statement, executing arbitrary system commands.
@@
 
-In our own experience, it is often not possible to convince all co-authors to switch to a different document processing environment, such as RMarkdown.
-That is, we have experienced that after writing up the first draft in Rmarkdown, one must eventually genereate a Word-file that is from then on used as static file serving as a basis for multiple iterations among the co-authors.
+In our own experience, it is often not possible to convince all co-authors to switch to a different document processing environment, such as R Markdown.
+That is, we have experienced that after writing up the first draft in R Markdown, one must eventually genereate a Word-file that is from then on used as static file serving as a basis for multiple iterations among the co-authors.
 Reteining reproduciblity in such situation requires tediues manuel synchronization.
-This labor can be saved with the [rodoc package](https://noamross.github.io/redoc/articles/mixed-workflows-with-redoc.html) [@redoc], which enables a two-way workflow between Word and Rmarkdown.
-The conversion between Rmarkdown and Word retains all changes and even supports words tracked changes feature.
+This labor can be saved with the [rodoc package](https://noamross.github.io/redoc/articles/mixed-workflows-with-redoc.html) [@redoc], which enables a two-way workflow between Word and R Markdown.
+The conversion between R Markdown and Word retains all changes and even supports words tracked changes feature.
 

### 2019-10-29 rr-flow.Rmd
make link clickable

@@
 This paper is fully reproducible using the workflow described in this paper.
-All materials, including the manuscript source and all R codes and files used to generate this manuscript, can be found on the accompanying GitHub repository <https://GitHub.com/aaronpeikert/reproducible-research/>.
+All materials, including the manuscript source and all R codes and files used to generate this manuscript, can be found on the accompanying GitHub repository [<https://github.com/aaronpeikert/reproducible-research/>](https://github.com/aaronpeikert/reproducible-research/).
 

### 2019-10-29 README.Rmd
explain how to rebuild everything #49

@@
 
+## Rebuild Everything
+
+In case you experience some unexpected behavior with this workflow, you should check that you have the most recent version (`git pull`), rebuild the docker image (`make build`) and force the rebuild of all targets (`make -B DOCKER`).
+
+```{bash, eval=FALSE}
+git pull && make build && make -B DOCKER=TRUE
+```
+
 # Session Info

### 2019-10-29 abstract.Rmd
correct spelling

@@
 While the research community increasingly recognizes dynamic document generation as a tool to ensure reproducibility, we demonstrate with practical examples that dynamic documents alone are not sufficient to ensure long-term computational reproducibility.
-By ensuring full computational reproducibility together with version management, dependence management and literate programming,  the workflow increases scientific productivity by facilitating later reuse of code and dat.
+By ensuring full computational reproducibility together with version management, dependence management and literate programming,  the workflow increases scientific productivity by facilitating later reuse of code and data.

### 2019-10-29 rr-flow.Rmd
correct spelling

@@
 R-based data analyses.
-Reproducibility means to us that statistica results together with their computational workflows are reported fully, transparently and sustainably remain available, such that the workflow can be re-run by a different person and/or later in time and the results will be identical to the ones intially reported [@herouxCompatibleReproducibilityTaxonomy2018; @claerboutElectronicDocumentsGive1992].
+Reproducibility means to us that statistical results together with their computational workflows are reported fully, transparently and sustainably remain available, such that the workflow can be re-run by a different person and/or later in time and the results will be identical to the ones intially reported [@herouxCompatibleReproducibilityTaxonomy2018; @claerboutElectronicDocumentsGive1992].
 The need to ensure reproducibility directly follows from commonly accepted rules of good scientific practice (such as the guidelines of the German Research Foundation; @dfg2019).
@@
 The most precise documentation is computer code, and the most productive computer code is code that runs automatically without human interaction.
-This principle is what guided our choice to use of R Markdown, Git, Make, and Docker.
+This principle is what guided our use of R Markdown, Git, Make, and Docker.
 
@@
 The translation of computational results into a human-readable summary, for example into a technical report, a presentation, or a journal manuscript, is time-consuming and error-prone.
-Typical errors result from copy\&paste mistakes, erroneous rounding, or missed updates of the manuscript when the computer coder has changed.
+Typical errors result from copy\&paste mistakes, erroneous rounding, or missed updates of the manuscript when the computer code has changed.
 To not only create fully reproducible results but also fully reproducible reports, one needs to resort to the Literate Programming paradigm [@knuth1984literate], in which human-readable language and computer-code are interspersed to create dynamic documents whose order follows the logic of thought rather then the order of the computer.
@@
 ```
+
 This outputs:
@@
 ```
+
 This outputs:
+
 ```{r, child="R/random36.md"}
@@
 These packages are taken from MRAN [@revolutionanalyticsReproducibilityUsingFixed2019], a repository for R packages fixed to the last date the R-version of the image was the most recent.
-Building upon these rocker-images, researchers can easily build there own Docker images with all required R packages.
+Building upon these rocker-images, researchers can easily build their own Docker images with all required R packages.
 The rocker project also provides images that include RStudio (`rocker/rstudio`), the `tidyverse` package (`rocker/tidyverse`) and the R Markdown package with LaTeX (`rocker/verse`).

### 2019-10-29 rr-flow.Rmd
https instead of ssh closes #57

@@
 A popular service for sharing materials via Git is [GitHub](https://github.com).
-Just sharing Git repositories with the public is always free, private repositories (only visible to persons you invite) are [free for researchers](https://help.GitHub.com/en/articles/applying-for-an-educator-or-researcher-discount).
+Just sharing Git repositories with the public is always free, private repositories (only visible to persons you invite) are [free for researchers](https://help.github.com/en/articles/applying-for-an-educator-or-researcher-discount).
 After creating a user account, one can create a new repository where GitHub advises you how to upload your repository from the terminal, e.g., for our repository (here with user name 'aaronpeikert' and repository name 'reproducible-research'):
@@
 ```{bash, eval=FALSE}
-git remote add origin git@github.com:aaronpeikert/reproducible-research.git
+git remote add origin https://github.com/aaronpeikert/reproducible-research.git
 git push -u origin master
@@
 ```{bash, eval=FALSE}
-https://github.com/aaronpeikert/reproducible-research.git
+git clone https://github.com/aaronpeikert/reproducible-research.git
 ```

### 2019-11-05 manuscript.Rmd
rename to manuscript close #58

--- 2019-10-29 rr-flow.Rmd
@@
 keywords          : "keywords"
-wordcount         : "`r tryCatch(wordcountaddin::word_count(here::here('rr-flow.Rmd')), error=function(err) 'NA')`"
+wordcount         : "`r tryCatch(wordcountaddin::word_count(here::here('manuscript.Rmd')), error=function(err) 'NA')`"
 
@@
 
-all: rr-flow.pdf
+all: manuscript.pdf
 
@@
 
-rr-flow.pdf: rr-flow.Rmd reproducible-research.bib
+manuscript.pdf: manuscript.Rmd reproducible-research.bib
 	$(run) Rscript -e 'rmarkdown::render("$(current_dir)/$<")'

### 2019-11-05 R/jou_version.R

@@ -1,6 +1,6 @@
 library(readr)
 library(here)
-doc <- read_lines(here("rr-flow.Rmd"))
+doc <- read_lines(here("manuscript.Rmd"))
 
 pos_classoption <- grep("^classoption", doc)
 doc[pos_classoption] <- 'classoption: "jou"'
@@ -13,4 +13,4 @@ pos_authornote <- grep("^authornote", doc)
 doc[pos_authornote] <- paste(doc[pos_authornote],
                              "Please note that code longer then 40 characters is continued on the next line and indended by four spaces.")
 
-write_lines(doc, here("rr-flow-jou.Rmd"))
+write_lines(doc, here("manuscript.Rmd"))

### 2019-11-07 abstract.Rmd
revision of the manuscript

--- 2019-10-28 abstract.Rmd
@@
-In this tutorial, we describe a principled workflow to ensure long-term reproducibility of R-based data analyses.
+In this tutorial, we describe a workflow to ensure long-term reproducibility of R-based data analyses.
 The workflow leverages established tools and practices from software engineering.
 It combines the benefits of various open-source software tools including R Markdown, Git, Make, and Docker, whose interplay ensures seamless integration of version management, APA-conformable dynamic report generation and full cross-platform and long-term computational reproducibility.
-The workflow ensures meeting the primary goals that 1) the reporting of statistical results is consistent with the actual statistical results (dynamic report generation), the analysis exactly reproduces at a later time even if the computing and software platform is updated or changed (computational reproducibility), and 3) changes at any time (during development and post-publication) are tracked, tagged, and documented while earlier versions of both data and code remain accessible.
-While the research community increasingly recognizes dynamic document generation as a tool to ensure reproducibility, we demonstrate with practical examples that dynamic documents alone are not sufficient to ensure long-term computational reproducibility.
-By ensuring full computational reproducibility together with version management, dependence management and literate programming,  the workflow increases scientific productivity by facilitating later reuse of code and dat.
+The workflow ensures meeting the primary goals that 1) the reporting of statistical results is consistent with the actual statistical results (dynamic report generation), the analysis exactly reproduces at a later time even if the computing platform or software is changed (computational reproducibility), and 3) changes at any time (during development and post-publication) are tracked, tagged, and documented while earlier versions of both data and code remain accessible.
+While the research community increasingly recognizes dynamic document generation and version management as tools to ensure reproducibility, we demonstrate with practical examples that these alone are not sufficient to ensure long-term computational reproducibility.
+Leveraging containerization, dependence management, version management, and literate programming, the workflow increases scientific productivity by facilitating later reproducibility and reuse of code and data.

### 2019-11-07 manuscript.Rmd
revision of the manuscript

@@
 
-In this tutorial, we describe a principled workflow to ensure long-term and cross-platform reproducibility of
+In this tutorial, we describe a workflow to ensure long-term and cross-platform reproducibility of
 R-based data analyses.
-Reproducibility means to us that statistical results together with their computational workflows are reported fully, transparently and sustainably remain available, such that the workflow can be re-run by a different person and/or later in time and the results will be identical to the ones intially reported [@herouxCompatibleReproducibilityTaxonomy2018; @claerboutElectronicDocumentsGive1992].
+Reproducibility means to us that statistical results together with their generating computational workflows are reported fully, and transparently, and remain sustainably available, such that a given workflow can be re-run by a different person later in time and the results will be identical to the ones intially reported [@herouxCompatibleReproducibilityTaxonomy2018; @claerboutElectronicDocumentsGive1992].
 The need to ensure reproducibility directly follows from commonly accepted rules of good scientific practice (such as the guidelines of the German Research Foundation; @dfg2019).
@@
 
-Here, we combine four software tools, whose interplay can guarantuee full and sustainable computational reproducibility of data analyses and their reporting.
-There are various tools to enhance reproducibility [@piccoloToolsTechniquesComputational2016], of which we believe four of them are particularly important: literate programming: [@ruleTenSimpleRules2019], version control [@barbaHardRoadReproducibility2016], dependency management [@askrenUsingMakeReproducible2016] and containerization [@clyburne-sherinComputationalReproducibilityContainers2018].
-Importantly, only a workflow using all four concepts in unison guarantuee confidence in reproducing a scientific report.
+Here, we combine four software tools, whose interplay can guarantuee full computational reproducibility of data analyses and their reporting.
+There are various ideas to enhance reproducibility [@piccoloToolsTechniquesComputational2016], of which we believe four of them are particularly important: literate programming: [@ruleTenSimpleRules2019], version control [@barbaHardRoadReproducibility2016], dependency management [@askrenUsingMakeReproducible2016] and containerization [@clyburne-sherinComputationalReproducibilityContainers2018].
+We argue that only a workflow using all four concepts in unison can guarantuee confidence in reproducing a scientific report.
 Even though there are various implementations of these concepts, we believe that for the R environment [@r2019], the following four implementations are best suited.
-For literate programming, R Markdown [@rmarkdown2019], for version control Git [@chaconProGit2014], for dependency management Make [@feldmanMakeProgramMaintaining1979], and for container Docker [@merkelDockerLightweightLinux2014].
-Each of them serves a meta-scientific valuable goal (reproducibility) and increases the productivity of the researcher.
-Each of these software solutions is highly flexible and powerful, so to master them fully, a considerable amount of practice is necessary.
-However, for our purposes, it is sufficient to master a minimal valuable subset of functions to ensure reproducibility.
+For literate programming, R Markdown [@rmarkdown2019], for version control, Git [@chaconProGit2014], for dependency management, Make [@feldmanMakeProgramMaintaining1979], and for containerization, Docker [@merkelDockerLightweightLinux2014].
+Each of these software solutions serves a meta-scientific valuable goal (reproducibility) and increases the productivity of the researcher.
+They all are highly flexible and powerful, so to master them fully, a considerable amount of practice is necessary.
+However, for our purposes, it is sufficient to master a minimal valuable subset of functions to ensure reproducibility of scientific analyses.
 
-To maximize chances of long-term reproducibility, proper documentation of all steps taken and their context is crucial since it answers the question "What steps have to be taken to reproduce the results?".
+<!--To maximize chances of long-term reproducibility, proper documentation of all steps taken and their context is crucial since it answers the question "What steps have to be taken to reproduce the results?".
 The most precise documentation is computer code, and the most productive computer code is code that runs automatically without human interaction.
-This principle is what guided our use of R Markdown, Git, Make, and Docker.
+This principle is what guided our use of R Markdown, Git, Make, and Docker.-->
 
@@
 This paper is fully reproducible using the workflow described in this paper.
-All materials, including the manuscript source and all R codes and files used to generate this manuscript, can be found on the accompanying GitHub repository [<https://github.com/aaronpeikert/reproducible-research/>](https://github.com/aaronpeikert/reproducible-research/).
+All materials, including the manuscript source and all R codes and files used to generate this manuscript, can be found on the accompanying GitHub repository [https://github.com/aaronpeikert/reproducible-research/](https://github.com/aaronpeikert/reproducible-research/).
 
@@
 
-The first step towards reproducibility is to have an R script or R Markdown file as primary entry point for the analysis, that runs on a local computer without error.
+The first step towards reproducibility is to create an R script or R Markdown file as primary entry point for the analysis, which runs on a local computer without error and performs the main statistical analyses.
 Next, one has to make sure that all files relevant for the analysis can be moved to another computer.
-To this end, it is essential that all files are within one folder (or, subfolders within it) and all paths are relative to that folder.
+To this end, it is recommendable that all files are within one folder (or, subfolders within it) and all paths are relative to that folder.
 A robust solution to the problem of making sure that file access does not break across computing platforms are [Rstudio projects](https://r4ds.had.co.nz/workflow-projects.html) and the `here` [@here2017] package.
@@
 In these cases, of course, unrestricted reproducibility is not guaranteed.
-If this way of dividing data and scripts into different databases is unavoidable, we recommend to validate all data files using checksums (also called a "hash", e.g., using the functions provided in package `digest` [@digest2019]) before analyzing them.
+If dividing data and scripts is unavoidable, we recommend to validate all data files using checksums (also called a "hash", e.g., using the functions provided in package `digest` [@digest2019]) before analyzing them.
 To do this, a checksum must be created and stored at the time of the original analysis.
-At the time of reproduction, the current checksum must be compared with the stored checksum.
+At the time of reproduction, the current checksum must be compared with the stored checksum to ensure data integrity.
 
@@
 if (checksum != "740e88404c04af1d67ffc87b008231c093539cc1"){
-  error("Wrong data file loaded or data has been tempered with!")
+  error("Wrong data file loaded or data has been changed!")
 }
@@
 R Markdown is a simple markup language to create dynamic documents with embedded chunks of R code that can be exported to standard formats such as documents (docx, rtf, epub), presentations (ppt, html) or websites (html) using the knitr package [@xie2015; @xie2019].
-Figure 1 illustrates R Markdown syntax and Figure 2 the resulting output.
 Several packages extend the functionality of knitr.
-Of particular note are the `papaja` package [@papaja2018], which additional functions to enable APA-conformable document formatting, including a journal-style final typeset format, and the `stargazer` package [@stargazer2018], which provides journal-ready tables and reports of statistical models.
-Here is a simple example of how `papaja` can be used to create dynamic reports in APA style:
+Of particular note are the `papaja` package [@papaja2018], which offers additional functions to enable APA-conformable document formatting, including a journal-style final typeset format, and the `stargazer` package [@stargazer2018], which provides journal-ready tables and reports of statistical models.
+Figure 1 illustrates R Markdown syntax using the papaja package and Figure 2 shows the resulting rendered document.
 
@@
 
-Reproducibility at a basic level assumes that results remain identical if neither script nor data have changed.
-It is often not trivial to find out whether anything has changed over time and if so to "go back in time." `git` enables you to do both.
-A good mental model for `git` is that it takes a sequence of snapshots of all files it is supposed to track.
-In the language of git, these snapshots are "commits".
+Fundamentally, reproducibility means that results remain identical if neither script nor data have changed.
+It is often not trivial to find out whether anything in a project has changed over time and if so to "go back in time." The program Git enables you to do both.
+A good mental model for Git is that it takes a sequence of snapshots of all files it is supposed to track.
+In the language of Git, these snapshots are "commits".
 A commit represents a complete copy of the state of all tracked files.
 Each commit has a unique identifier (a hash) and a human-readable description (commit message).
-Going back to one state is as easy as traversing the history of all commits; it is possible to visually track changes between different versions.
+Going back to one state is as easy as traversing the history of all commits and switch to a given one; it is possible to visually track changes between different versions.
 Finding a specific version can be made even easier by tagging snapshots, e.g., as "submission," "preprint," or "published." This collection of snapshots is called a "repository," which ideally tracks your entire R project.
 
-A typical git workflow in the terminal looks like that:
+A typical git workflow in the terminal looks like this:
 
@@
 
-So to keep track of all changes on your local computer, you only need to use `git add` & `git commit` or `git commit -a` to do both.
-These commands need to be executed in the `terminal`, which you can access from within Rstudio (`Shift + Alt + R`).
+To keep track of all changes on your local computer, you only need to use `git add` & `git commit` or `git commit -a` to do both.
+These commands need to be executed in the terminal, which you can access from within Rstudio (`Shift + Alt + R`).
 Rstudio also offers a graphical user interface for Git and, quite likely, this interface is sufficient for almost every interaction with Git.
@@
 
-Now you can inspect all changes (`git log`) and examine any previous state by stating the identifier of the commit to `git checkout`:
+In a given Git project, you can inspect all changes (`git log`) and examine any previous state by stating the identifier of the commit to `git checkout`:
 
@@
 
-Git also makes it especially easy to share and collaborate on a project with other researchers.
+Git also makes it particularly easy to share and collaborate on a project with other researchers.
 A popular service for sharing materials via Git is [GitHub](https://github.com).
@@
 
-`git push` or the green up-arrow in the git pane uploads local updates. To download this git repo to another computer type into the terminal:
+`git push` or the green up-arrow in the git pane (see Figure 3) uploads local updates. To download this git repo to another computer type into the terminal:
 
@@
 
-Git and GitHub can do even more to help you to collaborate with your fellow researchers, such as formal means to track issues and their status (open/closed/resolved) and further means to managing and merging multiple, parallel versions of code (such as branches, pull requests, merges) but this is beyond the already broad scope of this tutorial.
+Git and GitHub can do even more to support you collaborating with fellow researchers, for example, by providing a web interface to track issues and their status (open/closed/resolved) and further means to managing and merging multiple, parallel versions of code (such as branches, pull requests, or merges) but this is beyond the scope of this tutorial.
 In particular, GitHub's issue management can be leveraged as post-publication platform to discuss manuscripts and their results [^1].
 Another benefit of using Git and GitHub is that experimentation is highly encouraged since you can go back to any state quickly and even when you lose access to the file on your computer, everything is backed up on the GitHub servers.
+Further, one can reduce the chance of accumulating dead code (e.g., commented out lines) because it is safe to simply remove unneeded code blocks and track the removal in git.
 
@@
 
-Even when you have obtained a given version of a project to reproduce, and you can ensure that this version is unchanged, you may not know how to reproduce the results, because it may be unclear what files have to be run in which order.
+Even when you have obtained a given version of a project with the aim to reproduce reported result and you can ensure that this version is unchanged, you may not know exactly how to reproduce the results, because it may be unclear which files should be executed in which order.
 This is particularly the case when complex preprocessing pipelines are part of the computation and/or external dependencies are many.
@@
 
-A typical `Makefile` might look like:
+A typical `Makefile` might look like this:
 
@@
 
-The first line reads: to create `analysis.pdf` the `analysis.Rmd` needs to be rendered, which depends on `data/clean.csv`.
+The first line can be interpreted as follow: To create `analysis.pdf`, the file `analysis.Rmd` needs to be rendered, which depends on `data/clean.csv`.
 This dependency is itself a target.
@@
 Also, if one of the dependencies of a target is newer then the target itself, then Make updates everything that directly or indirectly depends on the target.
-Here, if `data\iris.csv` is newer then `data\clean.csv`, Make attempts to recreate `data\clean.csv` and `analysis.pdf`.
+Here, if `data\iris.csv` is newer then `data\clean.csv`, Make will attempt to recreate `data\clean.csv` first and then `analysis.pdf`.
 If there is a dependency which is missing, and there is no recipe to make it, `make` stops with an error.
@@
 However, succesful reproducibility relies on the crucial assumption that your computational environments are identically or sufficiently compatible, e.g., all software dependencies are installed (R and all additional R Packages) and no updates or other changes to the computational environment break or alter your analysis.
-As we will shortly see, ensuring full computational reproducibility requires one further level of documentation, that is, documentation of the computational context.
+As we will shortly see, ensuring full computational reproducibility requires one further level of documentation, that is, documentation of the computational environment.
 
@@
 Docker is a tool that allows to encapsulate, share, and re-create a computational environment on most operating systems (Windows, macOS & Linux).
-Docker achieves these goals by setting up a virtual computer on which it can execute well-defined series of commands (e.g., installing software).
-It then saves the resulting state of the virtual computer in what is called an "image." This image can be started and execute commands, e.g., running `Rscript` or `make`.
+Docker achieves these goals by setting up a virtual computer, on which it can execute well-defined series of commands (e.g., installing software).
+It then saves the resulting state of the virtual computer in what is called an "image." This image can be started and execute commands on the virtual computer, e.g., running `Rscript` or `make`.
 A running instance of an image is called a container.
 An image can be transferred and executed on any machine that has docker installed.
-Irrespetictive of the machine that is executing the container, the computational environment is the same for the programs running inside the container.
+Irrespective of the machine that is executing the container, the computational environment is the same for the programs running inside the container.
 The most important advantage over traditional virtual machines is that containers are lightweight, that is, they start fast and don't need much storage space.
@@
 Generally, with containers we would like to safe-guard against changes to the computational environment with unexpected consequences, e.g., changes in the functionality or default options in packages or even in the R environment itself.
-While the R programming language is considered stable and much effort is put into backward compatibility, sometimes even basic functions like `sample()`, to randomly sample from a set, change their behavior from one version to another.
+While the R programming language is considered stable and much effort is put into backward compatibility, sometimes even basic functions like `sample()` (to randomly sample from a set) change their behavior from one version to another.
 To ensure reproducibility in analyses based on a computers pseudo random number generator (PRNG), it is good practice to rely on fixed PRNG seeds, which are numeric values that set the PRNG into a deterministic state, i.e., the sequence of random numbers reproduces exactly.
-
 Consider the following R-code to randomly draw five numbers between 1 and 10:
@@
 ```
-
 The usual expectation is that this code delivers the same pseudo-random five numbers regardless of the operating system or R-Version (due to `set.seed()`).
@@
 Note that this is the intended behavior and the result of a [bugfix](https://bugs.r-project.org/bugzilla/show_bug.cgi?id=17494) in the random number generator implemented from R 3.6.0 upwards.
-Such changes may render analyses run on previous R versions strictly not reproducible if they contain, e.g., multiple imputations, bootstrapping, simulations studies, graphics with random jitter, Bayesian estimations using sampling algorithms (such as Markov Chain Monte Carlo), or similar techniques that involve random sampling.
+Now, such changes may render analyses run on previous R versions strictly not reproducible if they contain, e.g., multiple imputations, bootstrapping, simulations studies, graphics with random jitter, Bayesian estimations using sampling algorithms (such as Markov Chain Monte Carlo), or similar techniques that involve random sampling.
 We would like to illustrate this with a more concrete example (the full R code to reproduce this non-reproducibility is provided in the GitHub repository of this manuscript).
 Using a simulated dataset with two variables `x` and `y`, we ran a linear regression model with R's `lm()` function regressing `x` on `y`.
-Using the `boot` package [@boot], we bootstrapped the 95% confidence intervals around the regression estimate with 1000 bootstrap samples to evaluate whether zero was inside the confindence interval.
+Using the `boot` package [@boot], we bootstrapped the 95% confidence intervals around the regression coefficient estimate with 1000 bootstrap samples to evaluate whether zero was inside the confidence interval.
 To make the analysis reproducible, we set a random seed.
@@
 As we see from these R outputs, the confidence intervals once include zero and once do not include zero.
-Please note that one could discuss deeper issues about null hypothesis significance testing here but we would simply like to stress the point that computational reproducibility in the strict sense requires capturing the full computational environment.
+Please note that one could discuss deeper issues about null hypothesis significance testing here but with this example, we would simply like to stress the point that computational reproducibility in the strict sense requires capturing the full computational environment.
 
 Only rarely does an analysis depend only on base R.
-Often, a considerable number of packages is required, and they will increase the likelihood of breaking reproducibiity with every update to any of the packages of base R.
+Often, a considerable number of packages is required, and the addition of each package will increase the likelihood of breaking reproducibiity with every update to any of the packages of base R.
 The whole endeavor of reproducibility is therefore at stake, every time an update is rolled out.
-To ensure long-term reproducibility, `docker` replicates the original computational environment of an analysis exactly.
+To ensure long-term reproducibility, Docker replicates the original computational environment of an analysis exactly.
 Note that we do not intend to advocate that software should not be updated; updates typically promote bugfixes and provide new functionality; but full computational reproducibility is only achieved if the originally used software versions are documented.
 Among other things, this allows to trace back recent updates as to what change in what package caused a non-reproduction of the original results.
-Quite to the contrary, with solutions like docker, it gets more comfortable than ever to safely update to new versions, by just bumping the R Version number of the docker image.
-This convenience is possible because of the efforts of the [rocker project](https://GitHub.com/rocker-org/rocker) [@boettigerIntroductionRockerDocker2017a], which provides docker images preconfigured with an R-installation of given R versions.
+Quite to the contrary, with solutions like Docker, it gets more comfortable than ever to safely update to new versions, by just changing the R Version number of the Docker image.
+This convenience is possible because of the efforts of the [rocker project](https://GitHub.com/rocker-org/rocker) [@boettigerIntroductionRockerDocker2017a], which provides Docker images preconfigured with an R-installation of given R versions.
 These packages are taken from MRAN [@revolutionanalyticsReproducibilityUsingFixed2019], a repository for R packages fixed to the last date the R-version of the image was the most recent.
@@
 
-The basic rocker images are extendable by a Dockerfile, e.g., the basis for this articles docker image is the following Dockerfile:
+Building on a basic rocker image, we can specify further requirements for software in a Dockerfile. For example, the basis for this article's docker image is the following Dockerfile:
 
@@
 The command `docker build -t image-name` creates an image named `image-name` from the Dockerfile in the project.
-A way to automatically find out the dependencies and genereate a docker image from them is provided in the [liftR](https://cran.r-project.org/web/packages/liftr/vignettes/liftr-intro.html) package [@liftr].
+A way to automatically find out the dependencies and genereate a docker image from them is provided in the [`liftR`](https://cran.r-project.org/web/packages/liftr/vignettes/liftr-intro.html) package [@liftr].
 
-There are two principal ways to share a docker image, either by sharing the Dockerfile that creates the image, or by sharing the image itself, e.g., through a service like [dockerhub](https://hub.docker.com).
+There are two principled ways to share a docker image, either by sharing the Dockerfile that creates the image, or by sharing the image itself, e.g., through a service like [dockerhub](https://hub.docker.com).
 While both ways guarantee a replicable computational environment, sharing the Dockerfile is more transparent; in our workflow, we can use Git to track changes in the Dockerfile (such as updates to dependencies).
 
-There are two ways to use a rocker image. Both are based on the `docker run` command. The first way is to run a command inside the container. The call takes the form:
+To start a container and run commands in a container. Both are based on the `docker run` command. The first way is to run a command inside the container. The call takes the form:
 
@@
 
-The second way is to supply no commands and to interact with the container via the webbrowser and the rstudio server instance that runs in it.
+The second way is to start the container in the background (not supplying a specific command) and to interact with the container via the webbrowser and the rstudio server instance that runs in it.
 In order to do so, you need to supply a password to log into Rstudio server (`-e PASSWORD=<YOUR_PASS>`) and open a port (`-p 8787:8787`).
@@
 
-Since Docker commands grow long and are tedious to type manually, we advocate to use some automatic way to generate them.
+Since Docker commands tend to grow long and become tedious to type manually, we advocate to use some automatic way to generate them.
 Fortunately, one can use `make` to automatically generate the `docker` commands, e.g.
-the (simplified) Makefile for this paper allows command after `$(run)` to be condtionally passed through docker if one calls `make DOCKER=TRUE` (otherwise, they are run locally):
+the (simplified) Makefile for this paper allows command after `$(run)` to be conditionally passed through Docker if one calls `make DOCKER=TRUE` (otherwise, they are run locally):
 
@@
 
-Other than on R, RStudio, and R Markdown, our workflow relies on three software external pieces of software: `git`, `make`, and `docker`.
+Other than on R, RStudio, and R Markdown, our workflow relies on three pieces of software from outside the R environment: `git`, `make`, and `docker`.
 The installation process of these software packages depends on the operating system, e.g., on macOS, Make is allways availible, or on Linux, Git and Make are often allready installed.
@@
 The example shows a minimalistic analysis of the 'Considerations of Future Consequences (CFC) Scale'.
-The analysis demonstrates a succesful implementation of our workflow including downloading external data, comparing their state using a checksum, and running a confirmatory factor analysis on the first few items using lavaan [@lavaan2012].
+The analysis demonstrates a succesful implementation of our workflow including downloading external data, comparing their integrity using a checksum, and running a confirmatory factor analysis on the first few items using `lavaan` [@lavaan2012].
 Once all mentioned tools are installed on a computer, the following few steps are necessary to reproduce our analysis:
@@
 
-The overarching goal of this paper was to provide an easy to use workflow, that allows confidence in reproducibility.
-This workflow can be almost unfailingly trusted to be reproduced with the following commands:
+The overarching goal of this paper was to provide a complete and easy-to-use workflow that allows confidence in reproducibility of R-based data analyses.
+Analysis following our workflow can be reproduced with four commands (here shown for our paper):
 
@@
 
-That is, it reproduces a scientific report exactly without regard to the user, the operating system, the software, the timepoint or the recent changes to the involved files.
-To that end, it relies on tools that have been the foundation of reliable software development for years or in most cases, decades.
-As a by-product, it makes transparent how statistical results depend on the software that created them and enables, through that transparency, reuse by other researchers.
+That is, it reproduces a scientific report exactly without regard to the user, the operating system, the software, the timepoint or interim changes to the involved files.
+To that end, the proposed workflow relies on tools that have been the foundation of reliable software development for years or even decades.
+As a by-product, it makes transparent how statistical results depend on the software that created them and, by virtue of this transparency, facilitates later re-use by other researchers.
 
-Each tool reduces the likelihood of non-reproduction.
+Each tool in the workflow reduces the chances of non-reproduction.
 Dynamic reporting with R Markdown guarantuees consistency between computational results and their reporting; Version control with `git` guarantuees permanence and consistency across multiple versions of data and code; Dependency management with `make` guarantuees defined entry-points and dependency resolution; containerization with `docker` guarantuees reproducibility of the full computational environment.
@@
 There are R packages that implement parts of our workflow and, thus, lower the threshold to adoption when the full flexibility is not needed.
-The use of R Markdown within a project, tracked with `git` can be simplified with the [workflowR](https://GitHub.com/jdblischak/workflowr) package [@workflowr].
-The [drake](https://GitHub.com/ropensci/drake) package [@drake] is directly inspired by `make` but can only handle dependencies within R.
+The use of R Markdown within a project, tracked with `git` can be simplified with the [`workflowR`](https://GitHub.com/jdblischak/workflowr) package [@workflowr].
+The [`drake`](https://GitHub.com/ropensci/drake) package [@drake] is directly inspired by `make` but can only handle dependencies within R.
 The [liftR ](https://cran.r-project.org/web/packages/liftr/vignettes/liftr-intro.html) package [@liftr] and the [holepunch](https://karthik.GitHub.io/holepunch/) package [@holepunch] automatize the use of docker.
@@
 Another way to increase the chance of reproducibility is to develop an R package, which is a self-contained set of files with well-defined meta data, for each analysis.
-The reasoning is that by abiding the strict rules of package development and specifieng all dependencies a high degree of stability can be archieved (e.g., @marwickPackagingDataAnalytical2018).
+The reasoning is that, by abiding the strict rules of package development and specifying all dependencies, a high degree of stability can be achieved (e.g., @marwickPackagingDataAnalytical2018).
 
-However, beside these tools that ease the process of creating a workflow on the researchers side, we see increased interest in changing how research is published and used [@perkelToolkitDataTransparency2018], with *life code* [@perkelPioneeringLivecodeArticle2019] and *continuous integration* [@beaulieu-jonesReproducibilityComputationalWorkflows2017; @yenniDevelopingModernData2018].
-These techniques give us a glimpse of a paradigm shift from static to dynamic, interactive and living publications.
+Other than these tools that ease the process of creating workflows like ours, we noticed increased interest in changing how research is published and used [@perkelToolkitDataTransparency2018], with *life code* [@perkelPioneeringLivecodeArticle2019] and *continuous integration* [@beaulieu-jonesReproducibilityComputationalWorkflows2017; @yenniDevelopingModernData2018].
+These techniques give us a glimpse of a paradigm shift from static to dynamic, interactive and living publications that is yet to happen.
 
@@
 For example, already the installation of all tools is easier on POSIX-compatible platforms such as Unix, Linux, or macOS (but not Windows).
-However, once such a workflow is established as a default, it can be used with minimal changes to the familiar use of the R environment.
-It is even possible to move the entire development process to a docker environment.
+However, once a reproducible workflow is established as a default, it can be used with minimal changes from project to project.
 
 In our own experience, it is often not possible to convince all co-authors to switch to a different document processing environment, such as R Markdown.
-That is, we have experienced that after writing up the first draft in R Markdown, one must eventually genereate a Word-file that is from then on used as static file serving as a basis for multiple iterations among the co-authors.
-Reteining reproduciblity in such situation requires tediues manuel synchronization.
-This labor can be saved with the [rodoc package](https://noamross.github.io/redoc/articles/mixed-workflows-with-redoc.html) [@redoc], which enables a two-way workflow between Word and R Markdown.
-The conversion between R Markdown and Word retains all changes and even supports words tracked changes feature.
+That is, we have experienced that after writing up the first draft in R Markdown, we eventually had to genereate a Word-file that, from then on, was used as static file serving as a basis for multiple iterations among the co-authors.
+Retaining reproduciblity in such situation then requires tedious manual synchronization of files across formats.
+This annoyance may be reduced with the [`rodoc` package](https://noamross.github.io/redoc/articles/mixed-workflows-with-redoc.html) [@redoc], which enables a bidirectional synchronization between Word and R Markdown.
+Conversions between R Markdown and Word retain all changes and even support Word's track-changes feature.
 
@@
 How to best share a reproducible workflow?
-We believe that, ideally, a service provider should be found that guarantuees permanent and reliable hosting of the reproducible workflow, such as the Open Science Framework [@fosterOpenScienceFramework2017], that is independ of GitHub, Docker Hub & MRAN.
+We believe that, ideally, a non-commercial public service provider should be found that guarantuees permanent and reliable hosting of our reproducible workflows, such as the Open Science Framework [@fosterOpenScienceFramework2017].
+A provider mirroring and complementing the services offered by GitHub, Docker Hub, and MRAN.
 Second, to ensure that other users are legally able to benefit from the shared materials, authors must choose an appropriate license format.
@@
 Both licenses assure maximal freedom for future users while requiring the attribution of the original authors in derivative work.
-These licenses are also in line with the recommendations by the Reproducibile Research Standard [@stoddenEnhancingReproducibilityComputational2016; @stoddenEnablingReproducibleResearch2009].
-To facilitate an inclusive environment we recommend to name all contributer and include a [Code of Conduct](https://opensource.guide/code-of-conduct/).
+These licenses are also in line with the recommendations by the Reproducible Research Standard [@stoddenEnhancingReproducibilityComputational2016; @stoddenEnablingReproducibleResearch2009].
+To facilitate an inclusive environment we recommend to name all contributors and include a [Code of Conduct](https://opensource.guide/code-of-conduct/) in your project.
 
@@
 
-The presented workflow leverages various existing tools that are partly integrated into Rstudio already.
-Parts of the proposed workflow have been integrated into stand-alone packages (such as @workflowr or @holepunch) but they are either incomplete or rely on propietary service providers.
-We hope that with increasing awareness of the problems associated with computational reproducibility, the increased demand for unified and open solutions, will lead to an even better integration of existing tools to allow reproducible workflows to become a standard in the psychological research.
+The proposed workflow leverages various existing tools that are partly integrated into Rstudio already.
+Parts of the proposed workflow have been integrated into stand-alone packages (such as workflowr @[workflowr] or holepunch  @[holepunch]) but they are either incomplete in the sense of our proposal or rely heavily on propietary service providers.
+We hope that with increasing awareness of the challenges of computational reproducibility, the increased demand for unified and open solutions will lead to an even better integration of existing tools to allow reproducible workflows to become a standard in psychological research.
 

### 2019-11-08 manuscript.Rmd
fixes in bibliography added keywords

@@
 
-keywords          : "keywords"
+keywords          : "reproducibility, R, version management, literate programming, dependency management, containerization, open science"
 wordcount         : "`r tryCatch(wordcountaddin::word_count(here::here('manuscript.Rmd')), error=function(err) 'NA')`"

### 2019-11-11 README.Rmd
rename to manuscript close #58
describe windows workaround close #43

@@
 
-Open `rr-flow.Rmd` click on `Knit`.
+Open `manuscript.Rmd` click on `Knit`.
 
@@
 
+**Note: Windows user need to hand edit the `Makefile` and set current_path to the current directory and use `make all DOCKER=TRUE WINDOWS=TRUE`.
+We hope that future releases of Docker for Windows will not require that workaround.**
+
 ## Rebuild Everything

### 2019-11-11 manuscript.Rmd
describe windows workaround close #43

@@
 To use `docker` you need to start Docker Desktop.
+In the settings of docker you have to allow the sharing of your drive.
+Docker on Windows require the path to be special therfore you need to hand edit the `Makefile` and set `current_path` to the current directory and use `make all DOCKER=TRUE WINDOWS=TRUE`.
+We hope that future releases of Docker for Windows will not require that workaround.
 
@@
 
-To use `docker` you need to start Docker Desktop.
+To use `docker` you need to start Docker Desktop. In the settings of docker you have to allow the sharing of your drive.
 

### 2019-11-11 abstract.Rmd
add changes from Julia Delius

@@
 The workflow leverages established tools and practices from software engineering.
-It combines the benefits of various open-source software tools including R Markdown, Git, Make, and Docker, whose interplay ensures seamless integration of version management, APA-conformable dynamic report generation and full cross-platform and long-term computational reproducibility.
+It combines the benefits of various open-source software tools including R Markdown, Git, Make, and Docker, whose interplay ensures seamless integration of version management, dynamic report generation conforming to various journal styles and full cross-platform and long-term computational reproducibility.
 The workflow ensures meeting the primary goals that 1) the reporting of statistical results is consistent with the actual statistical results (dynamic report generation), the analysis exactly reproduces at a later time even if the computing platform or software is changed (computational reproducibility), and 3) changes at any time (during development and post-publication) are tracked, tagged, and documented while earlier versions of both data and code remain accessible.

### 2019-11-11 manuscript.Rmd
added thanks to JD in author notes
add changes from Julia Delius

--- 2019-11-08 manuscript.Rmd
@@
 ---
-title             : "A Reproducible Data Analysis Workflow with R Markdown, Git, Make, & Docker"
+title             : "A Reproducible Data Analysis Workflow with R Markdown, Git, Make, and Docker"
 shorttitle        : "Reproducible Data Analysis Workflow"
@@
     institution   : "Max Planck UCL Centre for Computational Psychiatry and Ageing Research, Berlin, Germany & London, U.K." 
-
-authornote: This paper is fully reproducible using the workflow described in this paper.
-All materials can be found on the accompanying GitHub repository <https://github.com/aaronpeikert/reproducible-research/>.
-To provide feedback or any other type of comment or questions regarding our workflow, please add an issue to the GitHub repository of our paper [https://github.com/aaronpeikert/reproducible-research/issues](https://github.com/aaronpeikert/reproducible-research/issues).
+authornote: |
+   We are grateful to Julia Delius for her helpful assistance in language and style editing.
+This paper is fully reproducible using the workflow described in this paper.
+All materials can be found on the accompanying GitHub repository: [https://github.com/aaronpeikert/reproducible-research/](https://github.com/aaronpeikert/reproducible-research/).
+To provide feedback or any other type of comment or questions regarding our workflow, please add an issue to the GitHub repository of our paper at [https://github.com/aaronpeikert/reproducible-research/issues](https://github.com/aaronpeikert/reproducible-research/issues).
 
@@
 R-based data analyses.
-Reproducibility means to us that statistical results together with their generating computational workflows are reported fully, and transparently, and remain sustainably available, such that a given workflow can be re-run by a different person later in time and the results will be identical to the ones intially reported [@herouxCompatibleReproducibilityTaxonomy2018; @claerboutElectronicDocumentsGive1992].
-The need to ensure reproducibility directly follows from commonly accepted rules of good scientific practice (such as the guidelines of the German Research Foundation; @dfg2019).
+To us Reproducibility means that statistical results as well as their generating computational workflows are reported fully, and transparently, and remain sustainably available, such that a given workflow can be re-run by a different person later in time and the results will be identical to the ones intially reported [@herouxCompatibleReproducibilityTaxonomy2018; @claerboutElectronicDocumentsGive1992].
+The need to ensure reproducibility directly follows from commonly accepted rules of good scientific practice [such as the guidelines of the German Research Foundation, @dfg2019].
 Ensuring reproducibility is a prerequisite for replicability and a means to increase trustworthiness of empirical results [@epskamp2019rep].
 
-Here, we combine four software tools, whose interplay can guarantuee full computational reproducibility of data analyses and their reporting.
-There are various ideas to enhance reproducibility [@piccoloToolsTechniquesComputational2016], of which we believe four of them are particularly important: literate programming: [@ruleTenSimpleRules2019], version control [@barbaHardRoadReproducibility2016], dependency management [@askrenUsingMakeReproducible2016] and containerization [@clyburne-sherinComputationalReproducibilityContainers2018].
-We argue that only a workflow using all four concepts in unison can guarantuee confidence in reproducing a scientific report.
-Even though there are various implementations of these concepts, we believe that for the R environment [@r2019], the following four implementations are best suited.
-For literate programming, R Markdown [@rmarkdown2019], for version control, Git [@chaconProGit2014], for dependency management, Make [@feldmanMakeProgramMaintaining1979], and for containerization, Docker [@merkelDockerLightweightLinux2014].
-Each of these software solutions serves a meta-scientific valuable goal (reproducibility) and increases the productivity of the researcher.
-They all are highly flexible and powerful, so to master them fully, a considerable amount of practice is necessary.
-However, for our purposes, it is sufficient to master a minimal valuable subset of functions to ensure reproducibility of scientific analyses.
+Here, we combine four software tools, whose interplay can guarantee full computational reproducibility of data analyses and their reporting.
+There are various ideas on how to enhance reproducibility [@piccoloToolsTechniquesComputational2016], four of which we believe to be particularly important: literate programming: [@ruleTenSimpleRules2019], version control [@barbaHardRoadReproducibility2016], dependency management [@askrenUsingMakeReproducible2016], and containerization [@clyburne-sherinComputationalReproducibilityContainers2018].
+We argue that only a workflow using all four concepts in unison can guarantee confidence in reproducing a scientific report.
+Various implementations of these concepts exis6, but we consider the following four best suited for the R environment [@r2019]: for literate programming, R Markdown [@rmarkdown2019], for version control, Git [@chaconProGit2014], for dependency management, Make [@feldmanMakeProgramMaintaining1979], and for containerization, Docker [@merkelDockerLightweightLinux2014].
+Each of these software solutions serves a valuable meta-scientific goal (reproducibility) and increases the researches' productivity.
+They all are highly flexible and powerful, so that mastering them fully, requires a considerable amount of practice.
+However, for our purposes, it is sufficient to master a valuable minimal subset of functions to ensure reproducibility of scientific analyses.
 
@@
 This paper is fully reproducible using the workflow described in this paper.
-All materials, including the manuscript source and all R codes and files used to generate this manuscript, can be found on the accompanying GitHub repository [https://github.com/aaronpeikert/reproducible-research/](https://github.com/aaronpeikert/reproducible-research/).
+All materials, including the manuscript source and all R codes and files used to generate this manuscript, can be found on the accompanying GitHub repository: [https://github.com/aaronpeikert/reproducible-research/](https://github.com/aaronpeikert/reproducible-research/).
 
@@
 
-The first step towards reproducibility is to create an R script or R Markdown file as primary entry point for the analysis, which runs on a local computer without error and performs the main statistical analyses.
-Next, one has to make sure that all files relevant for the analysis can be moved to another computer.
-To this end, it is recommendable that all files are within one folder (or, subfolders within it) and all paths are relative to that folder.
-A robust solution to the problem of making sure that file access does not break across computing platforms are [Rstudio projects](https://r4ds.had.co.nz/workflow-projects.html) and the `here` [@here2017] package.
+The first step towards reproducibility is to create an R script or R Markdown file as primary entry point for the analysis that runs on a local computer without error and performs the main statistical analyses.
+Next, one needs to make sure that all files relevant for the analysis can be moved to another computer.
+To this end, it is recommendable that all files are within one folder (or enclosed subfolders within it) and all paths are relative to that folder.
+A robust solution to the problem of making sure that file access does not break across computing platforms are [Rstudio projects](https://r4ds.had.co.nz/workflow-projects.html) and the `here` package [@here2017].
 
@@
 
-The folder, where all files reside that you need for analysis (code and data), is referred to as a 'project' (or sometimes 'research compendium').
-Working with projects is particularly comfortable with RStudio, an integrated development environment (IDE) for R.
-It is useful to organize a data-analysis project in way that strictly segregates (raw) data and code by placing them in directories called `data` and `R` [see Section 4 in @marwickPackagingDataAnalytical2018]; there are also tools that automatize the standardized creation of folder structures such as `workflowr` [@workflowr].
+The folder where all the files reside that you need for analysis (code and data), is referred to as a 'project' (or sometimes as a 'research compendium').
+Working with projects is particularly convinient with RStudio, an integrated development environment (IDE) for R.
+It is useful to organize a data analysis project in a way that strictly segregates (raw) data and code by placing them in directories called `data` and `R` [see Section 4 in @marwickPackagingDataAnalytical2018]; there are also tools that automatize the standardized creation of folder structures such as `workflowr` [@workflowr].
 
-Sometimes external requirements make it impossible that the data can be stored and shared together with the scripts.
+Sometimes external requirements make it impossible fot the data to be stored and shared together with the scripts.
 In most of the cases we have seen, these are either space constraints or privacy considerations.
 In these cases, of course, unrestricted reproducibility is not guaranteed.
-If dividing data and scripts is unavoidable, we recommend to validate all data files using checksums (also called a "hash", e.g., using the functions provided in package `digest` [@digest2019]) before analyzing them.
+If dividing data and scripts is unavoidable, we recommend validating all data files using checksums (also called a "hash", e.g., using the functions provided in package `digest` [@digest2019]) before analyzing them.
 To do this, a checksum must be created and stored at the time of the original analysis.
-At the time of reproduction, the current checksum must be compared with the stored checksum to ensure data integrity.
+At the time of reproduction, the current checksum must then be compared with the stored checksum to ensure data integrity.
 
@@
 Typical errors result from copy\&paste mistakes, erroneous rounding, or missed updates of the manuscript when the computer code has changed.
-To not only create fully reproducible results but also fully reproducible reports, one needs to resort to the Literate Programming paradigm [@knuth1984literate], in which human-readable language and computer-code are interspersed to create dynamic documents whose order follows the logic of thought rather then the order of the computer.
-R Markdown is a simple markup language to create dynamic documents with embedded chunks of R code that can be exported to standard formats such as documents (docx, rtf, epub), presentations (ppt, html) or websites (html) using the knitr package [@xie2015; @xie2019].
+In order to not only create fully reproducible results but also fully reproducible reports, one needs to resort to the Literate Programming paradigm [@knuth1984literate], in which human-readable language and computer code are intermingled to create dynamic documents whose order follows the logic of thought rather then the order of the computer.
+R Markdown is a simple markup language to create dynamic documents with embedded chunks of R code that can be exported to standard formats such as documents (docx, rtf, epub), presentations (ppt, html) or websites (html) using the `knitr` package [@xie2015; @xie2019].
 Several packages extend the functionality of knitr.
-Of particular note are the `papaja` package [@papaja2018], which offers additional functions to enable APA-conformable document formatting, including a journal-style final typeset format, and the `stargazer` package [@stargazer2018], which provides journal-ready tables and reports of statistical models.
-Figure 1 illustrates R Markdown syntax using the papaja package and Figure 2 shows the resulting rendered document.
+Of particular note are the `papaja` package [@papaja2018], which offers additional functions to enable American Psychological Association (APA) style document formatting, including a journal-style final typeset format, and the `stargazer` package [@stargazer2018], which provides journal-ready tables and reports of statistical models.
+Figure 1 illustrates R Markdown syntax using the `papaja` package and Figure 2 shows the resulting rendered document.
 
@@
 
-Fundamentally, reproducibility means that results remain identical if neither script nor data have changed.
-It is often not trivial to find out whether anything in a project has changed over time and if so to "go back in time." The program Git enables you to do both.
+Fundamentally, reproducibility means that results remain identical if neither the script nor th data have changed.
+It is often not trivial to find out whether anything in a project has changed over time and if so, to "go back in time." The Git program enables you to do both.
 A good mental model for Git is that it takes a sequence of snapshots of all files it is supposed to track.
-In the language of Git, these snapshots are "commits".
-A commit represents a complete copy of the state of all tracked files.
+In the language of Git, these snapshots are "commits." A commit represents a complete copy of the state of all tracked files.
 Each commit has a unique identifier (a hash) and a human-readable description (commit message).
-Going back to one state is as easy as traversing the history of all commits and switch to a given one; it is possible to visually track changes between different versions.
+Going back to one state is as easy as traversing the history of all commits and switching to a given one; it is possible to visually track changes between different versions.
 Finding a specific version can be made even easier by tagging snapshots, e.g., as "submission," "preprint," or "published." This collection of snapshots is called a "repository," which ideally tracks your entire R project.
@@
 
-To keep track of all changes on your local computer, you only need to use `git add` & `git commit` or `git commit -a` to do both.
+To keep track of all changes on your local computer, you only need to use `git add` and `git commit` or `git commit -a` to add and commit at the same time.
 These commands need to be executed in the terminal, which you can access from within Rstudio (`Shift + Alt + R`).
@@
 Just sharing Git repositories with the public is always free, private repositories (only visible to persons you invite) are [free for researchers](https://help.github.com/en/articles/applying-for-an-educator-or-researcher-discount).
-After creating a user account, one can create a new repository where GitHub advises you how to upload your repository from the terminal, e.g., for our repository (here with user name 'aaronpeikert' and repository name 'reproducible-research'):
+After creating a user account, one can create a new repository and GitHub provides information on how to upload your repository from the terminal, e.g., for our repository (here with user name 'aaronpeikert' and repository name 'reproducible-research'):
 
@@
 
-`git push` or the green up-arrow in the git pane (see Figure 3) uploads local updates. To download this git repo to another computer type into the terminal:
+`git push` or the green upward arrow in the Git pane (see Figure 3) uploads local updates. To download this Git repository to another computer, type into the terminal:
 
@@
 
-Git and GitHub can do even more to support you collaborating with fellow researchers, for example, by providing a web interface to track issues and their status (open/closed/resolved) and further means to managing and merging multiple, parallel versions of code (such as branches, pull requests, or merges) but this is beyond the scope of this tutorial.
-In particular, GitHub's issue management can be leveraged as post-publication platform to discuss manuscripts and their results [^1].
+Git and GitHub can do even more to support you when collaborating with fellow researchers, for example, by providing a web interface to track issues and their status (open/closed/resolved) and further means to manage and merge multiple, parallel versions of code (such as branches, pull requests, or merges), but this is beyond the scope of this tutorial.
+In particular, GitHub's issue management can be leveraged as a post-publication platform to discuss manuscripts and their results [^1].
 Another benefit of using Git and GitHub is that experimentation is highly encouraged since you can go back to any state quickly and even when you lose access to the file on your computer, everything is backed up on the GitHub servers.
-Further, one can reduce the chance of accumulating dead code (e.g., commented out lines) because it is safe to simply remove unneeded code blocks and track the removal in git.
+Further, one can reduce the liklihood of dead code accumulating (e.g., lines that have been commented out) because it is safe to simply remove unneeded code blocks and track the removal in git.
 
@@
 
-Even when you have obtained a given version of a project with the aim to reproduce reported result and you can ensure that this version is unchanged, you may not know exactly how to reproduce the results, because it may be unclear which files should be executed in which order.
+Even when you have obtained a given version of a project with the aim to reproduce reported result and you can confirm that this version is unchanged, you may not know exactly how to reproduce the results because it may be unclear which files should be executed in which order.
 This is particularly the case when complex preprocessing pipelines are part of the computation and/or external dependencies are many.
-Handling such dependencies is comfortable with Make because it allows you to create (computational) recipes to (re-)create files.
+Handling such dependencies is easy with Make because it allows you to create (computational) recipes to (re-)create files.
 
@@
 
-The first line can be interpreted as follow: To create `analysis.pdf`, the file `analysis.Rmd` needs to be rendered, which depends on `data/clean.csv`.
+The first line can be interpreted as follows: To create `analysis.pdf`, the file `analysis.Rmd`, which depends on `data/clean.csv`, needs to be rendered.
 This dependency is itself a target.
@@
 Here, if `data\iris.csv` is newer then `data\clean.csv`, Make will attempt to recreate `data\clean.csv` first and then `analysis.pdf`.
-If there is a dependency which is missing, and there is no recipe to make it, `make` stops with an error.
-It is a convention to have a target `all` which depends on everything and make it the first target in a `Makefile`, then the command `make` without any argument automatically creates everything.
-The button `Build All` from within RStudio triggers this process.
+If there is a dependency which is missing, and there is no recipe to make it, `make` stops with an error message.
+It is a convention to have the first target named `all`, which depends on everything.
+Then the command `make` without any argument automatically creates everything.
+The button `Build All` from within RStudio triggers this process (see Figure 4).
 
-![Build Pane in RStudio](Images/build-pane.png)
+![Build pane in RStudio](Images/build-pane.png)
 
@@
 
-However, succesful reproducibility relies on the crucial assumption that your computational environments are identically or sufficiently compatible, e.g., all software dependencies are installed (R and all additional R Packages) and no updates or other changes to the computational environment break or alter your analysis.
+However, successful reproducibility relies on the crucial assumption that your computational environments are identically or sufficiently compatible, e.g., all software dependencies are installed (R and all additional R Packages) and no updates or other changes to the computational environment break or alter your analysis.
 As we will shortly see, ensuring full computational reproducibility requires one further level of documentation, that is, documentation of the computational environment.
@@
 
-Docker is a tool that allows to encapsulate, share, and re-create a computational environment on most operating systems (Windows, macOS & Linux).
+Docker is a tool that allows encapsulation, sharing, and re-creation of a computational environment on most operating systems (Windows, macOS & Linux).
 Docker achieves these goals by setting up a virtual computer, on which it can execute well-defined series of commands (e.g., installing software).
@@
 Irrespective of the machine that is executing the container, the computational environment is the same for the programs running inside the container.
-The most important advantage over traditional virtual machines is that containers are lightweight, that is, they start fast and don't need much storage space.
-Docker achieves this by reusing large parts of the hosts operating system (how much is varying widely between Linux, macOS & Windows).
+The most important advantage over traditional virtual machines is that containers are lightweight, that is, they start rapidly and do not need much storage space.
+Docker achieves this by reusing large parts of the host's operating system (the extend varies between Linux, macOS and Windows).
 
 With the following example, we demonstrate why it is important to store the computational environment.
-Generally, with containers we would like to safe-guard against changes to the computational environment with unexpected consequences, e.g., changes in the functionality or default options in packages or even in the R environment itself.
-While the R programming language is considered stable and much effort is put into backward compatibility, sometimes even basic functions like `sample()` (to randomly sample from a set) change their behavior from one version to another.
-To ensure reproducibility in analyses based on a computers pseudo random number generator (PRNG), it is good practice to rely on fixed PRNG seeds, which are numeric values that set the PRNG into a deterministic state, i.e., the sequence of random numbers reproduces exactly.
-Consider the following R-code to randomly draw five numbers between 1 and 10:
+Generally, with containers we would like to safeguard against changes to the computational environment with unexpected consequences, e.g., changes in the functionality or default options in packages or even in the R environment itself.
+While the R programming language is considered stable and much effort is put into backward compatibility, even basic functions like `sample()` (to randomly sample from a set) sometimes change their behavior from one version to another.
+To ensure reproducibility in analyses based on a computer's pseudo random number generator (PRNG), it is good practice to rely on fixed PRNG seeds, which are numeric values that set the PRNG into a deterministic state, i.e., the sequence of random numbers reproduces exactly.
+Consider the following R code to randomly draw five numbers between 1 and 10:
 
@@
 ```
-The usual expectation is that this code delivers the same pseudo-random five numbers regardless of the operating system or R-Version (due to `set.seed()`).
-Using Docker, we can start an image which contains the R-Version 3.5.0, and execute the code there.
+
+The usual expectation is that this code delivers the same pseudo-random five numbers regardless of the operating system or R Version (due to `set.seed()`).
+Using Docker, we can start an image which contains the R Version 3.5.0, and execute the code there.
 
@@
 
-Surprisingly in an image with a more recent version of R (3.6.0), this results in another sample despite the identical seed:
+Surprisingly in an image with a more recent version of R (3.6.1), this results in another sample despite the identical seed:
 
@@
 
-Note that this is the intended behavior and the result of a [bugfix](https://bugs.r-project.org/bugzilla/show_bug.cgi?id=17494) in the random number generator implemented from R 3.6.0 upwards.
-Now, such changes may render analyses run on previous R versions strictly not reproducible if they contain, e.g., multiple imputations, bootstrapping, simulations studies, graphics with random jitter, Bayesian estimations using sampling algorithms (such as Markov Chain Monte Carlo), or similar techniques that involve random sampling.
+Note that this is intended behavior as it is the result of a [bugfix](https://bugs.r-project.org/bugzilla/show_bug.cgi?id=17494) in the random number generator implemented from R 3.6.0 upwards.
+Now, such changes may strictly render analyses run on previous R versions not reproducible if they contain, e.g., multiple imputations, bootstrapping, simulations studies, graphics with random jitter, Bayesian estimations using sampling algorithms (such as Markov Chain Monte Carlo), or similar techniques that involve random sampling.
 We would like to illustrate this with a more concrete example (the full R code to reproduce this non-reproducibility is provided in the GitHub repository of this manuscript).
-Using a simulated dataset with two variables `x` and `y`, we ran a linear regression model with R's `lm()` function regressing `x` on `y`.
+We ran a linear regression model on a simulated dataset with two variables `x` and `y` with R's `lm()` function regressing `x` on `y`.
 Using the `boot` package [@boot], we bootstrapped the 95% confidence intervals around the regression coefficient estimate with 1000 bootstrap samples to evaluate whether zero was inside the confidence interval.
@@
 
-Then, we ran the identical script with identical seed in R 3.6.1:
+Then, we ran the identical script with the identical seed in R 3.6.1:
 
@@
 
-As we see from these R outputs, the confidence intervals once include zero and once do not include zero.
+As we see from these R outputs, the confidence intervals once include zero and once do not.
 Please note that one could discuss deeper issues about null hypothesis significance testing here but with this example, we would simply like to stress the point that computational reproducibility in the strict sense requires capturing the full computational environment.
 
-Only rarely does an analysis depend only on base R.
-Often, a considerable number of packages is required, and the addition of each package will increase the likelihood of breaking reproducibiity with every update to any of the packages of base R.
-The whole endeavor of reproducibility is therefore at stake, every time an update is rolled out.
+Only rarely does an analysis depend on base R only.
+Often, a considerable number of packages is required, and the addition of each package will increase the likelihood of breaking reproducibility with every update to any of the packages of base R.
+The whole endeavor of reproducibility is therefore at stake every time an update is rolled out.
 To ensure long-term reproducibility, Docker replicates the original computational environment of an analysis exactly.
-Note that we do not intend to advocate that software should not be updated; updates typically promote bugfixes and provide new functionality; but full computational reproducibility is only achieved if the originally used software versions are documented.
-Among other things, this allows to trace back recent updates as to what change in what package caused a non-reproduction of the original results.
-Quite to the contrary, with solutions like Docker, it gets more comfortable than ever to safely update to new versions, by just changing the R Version number of the Docker image.
-This convenience is possible because of the efforts of the [rocker project](https://GitHub.com/rocker-org/rocker) [@boettigerIntroductionRockerDocker2017a], which provides Docker images preconfigured with an R-installation of given R versions.
-These packages are taken from MRAN [@revolutionanalyticsReproducibilityUsingFixed2019], a repository for R packages fixed to the last date the R-version of the image was the most recent.
-Building upon these rocker-images, researchers can easily build their own Docker images with all required R packages.
+Note that we do not intend to advocate that software should not be updated; updates typically promote bugfixes and provide new functionality; our point is that full computational reproducibility is only achieved if the originally used software versions are documented.
+Among other things, this makes it possible to trace back recent updates as to what change in what package caused a non-reproduction of the original results.
+Quite to the contrary, with solutions like Docker, it gets easier than ever to safely update to new versions just by changing the R version number of the Docker image.
+This convenience is possible because of the efforts of the [`rocker` project](https://GitHub.com/rocker-org/rocker) [@boettigerIntroductionRockerDocker2017a], which provides Docker images preconfigured with an R-installation of given R versions.
+These packages are taken from MRAN [@revolutionanalyticsReproducibilityUsingFixed2019], a repository for R packages fixed to the last date on which the R version of the image was the most recent.
+Building upon these rocker images, researchers can easily build their own Docker images with all required R packages.
 The rocker project also provides images that include RStudio (`rocker/rstudio`), the `tidyverse` package (`rocker/tidyverse`) and the R Markdown package with LaTeX (`rocker/verse`).
@@
 Those `RUN` statements can install dependencies that are not an R package, e.g., other programming languages like python or Matlab.
-The `WORKDIR` statement is not strictly necessary but saves time spent writing the working directory.
+The `WORKDIR` statement is not strictly necessary but saves time spent on writing the working directory.
 The command `docker build -t image-name` creates an image named `image-name` from the Dockerfile in the project.
-A way to automatically find out the dependencies and genereate a docker image from them is provided in the [`liftR`](https://cran.r-project.org/web/packages/liftr/vignettes/liftr-intro.html) package [@liftr].
+A way to identify the dependencies automatically and generate a docker image out of them is provided in the [`liftR`](https://cran.r-project.org/web/packages/liftr/vignettes/liftr-intro.html) package [@liftr].
 
-There are two principled ways to share a docker image, either by sharing the Dockerfile that creates the image, or by sharing the image itself, e.g., through a service like [dockerhub](https://hub.docker.com).
+There are two principled ways to share a docker image; either by sharing the Dockerfile that creates the image, or by sharing the image itself, e.g., through a service like [dockerhub](https://hub.docker.com).
 While both ways guarantee a replicable computational environment, sharing the Dockerfile is more transparent; in our workflow, we can use Git to track changes in the Dockerfile (such as updates to dependencies).
 
-To start a container and run commands in a container. Both are based on the `docker run` command. The first way is to run a command inside the container. The call takes the form:
+There are two options to interact with a container. Both options are based on the `docker run` command. The first way is to run a command inside the container. The call takes the form:
 
@@
 
-The `--rm` flag means to not save the state of the docker after the command finishes in an container.
+The `--rm` flag means that the state of the docker after the command finishes in an container is not saved.
 The `-it` flag tells Docker to accept inputs and return outputs to the terminal.
-Interacting with docker this way is especially useful when using `R` as a command because then you have access to the standard R terminal.
+Interacting with docker this way is especially useful when using `R` as a command because you then have access to the standard R terminal.
 This is the command to start an interactive R session inside a Docker image called `reproducible-research`:
@@
 
-![R-terminal running inside docker](Images/docker-r-terminal.png)
+![R terminal running inside docker](Images/docker-r-terminal.png)
 
-The second way is to start the container in the background (not supplying a specific command) and to interact with the container via the webbrowser and the rstudio server instance that runs in it.
-In order to do so, you need to supply a password to log into Rstudio server (`-e PASSWORD=<YOUR_PASS>`) and open a port (`-p 8787:8787`).
+The second option is to start the container in the background (not supplying a specific command) and to interact with the container via the webrowser and the RStudio server instance running in it.
+In order to do so, you need to supply a password to log into the Rstudio server (`-e PASSWORD=<YOUR_PASS>`) and open a port (`-p 8787:8787`).
 
@@
 
-The adress to connect to the Rstudio server is your ip adress (or, on Linux also `localhost`) in this scheme: `<IPADRESS>:8787`.
+The adress to connect to the RStudio server is your IP address (or `localhost` Linux) in this scheme: `<IPADRESS>:8787`.
 This offers a fully functioning Rstudio instance that runs in the image but is accessible through a local web browser.
 
-By default, programs inside the container cannot access files on the local computer; to enable access, a local folder needs to be linked explicitly:
+By default, programs inside the container cannot access files on the local computer, thus requiring an explicit link to a local folder to enable access (and on Mac and Windows this also have to be allowed in the settings) :
 
@@
 
-Since Docker commands tend to grow long and become tedious to type manually, we advocate to use some automatic way to generate them.
+Since Docker commands tend to grow long and become tedious to type manually, we recomend using some automatic way to generate them.
 Fortunately, one can use `make` to automatically generate the `docker` commands, e.g.
-the (simplified) Makefile for this paper allows command after `$(run)` to be conditionally passed through Docker if one calls `make DOCKER=TRUE` (otherwise, they are run locally):
+the (simplified) `Makefile` for this paper allows command after `$(run)` to be conditionally passed through Docker if one types `make DOCKER=TRUE` (otherwise, they are run locally):
 
 ```{bash, eval=FALSE}
-projekt := reproducible-research
-current_dir := $(shell git rev-parse --show-toplevel)
-home_dir := $(shell git rev-parse --show-toplevel)
-uid = $(shell id -u)
+projekt := $(notdir $(CURDIR))
+current_dir := $(CURDIR)
+home_dir := $(current_dir)
+uid = --user $(shell id -u)
 
@@
 Other than on R, RStudio, and R Markdown, our workflow relies on three pieces of software from outside the R environment: `git`, `make`, and `docker`.
-The installation process of these software packages depends on the operating system, e.g., on macOS, Make is allways availible, or on Linux, Git and Make are often allready installed.
-In the following section, we share what have experienced as the easiest way to install those packages.
-However, installation processes may be subject to change and we advise readers to also consult the documentations of the packages.
+The installation process of these software packages depends on the operating system, e.g., on macOS, Make is always available, while on Linux, Git and Make are often already installed.
+In the following section, we share what what we consider the easiest way to install those packages.
+However, installation processes may be subject to change, and we advise that readers also consult the documentations of the packages.
 
@@
 There is a package manager for Windows called chocolately, which you can install from: [https://chocolatey.org/install](https://chocolatey.org/install), providing all dependencies.
-After having installed chocolately (and restarted the computer), all dependencies can be installed in an **admin terminal** (Windows-Key, then type `cmd`, right-click *Run as administrator*) via:
+Having installed chocolately (and restarted the computer), all dependencies can be installed in an **admin terminal** (Windows key, then type `cmd`, right-click *Run as administrator*) via:
 
@@
 To use `docker` you need to start Docker Desktop.
+In the settings of docker you have to allow the sharing of your drive.
+Docker on Windows require the path to be special; therefore you need to hand edit the `Makefile` and set `current_path` to the current directory and use `make all DOCKER=TRUE WINDOWS=TRUE`.
+We hope that future releases of Docker for Windows will not require that workaround.
 
@@
 
-There is a host of diffferent Linux distributions and almost as many package managers.
-Still, to our knowledge there is no (recent) Linux edition, where `git`, `make` and `docker` are not availible.
+There is a host of different Linux distributions and almost as many package managers.
+Still, to our knowledge, there is no (recent) Linux edition, that does not include `git`, `make` and `docker`.
 For example, in Ubuntu Linux, installation is straightforward using the shipped package manager:
@@
 You may need elevated rights for the installation; in this case add `sudo` before the installation command.
-`docker` also needs elevated rights to run; therefore, we recomend to add the local user to the `docker` group, following the [documentation of docker](https://docs.docker.com/install/linux/linux-postinstall/#manage-docker-as-a-non-root-user).
+`docker` also needs elevated rights to run; therefore, we recommend adding the local user to the `docker` group, following the [documentation of docker](https://docs.docker.com/install/linux/linux-postinstall/#manage-docker-as-a-non-root-user).
 
 An alternative to `docker` on Linux is `singularity` [@kurtzerSingularityScientificContainers2017].
-Just replace any docker calls with `singularity docker`, because singularity fully supports docker-images.
-A possible advantage is that singularity works well in high-performance computing environments and on old Linux versions, the downside is that `singularity` is only availibe on Linux.
+Just replace any docker calls with `singularity docker`, because singularity fully supports docker images.
+A possible advantage is that singularity works well in high-performance computing environments and on old Linux versions, the downside is that `singularity` is only available on Linux.
 
@@
 
-We provide a reproducible analysis as working example via [GitHub](https://github.com/aaronpeikert/workflow-showcase) as a testbed for users.
+We provide a reproducible analysis as a working example via [GitHub](https://github.com/aaronpeikert/workflow-showcase) as a testbed for users.
 The example shows a minimalistic analysis of the 'Considerations of Future Consequences (CFC) Scale'.
-The analysis demonstrates a succesful implementation of our workflow including downloading external data, comparing their integrity using a checksum, and running a confirmatory factor analysis on the first few items using `lavaan` [@lavaan2012].
+The analysis demonstrates a successful implementation of our workflow including downloads of external data, comparison of their integrity using a checksum, and the running of a confirmatory factor analysis on the first few items using `lavaan` [@lavaan2012].
 Once all mentioned tools are installed on a computer, the following few steps are necessary to reproduce our analysis:
@@
 The overarching goal of this paper was to provide a complete and easy-to-use workflow that allows confidence in reproducibility of R-based data analyses.
-Analysis following our workflow can be reproduced with four commands (here shown for our paper):
+Analyses following our workflow can be reproduced with four commands (here shown for our paper):
 
@@
 
-That is, it reproduces a scientific report exactly without regard to the user, the operating system, the software, the timepoint or interim changes to the involved files.
-To that end, the proposed workflow relies on tools that have been the foundation of reliable software development for years or even decades.
+That is, this reproduces a scientific report exactly without regard to the user, the operating system, the software, the timepoint, or interim changes to the involved files.
+To that end, the proposed workflow relies on tools that have been the foundation of reliable software development for years or, even decades.
 As a by-product, it makes transparent how statistical results depend on the software that created them and, by virtue of this transparency, facilitates later re-use by other researchers.
 
-Each tool in the workflow reduces the chances of non-reproduction.
-Dynamic reporting with R Markdown guarantuees consistency between computational results and their reporting; Version control with `git` guarantuees permanence and consistency across multiple versions of data and code; Dependency management with `make` guarantuees defined entry-points and dependency resolution; containerization with `docker` guarantuees reproducibility of the full computational environment.
+Each tool in the workflow reduces the chances of non-reproduinility.
+Dynamic reporting with R Markdown guarantuees consistency between computational results and their reporting; Version control with `git` ensures permanence and consistency across multiple versions of data and code; Dependency management with `make` affirms defined entry-points and dependency resolution; containerization with `docker` secures reproducibility of the full computational environment.
 We believe that this tightening of loose ends does not restrict researchers but enables them to operate on a solid basis to deliver sound and sustainable research.
@@
 While our approach was designed to scale well with the complexity of a computing-intense project, we realise that this flexibility may not be straightforward to integrate into a researcher's everyday workflow.
-There are R packages that implement parts of our workflow and, thus, lower the threshold to adoption when the full flexibility is not needed.
+There are R packages that implement parts of our workflow and, thus, lower the threshold for adoption when the full flexibility provided by our workflow is not needed.
 The use of R Markdown within a project, tracked with `git` can be simplified with the [`workflowR`](https://GitHub.com/jdblischak/workflowr) package [@workflowr].
 The [`drake`](https://GitHub.com/ropensci/drake) package [@drake] is directly inspired by `make` but can only handle dependencies within R.
-The [liftR ](https://cran.r-project.org/web/packages/liftr/vignettes/liftr-intro.html) package [@liftr] and the [holepunch](https://karthik.GitHub.io/holepunch/) package [@holepunch] automatize the use of docker.
-The former is perfectly compatible with the described workflow and we recommend it to users uncomfortable with command-line use of docker.
-The later uses [binder](https://mybinder.org) [@binder2018] to move the analysis to the cloud, so that no local installation of docker is required.
-`holepunch` is well suited for simple analysis with low computational demands, because of memory and computing time limitations of binder.
+The [liftR ](https://cran.r-project.org/web/packages/liftr/vignettes/liftr-intro.html) package [@liftr] and the [`holepunch`](https://karthik.GitHub.io/holepunch/) package [@holepunch] automatize the use of docker.
+The former is perfectly compatible with the described workflow and we recommend it to users  who are not comfortable with command-line use of docker.
+`holepunch` uses [binder](https://mybinder.org) [@binder2018] to move the analysis to the cloud, so that no local installation of docker is required.
+`holepunch` is well suited for simple analysis with low computational demands, because binder's  memory and computing time is limited.
 There are several alternatives to docker that manage dependencies on R packages.
-[`packrat`](https://rstudio.GitHub.io/packrat/) [@packrat] is a way to freeze package version via local copies of packages in the project, however it does not garuantee a given base R version or system dependencies.
+[`packrat`](https://rstudio.GitHub.io/packrat/) [@packrat] is a way to freeze package version via local copies of packages in the project, but it does not gurantee a given base R version or system dependencies.
 Similar approaches are taken by [`jetpack`](https://GitHub.com/ankane/jetpack) [@jetpack], [`miniCRAN`](https://cran.r-project.org/web/packages/miniCRAN/vignettes/miniCRAN-introduction.html) [@miniCRAN] and [`checkpoint`](https://cran.r-project.org/web/packages/checkpoint/vignettes/checkpoint.html) [@checkpoint].
-Notable is also the package [`reprex`](https://GitHub.com/tidyverse/reprex) [**repr**oducible **ex**ample, @reprex], but its scope is limited.
+The package [`reprex`](https://GitHub.com/tidyverse/reprex) [**repr**oducible **ex**ample, @reprex] is also worth noting, but its scope is limited.
 
-Another way to increase the chance of reproducibility is to develop an R package, which is a self-contained set of files with well-defined meta data, for each analysis.
-The reasoning is that, by abiding the strict rules of package development and specifying all dependencies, a high degree of stability can be achieved (e.g., @marwickPackagingDataAnalytical2018).
+Another way to increase the chance of reproducibility is to develop an R package, which is a self-contained set of files with well-defined meta-data, for each analysis.
+The reasoning is that, abiding by the strict rules of package development and specifying all dependencies can archieve a high degree of stability can be achieved [e.g., @marwickPackagingDataAnalytical2018].
 
-Other than these tools that ease the process of creating workflows like ours, we noticed increased interest in changing how research is published and used [@perkelToolkitDataTransparency2018], with *life code* [@perkelPioneeringLivecodeArticle2019] and *continuous integration* [@beaulieu-jonesReproducibilityComputationalWorkflows2017; @yenniDevelopingModernData2018].
-These techniques give us a glimpse of a paradigm shift from static to dynamic, interactive and living publications that is yet to happen.
+Other than these tools, which ease the process of creating workflows like ours does, we have noticed increased interest in changing the way research is published and used [@perkelToolkitDataTransparency2018], with the emergence of *life code* [@perkelPioneeringLivecodeArticle2019] and *continuous integration* [@beaulieu-jonesReproducibilityComputationalWorkflows2017; @yenniDevelopingModernData2018].
+These techniques give us a glimpse of a paradigm shift from static to dynamic, interactive, and living publications that is yet to happen.
 
@@
 We are aware that implementing the proposed workflow is not straightforward and the difficulty of its implementation may vary by platform.
-For example, already the installation of all tools is easier on POSIX-compatible platforms such as Unix, Linux, or macOS (but not Windows).
+For example, already the installation of all tools is already easier on POSIX-compatible platforms such as Unix, Linux, or macOS (but not Windows).
 However, once a reproducible workflow is established as a default, it can be used with minimal changes from project to project.
@@
 In our own experience, it is often not possible to convince all co-authors to switch to a different document processing environment, such as R Markdown.
-That is, we have experienced that after writing up the first draft in R Markdown, we eventually had to genereate a Word-file that, from then on, was used as static file serving as a basis for multiple iterations among the co-authors.
-Retaining reproduciblity in such situation then requires tedious manual synchronization of files across formats.
+That is, we have experienced the case that after writing up the first draft in R Markdown, we eventually had to generate a Word file that, from then on, was used as static file serving as a basis for multiple iterations among the co-authors.
+Retaining reproduciblity in such situations then requires tedious manual synchronization of files across formats.
 This annoyance may be reduced with the [`rodoc` package](https://noamross.github.io/redoc/articles/mixed-workflows-with-redoc.html) [@redoc], which enables a bidirectional synchronization between Word and R Markdown.
@@
 
-How to best share a reproducible workflow?
-We believe that, ideally, a non-commercial public service provider should be found that guarantuees permanent and reliable hosting of our reproducible workflows, such as the Open Science Framework [@fosterOpenScienceFramework2017].
-A provider mirroring and complementing the services offered by GitHub, Docker Hub, and MRAN.
+How can one best share a reproducible workflow?
+We believe that, ideally, a non-commercial public service provider should be found that guarantees permanent and reliable hosting of our reproducible workflows, such as the Open Science Framework [@fosterOpenScienceFramework2017].
+A provider mirroring and complementing the services offered by GitHub, Docker Hub, and MRAN would be desirable.
 Second, to ensure that other users are legally able to benefit from the shared materials, authors must choose an appropriate license format.
 Typically, there is currently no single license that works for code, data, and media (such as text or figures).
-We encourage authors to choose appropriate license forms that do not hinder others from freely downloading, using, and modifying the shared workflows and materials while, at the same time, ensuring recognition for the time and effort invested in creating the workflow in the first place.
-In our experience, the Creative Commons license (CC-BY) is often appropriate for sharing texts, R Markdown files, generated figures, and other media whereas scripts and any other computer code is often best shared according to the MIT license.
+We encourage authors to choose appropriate license forms that do not hinder others from freely downloading, using, and modifying the shared workflows and materials while, at the same time, ensuring recognition for the time and effort invested into creating the workflow in the first place.
+In our experience, the Creative Commons license (CC-BY) is often appropriate for sharing texts, R Markdown files, generated figures, and other media, whereas scripts and any other computer code are often best shared according to the MIT license.
 Both licenses assure maximal freedom for future users while requiring the attribution of the original authors in derivative work.
 These licenses are also in line with the recommendations by the Reproducible Research Standard [@stoddenEnhancingReproducibilityComputational2016; @stoddenEnablingReproducibleResearch2009].
-To facilitate an inclusive environment we recommend to name all contributors and include a [Code of Conduct](https://opensource.guide/code-of-conduct/) in your project.
+To facilitate an inclusive environment we recommend naming all contributors and including a [Code of Conduct](https://opensource.guide/code-of-conduct/) in your project.
 
@@
 
-The proposed workflow leverages various existing tools that are partly integrated into Rstudio already.
+The proposed workflow leverages various existing tools that are partly integrated into RStudio already.
 Parts of the proposed workflow have been integrated into stand-alone packages (such as workflowr @[workflowr] or holepunch  @[holepunch]) but they are either incomplete in the sense of our proposal or rely heavily on propietary service providers.

### 2019-11-11 manuscript.Rmd
check spelling of software close #60

@@
 To this end, it is recommendable that all files are within one folder (or enclosed subfolders within it) and all paths are relative to that folder.
-A robust solution to the problem of making sure that file access does not break across computing platforms are [Rstudio projects](https://r4ds.had.co.nz/workflow-projects.html) and the `here` package [@here2017].
+A robust solution to the problem of making sure that file access does not break across computing platforms are [RStudio projects](https://r4ds.had.co.nz/workflow-projects.html) and the `here` package [@here2017].
 
@@
 
-A typical git workflow in the terminal looks like this:
+A typical Git workflow in the terminal looks like this:
 
@@
 To keep track of all changes on your local computer, you only need to use `git add` and `git commit` or `git commit -a` to add and commit at the same time.
-These commands need to be executed in the terminal, which you can access from within Rstudio (`Shift + Alt + R`).
-Rstudio also offers a graphical user interface for Git and, quite likely, this interface is sufficient for almost every interaction with Git.
+These commands need to be executed in the terminal, which you can access from within RStudio (`Shift + Alt + R`).
+RStudio also offers a graphical user interface for Git and, quite likely, this interface is sufficient for almost every interaction with Git.
 
@@
 Another benefit of using Git and GitHub is that experimentation is highly encouraged since you can go back to any state quickly and even when you lose access to the file on your computer, everything is backed up on the GitHub servers.
-Further, one can reduce the liklihood of dead code accumulating (e.g., lines that have been commented out) because it is safe to simply remove unneeded code blocks and track the removal in git.
+Further, one can reduce the liklihood of dead code accumulating (e.g., lines that have been commented out) because it is safe to simply remove unneeded code blocks and track the removal in Git.
 
@@
 To create `data/clean.csv`, `R/clean.R` and `data/iris.csv` are needed.
-If you type `make analysis.pdf`, `make` first checks whether the dependencies do exist and, if not, creates them.
-Here, if `data/clean.csv` does not exist, `make` creates it.
+If you type `make analysis.pdf`, Make first checks whether the dependencies do exist and, if not, creates them.
+Here, if `data/clean.csv` does not exist, Make creates it.
 Also, if one of the dependencies of a target is newer then the target itself, then Make updates everything that directly or indirectly depends on the target.
 Here, if `data\iris.csv` is newer then `data\clean.csv`, Make will attempt to recreate `data\clean.csv` first and then `analysis.pdf`.
-If there is a dependency which is missing, and there is no recipe to make it, `make` stops with an error message.
+If there is a dependency which is missing, and there is no recipe to make it, Make stops with an error message.
 It is a convention to have the first target named `all`, which depends on everything.
@@
 A running instance of an image is called a container.
-An image can be transferred and executed on any machine that has docker installed.
+An image can be transferred and executed on any machine that has Docker installed.
 Irrespective of the machine that is executing the container, the computational environment is the same for the programs running inside the container.
@@
 Quite to the contrary, with solutions like Docker, it gets easier than ever to safely update to new versions just by changing the R version number of the Docker image.
-This convenience is possible because of the efforts of the [`rocker` project](https://GitHub.com/rocker-org/rocker) [@boettigerIntroductionRockerDocker2017a], which provides Docker images preconfigured with an R-installation of given R versions.
+This convenience is possible because of the efforts of the [Rocker project](https://GitHub.com/rocker-org/rocker) [@boettigerIntroductionRockerDocker2017a], which provides Docker images preconfigured with an R-installation of given R versions.
 These packages are taken from MRAN [@revolutionanalyticsReproducibilityUsingFixed2019], a repository for R packages fixed to the last date on which the R version of the image was the most recent.
-Building upon these rocker images, researchers can easily build their own Docker images with all required R packages.
+Building upon these Rocker images, researchers can easily build their own Docker images with all required R packages.
 The rocker project also provides images that include RStudio (`rocker/rstudio`), the `tidyverse` package (`rocker/tidyverse`) and the R Markdown package with LaTeX (`rocker/verse`).
 Because our workflow relies on R Markdown, we suggest using the `rocker/verse` image (which also contains `rstudio` & `tidyverse`).
-These images are stored on [dockerhub](https://hub.docker.com) [@dockerinc.DockerHub2019].
+These images are stored on [Dockerhub](https://hub.docker.com) [@dockerinc.DockerHub2019].
 
-Building on a basic rocker image, we can specify further requirements for software in a Dockerfile. For example, the basis for this article's docker image is the following Dockerfile:
+Building on a basic Rocker image, we can specify further requirements for software in a Dockerfile. For example, the basis for this article's Docker image is the following Dockerfile:
 
@@
 
-The `FROM` statement specifies which docker image to use, in this case, the `rocker/verse` image with the tag 3.6.1 (referring to the R version 3.6.1).
-The `RUN` statement describes a command to execute, in this case, to run an R script `install2.r` which is available on all rocker images, to install the specified packages (here, `pacman`,  `here` & `pander`).
+The `FROM` statement specifies which Docker image to use, in this case, the `rocker/verse` image with the tag 3.6.1 (referring to the R version 3.6.1).
+The `RUN` statement describes a command to execute, in this case, to run an R script `install2.r` which is available on all Rocker images, to install the specified packages (here, `pacman`,  `here` & `pander`).
 A Dockerfile allows more than one `RUN` statement, executing arbitrary system commands.
@@
 
-There are two principled ways to share a docker image; either by sharing the Dockerfile that creates the image, or by sharing the image itself, e.g., through a service like [dockerhub](https://hub.docker.com).
+There are two principled ways to share a Docker image; either by sharing the Dockerfile that creates the image, or by sharing the image itself, e.g., through a service like [Dockerhub](https://hub.docker.com).
 While both ways guarantee a replicable computational environment, sharing the Dockerfile is more transparent; in our workflow, we can use Git to track changes in the Dockerfile (such as updates to dependencies).
@@
 
-The `--rm` flag means that the state of the docker after the command finishes in an container is not saved.
+The `--rm` flag means that the state of the container after the command finishes is not saved.
 The `-it` flag tells Docker to accept inputs and return outputs to the terminal.
@@
 
-![R terminal running inside docker](Images/docker-r-terminal.png)
+![R terminal running inside Docker](Images/docker-r-terminal.png)
 
 The second option is to start the container in the background (not supplying a specific command) and to interact with the container via the webrowser and the RStudio server instance running in it.
-In order to do so, you need to supply a password to log into the Rstudio server (`-e PASSWORD=<YOUR_PASS>`) and open a port (`-p 8787:8787`).
+In order to do so, you need to supply a password to log into the RStudio server (`-e PASSWORD=<YOUR_PASS>`) and open a port (`-p 8787:8787`).
 
@@
 The adress to connect to the RStudio server is your IP address (or `localhost` Linux) in this scheme: `<IPADRESS>:8787`.
-This offers a fully functioning Rstudio instance that runs in the image but is accessible through a local web browser.
+This offers a fully functioning RStudio instance that runs in the image but is accessible through a local web browser.
 
-By default, programs inside the container cannot access files on the local computer, thus requiring an explicit link to a local folder to enable access (and on Mac and Windows this also have to be allowed in the settings) :
+By default, programs inside the container cannot access files on the local computer, thus requiring an explicit link to a local folder to enable access (and on macOS and Windows this also have to be allowed in the settings) :
 
@@
 
-The main directory for RStudio inside docker is `/home/rstudio`, so the final call to start Rstudio inside a docker image may look like this:
+The main directory for RStudio inside the container is `/home/rstudio`, so the final call to start RStudio inside a Docker container may look like this:
 
@@
 Since Docker commands tend to grow long and become tedious to type manually, we recomend using some automatic way to generate them.
-Fortunately, one can use `make` to automatically generate the `docker` commands, e.g.
+Fortunately, one can use Make to automatically generate the `docker` commands, e.g.
 the (simplified) `Makefile` for this paper allows command after `$(run)` to be conditionally passed through Docker if one types `make DOCKER=TRUE` (otherwise, they are run locally):
@@
 
-Other than on R, RStudio, and R Markdown, our workflow relies on three pieces of software from outside the R environment: `git`, `make`, and `docker`.
+Other than on R, RStudio, and R Markdown, our workflow relies on three pieces of software from outside the R environment: Git, Make, and Docker.
 The installation process of these software packages depends on the operating system, e.g., on macOS, Make is always available, while on Linux, Git and Make are often already installed.
@@
 Note that you have to have Windows Pro, Enterprise, Education or Server installed, as Microsoft prevents the use of Docker on Windows Home.
-There is a package manager for Windows called chocolately, which you can install from: [https://chocolatey.org/install](https://chocolatey.org/install), providing all dependencies.
-Having installed chocolately (and restarted the computer), all dependencies can be installed in an **admin terminal** (Windows key, then type `cmd`, right-click *Run as administrator*) via:
+There is a package manager for Windows called Chocolatey, which you can install from: [https://chocolatey.org/install](https://chocolatey.org/install), providing all dependencies.
+Having installed Chocolatey (and restarted the computer), all dependencies can be installed in an **admin terminal** (Windows key, then type `cmd`, right-click *Run as administrator*) via:
 
@@
 To use `docker` you need to start Docker Desktop.
-In the settings of docker you have to allow the sharing of your drive.
+In the settings of Docker Desktop you have to allow the sharing of your drive.
 Docker on Windows require the path to be special; therefore you need to hand edit the `Makefile` and set `current_path` to the current directory and use `make all DOCKER=TRUE WINDOWS=TRUE`.
@@
 
-As make already ships with macOS, you only need `git` and `docker-desktop`.
-We suggest using the package manager `homebrew`, which you can install from [https://docs.brew.sh/Installation](https://docs.brew.sh/Installation), to install `docker` (git will be installed during the installation of `homebrew`):
+As Make already ships with macOS, you only need Git and Docker.
+We suggest using the package manager Homebrew, which you can install from [https://docs.brew.sh/Installation](https://docs.brew.sh/Installation), to install Docker (Git will be installed during the installation of Homebrew):
 
@@
 There is a host of different Linux distributions and almost as many package managers.
-Still, to our knowledge, there is no (recent) Linux edition, that does not include `git`, `make` and `docker`.
+Still, to our knowledge, there is no (recent) Linux edition, that does not include Git, Make and Docker.
 For example, in Ubuntu Linux, installation is straightforward using the shipped package manager:
@@
 You may need elevated rights for the installation; in this case add `sudo` before the installation command.
-`docker` also needs elevated rights to run; therefore, we recommend adding the local user to the `docker` group, following the [documentation of docker](https://docs.docker.com/install/linux/linux-postinstall/#manage-docker-as-a-non-root-user).
+`docker` also needs elevated rights to run; therefore, we recommend adding the local user to the `docker` group, following the [documentation of Docker](https://docs.docker.com/install/linux/linux-postinstall/#manage-docker-as-a-non-root-user).
 
-An alternative to `docker` on Linux is `singularity` [@kurtzerSingularityScientificContainers2017].
+An alternative to Docker on Linux is Singularity [@kurtzerSingularityScientificContainers2017].
 Just replace any docker calls with `singularity docker`, because singularity fully supports docker images.
-A possible advantage is that singularity works well in high-performance computing environments and on old Linux versions, the downside is that `singularity` is only available on Linux.
+A possible advantage is that singularity works well in high-performance computing environments and on old Linux versions, the downside is that Singularity is only available on Linux.
 
@@
 Each tool in the workflow reduces the chances of non-reproduinility.
-Dynamic reporting with R Markdown guarantuees consistency between computational results and their reporting; Version control with `git` ensures permanence and consistency across multiple versions of data and code; Dependency management with `make` affirms defined entry-points and dependency resolution; containerization with `docker` secures reproducibility of the full computational environment.
+Dynamic reporting with R Markdown guarantuees consistency between computational results and their reporting; Version control with Git ensures permanence and consistency across multiple versions of data and code; Dependency management with Make affirms defined entry-points and dependency resolution; containerization with Docker secures reproducibility of the full computational environment.
 We believe that this tightening of loose ends does not restrict researchers but enables them to operate on a solid basis to deliver sound and sustainable research.
@@
 There are R packages that implement parts of our workflow and, thus, lower the threshold for adoption when the full flexibility provided by our workflow is not needed.
-The use of R Markdown within a project, tracked with `git` can be simplified with the [`workflowR`](https://GitHub.com/jdblischak/workflowr) package [@workflowr].
-The [`drake`](https://GitHub.com/ropensci/drake) package [@drake] is directly inspired by `make` but can only handle dependencies within R.
-The [liftR ](https://cran.r-project.org/web/packages/liftr/vignettes/liftr-intro.html) package [@liftr] and the [`holepunch`](https://karthik.GitHub.io/holepunch/) package [@holepunch] automatize the use of docker.
-The former is perfectly compatible with the described workflow and we recommend it to users  who are not comfortable with command-line use of docker.
-`holepunch` uses [binder](https://mybinder.org) [@binder2018] to move the analysis to the cloud, so that no local installation of docker is required.
+The use of R Markdown within a project, tracked with Git can be simplified with the [`workflowR`](https://github.com/jdblischak/workflowr) package [@workflowr].
+The [`drake`](https://github.com/ropensci/drake) package [@drake] is directly inspired by Make but can only handle dependencies within R.
+The [liftR ](https://cran.r-project.org/web/packages/liftr/vignettes/liftr-intro.html) package [@liftr] and the [`holepunch`](https://karthik.github.io/holepunch/) package [@holepunch] automatize the use of Docker.
+The former is perfectly compatible with the described workflow and we recommend it to users  who are not comfortable with command-line use of Docker.
+`holepunch` uses [binder](https://mybinder.org) [@binder2018] to move the analysis to the cloud, so that no local installation of Docker is required.
 `holepunch` is well suited for simple analysis with low computational demands, because binder's  memory and computing time is limited.
-There are several alternatives to docker that manage dependencies on R packages.
-[`packrat`](https://rstudio.GitHub.io/packrat/) [@packrat] is a way to freeze package version via local copies of packages in the project, but it does not gurantee a given base R version or system dependencies.
-Similar approaches are taken by [`jetpack`](https://GitHub.com/ankane/jetpack) [@jetpack], [`miniCRAN`](https://cran.r-project.org/web/packages/miniCRAN/vignettes/miniCRAN-introduction.html) [@miniCRAN] and [`checkpoint`](https://cran.r-project.org/web/packages/checkpoint/vignettes/checkpoint.html) [@checkpoint].
-The package [`reprex`](https://GitHub.com/tidyverse/reprex) [**repr**oducible **ex**ample, @reprex] is also worth noting, but its scope is limited.
+There are several alternatives to Docker that manage dependencies on R packages.
+[`packrat`](https://rstudio.github.io/packrat/) [@packrat] is a way to freeze package version via local copies of packages in the project, but it does not gurantee a given base R version or system dependencies.
+Similar approaches are taken by [`jetpack`](https://github.com/ankane/jetpack) [@jetpack], [`miniCRAN`](https://cran.r-project.org/web/packages/miniCRAN/vignettes/miniCRAN-introduction.html) [@miniCRAN] and [`checkpoint`](https://cran.r-project.org/web/packages/checkpoint/vignettes/checkpoint.html) [@checkpoint].
+The package [`reprex`](https://github.com/tidyverse/reprex) [**repr**oducible **ex**ample, @reprex] is also worth noting, but its scope is limited.
 

### 2019-11-11 R/rmarkdown.Rmd
sync image and source of markdown example close #59

--- 2019-09-03 R/rmarkdown.Rmd
@@
 ---
-title: "An Example"
-author: "Aaron Peikert & Andreas Brandmaier"
-date: "9/3/2019"
+title: "A simple R markdown example"
+author: "Aaron Peikert & Andreas M. Brandmaier"
+date: '`r format(Sys.time(), "%B %d, %Y")`'
 output: pdf_document
@@
 
-The data was never before anylized in this way. Except in `help(t.test)`.
+This is an example of students' sleep data taken from `help(t.test)`.
 
@@
 I can now assert that what I *believe* to be true
---- that there is a difference in mean between the groups ---
-is `r ifelse(result$p.value > .025, "**not**", "")` supported by data.
+--- that there is a difference in means between the groups ---
+is `r ifelse(result$p.value > .025, "**not**", "")` the supported by the data.

### 2019-11-12 R/rmarkdown.Rmd
fixed some typos (thanks to Julia Rohrer)

@@
 --- that there is a difference in means between the groups ---
-is `r ifelse(result$p.value > .025, "**not**", "")` the supported by the data.
+is `r ifelse(result$p.value > .025, "**not**", "")` supported by the data.

### 2019-11-12 manuscript.Rmd
fixed some typos (thanks to Julia Rohrer)

@@
 We argue that only a workflow using all four concepts in unison can guarantee confidence in reproducing a scientific report.
-Various implementations of these concepts exis6, but we consider the following four best suited for the R environment [@r2019]: for literate programming, R Markdown [@rmarkdown2019], for version control, Git [@chaconProGit2014], for dependency management, Make [@feldmanMakeProgramMaintaining1979], and for containerization, Docker [@merkelDockerLightweightLinux2014].
+Various implementations of these concepts exist, but we consider the following four best suited for the R environment [@r2019]: for literate programming, R Markdown [@rmarkdown2019], for version control, Git [@chaconProGit2014], for dependency management, Make [@feldmanMakeProgramMaintaining1979], and for containerization, Docker [@merkelDockerLightweightLinux2014].
 Each of these software solutions serves a valuable meta-scientific goal (reproducibility) and increases the researches' productivity.
@@
 The folder where all the files reside that you need for analysis (code and data), is referred to as a 'project' (or sometimes as a 'research compendium').
-Working with projects is particularly convinient with RStudio, an integrated development environment (IDE) for R.
+Working with projects is particularly convenient with RStudio, an integrated development environment (IDE) for R.
 It is useful to organize a data analysis project in a way that strictly segregates (raw) data and code by placing them in directories called `data` and `R` [see Section 4 in @marwickPackagingDataAnalytical2018]; there are also tools that automatize the standardized creation of folder structures such as `workflowr` [@workflowr].
@@
 
-Each tool in the workflow reduces the chances of non-reproduinility.
+Each tool in the workflow reduces the chances of non-reproducibility.
 Dynamic reporting with R Markdown guarantuees consistency between computational results and their reporting; Version control with Git ensures permanence and consistency across multiple versions of data and code; Dependency management with Make affirms defined entry-points and dependency resolution; containerization with Docker secures reproducibility of the full computational environment.
@@
 The proposed workflow leverages various existing tools that are partly integrated into RStudio already.
-Parts of the proposed workflow have been integrated into stand-alone packages (such as workflowr @[workflowr] or holepunch  @[holepunch]) but they are either incomplete in the sense of our proposal or rely heavily on propietary service providers.
+Parts of the proposed workflow have been integrated into stand-alone packages (such as workflowr [@workflowr] or holepunch  [@holepunch]) but they are either incomplete in the sense of our proposal or rely heavily on propietary service providers.
 We hope that with increasing awareness of the challenges of computational reproducibility, the increased demand for unified and open solutions will lead to an even better integration of existing tools to allow reproducible workflows to become a standard in psychological research.

### 2019-11-12 abstract.Rmd
added missing #2

@@
 It combines the benefits of various open-source software tools including R Markdown, Git, Make, and Docker, whose interplay ensures seamless integration of version management, dynamic report generation conforming to various journal styles and full cross-platform and long-term computational reproducibility.
-The workflow ensures meeting the primary goals that 1) the reporting of statistical results is consistent with the actual statistical results (dynamic report generation), the analysis exactly reproduces at a later time even if the computing platform or software is changed (computational reproducibility), and 3) changes at any time (during development and post-publication) are tracked, tagged, and documented while earlier versions of both data and code remain accessible.
+The workflow ensures meeting the primary goals that 1) the reporting of statistical results is consistent with the actual statistical results (dynamic report generation), 2) the analysis exactly reproduces at a later time even if the computing platform or software is changed (computational reproducibility), and 3) changes at any time (during development and post-publication) are tracked, tagged, and documented while earlier versions of both data and code remain accessible.
 While the research community increasingly recognizes dynamic document generation and version management as tools to ensure reproducibility, we demonstrate with practical examples that these alone are not sufficient to ensure long-term computational reproducibility.

### 2019-11-12 manuscript.Rmd
Update manuscript.Rmd
a few typos

@@
 
-Sometimes external requirements make it impossible fot the data to be stored and shared together with the scripts.
+Sometimes external requirements make it impossible for the data to be stored and shared together with the scripts.
 In most of the cases we have seen, these are either space constraints or privacy considerations.
@@
 
-Fundamentally, reproducibility means that results remain identical if neither the script nor th data have changed.
+Fundamentally, reproducibility means that results remain identical if neither the script nor the data have changed.
 It is often not trivial to find out whether anything in a project has changed over time and if so, to "go back in time." The Git program enables you to do both.
@@
 Each tool in the workflow reduces the chances of non-reproducibility.
-Dynamic reporting with R Markdown guarantuees consistency between computational results and their reporting; Version control with Git ensures permanence and consistency across multiple versions of data and code; Dependency management with Make affirms defined entry-points and dependency resolution; containerization with Docker secures reproducibility of the full computational environment.
+Dynamic reporting with R Markdown guarantees consistency between computational results and their reporting; Version control with Git ensures permanence and consistency across multiple versions of data and code; Dependency management with Make affirms defined entry-points and dependency resolution; containerization with Docker secures reproducibility of the full computational environment.
 We believe that this tightening of loose ends does not restrict researchers but enables them to operate on a solid basis to deliver sound and sustainable research.

### 2019-11-12 manuscript.Rmd
typo fixes manuscript
fix three typos including wrong package name rodoc > redoc

@@
 Retaining reproduciblity in such situations then requires tedious manual synchronization of files across formats.
-This annoyance may be reduced with the [`rodoc` package](https://noamross.github.io/redoc/articles/mixed-workflows-with-redoc.html) [@redoc], which enables a bidirectional synchronization between Word and R Markdown.
+This annoyance may be reduced with the [`redoc` package](https://noamross.github.io/redoc/articles/mixed-workflows-with-redoc.html) [@redoc], which enables a bidirectional synchronization between Word and R Markdown.
 Conversions between R Markdown and Word retain all changes and even support Word's track-changes feature.

### 2019-11-12 manuscript.Rmd
typo fixes manuscript
fix three typos including wrong package name rodoc > redoc

@@
 Each tool in the workflow reduces the chances of non-reproducibility.
-Dynamic reporting with R Markdown guarantees consistency between computational results and their reporting; Version control with Git ensures permanence and consistency across multiple versions of data and code; Dependency management with Make affirms defined entry-points and dependency resolution; containerization with Docker secures reproducibility of the full computational environment.
+Dynamic reporting with R Markdown guarantuees consistency between computational results and their reporting; Version control with Git ensures permanence and consistency across multiple versions of data and code; Dependency management with Make affirms defined entry-points and dependency resolution; containerization with Docker secures reproducibility of the full computational environment.
 We believe that this tightening of loose ends does not restrict researchers but enables them to operate on a solid basis to deliver sound and sustainable research.

### 2019-12-02 manuscript.Rmd
change sha - thanks to Felix Schönbrodt for spotting

--- 2019-11-12 manuscript.Rmd
@@
 checksum <- digest::sha1(x)
-if (checksum != "740e88404c04af1d67ffc87b008231c093539cc1"){
-  error("Wrong data file loaded or data has been changed!")
+if (checksum != "bd4b9197ff7f8e32e56b9d62c3ee7718f367b485"){
+  warning("Wrong data file loaded or data has been changed!")
 }

### 2019-12-12 manuscript.Rmd
adress #3

@@
 
-There are two principled ways to share a Docker image; either by sharing the Dockerfile that creates the image, or by sharing the image itself, e.g., through a service like [Dockerhub](https://hub.docker.com).
+There are two ways to share a Docker image; either by sharing the Dockerfile that creates the image, or by sharing the image itself, e.g., through a service like [Dockerhub](https://hub.docker.com).
 While both ways guarantee a replicable computational environment, sharing the Dockerfile is more transparent; in our workflow, we can use Git to track changes in the Dockerfile (such as updates to dependencies).

### 2019-12-12 manuscript.Rmd
name correct directory
adjust hashing #4

@@
 x <- data.frame(VAR1=c(1,2,3,4),VAR2=c(0,4,6,9) )
-checksum <- digest::sha1(x)
-if (checksum != "bd4b9197ff7f8e32e56b9d62c3ee7718f367b485"){
-  warning("Wrong data file loaded or data has been changed!")
+checksum <- digest::digest(x, "sha1")
+if (checksum != "8e222b9b3616595dc3f5946f9eac101fbf1336dc"){
+  warning("Mismatch between original and current data file!")
 }
@@
 git clone https://github.com/aaronpeikert/workflow-showcase.git
-cd reproducible-research
+cd workflow-showcase
 make build

### 2020-02-07 manuscript.Rmd
correct mistake regarding drake
closes #7

@@
 The use of R Markdown within a project, tracked with Git can be simplified with the [`workflowR`](https://github.com/jdblischak/workflowr) package [@workflowr].
-The [`drake`](https://github.com/ropensci/drake) package [@drake] is directly inspired by Make but can only handle dependencies within R.
+The [`drake`](https://github.com/ropensci/drake) package [@drake] is directly inspired by Make and takes an R centric approach, making it especially suited for projects only involving R.
 The [liftR ](https://cran.r-project.org/web/packages/liftr/vignettes/liftr-intro.html) package [@liftr] and the [`holepunch`](https://karthik.github.io/holepunch/) package [@holepunch] automatize the use of Docker.

### 2020-02-13 manuscript.Rmd
accept @brandmaier's suggestion

--- 2019-12-12 manuscript.Rmd
@@
 The use of R Markdown within a project, tracked with Git can be simplified with the [`workflowR`](https://github.com/jdblischak/workflowr) package [@workflowr].
-The [`drake`](https://github.com/ropensci/drake) package [@drake] is directly inspired by Make but can only handle dependencies within R.
+The [`drake`](https://github.com/ropensci/drake) package [@drake] is directly inspired by Make and takes an R centric approach, making it especially suited for projects only involving R, but can also handle external dependencies.
 The [liftR ](https://cran.r-project.org/web/packages/liftr/vignettes/liftr-intro.html) package [@liftr] and the [`holepunch`](https://karthik.github.io/holepunch/) package [@holepunch] automatize the use of Docker.

### 2020-03-13 manuscript.Rmd
describe GitHub releases closes #14

--- 2019-12-12 manuscript.Rmd
@@
 Going back to one state is as easy as traversing the history of all commits and switching to a given one; it is possible to visually track changes between different versions.
-Finding a specific version can be made even easier by tagging snapshots, e.g., as "submission," "preprint," or "published." This collection of snapshots is called a "repository," which ideally tracks your entire R project.
+This collection of snapshots is called a "repository," which ideally tracks your entire R project.
 
@@
 
+GitHub allows you to archive a specific version of your repository in the form of a release.
+A release tags a commit, e.g., as "submission,"[^sub] "preprint," or "published," and archives also "binary" products of your code, e.g.
+the resulting pdf of the manuscript or the docker image (see [Containerization]).
+From such a release a DOI can be created, making it easier to reference it (see this [GitHub Guide](https://guides.github.com/activities/citable-code/)[^doi]).
+
 [^1]: To comment on our paper, please add an issue to the GitHub repository of our paper: <https://github.com/aaronpeikert/reproducible-research/issues>.
+[^doi]: [https://guides.github.com/activities/citable-code/](https://guides.github.com/activities/citable-code/)
+[^sub]: We created an release for the submission: [https://github.com/aaronpeikert/reproducible-research/releases/tag/v0.1.1-submission](https://github.com/aaronpeikert/reproducible-research/releases/tag/v0.1.1-submission)
 

### 2020-03-13 manuscript.Rmd
advisee to archieve docker image closes #11

@@
 While both ways guarantee a replicable computational environment, sharing the Dockerfile is more transparent; in our workflow, we can use Git to track changes in the Dockerfile (such as updates to dependencies).
+A possible downside is that in order to create an image from a Dockerfile, all software repositories need to be still availible.
+Hence, to garuntee reproducibility for decades to come, it is best to archieve the binary image at major points of the projets progress e.g.
+publication with a release on GitHub (see [Version Management] for details).
 

### 2020-03-15 manuscript.Rmd
mention GitLab closes #13

--- 2020-02-13 manuscript.Rmd
@@
 A popular service for sharing materials via Git is [GitHub](https://github.com).
-Just sharing Git repositories with the public is always free, private repositories (only visible to persons you invite) are [free for researchers](https://help.github.com/en/articles/applying-for-an-educator-or-researcher-discount).
+Alternatively, institutions can host an equally feature-rich open-source software called [GitLab](https://gitlab.com) themself, avoiding the reliance on commercial service providers.
+Just sharing Git repositories on GitHub with the public is always free, private repositories (only visible to persons you invite) are [free for researchers](https://help.github.com/en/articles/applying-for-an-educator-or-researcher-discount).
 After creating a user account, one can create a new repository and GitHub provides information on how to upload your repository from the terminal, e.g., for our repository (here with user name 'aaronpeikert' and repository name 'reproducible-research'):
@@
 The use of R Markdown within a project, tracked with Git can be simplified with the [`workflowR`](https://github.com/jdblischak/workflowr) package [@workflowr].
-The [`drake`](https://github.com/ropensci/drake) package [@drake] is directly inspired by Make and takes an R centric approach, making it especially suited for projects only involving R, but can also handle external dependencies.
+The [`drake`](https://github.com/ropensci/drake) package [@drake] is directly inspired by Make but can only handle dependencies within R.
 The [liftR ](https://cran.r-project.org/web/packages/liftr/vignettes/liftr-intro.html) package [@liftr] and the [`holepunch`](https://karthik.github.io/holepunch/) package [@holepunch] automatize the use of Docker.

### 2020-03-15 manuscript.Rmd
replace packrat with renv its sucsessor closes #12

@@
 A popular service for sharing materials via Git is [GitHub](https://github.com).
-Alternatively, institutions can host an equally feature-rich open-source software called [GitLab](https://gitlab.com) themself, avoiding the reliance on commercial service providers.
-Just sharing Git repositories on GitHub with the public is always free, private repositories (only visible to persons you invite) are [free for researchers](https://help.github.com/en/articles/applying-for-an-educator-or-researcher-discount).
+Just sharing Git repositories with the public is always free, private repositories (only visible to persons you invite) are [free for researchers](https://help.github.com/en/articles/applying-for-an-educator-or-researcher-discount).
 After creating a user account, one can create a new repository and GitHub provides information on how to upload your repository from the terminal, e.g., for our repository (here with user name 'aaronpeikert' and repository name 'reproducible-research'):
@@
 There are several alternatives to Docker that manage dependencies on R packages.
-[`packrat`](https://rstudio.github.io/packrat/) [@packrat] is a way to freeze package version via local copies of packages in the project, but it does not gurantee a given base R version or system dependencies.
+[`renv`](https://rstudio.github.io/renv/articles/renv.html) [@renv] is a way to freeze package version via local copies of packages in the project, but it does not gurantee a given base R version or system dependencies.
 Similar approaches are taken by [`jetpack`](https://github.com/ankane/jetpack) [@jetpack], [`miniCRAN`](https://cran.r-project.org/web/packages/miniCRAN/vignettes/miniCRAN-introduction.html) [@miniCRAN] and [`checkpoint`](https://cran.r-project.org/web/packages/checkpoint/vignettes/checkpoint.html) [@checkpoint].

### 2020-03-15 manuscript.Rmd
include hash in manuscript closes #15

@@
 All materials can be found on the accompanying GitHub repository: [https://github.com/aaronpeikert/reproducible-research/](https://github.com/aaronpeikert/reproducible-research/).
+This particular document was created from the commit with the hash `"`r system2("git", "rev-parse --short HEAD", stdout = TRUE)`"`.
 To provide feedback or any other type of comment or questions regarding our workflow, please add an issue to the GitHub repository of our paper at [https://github.com/aaronpeikert/reproducible-research/issues](https://github.com/aaronpeikert/reproducible-research/issues).
@@
 There are several alternatives to Docker that manage dependencies on R packages.
-[`renv`](https://rstudio.github.io/renv/articles/renv.html) [@renv] is a way to freeze package version via local copies of packages in the project, but it does not gurantee a given base R version or system dependencies.
+[`packrat`](https://rstudio.github.io/packrat/) [@packrat] is a way to freeze package version via local copies of packages in the project, but it does not gurantee a given base R version or system dependencies.
 Similar approaches are taken by [`jetpack`](https://github.com/ankane/jetpack) [@jetpack], [`miniCRAN`](https://cran.r-project.org/web/packages/miniCRAN/vignettes/miniCRAN-introduction.html) [@miniCRAN] and [`checkpoint`](https://cran.r-project.org/web/packages/checkpoint/vignettes/checkpoint.html) [@checkpoint].

### 2020-04-09 manuscript.Rmd
added Turing Way citation to Introduction

@@
 All materials can be found on the accompanying GitHub repository: [https://github.com/aaronpeikert/reproducible-research/](https://github.com/aaronpeikert/reproducible-research/).
-This particular document was created from the commit with the hash `"`r system2("git", "rev-parse --short HEAD", stdout = TRUE)`"`.
 To provide feedback or any other type of comment or questions regarding our workflow, please add an issue to the GitHub repository of our paper at [https://github.com/aaronpeikert/reproducible-research/issues](https://github.com/aaronpeikert/reproducible-research/issues).
@@
 
-In this tutorial, we describe a workflow to ensure long-term and cross-platform reproducibility of
-R-based data analyses.
-To us Reproducibility means that statistical results as well as their generating computational workflows are reported fully, and transparently, and remain sustainably available, such that a given workflow can be re-run by a different person later in time and the results will be identical to the ones intially reported [@herouxCompatibleReproducibilityTaxonomy2018; @claerboutElectronicDocumentsGive1992].
+In this tutorial, we describe a workflow to ensure long-term and cross-platform reproducibility of data analyses in R [@baseR].
+Reproducibility is the ability to draw identical conclusion from the same statistical analysis and the same data.
+In practice, reproducibility means to us that statistical results as well as their generating computational workflows are reported fully, and transparently, and remain sustainably available, such that a given workflow can be re-run by a different person, later in time, and that the results will be identical to the ones intially reported [@turingway2020; @herouxCompatibleReproducibilityTaxonomy2018; @claerboutElectronicDocumentsGive1992].
 The need to ensure reproducibility directly follows from commonly accepted rules of good scientific practice [such as the guidelines of the German Research Foundation, @dfg2019].
-Ensuring reproducibility is a prerequisite for replicability and a means to increase trustworthiness of empirical results [@epskamp2019rep].
+Ensuring reproducibility is a prerequisite for replicability (the ability to reach a similar conclusion from the same analysis and new data), and a means to increase trustworthiness of empirical results [@epskamp2019rep].
 
@@
 There are various ideas on how to enhance reproducibility [@piccoloToolsTechniquesComputational2016], four of which we believe to be particularly important: literate programming: [@ruleTenSimpleRules2019], version control [@barbaHardRoadReproducibility2016], dependency management [@askrenUsingMakeReproducible2016], and containerization [@clyburne-sherinComputationalReproducibilityContainers2018].
-We argue that only a workflow using all four concepts in unison can guarantee confidence in reproducing a scientific report.
+We argue that only a workflow using all four concepts in unison can guarantee confidence in reproducing a scientific report [see @turingway2020 for similar arguments].
 Various implementations of these concepts exist, but we consider the following four best suited for the R environment [@r2019]: for literate programming, R Markdown [@rmarkdown2019], for version control, Git [@chaconProGit2014], for dependency management, Make [@feldmanMakeProgramMaintaining1979], and for containerization, Docker [@merkelDockerLightweightLinux2014].

### 2020-04-09 manuscript.Rmd
removed Disclosures (this was previously included for the AMPPS submission)

@@
 This principle is what guided our use of R Markdown, Git, Make, and Docker.-->
-
-# Disclosures
-
-This paper is fully reproducible using the workflow described in this paper.
-All materials, including the manuscript source and all R codes and files used to generate this manuscript, can be found on the accompanying GitHub repository: [https://github.com/aaronpeikert/reproducible-research/](https://github.com/aaronpeikert/reproducible-research/).
 

### 2020-04-09 manuscript.Rmd
reformatted bib file with bibdesk

--- 2020-04-09 manuscript.Rmd
@@
 To this end, it is recommendable that all files are within one folder (or enclosed subfolders within it) and all paths are relative to that folder.
-A robust solution to the problem of making sure that file access does not break across computing platforms are [RStudio projects](https://r4ds.had.co.nz/workflow-projects.html) and the `here` package [@here2017].
+A robust solution to the problem of making sure that file access does not break across computing platforms are [RStudio projects](https://r4ds.had.co.nz/workflow-projects.html) and the `here` package [@here2017] to manage file access.
 

### 2020-04-09 manuscript.Rmd
added new intro paragraph

@@
 
-# Project organization
+# The Reproducible Workflow in a nutshell
 
-The first step towards reproducibility is to create an R script or R Markdown file as primary entry point for the analysis that runs on a local computer without error and performs the main statistical analyses.
-Next, one needs to make sure that all files relevant for the analysis can be moved to another computer.
-To this end, it is recommendable that all files are within one folder (or enclosed subfolders within it) and all paths are relative to that folder.
-A robust solution to the problem of making sure that file access does not break across computing platforms are [RStudio projects](https://r4ds.had.co.nz/workflow-projects.html) and the `here` package [@here2017] to manage file access.
-
-```{r, eval=FALSE}
-# BAD
-iris <- read.csv("/home/aaron/reproducible-research/data/iris.csv")
-# GOOD
-iris <- read.csv("data/iris.csv")
-# BETTER
-iris <- read.csv(here("data", "iris.csv"))
-```
-
-The folder where all the files reside that you need for analysis (code and data), is referred to as a 'project' (or sometimes as a 'research compendium').
-Working with projects is particularly convenient with RStudio, an integrated development environment (IDE) for R.
-It is useful to organize a data analysis project in a way that strictly segregates (raw) data and code by placing them in directories called `data` and `R` [see Section 4 in @marwickPackagingDataAnalytical2018]; there are also tools that automatize the standardized creation of folder structures such as `workflowr` [@workflowr].
-
-Sometimes external requirements make it impossible for the data to be stored and shared together with the scripts.
-In most of the cases we have seen, these are either space constraints or privacy considerations.
-In these cases, of course, unrestricted reproducibility is not guaranteed.
-If dividing data and scripts is unavoidable, we recommend validating all data files using checksums (also called a "hash", e.g., using the functions provided in package `digest` [@digest2019]) before analyzing them.
-To do this, a checksum must be created and stored at the time of the original analysis.
-At the time of reproduction, the current checksum must then be compared with the stored checksum to ensure data integrity.
-
-```{r}
-x <- data.frame(VAR1=c(1,2,3,4),VAR2=c(0,4,6,9) )
-checksum <- digest::digest(x, "sha1")
-if (checksum != "8e222b9b3616595dc3f5946f9eac101fbf1336dc"){
-  warning("Mismatch between original and current data file!")
-}
-```
+![](Images/placeholder.png)
+Figure XXX gives an overview of how the four components of our workflow interface to ensure computational reproducibility.
+We begin with a minimal description of the roles of the four components.
+In the remainder of this tutorial, we will further detail each of the four components of our workflow.
+The first component is version control.
+Version control manages  changes to files (e.g., data and code) over time so that you can access specific versions later.
+Version control offers snapshots of your workflow at different time points identified by a unique identifier.
+The different frames in the z-axis of the image illustrate the different versions of the workflow.
+How different parts of a analysis and a report of it relate and in what order they need to be executed is monitored using dependency management.
+The arrows visualize dependencies, such as an analysis depending on the availability of a particular data file.
+Third, all computer code (such as a statistical analysis in R) is executed in a virtual environment that guarantuees exact reproduction of results independent of the host operating system, R version, and installed package versions.
+Finally, literate programming interweaves human-readable code and automatically produced results (such as p values, effect sizes, confidence intervals) to eliminate inconsistency errors such as those arising from careless copy and paste actions.
 
@@
 
-## Reproducing an Analysis
+# Project organization
 
-We provide a reproducible analysis as a working example via [GitHub](https://github.com/aaronpeikert/workflow-showcase) as a testbed for users.
+Finally, we conclude with some notes on project organization, which we think makes migrating projects to a reproducible workflow easier.
+The first step towards reproducibility is to create an R script or R Markdown file as primary entry point for the analysis that runs on a local computer without error and performs the main statistical analyses.
+Next, one needs to make sure that all files relevant for the analysis can be moved to another computer.
+To this end, it is recommendable that all files are within one folder (or enclosed subfolders within it) and all paths are relative to that folder.
+A robust solution to the problem of making sure that file access does not break across computing platforms are [RStudio projects](https://r4ds.had.co.nz/workflow-projects.html) and the `here` package [@here2017] to manage file access.
+
+```{r, eval=FALSE}
+# BAD
+iris <- read.csv("/home/aaron/reproducible-research/data/iris.csv")
+# GOOD
+iris <- read.csv("data/iris.csv")
+# BETTER
+iris <- read.csv(here("data", "iris.csv"))
+```
+
+The folder where all the files reside that you need for analysis (code and data), is referred to as a 'project' (or sometimes as a 'research compendium').
+Working with projects is particularly convenient with RStudio, an integrated development environment (IDE) for R.
+It is useful to organize a data analysis project in a way that strictly segregates (raw) data and code by placing them in directories called `data` and `R` [see Section 4 in @marwickPackagingDataAnalytical2018]; there are also tools that automatize the standardized creation of folder structures such as `workflowr` [@workflowr].
+
+Sometimes external requirements make it impossible for the data to be stored and shared together with the scripts.
+In most of the cases we have seen, these are either space constraints or privacy considerations.
+In these cases, of course, unrestricted reproducibility is not guaranteed.
+If dividing data and scripts is unavoidable, we recommend validating all data files using checksums (also called a "hash", e.g., using the functions provided in package `digest` [@digest2019]) before analyzing them.
+To do this, a checksum must be created and stored at the time of the original analysis.
+At the time of reproduction, the current checksum must then be compared with the stored checksum to ensure data integrity.
+
+```{r}
+x <- data.frame(VAR1=c(1,2,3,4),VAR2=c(0,4,6,9) )
+checksum <- digest::digest(x, "sha1")
+if (checksum != "8e222b9b3616595dc3f5946f9eac101fbf1336dc"){
+  warning("Mismatch between original and current data file!")
+}
+```
+
+## Use Case: Reproducing an Analysis
+
+We provide a reproducible analysis as a working example via [GitHub](https://github.com/aaronpeikert/workflow-showcase).
+We encourage interested readers to try to reproduce this example as a practical exercise.
 The example shows a minimalistic analysis of the 'Considerations of Future Consequences (CFC) Scale'.
-The analysis demonstrates a successful implementation of our workflow including downloads of external data, comparison of their integrity using a checksum, and the running of a confirmatory factor analysis on the first few items using `lavaan` [@lavaan2012].
-Once all mentioned tools are installed on a computer, the following few steps are necessary to reproduce our analysis:
+The analysis demonstrates a complete implementation of our workflow including downloads of external data, comparison of their integrity using a checksum, and  a confirmatory factor analysis on the first few items using the R package `lavaan` [@lavaan2012].
+Once all required tools are installed on a computer, the following four command-line commands are sufficient to reproduce our demo analysis:
 
 ```{bash, eval=FALSE}
+# (1) download the complete repository containing the workflow to your computer
 git clone https://github.com/aaronpeikert/workflow-showcase.git
+# (2) enter the directory containing the cloned workflow
 cd workflow-showcase
+# (3) building the docker container
 make build
+# (4) running the analysis and producing the final PDF inside the container
 make all DOCKER=TRUE

### 2020-04-10 manuscript.Rmd

@@
 # Literate Programming
-
-The translation of computational results into a human-readable summary, for example into a technical report, a presentation, or a journal manuscript, is time-consuming and error-prone.
-Typical errors result from copy\&paste mistakes, erroneous rounding, or missed updates of the manuscript when the computer code has changed.
-In order to not only create fully reproducible results but also fully reproducible reports, one needs to resort to the Literate Programming paradigm [@knuth1984literate], in which human-readable language and computer code are intermingled to create dynamic documents whose order follows the logic of thought rather then the order of the computer.
-R Markdown is a simple markup language to create dynamic documents with embedded chunks of R code that can be exported to standard formats such as documents (docx, rtf, epub), presentations (ppt, html) or websites (html) using the `knitr` package [@xie2015; @xie2019].
-Several packages extend the functionality of knitr.
-Of particular note are the `papaja` package [@papaja2018], which offers additional functions to enable American Psychological Association (APA) style document formatting, including a journal-style final typeset format, and the `stargazer` package [@stargazer2018], which provides journal-ready tables and reports of statistical models.
-Figure 1 illustrates R Markdown syntax using the `papaja` package and Figure 2 shows the resulting rendered document.
 

### 2020-05-12 manuscript.Rmd
replaced literate programming with dynamic document generation

@@
 
-keywords          : "reproducibility, R, version management, literate programming, dependency management, containerization, open science"
+keywords          : "reproducibility, R, version management, dynamic document generation, dependency management, containerization, open science"
 wordcount         : "`r tryCatch(wordcountaddin::word_count(here::here('manuscript.Rmd')), error=function(err) 'NA')`"
@@
 Here, we combine four software tools, whose interplay can guarantee full computational reproducibility of data analyses and their reporting.
-There are various ideas on how to enhance reproducibility [@piccoloToolsTechniquesComputational2016], four of which we believe to be particularly important: literate programming: [@ruleTenSimpleRules2019], version control [@barbaHardRoadReproducibility2016], dependency management [@askrenUsingMakeReproducible2016], and containerization [@clyburne-sherinComputationalReproducibilityContainers2018].
+There are various ideas on how to enhance reproducibility [@piccoloToolsTechniquesComputational2016], four of which we believe to be particularly important: dynamic document generation: [@ruleTenSimpleRules2019], version control [@barbaHardRoadReproducibility2016], dependency management [@askrenUsingMakeReproducible2016], and containerization [@clyburne-sherinComputationalReproducibilityContainers2018].
 We argue that only a workflow using all four concepts in unison can guarantee confidence in reproducing a scientific report [see @turingway2020 for similar arguments].
-Various implementations of these concepts exist, but we consider the following four best suited for the R environment [@r2019]: for literate programming, R Markdown [@rmarkdown2019], for version control, Git [@chaconProGit2014], for dependency management, Make [@feldmanMakeProgramMaintaining1979], and for containerization, Docker [@merkelDockerLightweightLinux2014].
+Various implementations of these concepts exist, but we consider the following four best suited for the R environment [@r2019]: for dynamic document generation, R Markdown [@rmarkdown2019], for version control, Git [@chaconProGit2014], for dependency management, Make [@feldmanMakeProgramMaintaining1979], and for containerization, Docker [@merkelDockerLightweightLinux2014].
 Each of these software solutions serves a valuable meta-scientific goal (reproducibility) and increases the researches' productivity.
@@
 Third, all computer code (such as a statistical analysis in R) is executed in a virtual environment that guarantuees exact reproduction of results independent of the host operating system, R version, and installed package versions.
-Finally, literate programming interweaves human-readable code and automatically produced results (such as p values, effect sizes, confidence intervals) to eliminate inconsistency errors such as those arising from careless copy and paste actions.
+Finally, dynamic document generation (also known as the literate programming paradigm) interweaves human-readable code and automatically produced results (such as point estimates, p values, or confidence intervals) to eliminate inconsistency errors such as those arising from careless copy and paste actions.
 
-# Literate Programming
+# Dynamic Document Generation
 

### 2020-05-12 manuscript.Rmd
various changes in response to reviewers

@@
 
-# The Reproducible Workflow in a nutshell
+# The Reproducible Workflow in a Nutshell
 
-![](Images/placeholder.png)
-Figure XXX gives an overview of how the four components of our workflow interface to ensure computational reproducibility.
+```{r schematic, fig.cap="Schematic illustration of the interplay of the four components central to a reproducible workflow: version control (git), dependency management (Make), containerization (Docker), and dynamic document generation (R Markdown).
+"}
+knitr::include_graphics("Images/placeholder.png")
+```
+
+Figure @\ref{fig:schematic} gives an overview of how the four components of our workflow interface to ensure computational reproducibility.
 We begin with a minimal description of the roles of the four components.
@@
 
-Fundamentally, reproducibility means that results remain identical if neither the script nor the data have changed.
-It is often not trivial to find out whether anything in a project has changed over time and if so, to "go back in time." The Git program enables you to do both.
+Fundamentally, reproducibility means that computational results remain identical if neither the script nor the data have changed.
+It is often not trivial to find out whether any element in a project has changed over time and if so, to "go back in time." The Git program enables you to do both.
 A good mental model for Git is that it takes a sequence of snapshots of all files it is supposed to track.
 In the language of Git, these snapshots are "commits." A commit represents a complete copy of the state of all tracked files.
-Each commit has a unique identifier (a hash) and a human-readable description (commit message).
-Going back to one state is as easy as traversing the history of all commits and switching to a given one; it is possible to visually track changes between different versions.
-This collection of snapshots is called a "repository," which ideally tracks your entire R project.
+Each commit has a short, unique identifier (a hash code) and a human-readable description (commit message).
+Going back to one state is as easy as traversing the history of all commits and switching the repository to a given previous state; it is possible to visually track changes between different versions (similar to the track changes function offered in Microsoft Word).
+The collection of all snapshots is called a "repository," which ideally tracks your entire R project.
 
@@
 git init # to initialize git in the current directory
-git add ./data/iris.csv ./R/analysis.R # track specific files/changes
+git add ./data/iris.csv ./R/analysis.R # track specific files
 git commit -m "add data & analysis" # take snapshot with comment
@@
 These commands need to be executed in the terminal, which you can access from within RStudio (`Shift + Alt + R`).
-RStudio also offers a graphical user interface for Git and, quite likely, this interface is sufficient for almost every interaction with Git.
+RStudio also offers a graphical user interface for Git.
+For most basic operations, this interface is convenient and sufficient.
 
@@
 ```{bash, eval = FALSE}
-git log
-git checkout 77db06f78e
+git log # inspect all changes
+git checkout 77db06f78e # revert local directory to previous version with hash code '77db06f78e'
 ```
@@
 A popular service for sharing materials via Git is [GitHub](https://github.com).
-Alternatively, institutions can host an equally feature-rich open-source software called [GitLab](https://gitlab.com) themself, avoiding the reliance on commercial service providers.
-Just sharing Git repositories on GitHub with the public is always free, private repositories (only visible to persons you invite) are [free for researchers](https://help.github.com/en/articles/applying-for-an-educator-or-researcher-discount).
+Alternatively, institutions can host an equally feature-rich open-source service called [GitLab](https://gitlab.com), avoiding the reliance on commercial service providers.
+Just sharing Git repositories on GitHub with the public is always free, private repositories (only visible to persons you invite) are [free for researchers](https://help.github.com/en/articles/applying-for-an-educator-or-researcher-discount) or have limited features.
 After creating a user account, one can create a new repository and GitHub provides information on how to upload your repository from the terminal, e.g., for our repository (here with user name 'aaronpeikert' and repository name 'reproducible-research'):
@@
 ```{bash, eval=FALSE}
+# link remote github repository to local directory
 git remote add origin https://github.com/aaronpeikert/reproducible-research.git
+# push all changes from local repository to the remote repository
 git push -u origin master
@@
 
-`git push` or the green upward arrow in the Git pane (see Figure 3) uploads local updates. To download this Git repository to another computer, type into the terminal:
+`git push` or the green upward arrow in the Git pane (see Figure 3) uploads local updates. To download the remote Git repository on another computer, type into the terminal:
 
@@
 In particular, GitHub's issue management can be leveraged as a post-publication platform to discuss manuscripts and their results [^1].
-Another benefit of using Git and GitHub is that experimentation is highly encouraged since you can go back to any state quickly and even when you lose access to the file on your computer, everything is backed up on the GitHub servers.
-Further, one can reduce the liklihood of dead code accumulating (e.g., lines that have been commented out) because it is safe to simply remove unneeded code blocks and track the removal in Git.
+Another benefit of using Git and GitHub is that experimentation is highly encouraged since you can go back to any state quickly.
+Even when you lose access to the file on your computer, everything can be backed up on a remote git server.
+Further, one can reduce the liklihood of dead code accumulating (e.g., lines that have been commented out) because it is safe to simply remove unneeded code blocks and track their removal in Git.
 
-GitHub allows you to archive a specific version of your repository in the form of a release.
-A release tags a commit, e.g., as "submission,"[^sub] "preprint," or "published," and archives also "binary" products of your code, e.g.
+GitHub allows you to archive and label a specific version of your repository in the form of a release.
+A release tags a particular commit with an arbitrary label, e.g., as "submission,"[^sub] "preprint," or "published," and archives also "binary" products of your code, e.g.
 the resulting pdf of the manuscript or the docker image (see [Containerization]).
@@
 Even when you have obtained a given version of a project with the aim to reproduce reported result and you can confirm that this version is unchanged, you may not know exactly how to reproduce the results because it may be unclear which files should be executed in which order.
-This is particularly the case when complex preprocessing pipelines are part of the computation and/or external dependencies are many.
-Handling such dependencies is easy with Make because it allows you to create (computational) recipes to (re-)create files.
+This is particularly the case when complex preprocessing pipelines are part of the computation and/or external dependencies (for example, a program needs to be called outside the R environment) are present.
+Handling such dependencies is easy with Make because it allows you to manage dependencies by creating (computational) recipes to create or recreate files.
 
-A `Makefile` obeys the following scheme:
-
-```{bash, eval=FALSE}
-target: dependency1 dependency2
-  command-to-create-target
-```
-
+A `Makefile` obeys a simple scheme.
+It contains a list of recipes.
+Each recipe has a target (the name of the recipe) followed by a list of dependent targets or files.
+If any of the dependencies have changed since the last time the target was built, the recipe's commands are executed to recreate the target file.
 A typical `Makefile` might look like this:
@@
 
-The first line can be interpreted as follows: To create `analysis.pdf`, the file `analysis.Rmd`, which depends on `data/clean.csv`, needs to be rendered.
-This dependency is itself a target.
+The first line can be interpreted as follows: To create `analysis.pdf` (the first target in the file), the file `analysis.Rmd`, which depends on `data/clean.csv`, needs to be rendered.
+This dependency is itself a target (the second target in the file).
 To create `data/clean.csv`, `R/clean.R` and `data/iris.csv` are needed.
 If you type `make analysis.pdf`, Make first checks whether the dependencies do exist and, if not, creates them.
-Here, if `data/clean.csv` does not exist, Make creates it.
+Here, if `data/clean.csv` does not exist, Make creates it by executing the second target.
 Also, if one of the dependencies of a target is newer then the target itself, then Make updates everything that directly or indirectly depends on the target.
 Here, if `data\iris.csv` is newer then `data\clean.csv`, Make will attempt to recreate `data\clean.csv` first and then `analysis.pdf`.
-If there is a dependency which is missing, and there is no recipe to make it, Make stops with an error message.
+If there is a dependency which is missing, and there is no target to make it, Make stops with an error message.
 It is a convention to have the first target named `all`, which depends on everything.
-Then the command `make` without any argument automatically creates everything.
+Then the command `make` without any argument automatically creates everything possible in the project.
 The button `Build All` from within RStudio triggers this process (see Figure 4).
@@
 
-If you have followed our workflow as presented thus far, a fellow researcher is only three commands away from fully reproducing your analysis:
+If you have followed our workflow as presented thus far, a fellow researcher is only three commands away from fully reproducing your analysis.
+They would simply have to type the following commands in the terminal:
 
 ```{bash, eval=FALSE}
+# obtain a local copy from the remote repository
 git clone https://github.com/sirfisher/iris.git
+# enter the local copy
 cd iris
+# create all files from that project
 make all
@@
 
-However, successful reproducibility relies on the crucial assumption that your computational environments are identically or sufficiently compatible, e.g., all software dependencies are installed (R and all additional R Packages) and no updates or other changes to the computational environment break or alter your analysis.
+If you execute the above on your system, there is a good chance that you cannot reproduce our manuscript and the `make all` command exits with an error.
+Successful reproducibility relies on the crucial assumption that your computational environments are identically or sufficiently compatible, e.g., all software dependencies are installed (R and all additional R Packages) and no updates or other changes to the computational environment break or alter your analysis.
 As we will shortly see, ensuring full computational reproducibility requires one further level of documentation, that is, documentation of the computational environment.

### 2020-05-12 manuscript.Rmd
- further revisions - added comments to Docker file

@@
 # obtain a local copy from the remote repository
-git clone https://github.com/sirfisher/iris.git
+git clone https://github.com/aaronpeikert/reproducible-research.git
 # enter the local copy
@@
 
-Surprisingly in an image with a more recent version of R (3.6.1), this results in another sample despite the identical seed:
+Executing the code in an image with a more recent version of R (3.6.1), this results in another sample despite the identical random seed:
 
@@
 Only rarely does an analysis depend on base R only.
-Often, a considerable number of packages is required, and the addition of each package will increase the likelihood of breaking reproducibility with every update to any of the packages of base R.
+Typically, a considerable number of packages is required that each may depend on multiple other packages, and the addition of each package will increase the likelihood of breaking reproducibility with every update to any of the packages of base R (the resulting frustration is sometimes refered to as _dependency hell_).
 The whole endeavor of reproducibility is therefore at stake every time an update is rolled out.
 To ensure long-term reproducibility, Docker replicates the original computational environment of an analysis exactly.
-Note that we do not intend to advocate that software should not be updated; updates typically promote bugfixes and provide new functionality; our point is that full computational reproducibility is only achieved if the originally used software versions are documented.
-Among other things, this makes it possible to trace back recent updates as to what change in what package caused a non-reproduction of the original results.
-Quite to the contrary, with solutions like Docker, it gets easier than ever to safely update to new versions just by changing the R version number of the Docker image.
-This convenience is possible because of the efforts of the [Rocker project](https://GitHub.com/rocker-org/rocker) [@boettigerIntroductionRockerDocker2017a], which provides Docker images preconfigured with an R-installation of given R versions.
+Note that we do not intend to advocate that software should not be updated; updates typically promote bugfixes and provide new functionality; our point is that full computational reproducibility is only achieved if the software versions used originally are precisely documented.
+Among other things, this makes it possible to trace back update histories to discover which change in which package caused the non-reproducibility.
+Quite to the contrary, with containerization, it gets easier than ever to safely update to new versions just by changing the R version number of the Docker image (and reverting back if this update breaks code).
+This convenience is possible because of the efforts of the [Rocker project](https://GitHub.com/rocker-org/rocker) [@boettigerIntroductionRockerDocker2017a], which provides Docker images pre-configured with an installation of selected R versions.
 These packages are taken from MRAN [@revolutionanalyticsReproducibilityUsingFixed2019], a repository for R packages fixed to the last date on which the R version of the image was the most recent.
@@
 
-Building on a basic Rocker image, we can specify further requirements for software in a Dockerfile. For example, the basis for this article's Docker image is the following Dockerfile:
+Building on a basic Rocker image, we can specify further software dependencies in a Dockerfile. For example, the basis for this manuscript's Docker image is the following Dockerfile:
 
 ```{bash, eval=FALSE}
+# Define the R version to be installed from rocker project
 FROM rocker/verse:3.6.1
+# install CRAN R packages: pacman, here, and pander
 RUN install2.r --error --skipinstalled\
   pacman here pander
+# install additional R packages from github: papaja and wordcountaddin
 RUN installGithub.r\
   crsh/papaja benmarwick/wordcountaddin
+# set the working directory inside the container
 WORKDIR /home/rstudio

### 2020-05-12 manuscript.Rmd
more changes

@@
 There are two ways to share a Docker image; either by sharing the Dockerfile that creates the image, or by sharing the image itself, e.g., through a service like [Dockerhub](https://hub.docker.com).
-While both ways guarantee a replicable computational environment, sharing the Dockerfile is more transparent; in our workflow, we can use Git to track changes in the Dockerfile (such as updates to dependencies).
-A possible downside is that in order to create an image from a Dockerfile, all software repositories need to be still availible.
-Hence, to garuntee reproducibility for decades to come, it is best to archieve the binary image at major points of the projets progress e.g.
-publication with a release on GitHub (see [Version Management] for details).
+While both ways guarantee a replicable computational environment, sharing the Dockerfile is more transparent and more space-saving; in our workflow, we can use Git to track changes in the Dockerfile (such as updates to dependencies).
+A possible downside is that in order to create an image from a Dockerfile, all software repositories need to be still available.
+Hence, to guarantee longterm reproducibility, it is best to archive the complete binary image at major points of the projects' progress, for example, on publication (ideally, using a release tag; see [Version Management] for details).
 
-There are two options to interact with a container. Both options are based on the `docker run` command. The first way is to run a command inside the container. The call takes the form:
+There are two options to execute commands in a container. Both options are based on the `docker run` command. The first way is to run a command inside the container. The call takes the form:
 
 ```{bash, eval=FALSE}
+# execute a given command in a given container image; do not save
+# the state of the container; accept inputs from and return outputs to terminal
 docker run --rm -it <IMAGENAME> <COMMAND>
@@
 Interacting with docker this way is especially useful when using `R` as a command because you then have access to the standard R terminal.
-This is the command to start an interactive R session inside a Docker image called `reproducible-research`:
+This is the command to start an interactive R session inside a Docker image called `reproducible-research` (see Figure \@ref{fig:docker-r-terminal} for a screenshot):
 
 ```{bash, eval=FALSE}
+# start an interactive R session in the container named 'reproducible-research'
 docker run --rm -it reproducible-research R
@@
 
-![R terminal running inside Docker](Images/docker-r-terminal.png)
+```{r docker-r-terminal, fig.cap="R terminal running inside Docker"}
+knitr::include_graphics("Images/docker-r-terminal.png")
+```
 
-The second option is to start the container in the background (not supplying a specific command) and to interact with the container via the webrowser and the RStudio server instance running in it.
-In order to do so, you need to supply a password to log into the RStudio server (`-e PASSWORD=<YOUR_PASS>`) and open a port (`-p 8787:8787`).
+The second option is to start the container in the background and to interact with the container via the webrowser and the RStudio server instance running in it.
+In order to do so, you need to supply a password to log into the RStudio server (`-e PASSWORD=<YOUR_PASS>`) and open a network service on a specified port (`-p 8787:8787`).
 
@@
 
-The main directory for RStudio inside the container is `/home/rstudio`, so the final call to start RStudio inside a Docker container may look like this:
+The main directory for RStudio inside the container is `/home/rstudio`, so the complete call to start RStudio inside a Docker container may look like this in the local terminal:
 
 ```{bash, eval=FALSE}
+# start docker in the background, open a local web service with a virtual
+# Rstudio instance and enable access to selected local directories
 docker run --rm -it -e PASSWORD=<YOUR_PASS> -p 8787:8787 -v
@@
 
-![RStudio running inside Docker](Images/docker-rstudio.png)
+```{r rstudio-in-docker, fig.cap="RStudio running inside Docker"}
+knitr::include_graphics("Images/docker-rstudio.png")
+```
+
+Figure \@ref{fig:rstudio-in-docker} shows a screenshot of Rstudio running inside Docker accessed from a local web browser.
 

### 2020-05-12 manuscript.Rmd
fixed typo in Dockerfile

@@
 ```{bash, eval=FALSE}
-projekt := $(notdir $(CURDIR))
+project := $(notdir $(CURDIR))
 current_dir := $(CURDIR)
@@
 ifeq ($(DOCKER),TRUE)
-	run:=docker run --rm --user $(uid) -v $(home_dir):/home/rstudio $(projekt)
+	run:=docker run --rm --user $(uid) -v $(home_dir):/home/rstudio $(project)
 	current_dir=/home/rstudio

### 2020-05-12 manuscript.Rmd
clarified checksums

@@
 ```{bash, eval=FALSE}
+# install docker using choco
 choco install -y git make docker-desktop
@@
 ```{bash, eval=FALSE}
+# install docker via brew
 brew cask docker
@@
 ```{bash, eval=FALSE}
+# install docker via advanced package tool
 apt install git make docker 
@@
 In most of the cases we have seen, these are either space constraints or privacy considerations.
-In these cases, of course, unrestricted reproducibility is not guaranteed.
+In these cases, unrestricted reproducibility is not guaranteed.
 If dividing data and scripts is unavoidable, we recommend validating all data files using checksums (also called a "hash", e.g., using the functions provided in package `digest` [@digest2019]) before analyzing them.
-To do this, a checksum must be created and stored at the time of the original analysis.
+A checksum is a short fixed-length fingerprint (often displayed in hexadecimal system) of a file with the purpose of verifying integrity of a digital object.
+Fingerprints are computed from digital objects such that they change with high probability if data is changed only little.
+To use checksum validation, checksums for all data files must be created and stored at the time of the original analysis.
 At the time of reproduction, the current checksum must then be compared with the stored checksum to ensure data integrity.
@@
 ```{r}
+# create a dummy data.frame with two columns
 x <- data.frame(VAR1=c(1,2,3,4),VAR2=c(0,4,6,9) )
-checksum <- digest::digest(x, "sha1")
-if (checksum != "8e222b9b3616595dc3f5946f9eac101fbf1336dc"){
+# compute checksum using md5
+checksum <- digest::digest(x, "md5")
+if (checksum != "5ba412f5a26f43842971dd74954fcdeb"){
   warning("Mismatch between original and current data file!")

### 2020-05-12 manuscript.Rmd
removed the idea of creating R packages - I think this is replaced by worcs and related approaches

@@
 
-Another way to increase the chance of reproducibility is to develop an R package, which is a self-contained set of files with well-defined meta-data, for each analysis.
-The reasoning is that, abiding by the strict rules of package development and specifying all dependencies can archieve a high degree of stability can be achieved [e.g., @marwickPackagingDataAnalytical2018].
-
 Other than these tools, which ease the process of creating workflows like ours does, we have noticed increased interest in changing the way research is published and used [@perkelToolkitDataTransparency2018], with the emergence of *life code* [@perkelPioneeringLivecodeArticle2019] and *continuous integration* [@beaulieu-jonesReproducibilityComputationalWorkflows2017; @yenniDevelopingModernData2018].

### 2020-05-12 manuscript.Rmd
added worcs references

@@
 The package [`reprex`](https://github.com/tidyverse/reprex) [**repr**oducible **ex**ample, @reprex] is also worth noting, but its scope is limited.
+A particular noteworthy approach is provided by the [`worcs`](https://cran.r-project.org/web/packages/worcs) package [@worcs, @worcspaper], which is an R project template that creates a standardized file structure for code and data supporting version management with git, package management with renv and dynamic document generation with R Markdown.
+We acknowledge that worcs is much easier to install and provides a one-click solution for the creation of reproducible projects.
+It achieves a high standard of reproducibility but does not guarantuee full computational reproducibility and is limited to dependency management within the R environment.
 

### 2020-05-13 manuscript.Rmd
removed 'author contributions' and 'conflicts of interest' as this is not required by QCMB

@@
 We are aware that implementing the proposed workflow is not straightforward and the difficulty of its implementation may vary by platform.
-For example, already the installation of all tools is already easier on POSIX-compatible platforms such as Unix, Linux, or macOS (but not Windows).
-However, once a reproducible workflow is established as a default, it can be used with minimal changes from project to project.
+For example, the installation of all tools is already easier on POSIX-compatible platforms such as Unix, Linux, or macOS (but not Windows).
+However, once a reproducible workflow is established as a default, it can be used with minimal changes for every R project.
 
@@
 The proposed workflow leverages various existing tools that are partly integrated into RStudio already.
-Parts of the proposed workflow have been integrated into stand-alone packages (such as workflowr [@workflowr] or holepunch  [@holepunch]) but they are either incomplete in the sense of our proposal or rely heavily on propietary service providers.
+Parts of the proposed workflow have been integrated into stand-alone packages (such as worcs [@worcs], workflowr [@workflowr], holepunch  [@holepunch]) but they are either incomplete in the sense of our proposal or rely heavily on propietary service providers.
 We hope that with increasing awareness of the challenges of computational reproducibility, the increased demand for unified and open solutions will lead to an even better integration of existing tools to allow reproducible workflows to become a standard in psychological research.
-
-\newpage
-
-## Author Contributions
-
-Aaron Peikert and Andres Brandmaier jointly generated the idea for the manuscript.
-Aaron Peikert implemented the workflow.
-Aaron Peikert and Andreas Brandmaier jointly generated the R code examples.
-Aaron Peikert wrote the first draft of the manuscript, and both authors critically edited it.
-Both authors approved the final submitted version of the manuscript.
-
-## Conflicts of Interest
-
-The authors declare that there were no conflicts of interest with respect to the authorship or the publication of this article.
 

### 2020-05-13 manuscript.Rmd
fixed rendering problem in RMarkdown because of trailing whitespace in abstract

@@
     institution   : "Max Planck UCL Centre for Computational Psychiatry and Ageing Research, Berlin, Germany & London, U.K."
+
 authornote: |
@@
 
-abstract: "`r readr::read_file('abstract.Rmd')`"
+abstract: "`r tryCatch(trimws(readr::read_file('abstract.Rmd')))`"
 
 keywords          : "reproducibility, R, version management, dynamic document generation, dependency management, containerization, open science"
+
 wordcount         : "`r tryCatch(wordcountaddin::word_count(here::here('manuscript.Rmd')), error=function(err) 'NA')`"
@@
 
-<div id = "refs"></div>
+<div id="refs" custom-style="Bibliography"></div>
 \endgroup

### 2020-05-13 manuscript.Rmd
added comments to file path examples

@@
 RStudio also offers a graphical user interface for Git.
-For most basic operations, this interface is convenient and sufficient.
+For most basic operations, this interface is convenient and sufficient (see Figure \@ref{fig:gitpane}).
 
-![Git Pane in RStudio](Images/git-pane.png)
+```{r gitpane, fig.cap="Git Pane providing easy access to basic functions in RStudio"}
+knitr::include_graphics("Images/git-pane.png")
+```
 
@@
 ```{r, eval=FALSE}
-# BAD
+# BAD because of path that is specific to the computer
 iris <- read.csv("/home/aaron/reproducible-research/data/iris.csv")
-# GOOD
+# GOOD because it is a relative path but slash depends on OS
 iris <- read.csv("data/iris.csv")
-# BETTER
+# BETTER because truly compatible across OS
 iris <- read.csv(here("data", "iris.csv"))

### 2020-05-13 manuscript.Rmd
moving figures to proper figures with refs

@@
 
-```{r schematic, fig.cap="Schematic illustration of the interplay of the four components central to a reproducible workflow: version control (git), dependency management (Make), containerization (Docker), and dynamic document generation (R Markdown).
+```{r schematic, eval = TRUE, echo = FALSE, fig.cap="Schematic illustration of the interplay of the four components central to a reproducible workflow: version control (git), dependency management (Make), containerization (Docker), and dynamic document generation (R Markdown).
 "}
@@
 
-![Source of an R Markdown](Images/rmarkdown.png)
+```{r source-rmarkdown, eval = TRUE, echo = FALSE,  fig.cap="Exemplary source of an R Markdown showing a combination of executable R code and English manuscript text."}
+knitr::include_graphics("Images/rmarkdown.png")
+```
 
-![Rendered result](R/rmarkdown.pdf)
+```{r rendered-result, eval = TRUE, echo = FALSE,  fig.cap="Rendered result of the source code shown in Figure \ref{fig:source-rmarkdown}."}
+knitr::include_graphics("R/rmarkdown.pdf")
+```
 
@@
 
-```{r gitpane, fig.cap="Git Pane providing easy access to basic functions in RStudio"}
+```{r gitpane, eval = TRUE, echo = FALSE, fig.cap="Git Pane providing easy access to basic functions in RStudio"}
 knitr::include_graphics("Images/git-pane.png")
@@
 
-```{r docker-r-terminal, fig.cap="R terminal running inside Docker"}
+```{r docker-r-terminal, eval = TRUE, echo = FALSE, fig.cap="R terminal running inside Docker"}
 knitr::include_graphics("Images/docker-r-terminal.png")
@@
 
-```{r rstudio-in-docker, fig.cap="RStudio running inside Docker"}
+```{r rstudio-in-docker, eval = TRUE, echo = FALSE, fig.cap="RStudio running inside Docker"}
 knitr::include_graphics("Images/docker-rstudio.png")

### 2020-05-13 manuscript.Rmd
restored lost paragraph on dynamic document generation (commit: 3e5cf89ff21253ac91bcec554d806645456e3231)

@@
 # Dynamic Document Generation
+
+The translation of computational results into a human-readable summary, for example into a technical report, a presentation, or a journal manuscript, is time-consuming and error-prone.
+Typical errors result from copy\&paste mistakes, erroneous rounding, or missed updates of the manuscript when the computer code has changed.
+In order to not only create fully reproducible results but also fully reproducible reports, one needs to resort to the Literate Programming paradigm [@knuth1984literate], in which human-readable language and computer code are intermingled to create dynamic documents whose order follows the logic of thought rather then the order of the computer.
+R Markdown is a simple markup language to create dynamic documents with embedded chunks of R code that can be exported to standard formats such as documents (docx, rtf, epub), presentations (ppt, html) or websites (html) using the `knitr` package [@xie2015; @xie2019].
+Several packages extend the functionality of knitr.
+Of particular note are the `papaja` package [@papaja2018], which offers additional functions to enable American Psychological Association (APA) style document formatting, including a journal-style final typeset format, and the `stargazer` package [@stargazer2018], which provides journal-ready tables and reports of statistical models.
+Figure \@ref{fig:source-rmarkdown} illustrates R Markdown syntax using the `papaja` package and Figure \@ref{fig:rendered-result} shows the resulting rendered document.
 

### 2020-05-13 manuscript.Rmd
all figures now using include_graphics from knitr

@@
 Then the command `make` without any argument automatically creates everything possible in the project.
-The button `Build All` from within RStudio triggers this process (see Figure 4).
+The button `Build All` from within RStudio triggers this process (see Figure \@ref{fig:buildpane-rstudio}).
 
-![Build pane in RStudio](Images/build-pane.png)
+```{r buildpane-rstudio, fig.cap="Build pane in RStudio with access to Makefile target 'all'."}
+knitr::include_graphics("Images/build-pane.png")
+```
 

### 2020-05-13 manuscript.Rmd
revision of Make example to increase clarity

@@
 If any of the dependencies have changed since the last time the target was built, the recipe's commands are executed to recreate the target file.
-A typical `Makefile` might look like this:
+We illustrate the use of Makefiles with an example.
+Assume the final product is a report (`report.pdf`).
+This report is written in Rmarkdown (`report.Rmd`) and includes dynamically generated plots from a raw data file (`data/iris.csv`) that needs to be preprocessed first using a separate script (`R/preprocess.R`) into a processed data file (`ìris-processed.csv`).
+A `Makefile` for these dependencies may look like this:
 
 ```{bash, eval=FALSE}
-analysis.pdf: data/clean.csv
-  Rscript -e 'rmarkdown::render("analysis.Rmd")'
+report.pdf: data/iris-processed.csv report.Rmd
+  Rscript -e 'rmarkdown::render("report.Rmd")'
 
-data/clean.csv: R/clean.R data/iris.csv
-  Rscript -e 'source("R/clean.R")'
+data/iris-processed.csv: R/preprocess.R data/iris.csv
+  Rscript -e 'source("R/preprocess.R")'
 ```
 
-The first line can be interpreted as follows: To create `analysis.pdf` (the first target in the file), the file `analysis.Rmd`, which depends on `data/clean.csv`, needs to be rendered.
+The first line can be interpreted as follows: To create `report.pdf` (the first target in the file), the file `report.Rmd`, which depends on `data/iris-processed.csv`, needs to be rendered.
 This dependency is itself a target (the second target in the file).
-To create `data/clean.csv`, `R/clean.R` and `data/iris.csv` are needed.
-If you type `make analysis.pdf`, Make first checks whether the dependencies do exist and, if not, creates them.
-Here, if `data/clean.csv` does not exist, Make creates it by executing the second target.
+To create `data/iris-processed.csv`, `R/preprocess.R` and `data/iris.csv` are needed.
+If you type `make report.pdf`, Make first checks whether the dependencies do exist and, if not, creates them.
+Here, if `data/iris-processed.csv` does not exist, Make creates it by executing the second target.
 Also, if one of the dependencies of a target is newer then the target itself, then Make updates everything that directly or indirectly depends on the target.
-Here, if `data\iris.csv` is newer then `data\clean.csv`, Make will attempt to recreate `data\clean.csv` first and then `analysis.pdf`.
-If there is a dependency which is missing, and there is no target to make it, Make stops with an error message.
+Here, if `data\iris.csv` is newer then `data\iris-processed.csv`, Make will attempt to recreate `data\iris-processed.csv` first and then `report.pdf`.
+If there is a dependency missing, and there is no target to make it, Make stops with an error message.
 It is a convention to have the first target named `all`, which depends on everything.

### 2020-05-13 manuscript.Rmd
added thanks note to GitHub contributors

@@
    We are grateful to Julia Delius for her helpful assistance in language and style editing.
+We thank all contributors of our GitHub repository for helpful feedback on the paper.
 This paper is fully reproducible using the workflow described in this paper.
@@
 The package [`reprex`](https://github.com/tidyverse/reprex) [**repr**oducible **ex**ample, @reprex] is also worth noting, but its scope is limited.
-A particular noteworthy approach is provided by the [`worcs`](https://cran.r-project.org/web/packages/worcs) package [@worcs, @worcspaper], which is an R project template that creates a standardized file structure for code and data supporting version management with git, package management with renv and dynamic document generation with R Markdown.
+A particularly noteworthy approach is provided by the [`worcs`](https://cran.r-project.org/web/packages/worcs) package [@worcs, @worcspaper], which is an R project template that creates a standardized file structure for code and data supporting version management with git, package management with renv and dynamic document generation with R Markdown.
 We acknowledge that worcs is much easier to install and provides a one-click solution for the creation of reproducible projects.
@@
 The proposed workflow leverages various existing tools that are partly integrated into RStudio already.
-Parts of the proposed workflow have been integrated into stand-alone packages (such as worcs [@worcs], workflowr [@workflowr], holepunch  [@holepunch]) but they are either incomplete in the sense of our proposal or rely heavily on propietary service providers.
+Parts of the proposed workflow have been integrated into stand-alone packages (such as worcs [@worcs], workflowr [@workflowr], holepunch  [@holepunch]), which are particularly recommenable to beginners but they are either incomplete in the sense of our proposal or rely on proprietary service providers.
 We hope that with increasing awareness of the challenges of computational reproducibility, the increased demand for unified and open solutions will lead to an even better integration of existing tools to allow reproducible workflows to become a standard in psychological research.

### 2020-05-13 manuscript.Rmd
add reviewer thanks and added contributors

@@
    We are grateful to Julia Delius for her helpful assistance in language and style editing.
-We thank all contributors of our GitHub repository for helpful feedback on the paper.
+We thank the reviewers and all contributors of our GitHub repository for helpful feedback on the paper.
 This paper is fully reproducible using the workflow described in this paper.
@@
 
-The `--rm` flag means that the state of the container after the command finishes is not saved.
-The `-it` flag tells Docker to accept inputs and return outputs to the terminal.
-Interacting with docker this way is especially useful when using `R` as a command because you then have access to the standard R terminal.
-This is the command to start an interactive R session inside a Docker image called `reproducible-research` (see Figure \@ref{fig:docker-r-terminal} for a screenshot):
+The `--rm` flag means that the state of the container after the command will have finished is not going to be saved.
+The `-it` flag tells Docker to run the command interactively, that is, to accept keyboard inputs and return outputs to the terminal.
+For example, this is the command to start an interactive R session inside a Docker image called `reproducible-research` (see Figure \@ref{fig:docker-r-terminal} for a screenshot):
 

### 2020-05-13 manuscript.Rmd
link to github repo for installation instruction links

@@
 Other than on R, RStudio, and R Markdown, our workflow relies on three pieces of software from outside the R environment: Git, Make, and Docker.
-The installation process of these software packages depends on the operating system, e.g., on macOS, Make is always available, while on Linux, Git and Make are often already installed.
-In the following section, we share what what we consider the easiest way to install those packages.
-However, installation processes may be subject to change, and we advise that readers also consult the documentations of the packages.
+The smoothness of the installation process of these software packages varies across operating systems.
+For example, on macOS, Make is always available, whereas Linux systems are typically shipped with both Git and Make.
+In the following section, we share what what we consider the easiest way to install those packages across common operating systems.
+However, installation processes may be subject to change, and we advise that readers also consult the documentations of the packages or see our collection of links to tutorials and installation instructions on our GitHub repository.
 

### 2020-05-13 manuscript.Rmd
improve code comments

@@
 ```{bash, eval=FALSE}
+# -- type this in a terminal --
 git init # to initialize git in the current directory
@@
 ```{bash, eval = FALSE}
+# -- type this in a terminal --
 git log # inspect all changes
@@
 ```{bash, eval=FALSE}
+# -- type this in a terminal --
+
 # link remote github repository to local directory
@@
 ```{bash, eval=FALSE}
+# -- type this in a terminal --
 git clone https://github.com/aaronpeikert/reproducible-research.git
@@
 ```{bash, eval=FALSE}
+# -- this is a Makefile --
+
 report.pdf: data/iris-processed.csv report.Rmd
@@
 ```{bash, eval=FALSE}
+# -- type this in a terminal --
+
 # obtain a local copy from the remote repository
@@
 ```{r,eval=FALSE}
+# -- R code --
 set.seed(1234)
@@
 ```{bash, eval=FALSE}
+# -- this is a Dockerfile --
+
 # Define the R version to be installed from rocker project
@@
 ```{bash, eval=FALSE}
+# -- type this in a terminal --
+
 # execute a given command in a given container image; do not save
@@
 ```{bash, eval=FALSE}
+# -- type this in a terminal --
+
 # start an interactive R session in the container named 'reproducible-research'
@@
 ```{bash, eval=FALSE}
+# -- this is a Makefile --
+
 project := $(notdir $(CURDIR))
@@
 
-Windows systems typically require the biggest efforts to install all necessary software.
-Note that you have to have Windows Pro, Enterprise, Education or Server installed, as Microsoft prevents the use of Docker on Windows Home.
-There is a package manager for Windows called Chocolatey, which you can install from: [https://chocolatey.org/install](https://chocolatey.org/install), providing all dependencies.
+Windows systems typically require the biggest efforts to install all necessary pieces of software.
+Note that you have to have either Windows Pro, Enterprise, Education, or Server installed, as Microsoft prevents the use of Docker on Windows Home.
+There is a package manager for Windows called Chocolatey, which you can install from: [https://chocolatey.org/install](https://chocolatey.org/install).
+Chocolatey provides all software packages needed for our workflow in one place..
 Having installed Chocolatey (and restarted the computer), all dependencies can be installed in an **admin terminal** (Windows key, then type `cmd`, right-click *Run as administrator*) via:
@@
 ```{bash, eval=FALSE}
-# install docker using choco
+# -- type this in a terminal --
+
+# install docker, make, and git using choco
 choco install -y git make docker-desktop
@@
 ```{bash, eval=FALSE}
+# -- type this in a terminal --
+
 # install docker via brew
@@
 ```{bash, eval=FALSE}
+# -- type this in a terminal --
+
 # install docker via advanced package tool
@@
 Just replace any docker calls with `singularity docker`, because singularity fully supports docker images.
-A possible advantage is that singularity works well in high-performance computing environments and on old Linux versions, the downside is that Singularity is only available on Linux.
+A possible advantage is that singularity works well in high-performance computing environments and on old Linux versions, the downside is that Singularity is currently only available on Linux.
 
@@
 ```{r, eval=FALSE}
+# -- R code --
+
 # BAD because of path that is specific to the computer
@@
 ```{bash, eval=FALSE}
-# (1) download the complete repository containing the workflow to your computer
+# -- type this in a terminal --
+
+# (1) obtain a local copy of the remote repository
 git clone https://github.com/aaronpeikert/workflow-showcase.git
-# (2) enter the directory containing the cloned workflow
+# (2) enter the directory containing the project
 cd workflow-showcase
-# (3) building the docker container
+# (3) build the docker container
 make build
-# (4) running the analysis and producing the final PDF inside the container
+# (4) run the analysis and produce the final PDF inside the container
 make all DOCKER=TRUE
@@
 ```{bash, eval=FALSE}
+# -- type this in a terminal --
+
+# (1) obtain a local copy of the remote repository
 git clone https://github.com/aaronpeikert/reproducible-research.git
+# (2) enter the directory containing the project
 cd reproducible-research
+# (3) build the docker container
 make build
+# (4) run the analysis and produce the final PDF inside the container
 make all DOCKER=TRUE

### 2020-05-14 manuscript.Rmd
added new sentence addressing the importance of reproducibility and openness for efficiency of knowledge accumulation

@@
 Ensuring reproducibility is a prerequisite for replicability (the ability to reach a similar conclusion from the same analysis and new data), and a means to increase trustworthiness of empirical results [@epskamp2019rep].
+Transparency and accessibility are  central scientific values and open, reproducible code will increase the efficiency and veracity of knowledge accumulation [@nosek2012scientific].
 
@@
 An alternative to Docker on Linux is Singularity [@kurtzerSingularityScientificContainers2017].
-Just replace any docker calls with `singularity docker`, because singularity fully supports docker images.
+To use it, just replace any docker calls with `singularity docker`, because singularity fully supports docker images.
 A possible advantage is that singularity works well in high-performance computing environments and on old Linux versions, the downside is that Singularity is currently only available on Linux.
@@
 Next, one needs to make sure that all files relevant for the analysis can be moved to another computer.
-To this end, it is recommendable that all files are within one folder (or enclosed subfolders within it) and all paths are relative to that folder.
+To this end, it is recommendable that all files are within one folder (or enclosed subfolders within it) and all paths are relative to that folder because absolute paths are specific to a given computer.
 A robust solution to the problem of making sure that file access does not break across computing platforms are [RStudio projects](https://r4ds.had.co.nz/workflow-projects.html) and the `here` package [@here2017] to manage file access.
+The here package solves two common issues with relative paths.
+First, it takes care of the fact that path separator characters vary across operating systems (typically, slash or backslash).
+Second, anchor points of relative paths may differ depending on context.
+For example, knitr interprets paths relative to the dynamic document whereas R has a current working directory that may change over the course of an R ssession.
+The `here` package provides consistent paths relative to the project directory.
+The following are three examples of refering to local files ranging from absolute paths with system-specific path separators (bad) to relative paths using the `here` package:
 

### 2020-05-14 README.Rmd
removed current-dir-related check in Makefile update to README

@@
 library(here)
+library(readr)
+library(pander)
 ```
@@
 ```{r abstract, child = 'abstract.Rmd'}
+```
+
+# Resources
+
+```{r, include=FALSE}
+resources <- read_csv(here("data", "resources.csv"), trim_ws = FALSE)
+```
+
+```{r, echo=FALSE, asis=TRUE}
+pander(
+  resources,
+  split.tables = Inf,
+  split.cells = Inf,
+  missing = "",
+  keep.line.breaks = TRUE,
+  style = "multiline"
+)
 ```
@@
 
-**Note: Windows user need to hand edit the `Makefile` and set current_path to the current directory and use `make all DOCKER=TRUE WINDOWS=TRUE`.
+**Note: Windows user need to manually edit the `Makefile` and set current_path to the current directory and use `make all DOCKER=TRUE WINDOWS=TRUE`.
 We hope that future releases of Docker for Windows will not require that workaround.**

### 2020-05-14 manuscript.Rmd
include nutshell figure

@@
 Ensuring reproducibility is a prerequisite for replicability (the ability to reach a similar conclusion from the same analysis and new data), and a means to increase trustworthiness of empirical results [@epskamp2019rep].
-Transparency and accessibility are  central scientific values and open, reproducible code will increase the efficiency and veracity of knowledge accumulation [@nosek2012scientific].
+Transparency and accessibility are  central scientific values and open, reproducible projects will increase the efficiency and veracity of knowledge accumulation [@nosek2012scientific].
 
@@
 "}
-knitr::include_graphics("Images/placeholder.png")
+knitr::include_graphics("Images/nutshell.svg", auto_pdf = TRUE)
 ```

### 2020-05-14 README.Rmd
clarifications and typos in README

@@
 
-This is the accompanying GitHub repository to a work in progress paper by Aaron Peikert[![ORCID iD](https://orcid.org/sites/default/files/images/orcid_16x16.png)](https://orcid.org/0000-0001-7813-818X) and Andreas Brandmaier [![ORCID iD](https://orcid.org/sites/default/files/images/orcid_16x16.png)](http://orcid.org/0000-0001-8765-6982).
+This is the accompanying GitHub repository to a work in progress paper by Aaron Peikert[![ORCID iD](https://orcid.org/sites/default/files/images/orcid_16x16.png)](https://orcid.org/0000-0001-7813-818X) and Andreas M.
+Brandmaier [![ORCID iD](https://orcid.org/sites/default/files/images/orcid_16x16.png)](http://orcid.org/0000-0001-8765-6982).
 
@@
 
-## Usual Way
+The following paragraphs describe how you can obtain a copy of the source files of our manuscript describing reproducible workflows, and create the PDF.
+Either, you can go the 'standard' way of downloading a local copy of the repository and knit the manuscript file in R, or you can use the reproducible workflow as suggested and use Make to create a container and build the final PDF file in exactly the same virtual computational environment that we used to render the PDF.
+
+## Standard Way
 
@@
 
-## Using Workflow
+## Using a Reproducible Workflow
 

### 2020-05-15 manuscript.Rmd
clarifications and typos in README
extend mentioning of singularity closes #30

@@
 "}
-knitr::include_graphics("Images/nutshell.svg", auto_pdf = TRUE)
+knitr::include_graphics("Images/placeholder.png")
 ```
@@
 A way to identify the dependencies automatically and generate a docker image out of them is provided in the [`liftR`](https://cran.r-project.org/web/packages/liftr/vignettes/liftr-intro.html) package [@liftr].
+
+This flexibility to fully control the software environment is of particular interest for software infrastructures where the user cannot install software themself, because of limited rights e.g.
+on cloud computing platforms or High-Performance Computing Clusters.
+However, executing docker itself needs special rights which are seldomly granted on High-Performance Computing Cluster.
+For this case, Singularity provides a fully compatible alternative (see Section under [Linux]).
 

### 2020-05-15 manuscript.Rmd
report -> manuscript, preprocess -> prepare/prepped #25

@@
 We illustrate the use of Makefiles with an example.
-Assume the final product is a report (`report.pdf`).
-This report is written in Rmarkdown (`report.Rmd`) and includes dynamically generated plots from a raw data file (`data/iris.csv`) that needs to be preprocessed first using a separate script (`R/preprocess.R`) into a processed data file (`ìris-processed.csv`).
+Assume the final product is a manuscript (`manuscript.pdf`).
+This manuscript is written in Rmarkdown (`manuscript.Rmd`) and includes dynamically generated plots from a raw data file (`data/iris.csv`) that needs to be preprocessed first using a separate script (`R/prepare_data.R`) into a prepared data file (`ìris_prepped.csv`).
 A `Makefile` for these dependencies may look like this:
@@
 
-report.pdf: data/iris-processed.csv report.Rmd
-  Rscript -e 'rmarkdown::render("report.Rmd")'
+manuscript.pdf: data/ìris_prepped manuscript.Rmd
+  Rscript -e 'rmarkdown::render("manuscript.Rmd")'
 
-data/iris-processed.csv: R/preprocess.R data/iris.csv
-  Rscript -e 'source("R/preprocess.R")'
+data/ìris_prepped: R/prepare_data.R data/iris.csv
+  Rscript -e 'source("R/prepare_data.R")'
 ```
 
-The first line can be interpreted as follows: To create `report.pdf` (the first target in the file), the file `report.Rmd`, which depends on `data/iris-processed.csv`, needs to be rendered.
+The first line can be interpreted as follows: To create `manuscript.pdf` (the first target in the file), the file `manuscript.Rmd`, which depends on `data/ìris_prepped`, needs to be rendered.
 This dependency is itself a target (the second target in the file).
-To create `data/iris-processed.csv`, `R/preprocess.R` and `data/iris.csv` are needed.
-If you type `make report.pdf`, Make first checks whether the dependencies do exist and, if not, creates them.
-Here, if `data/iris-processed.csv` does not exist, Make creates it by executing the second target.
+To create `data/ìris_prepped`, `R/prepare_data.R` and `data/iris.csv` are needed.
+If you type `make manuscript.pdf`, Make first checks whether the dependencies do exist and, if not, creates them.
+Here, if `data/ìris_prepped` does not exist, Make creates it by executing the second target.
 Also, if one of the dependencies of a target is newer then the target itself, then Make updates everything that directly or indirectly depends on the target.
-Here, if `data\iris.csv` is newer then `data\iris-processed.csv`, Make will attempt to recreate `data\iris-processed.csv` first and then `report.pdf`.
+Here, if `data\iris.csv` is newer then `data\ìris_prepped`, Make will attempt to recreate `data\ìris_prepped` first and then `manuscript.pdf`.
 If there is a dependency missing, and there is no target to make it, Make stops with an error message.

### 2020-05-16 manuscript.Rmd
clarified section on singularity vs docker

@@
 
-This flexibility to fully control the software environment is of particular interest for software infrastructures where the user cannot install software themself, because of limited rights e.g.
-on cloud computing platforms or High-Performance Computing Clusters.
-However, executing docker itself needs special rights which are seldomly granted on High-Performance Computing Cluster.
-For this case, Singularity provides a fully compatible alternative (see Section under [Linux]).
+The flexibility to fully control the software environment is of particular interest for software infrastructures where users cannot install software because of limited access rights, for example, on cloud computing platforms or high-performance computing clusters.
+However, running docker needs unrestricted access rights to the system, which are rarely granted on those platforms.
+For this case, Singularity provides a fully compatible alternative (see Section under [Linux]) that can be executed with limited access rights.
 

### 2020-05-17 manuscript.Rmd
fix small typo

@@
 Another benefit of using Git and GitHub is that experimentation is highly encouraged since you can go back to any state quickly.
-Even when you lose access to the file on your computer, everything can be backed up on a remote git server.
+Even when you lose access to the file on your computer, everything can be backed up on a remote Git server (like GitHub or GitLab).
 Further, one can reduce the liklihood of dead code accumulating (e.g., lines that have been commented out) because it is safe to simply remove unneeded code blocks and track their removal in Git.
@@
 The flexibility to fully control the software environment is of particular interest for software infrastructures where users cannot install software because of limited access rights, for example, on cloud computing platforms or high-performance computing clusters.
-However, running docker needs unrestricted access rights to the system, which are rarely granted on those platforms.
+However, docker needs unrestricted access rights to the system, which are rarely granted on high-performance computing clusters.
 For this case, Singularity provides a fully compatible alternative (see Section under [Linux]) that can be executed with limited access rights.

### 2020-05-17 manuscript.Rmd
fix several typos

@@
 
-Even when you have obtained a given version of a project with the aim to reproduce reported result and you can confirm that this version is unchanged, you may not know exactly how to reproduce the results because it may be unclear which files should be executed in which order.
+Even when you have obtained a given version of a project with the aim to reproduce reported results and you can confirm that this version is unchanged, you may not know exactly how to reproduce the results because it may be unclear which files should be executed in which order.
 This is particularly the case when complex preprocessing pipelines are part of the computation and/or external dependencies (for example, a program needs to be called outside the R environment) are present.
@@
 Here, if `data/ìris_prepped` does not exist, Make creates it by executing the second target.
-Also, if one of the dependencies of a target is newer then the target itself, then Make updates everything that directly or indirectly depends on the target.
-Here, if `data\iris.csv` is newer then `data\ìris_prepped`, Make will attempt to recreate `data\ìris_prepped` first and then `manuscript.pdf`.
+Also, if one of the dependencies of a target is newer than the target itself, Make updates everything that directly or indirectly depends on the target.
+Here, if `data\iris.csv` is newer than `data\ìris_prepped`, Make will attempt to recreate `data\ìris_prepped` first and then `manuscript.pdf`.
 If there is a dependency missing, and there is no target to make it, Make stops with an error message.
 It is a convention to have the first target named `all`, which depends on everything.
-Then the command `make` without any argument automatically creates everything possible in the project.
+Subsequently, the command `make` without any argument automatically creates everything possible in the project.
 The button `Build All` from within RStudio triggers this process (see Figure \@ref{fig:buildpane-rstudio}).
@@
 
-If you execute the above on your system, there is a good chance that you cannot reproduce our manuscript and the `make all` command exits with an error.
+If you execute the above on your system, there is a good chance that you cannot reproduce our manuscript and the `make all` command results in an error.
 Successful reproducibility relies on the crucial assumption that your computational environments are identically or sufficiently compatible, e.g., all software dependencies are installed (R and all additional R Packages) and no updates or other changes to the computational environment break or alter your analysis.
@@
 
-With the following example, we demonstrate why it is important to store the computational environment.
-Generally, with containers we would like to safeguard against changes to the computational environment with unexpected consequences, e.g., changes in the functionality or default options in packages or even in the R environment itself.
+With the following example, we demonstrate the importance of storing computational environment.
+Generally, with containers we would like to safeguard against changes to the computational environment resulting in unexpected consequences, e.g., changes in the functionality or default options in packages or even in the R environment itself.
 While the R programming language is considered stable and much effort is put into backward compatibility, even basic functions like `sample()` (to randomly sample from a set) sometimes change their behavior from one version to another.
@@
 
-Executing the code in an image with a more recent version of R (3.6.1), this results in another sample despite the identical random seed:
+When executing the code in an image with a more recent version of R (3.6.1), the function results in another sample despite the identical random seed:
 
@@
 
-Note that this is intended behavior as it is the result of a [bugfix](https://bugs.r-project.org/bugzilla/show_bug.cgi?id=17494) in the random number generator implemented from R 3.6.0 upwards.
+Note, that this is intended behavior as it is the result of a [bugfix](https://bugs.r-project.org/bugzilla/show_bug.cgi?id=17494) in the random number generator implemented from R 3.6.0 upwards.
 Now, such changes may strictly render analyses run on previous R versions not reproducible if they contain, e.g., multiple imputations, bootstrapping, simulations studies, graphics with random jitter, Bayesian estimations using sampling algorithms (such as Markov Chain Monte Carlo), or similar techniques that involve random sampling.
@@
 We ran a linear regression model on a simulated dataset with two variables `x` and `y` with R's `lm()` function regressing `x` on `y`.
-Using the `boot` package [@boot], we bootstrapped the 95% confidence intervals around the regression coefficient estimate with 1000 bootstrap samples to evaluate whether zero was inside the confidence interval.
+Using the `boot` package [@boot], we bootstrapped the 95% confidence intervals around the regression coefficient estimate with 1000 bootstrap samples to evaluate whether the estimated confidence interval included zero.
 To make the analysis reproducible, we set a random seed.
@@
 
-Then, we ran the identical script with the identical seed in R 3.6.1:
+Subsequently, we ran the identical script with the identical seed in R 3.6.1:
 
@@
 
-As we see from these R outputs, the confidence intervals once include zero and once do not.
-Please note that one could discuss deeper issues about null hypothesis significance testing here but with this example, we would simply like to stress the point that computational reproducibility in the strict sense requires capturing the full computational environment.
+As we see from these R outputs, the latter of the estimated confidence intervals does include zero while the former does not.
+Please note that one could discuss deeper issues about null hypothesis significance testing here but with this example, we would simply like to stress that computational reproducibility in the strict sense requires capturing the full computational environment.
 
 Only rarely does an analysis depend on base R only.
-Typically, a considerable number of packages is required that each may depend on multiple other packages, and the addition of each package will increase the likelihood of breaking reproducibility with every update to any of the packages of base R (the resulting frustration is sometimes refered to as _dependency hell_).
+Typically, a considerable number of packages is required that each may depend on multiple other packages, and the addition of each package will increase the likelihood of breaking reproducibility with every update to any of the packages or base R (the resulting frustration is sometimes refered to as _dependency hell_).
 The whole endeavor of reproducibility is therefore at stake every time an update is rolled out.
 To ensure long-term reproducibility, Docker replicates the original computational environment of an analysis exactly.
-Note that we do not intend to advocate that software should not be updated; updates typically promote bugfixes and provide new functionality; our point is that full computational reproducibility is only achieved if the software versions used originally are precisely documented.
+Note, that we do not intend to advocate that software should not be updated; updates typically promote bugfixes and provide new functionality; our point is that full computational reproducibility is only achieved if the software versions used originally are precisely documented.
 Among other things, this makes it possible to trace back update histories to discover which change in which package caused the non-reproducibility.
@@
 
-The adress to connect to the RStudio server is your IP address (or `localhost` Linux) in this scheme: `<IPADRESS>:8787`.
+The adress to connect to the RStudio server is your IP address (or `localhost` on Linux) in this scheme: `<IPADRESS>:8787`.
 This offers a fully functioning RStudio instance that runs in the image but is accessible through a local web browser.
 
-By default, programs inside the container cannot access files on the local computer, thus requiring an explicit link to a local folder to enable access (and on macOS and Windows this also have to be allowed in the settings) :
+By default, programs inside the container cannot access files on the local computer, thus requiring an explicit link to a local folder to enable access (and on macOS and Windows this also has to be allowed in the settings) :
 
@@
 Fortunately, one can use Make to automatically generate the `docker` commands, e.g.
-the (simplified) `Makefile` for this paper allows command after `$(run)` to be conditionally passed through Docker if one types `make DOCKER=TRUE` (otherwise, they are run locally):
+the (simplified) `Makefile` for this paper allows the command after `$(run)` to be conditionally passed through Docker if one types `make DOCKER=TRUE` (otherwise, they are run locally):
 
@@
 For example, on macOS, Make is always available, whereas Linux systems are typically shipped with both Git and Make.
-In the following section, we share what what we consider the easiest way to install those packages across common operating systems.
+In the following section, we share what we consider the easiest way to install those packages across common operating systems.
 However, installation processes may be subject to change, and we advise that readers also consult the documentations of the packages or see our collection of links to tutorials and installation instructions on our GitHub repository.
@@
 Windows systems typically require the biggest efforts to install all necessary pieces of software.
-Note that you have to have either Windows Pro, Enterprise, Education, or Server installed, as Microsoft prevents the use of Docker on Windows Home.
+Note, that you must have either Windows Pro, Enterprise, Education, or Server installed, as Microsoft prevents the use of Docker on Windows Home.
 There is a package manager for Windows called Chocolatey, which you can install from: [https://chocolatey.org/install](https://chocolatey.org/install).
@@
 In the settings of Docker Desktop you have to allow the sharing of your drive.
-Docker on Windows require the path to be special; therefore you need to hand edit the `Makefile` and set `current_path` to the current directory and use `make all DOCKER=TRUE WINDOWS=TRUE`.
+Docker on Windows requires the path to be special; therefore you need to hand edit the `Makefile` and set `current_path` to the current directory and use `make all DOCKER=TRUE WINDOWS=TRUE`.
 We hope that future releases of Docker for Windows will not require that workaround.
@@
 
-To use `docker` you need to start Docker Desktop. In the settings of docker you have to allow the sharing of your drive.
+To use `docker`, you need to start Docker Desktop. In the settings of docker you have to allow the sharing of your drive.
 
@@
 The `here` package provides consistent paths relative to the project directory.
-The following are three examples of refering to local files ranging from absolute paths with system-specific path separators (bad) to relative paths using the `here` package:
+The following three examples refer to local files ranging from absolute paths with system-specific path separators (bad) to relative paths using the `here` package:
 
@@
 
-Sometimes external requirements make it impossible for the data to be stored and shared together with the scripts.
+Sometimes external requirements make it impossible for the data to be stored and shared with the scripts.
 In most of the cases we have seen, these are either space constraints or privacy considerations.
@@
 To use checksum validation, checksums for all data files must be created and stored at the time of the original analysis.
-At the time of reproduction, the current checksum must then be compared with the stored checksum to ensure data integrity.
+At the time of reproduction, the current checksum must be compared with the stored checksum to ensure data integrity.
 
@@
 That is, this reproduces a scientific report exactly without regard to the user, the operating system, the software, the timepoint, or interim changes to the involved files.
-To that end, the proposed workflow relies on tools that have been the foundation of reliable software development for years or, even decades.
+To that end, the proposed workflow relies on tools that have been the foundation of reliable software development for years or even decades.
 As a by-product, it makes transparent how statistical results depend on the software that created them and, by virtue of this transparency, facilitates later re-use by other researchers.
@@
 
-While our approach was designed to scale well with the complexity of a computing-intense project, we realise that this flexibility may not be straightforward to integrate into a researcher's everyday workflow.
+While our approach was designed to scale well with the complexity of a computationally intense project, we realise that this flexibility may not be straightforward to integrate into a researcher's everyday workflow.
 There are R packages that implement parts of our workflow and, thus, lower the threshold for adoption when the full flexibility provided by our workflow is not needed.
@@
 `holepunch` uses [binder](https://mybinder.org) [@binder2018] to move the analysis to the cloud, so that no local installation of Docker is required.
-`holepunch` is well suited for simple analysis with low computational demands, because binder's  memory and computing time is limited.
+`holepunch` is well suited for simple analyses with low computational demands, because binder's  memory and computing time is limited.
 There are several alternatives to Docker that manage dependencies on R packages.
@@
 Second, to ensure that other users are legally able to benefit from the shared materials, authors must choose an appropriate license format.
-Typically, there is currently no single license that works for code, data, and media (such as text or figures).
-We encourage authors to choose appropriate license forms that do not hinder others from freely downloading, using, and modifying the shared workflows and materials while, at the same time, ensuring recognition for the time and effort invested into creating the workflow in the first place.
-In our experience, the Creative Commons license (CC-BY) is often appropriate for sharing texts, R Markdown files, generated figures, and other media, whereas scripts and any other computer code are often best shared according to the MIT license.
+Typically, there is no single license that works for code, data, and media (such as text or figures).
+We encourage authors to choose appropriate license forms that do not hinder others from freely downloading, using, and modifying the shared workflows and materials while, at the same time, ensuring recognition for the time and effort invested in creating the workflow in the first place.
+In our experience, the Creative Commons license (CC-BY) is often appropriate for sharing texts, R Markdown files, generated figures, and other media, whereas scripts and any other computer code are often best shared under the MIT license.
 Both licenses assure maximal freedom for future users while requiring the attribution of the original authors in derivative work.
 These licenses are also in line with the recommendations by the Reproducible Research Standard [@stoddenEnhancingReproducibilityComputational2016; @stoddenEnablingReproducibleResearch2009].
-To facilitate an inclusive environment we recommend naming all contributors and including a [Code of Conduct](https://opensource.guide/code-of-conduct/) in your project.
+To facilitate an inclusive environment, we recommend naming all contributors and including a [Code of Conduct](https://opensource.guide/code-of-conduct/) in your project.
 

### 2020-05-17 manuscript.Rmd
close #31

@@
 
-Figure @\ref{fig:schematic} gives an overview of how the four components of our workflow interface to ensure computational reproducibility.
+Figure \ref{fig:schematic} gives an overview of how the four components of our workflow interface to ensure computational reproducibility.
 We begin with a minimal description of the roles of the four components.
@@
 Of particular note are the `papaja` package [@papaja2018], which offers additional functions to enable American Psychological Association (APA) style document formatting, including a journal-style final typeset format, and the `stargazer` package [@stargazer2018], which provides journal-ready tables and reports of statistical models.
-Figure \@ref{fig:source-rmarkdown} illustrates R Markdown syntax using the `papaja` package and Figure \@ref{fig:rendered-result} shows the resulting rendered document.
+Figure \ref{fig:source-rmarkdown} illustrates R Markdown syntax using the `papaja` package and Figure \ref{fig:rendered-result} shows the resulting rendered document.
 
@@
 RStudio also offers a graphical user interface for Git.
-For most basic operations, this interface is convenient and sufficient (see Figure \@ref{fig:gitpane}).
+For most basic operations, this interface is convenient and sufficient (see Figure \ref{fig:gitpane}).
 
@@
 
-`git push` or the green upward arrow in the Git pane (see Figure 3) uploads local updates. To download the remote Git repository on another computer, type into the terminal:
+`git push` or the green upward arrow in the Git pane (see Figure \ref{fig:gitpane}) uploads local updates. To download the remote Git repository on another computer, type into the terminal:
 
@@
 Subsequently, the command `make` without any argument automatically creates everything possible in the project.
-The button `Build All` from within RStudio triggers this process (see Figure \@ref{fig:buildpane-rstudio}).
+The button `Build All` from within RStudio triggers this process (see Figure \ref{fig:buildpane-rstudio}).
 
@@
 The `-it` flag tells Docker to run the command interactively, that is, to accept keyboard inputs and return outputs to the terminal.
-For example, this is the command to start an interactive R session inside a Docker image called `reproducible-research` (see Figure \@ref{fig:docker-r-terminal} for a screenshot):
+For example, this is the command to start an interactive R session inside a Docker image called `reproducible-research` (see Figure \ref{fig:docker-r-terminal} for a screenshot):
 
@@
 
-Figure \@ref{fig:rstudio-in-docker} shows a screenshot of Rstudio running inside Docker accessed from a local web browser.
+Figure \ref{fig:rstudio-in-docker} shows a screenshot of Rstudio running inside Docker accessed from a local web browser.
 

### 2020-05-18 manuscript.Rmd

@@
 We encourage authors to choose appropriate license forms that do not hinder others from freely downloading, using, and modifying the shared workflows and materials while, at the same time, ensuring recognition for the time and effort invested in creating the workflow in the first place.
-In our experience, the Creative Commons license (CC-BY) is often appropriate for sharing texts, R Markdown files, generated figures, and other media, whereas scripts and any other computer code are often best shared under the MIT license.
+In our experience, the [Creative Commons license (CC-BY)](https://creativecommons.org/licenses/by/4.0/) is often appropriate for sharing texts, R Markdown files, generated figures, and other media, whereas scripts and any other computer code are often best shared under the [MIT license](https://opensource.org/licenses/MIT).
 Both licenses assure maximal freedom for future users while requiring the attribution of the original authors in derivative work.
 These licenses are also in line with the recommendations by the Reproducible Research Standard [@stoddenEnhancingReproducibilityComputational2016; @stoddenEnablingReproducibleResearch2009].
+A great resource to choose a license is [choosealicense.com](https://choosealicense.com), however, no resource, including our recommendation, replaces legal advice.
 To facilitate an inclusive environment, we recommend naming all contributors and including a [Code of Conduct](https://opensource.guide/code-of-conduct/) in your project.

### 2020-05-18 manuscript.Rmd
- added all target to Makefile - added .csv to iris_prepped

--- 2020-05-17 manuscript.Rmd
@@
 It contains a list of recipes.
-Each recipe has a target (the name of the recipe) followed by a list of dependent targets or files.
+Each recipe has a target (the name of the recipe) followed by a colon and a list of dependent targets or files.
 If any of the dependencies have changed since the last time the target was built, the recipe's commands are executed to recreate the target file.
@@
 # -- this is a Makefile --
+all: manuscript.pdf
 
-manuscript.pdf: data/ìris_prepped manuscript.Rmd
+manuscript.pdf: data/ìris_prepped.csv manuscript.Rmd
   Rscript -e 'rmarkdown::render("manuscript.Rmd")'
 
-data/ìris_prepped: R/prepare_data.R data/iris.csv
+data/ìris_prepped.csv: R/prepare_data.R data/iris.csv
   Rscript -e 'source("R/prepare_data.R")'
@@
 
-The first line can be interpreted as follows: To create `manuscript.pdf` (the first target in the file), the file `manuscript.Rmd`, which depends on `data/ìris_prepped`, needs to be rendered.
-This dependency is itself a target (the second target in the file).
-To create `data/ìris_prepped`, `R/prepare_data.R` and `data/iris.csv` are needed.
+The first line after the comment is the first (default) target called 'all', which depends on manuscript.pdf, which itself is a target.
+If Make is called without an argument, the first target is built.
+To create `manuscript.pdf` (the second target in the file), the file `manuscript.Rmd` needs to be rendered, which depends on `data/ìris_prepped.csv`.
+This dependency is itself a target (the third target in the file).
+To create `data/ìris_prepped.csv`, `R/prepare_data.R` and `data/iris.csv` are needed.
 If you type `make manuscript.pdf`, Make first checks whether the dependencies do exist and, if not, creates them.
-Here, if `data/ìris_prepped` does not exist, Make creates it by executing the second target.
+Here, if `data/ìris_prepped.csv` does not exist, Make creates it by executing the third target.
 Also, if one of the dependencies of a target is newer than the target itself, Make updates everything that directly or indirectly depends on the target.
-Here, if `data\iris.csv` is newer than `data\ìris_prepped`, Make will attempt to recreate `data\ìris_prepped` first and then `manuscript.pdf`.
+Here, if `data\iris.csv` is newer than `data\ìris_prepped.csv`, Make will attempt to recreate `data\ìris_prepped.csv` first and then `manuscript.pdf`.
 If there is a dependency missing, and there is no target to make it, Make stops with an error message.

### 2020-05-18 manuscript.Rmd
run grammerly closes #16 #36

@@
 In this tutorial, we describe a workflow to ensure long-term and cross-platform reproducibility of data analyses in R [@baseR].
-Reproducibility is the ability to draw identical conclusion from the same statistical analysis and the same data.
-In practice, reproducibility means to us that statistical results as well as their generating computational workflows are reported fully, and transparently, and remain sustainably available, such that a given workflow can be re-run by a different person, later in time, and that the results will be identical to the ones intially reported [@turingway2020; @herouxCompatibleReproducibilityTaxonomy2018; @claerboutElectronicDocumentsGive1992].
+Reproducibility is the ability to draw an identical conclusion from the same statistical analysis and the same data.
+In practice, reproducibility means to us that statistical results as well as their generating computational workflows are reported fully, and transparently, and remain sustainably available, such that a given workflow can be re-run by a different person, later in time, and that the results will be identical to the ones initially reported [@turingway2020; @herouxCompatibleReproducibilityTaxonomy2018; @claerboutElectronicDocumentsGive1992].
 The need to ensure reproducibility directly follows from commonly accepted rules of good scientific practice [such as the guidelines of the German Research Foundation, @dfg2019].
-Ensuring reproducibility is a prerequisite for replicability (the ability to reach a similar conclusion from the same analysis and new data), and a means to increase trustworthiness of empirical results [@epskamp2019rep].
-Transparency and accessibility are  central scientific values and open, reproducible projects will increase the efficiency and veracity of knowledge accumulation [@nosek2012scientific].
+Ensuring reproducibility is a prerequisite for replicability (the ability to reach a similar conclusion from the same analysis and new data), and a means to increase the trustworthiness of empirical results [@epskamp2019rep].
+Transparency and accessibility are central scientific values, and open, reproducible projects will increase the efficiency and veracity of knowledge accumulation [@nosek2012scientific].
 
@@
 They all are highly flexible and powerful, so that mastering them fully, requires a considerable amount of practice.
-However, for our purposes, it is sufficient to master a valuable minimal subset of functions to ensure reproducibility of scientific analyses.
+However, for our purposes, it is sufficient to master a valuable minimal subset of functions to ensure the reproducibility of scientific analyses.
 
@@
 Typical errors result from copy\&paste mistakes, erroneous rounding, or missed updates of the manuscript when the computer code has changed.
-In order to not only create fully reproducible results but also fully reproducible reports, one needs to resort to the Literate Programming paradigm [@knuth1984literate], in which human-readable language and computer code are intermingled to create dynamic documents whose order follows the logic of thought rather then the order of the computer.
+In order to create not only fully reproducible results but also fully reproducible reports, one needs to resort to the Literate Programming paradigm [@knuth1984literate], in which human-readable language and computer code are intermingled to create dynamic documents whose order follows the logic of thought rather then the order of the computer.
 R Markdown is a simple markup language to create dynamic documents with embedded chunks of R code that can be exported to standard formats such as documents (docx, rtf, epub), presentations (ppt, html) or websites (html) using the `knitr` package [@xie2015; @xie2019].
@@
 # -- type this in a terminal --
-git init # to initialize git in the current directory
+git init # to initialize Git in the current directory
 git add ./data/iris.csv ./R/analysis.R # track specific files
@@
 
-Even when you have obtained a given version of a project with the aim to reproduce reported results and you can confirm that this version is unchanged, you may not know exactly how to reproduce the results because it may be unclear which files should be executed in which order.
+Even when you have obtained a given version of a project with the aim to reproduce reported results, and you can confirm that this version is unchanged, you may not know exactly how to reproduce the results because it may be unclear which files should be executed in which order.
 This is particularly the case when complex preprocessing pipelines are part of the computation and/or external dependencies (for example, a program needs to be called outside the R environment) are present.
@@
 Docker is a tool that allows encapsulation, sharing, and re-creation of a computational environment on most operating systems (Windows, macOS & Linux).
-Docker achieves these goals by setting up a virtual computer, on which it can execute well-defined series of commands (e.g., installing software).
+Docker achieves these goals by setting up a virtual computer, on which it can execute a well-defined series of commands (e.g., installing software).
 It then saves the resulting state of the virtual computer in what is called an "image." This image can be started and execute commands on the virtual computer, e.g., running `Rscript` or `make`.
@@
 Irrespective of the machine that is executing the container, the computational environment is the same for the programs running inside the container.
-The most important advantage over traditional virtual machines is that containers are lightweight, that is, they start rapidly and do not need much storage space.
+The most important advantage over traditional virtual machines is that containers are lightweight; that is, they start rapidly and do not need much storage space.
 Docker achieves this by reusing large parts of the host's operating system (the extend varies between Linux, macOS and Windows).
@@
 With the following example, we demonstrate the importance of storing computational environment.
-Generally, with containers we would like to safeguard against changes to the computational environment resulting in unexpected consequences, e.g., changes in the functionality or default options in packages or even in the R environment itself.
-While the R programming language is considered stable and much effort is put into backward compatibility, even basic functions like `sample()` (to randomly sample from a set) sometimes change their behavior from one version to another.
+Generally, with containers, we would like to safeguard against changes to the computational environment resulting in unexpected consequences, e.g., changes in the functionality or default options in packages or even in the R environment itself.
+While the R programming language is considered stable and much effort is put into backward compatibility, even basic functions like `sample()` (to randomly sample from a set) sometimes change their behaviour from one version to another.
 To ensure reproducibility in analyses based on a computer's pseudo random number generator (PRNG), it is good practice to rely on fixed PRNG seeds, which are numeric values that set the PRNG into a deterministic state, i.e., the sequence of random numbers reproduces exactly.
@@
 
-Note, that this is intended behavior as it is the result of a [bugfix](https://bugs.r-project.org/bugzilla/show_bug.cgi?id=17494) in the random number generator implemented from R 3.6.0 upwards.
+Note, that this is intended behaviour as it is the result of a [bugfix](https://bugs.r-project.org/bugzilla/show_bug.cgi?id=17494) in the random number generator implemented from R 3.6.0 upwards.
 Now, such changes may strictly render analyses run on previous R versions not reproducible if they contain, e.g., multiple imputations, bootstrapping, simulations studies, graphics with random jitter, Bayesian estimations using sampling algorithms (such as Markov Chain Monte Carlo), or similar techniques that involve random sampling.
@@
 As we see from these R outputs, the latter of the estimated confidence intervals does include zero while the former does not.
-Please note that one could discuss deeper issues about null hypothesis significance testing here but with this example, we would simply like to stress that computational reproducibility in the strict sense requires capturing the full computational environment.
+Please note that one could discuss deeper issues about null hypothesis significance testing here, but with this example, we would simply like to stress that computational reproducibility in the strict sense requires capturing the full computational environment.
 
 Only rarely does an analysis depend on base R only.
-Typically, a considerable number of packages is required that each may depend on multiple other packages, and the addition of each package will increase the likelihood of breaking reproducibility with every update to any of the packages or base R (the resulting frustration is sometimes refered to as _dependency hell_).
-The whole endeavor of reproducibility is therefore at stake every time an update is rolled out.
+Typically, a considerable number of packages is required that each may depend on multiple other packages, and the addition of each package will increase the likelihood of breaking reproducibility with every update to any of the packages or base R (the resulting frustration is sometimes referred to as _dependency hell_).
+The whole endeavour of reproducibility is therefore at stake every time an update is rolled out.
 To ensure long-term reproducibility, Docker replicates the original computational environment of an analysis exactly.
@@
 The flexibility to fully control the software environment is of particular interest for software infrastructures where users cannot install software because of limited access rights, for example, on cloud computing platforms or high-performance computing clusters.
-However, docker needs unrestricted access rights to the system, which are rarely granted on high-performance computing clusters.
+However, Docker needs unrestricted access rights to the system, which are rarely granted on high-performance computing clusters.
 For this case, Singularity provides a fully compatible alternative (see Section under [Linux]) that can be executed with limited access rights.
 
-There are two ways to share a Docker image; either by sharing the Dockerfile that creates the image, or by sharing the image itself, e.g., through a service like [Dockerhub](https://hub.docker.com).
+There are two ways to share a Docker image; either by sharing the Dockerfile that creates the image or by sharing the image itself, e.g., through a service like [Dockerhub](https://hub.docker.com).
 While both ways guarantee a replicable computational environment, sharing the Dockerfile is more transparent and more space-saving; in our workflow, we can use Git to track changes in the Dockerfile (such as updates to dependencies).
@@
 
-The second option is to start the container in the background and to interact with the container via the webrowser and the RStudio server instance running in it.
+The second option is to start the container in the background and to interact with the container via the web browser and the RStudio server instance running in it.
 In order to do so, you need to supply a password to log into the RStudio server (`-e PASSWORD=<YOUR_PASS>`) and open a network service on a specified port (`-p 8787:8787`).
@@
 
-The adress to connect to the RStudio server is your IP address (or `localhost` on Linux) in this scheme: `<IPADRESS>:8787`.
+The address to connect to the RStudio server is your IP address (or `localhost` on Linux) in this scheme: `<IPADRESS>:8787`.
 This offers a fully functioning RStudio instance that runs in the image but is accessible through a local web browser.
@@
 
-Since Docker commands tend to grow long and become tedious to type manually, we recomend using some automatic way to generate them.
+Since Docker commands tend to grow long and become tedious to type manually, we recommend using some automatic way to generate them.
 Fortunately, one can use Make to automatically generate the `docker` commands, e.g.
@@
 There is a package manager for Windows called Chocolatey, which you can install from: [https://chocolatey.org/install](https://chocolatey.org/install).
-Chocolatey provides all software packages needed for our workflow in one place..
+Chocolatey provides all software packages needed for our workflow in one place.
 Having installed Chocolatey (and restarted the computer), all dependencies can be installed in an **admin terminal** (Windows key, then type `cmd`, right-click *Run as administrator*) via:
@@
 
-# install docker, make, and git using choco
+# install Docker, Make, and Git using Chocolatey
 choco install -y git make docker-desktop
@@
 To use `docker` you need to start Docker Desktop.
-In the settings of Docker Desktop you have to allow the sharing of your drive.
-Docker on Windows requires the path to be special; therefore you need to hand edit the `Makefile` and set `current_path` to the current directory and use `make all DOCKER=TRUE WINDOWS=TRUE`.
+In the settings of Docker Desktop, you have to allow the sharing of your drive.
+Docker on Windows requires the path to be special; therefore, you need to hand-edit the `Makefile` and set `current_path` to the current directory and use `make all DOCKER=TRUE WINDOWS=TRUE`.
 We hope that future releases of Docker for Windows will not require that workaround.
@@
 
-# install docker via brew
+# install Docker via Homebrew
 brew cask docker
@@
 
-To use `docker`, you need to start Docker Desktop. In the settings of docker you have to allow the sharing of your drive.
+To use `docker`, you need to start Docker Desktop. In the settings of Docker you have to allow the sharing of your drive.
 
@@
 
-# install docker via advanced package tool
+# install Docker via advanced package tool
 apt install git make docker
@@
 For other distributions, replace `apt` with the native package manager.
-You may need elevated rights for the installation; in this case add `sudo` before the installation command.
+You may need elevated rights for the installation; in this case, add `sudo` before the installation command.
 `docker` also needs elevated rights to run; therefore, we recommend adding the local user to the `docker` group, following the [documentation of Docker](https://docs.docker.com/install/linux/linux-postinstall/#manage-docker-as-a-non-root-user).
@@
 An alternative to Docker on Linux is Singularity [@kurtzerSingularityScientificContainers2017].
-To use it, just replace any docker calls with `singularity docker`, because singularity fully supports docker images.
-A possible advantage is that singularity works well in high-performance computing environments and on old Linux versions, the downside is that Singularity is currently only available on Linux.
+To use it, just replace any docker calls with `singularity docker` because Singularity fully supports docker images.
+A possible advantage is that Singularity works well in high-performance computing environments and on old Linux versions, the downside is that Singularity is currently only available on Linux.
 
@@
 Finally, we conclude with some notes on project organization, which we think makes migrating projects to a reproducible workflow easier.
-The first step towards reproducibility is to create an R script or R Markdown file as primary entry point for the analysis that runs on a local computer without error and performs the main statistical analyses.
-Next, one needs to make sure that all files relevant for the analysis can be moved to another computer.
+The first step towards reproducibility is to create an R script or R Markdown file as the primary entry point for the analysis that runs on a local computer without error and performs the main statistical analyses.
+Next, one needs to make sure that all files relevant to the analysis can be moved to another computer.
 To this end, it is recommendable that all files are within one folder (or enclosed subfolders within it) and all paths are relative to that folder because absolute paths are specific to a given computer.
@@
 Second, anchor points of relative paths may differ depending on context.
-For example, knitr interprets paths relative to the dynamic document whereas R has a current working directory that may change over the course of an R ssession.
+For example, knitr interprets paths relative to the dynamic document, whereas R has a current working directory that may change over the course of an R session.
 The `here` package provides consistent paths relative to the project directory.
@@
 If dividing data and scripts is unavoidable, we recommend validating all data files using checksums (also called a "hash", e.g., using the functions provided in package `digest` [@digest2019]) before analyzing them.
-A checksum is a short fixed-length fingerprint (often displayed in hexadecimal system) of a file with the purpose of verifying integrity of a digital object.
-Fingerprints are computed from digital objects such that they change with high probability if data is changed only little.
+A checksum is a short fixed-length fingerprint (often displayed in the hexadecimal system) of a file with the purpose of verifying the integrity of a digital object.
+Fingerprints are computed from digital objects such that they change with high probability if data is changed only a little.
 To use checksum validation, checksums for all data files must be created and stored at the time of the original analysis.
@@
 
-The overarching goal of this paper was to provide a complete and easy-to-use workflow that allows confidence in reproducibility of R-based data analyses.
+The overarching goal of this paper was to provide a complete and easy-to-use workflow that allows confidence in the reproducibility of R-based data analyses.
 Analyses following our workflow can be reproduced with four commands (here shown for our paper):
@@
 To that end, the proposed workflow relies on tools that have been the foundation of reliable software development for years or even decades.
-As a by-product, it makes transparent how statistical results depend on the software that created them and, by virtue of this transparency, facilitates later re-use by other researchers.
+As a by-product, it makes transparent how statistical results depend on the software that created them and, by virtue of this transparency, facilitates later reuse by other researchers.
 
@@
 
-While our approach was designed to scale well with the complexity of a computationally intense project, we realise that this flexibility may not be straightforward to integrate into a researcher's everyday workflow.
+While our approach was designed to scale well with the complexity of a computationally intense project, we realize that this flexibility may not be straightforward to integrate into a researcher's everyday workflow.
 There are R packages that implement parts of our workflow and, thus, lower the threshold for adoption when the full flexibility provided by our workflow is not needed.
@@
 The [liftR ](https://cran.r-project.org/web/packages/liftr/vignettes/liftr-intro.html) package [@liftr] and the [`holepunch`](https://karthik.github.io/holepunch/) package [@holepunch] automatize the use of Docker.
-The former is perfectly compatible with the described workflow and we recommend it to users  who are not comfortable with command-line use of Docker.
+The former is perfectly compatible with the described workflow, and we recommend it to users who are not comfortable with command-line use of Docker.
 `holepunch` uses [binder](https://mybinder.org) [@binder2018] to move the analysis to the cloud, so that no local installation of Docker is required.
-`holepunch` is well suited for simple analyses with low computational demands, because binder's  memory and computing time is limited.
+`holepunch` is well suited for simple analyses with low computational demands because binder's memory and computing time is limited.
 There are several alternatives to Docker that manage dependencies on R packages.
-[`renv`](https://rstudio.github.io/renv/articles/renv.html) [@renv] is a way to freeze package version via local copies of packages in the project, but it does not gurantee a given base R version or system dependencies.
+[`renv`](https://rstudio.github.io/renv/articles/renv.html) [@renv] is a way to freeze package version via local copies of packages in the project, but it does not guarantee a given base R version or system dependencies.
 Similar approaches are taken by [`jetpack`](https://github.com/ankane/jetpack) [@jetpack], [`miniCRAN`](https://cran.r-project.org/web/packages/miniCRAN/vignettes/miniCRAN-introduction.html) [@miniCRAN] and [`checkpoint`](https://cran.r-project.org/web/packages/checkpoint/vignettes/checkpoint.html) [@checkpoint].
 The package [`reprex`](https://github.com/tidyverse/reprex) [**repr**oducible **ex**ample, @reprex] is also worth noting, but its scope is limited.
-A particularly noteworthy approach is provided by the [`worcs`](https://cran.r-project.org/web/packages/worcs) package [@worcs, @worcspaper], which is an R project template that creates a standardized file structure for code and data supporting version management with git, package management with renv and dynamic document generation with R Markdown.
+A particularly noteworthy approach is provided by the [`worcs`](https://cran.r-project.org/web/packages/worcs) package [@worcs, @worcspaper], which is an R project template that creates a standardized file structure for code and data supporting version management with Git, package management with renv and dynamic document generation with R Markdown.
 We acknowledge that worcs is much easier to install and provides a one-click solution for the creation of reproducible projects.
-It achieves a high standard of reproducibility but does not guarantuee full computational reproducibility and is limited to dependency management within the R environment.
+It achieves a high standard of reproducibility but does not guarantee full computational reproducibility and is limited to dependency management within the R environment.
 
-Other than these tools, which ease the process of creating workflows like ours does, we have noticed increased interest in changing the way research is published and used [@perkelToolkitDataTransparency2018], with the emergence of *life code* [@perkelPioneeringLivecodeArticle2019] and *continuous integration* [@beaulieu-jonesReproducibilityComputationalWorkflows2017; @yenniDevelopingModernData2018].
+Other than these tools, which ease the process of creating workflows like ours does, we have noticed an increased interest in changing the way research is published and used [@perkelToolkitDataTransparency2018], with the emergence of *life code* [@perkelPioneeringLivecodeArticle2019] and *continuous integration* [@beaulieu-jonesReproducibilityComputationalWorkflows2017; @yenniDevelopingModernData2018].
 These techniques give us a glimpse of a paradigm shift from static to dynamic, interactive, and living publications that is yet to happen.
@@
 
-We are aware that implementing the proposed workflow is not straightforward and the difficulty of its implementation may vary by platform.
+We are aware that implementing the proposed workflow is not straightforward, and the difficulty of its implementation may vary by platform.
 For example, the installation of all tools is already easier on POSIX-compatible platforms such as Unix, Linux, or macOS (but not Windows).
@@
 That is, we have experienced the case that after writing up the first draft in R Markdown, we eventually had to generate a Word file that, from then on, was used as static file serving as a basis for multiple iterations among the co-authors.
-Retaining reproduciblity in such situations then requires tedious manual synchronization of files across formats.
+Retaining reproducibility in such situations then requires tedious manual synchronization of files across formats.
 This annoyance may be reduced with the [`redoc` package](https://noamross.github.io/redoc/articles/mixed-workflows-with-redoc.html) [@redoc], which enables a bidirectional synchronization between Word and R Markdown.
@@
 The proposed workflow leverages various existing tools that are partly integrated into RStudio already.
-Parts of the proposed workflow have been integrated into stand-alone packages (such as worcs [@worcs], workflowr [@workflowr], holepunch  [@holepunch]), which are particularly recommenable to beginners but they are either incomplete in the sense of our proposal or rely on proprietary service providers.
-We hope that with increasing awareness of the challenges of computational reproducibility, the increased demand for unified and open solutions will lead to an even better integration of existing tools to allow reproducible workflows to become a standard in psychological research.
+Parts of the proposed workflow have been integrated into stand-alone packages (such as worcs [@worcs], workflowr [@workflowr], holepunch  [@holepunch]), which are particularly recommendable to beginners but they are either incomplete in the sense of our proposal or rely on proprietary service providers.
+We hope that with increasing awareness of the challenges of computational reproducibility, the increased demand for unified and open solutions will lead to better integration of existing tools to allow reproducible workflows to become a standard in psychological research.
 

### 2020-05-18 abstract.Rmd
small style changes at abstract and introduction

--- 2019-11-11 abstract.Rmd
@@
 It combines the benefits of various open-source software tools including R Markdown, Git, Make, and Docker, whose interplay ensures seamless integration of version management, dynamic report generation conforming to various journal styles and full cross-platform and long-term computational reproducibility.
-The workflow ensures meeting the primary goals that 1) the reporting of statistical results is consistent with the actual statistical results (dynamic report generation), the analysis exactly reproduces at a later time even if the computing platform or software is changed (computational reproducibility), and 3) changes at any time (during development and post-publication) are tracked, tagged, and documented while earlier versions of both data and code remain accessible.
+The workflow ensures meeting the primary goals that 1) the reporting of statistical results is consistent with the actual statistical results (dynamic report generation), 2) the analysis exactly reproduces at a later point in time even if the computing platform or software is changed (computational reproducibility), and 3) changes at any time (during development and post-publication) are tracked, tagged, and documented while earlier versions of both data and code remain accessible.
 While the research community increasingly recognizes dynamic document generation and version management as tools to ensure reproducibility, we demonstrate with practical examples that these alone are not sufficient to ensure long-term computational reproducibility.

### 2020-05-18 manuscript.Rmd
small style changes at abstract and introduction

@@
 Reproducibility is the ability to draw an identical conclusion from the same statistical analysis and the same data.
-In practice, reproducibility means to us that statistical results as well as their generating computational workflows are reported fully, and transparently, and remain sustainably available, such that a given workflow can be re-run by a different person, later in time, and that the results will be identical to the ones initially reported [@turingway2020; @herouxCompatibleReproducibilityTaxonomy2018; @claerboutElectronicDocumentsGive1992].
+For us a statstical result is only reproducible if their generating computational workflows are reported fully, and transparently, and remain sustainably available, such that a given workflow can be re-run by a different person, later in time, and that the results will be identical to the ones initially reported [@turingway2020; @herouxCompatibleReproducibilityTaxonomy2018; @claerboutElectronicDocumentsGive1992].
 The need to ensure reproducibility directly follows from commonly accepted rules of good scientific practice [such as the guidelines of the German Research Foundation, @dfg2019].

### 2020-05-19 abstract.Rmd
replaced literate programming

@@
 While the research community increasingly recognizes dynamic document generation and version management as tools to ensure reproducibility, we demonstrate with practical examples that these alone are not sufficient to ensure long-term computational reproducibility.
-Leveraging containerization, dependence management, version management, and literate programming, the workflow increases scientific productivity by facilitating later reproducibility and reuse of code and data.
+Leveraging containerization, dependence management, version management, and dynamic document generation, the workflow increases scientific productivity by facilitating later reproducibility and reuse of code and data.

### 2020-05-19 abstract.Rmd
revision for resubmission

--- 2020-05-18 abstract.Rmd
@@
 While the research community increasingly recognizes dynamic document generation and version management as tools to ensure reproducibility, we demonstrate with practical examples that these alone are not sufficient to ensure long-term computational reproducibility.
-Leveraging containerization, dependence management, version management, and literate programming, the workflow increases scientific productivity by facilitating later reproducibility and reuse of code and data.
+Combining containerization, dependence management, version management, and dynamic document generation, the proposed workflow increases scientific productivity by facilitating later reproducibility and reuse of code and data.

### 2020-05-19 manuscript.Rmd
changed to "thie document is based on" because it is likely modified until final print.
revision for resubmission

@@
 All materials can be found on the accompanying GitHub repository: [https://github.com/aaronpeikert/reproducible-research/](https://github.com/aaronpeikert/reproducible-research/).
-This particular document was created from the commit with the hash `"`r system2("git", "rev-parse --short HEAD", stdout = TRUE)`"`.
+This document is based on the commit identified by hash `"`r system2("git", "rev-parse --short HEAD", stdout = TRUE)`"`.
 To provide feedback or any other type of comment or questions regarding our workflow, please add an issue to the GitHub repository of our paper at [https://github.com/aaronpeikert/reproducible-research/issues](https://github.com/aaronpeikert/reproducible-research/issues).
@@
 Reproducibility is the ability to draw an identical conclusion from the same statistical analysis and the same data.
-For us a statstical result is only reproducible if their generating computational workflows are reported fully, and transparently, and remain sustainably available, such that a given workflow can be re-run by a different person, later in time, and that the results will be identical to the ones initially reported [@turingway2020; @herouxCompatibleReproducibilityTaxonomy2018; @claerboutElectronicDocumentsGive1992].
-The need to ensure reproducibility directly follows from commonly accepted rules of good scientific practice [such as the guidelines of the German Research Foundation, @dfg2019].
-Ensuring reproducibility is a prerequisite for replicability (the ability to reach a similar conclusion from the same analysis and new data), and a means to increase the trustworthiness of empirical results [@epskamp2019rep].
+For us,  statistical results are only reproducible if their generating, computational workflow is reported fully and transparently, and remains sustainably available, such that the workflow can be re-run by a different person or later in time, and that the results will be identical to the ones initially reported [@turingway2020; @herouxCompatibleReproducibilityTaxonomy2018; @claerboutElectronicDocumentsGive1992].
+The need to ensure reproducibility directly follows from commonly accepted rules of good scientific practice [such as the guidelines of the German Research Foundation; @dfg2019].
+Ensuring reproducibility is a prerequisite for replicability (the ability to reach a similar conclusion from the same analysis and _new_ data), and a means to increase the trustworthiness of empirical results [@epskamp2019rep].
 Transparency and accessibility are central scientific values, and open, reproducible projects will increase the efficiency and veracity of knowledge accumulation [@nosek2012scientific].
@@
 We argue that only a workflow using all four concepts in unison can guarantee confidence in reproducing a scientific report [see @turingway2020 for similar arguments].
-Various implementations of these concepts exist, but we consider the following four best suited for the R environment [@r2019]: for dynamic document generation, R Markdown [@rmarkdown2019], for version control, Git [@chaconProGit2014], for dependency management, Make [@feldmanMakeProgramMaintaining1979], and for containerization, Docker [@merkelDockerLightweightLinux2014].
+Various implementations of these concepts exist, but we consider the following four best suited for analyses centered on the R environment [@r2019] but also allowing for external dependencies: R Markdown [@rmarkdown2019] for dynamic document generation, Git [@chaconProGit2014] for version control, Make [@feldmanMakeProgramMaintaining1979] for dependency management, and Docker [@merkelDockerLightweightLinux2014] for containerization.
 Each of these software solutions serves a valuable meta-scientific goal (reproducibility) and increases the researches' productivity.
@@
 
-<!--To maximize chances of long-term reproducibility, proper documentation of all steps taken and their context is crucial since it answers the question "What steps have to be taken to reproduce the results?".
-The most precise documentation is computer code, and the most productive computer code is code that runs automatically without human interaction.
-This principle is what guided our use of R Markdown, Git, Make, and Docker.-->
-
 # The Reproducible Workflow in a Nutshell
 
-```{r schematic, eval = TRUE, echo = FALSE, fig.cap="Schematic illustration of the interplay of the four components central to a reproducible workflow: version control (git), dependency management (Make), containerization (Docker), and dynamic document generation (R Markdown).
+```{r schematic, eval = TRUE, echo = FALSE, fig.cap="Schematic illustration of the interplay of the four components (in dashed boxes) central to the reproducible workflow: version control (git), dependency management (Make), containerization (Docker), and dynamic document generation (R Markdown).
+Git tracks changes to the project over time.
+Make manages dependencies among the files.
+Docker provides a container, in which the final report is built using dynamic document generation in R Markdown.
 "}
@@
 
-Figure \ref{fig:schematic} gives an overview of how the four components of our workflow interface to ensure computational reproducibility.
-We begin with a minimal description of the roles of the four components.
+Figure \ref{fig:schematic} gives an overview of how the four components of our workflow interact to ensure computational reproducibility.
+Before we describe the four components in more detail, we begin with a minimal description of the roles of each component.
 In the remainder of this tutorial, we will further detail each of the four components of our workflow.
 The first component is version control.
-Version control manages  changes to files (e.g., data and code) over time so that you can access specific versions later.
+Version control manages changes to files (e.g., data and code) over time so that you can recall specific versions of files later or revert the entire project to a past state.
 Version control offers snapshots of your workflow at different time points identified by a unique identifier.
-The different frames in the z-axis of the image illustrate the different versions of the workflow.
-How different parts of a analysis and a report of it relate and in what order they need to be executed is monitored using dependency management.
+The different frames in the z-axis of Figure  \ref{fig:schematic} illustrate the different versions of the project.
+How different parts of a analysis and a report of it relate and in what order they need to be executed is documented using dependency management.
 The arrows visualize dependencies, such as an analysis depending on the availability of a particular data file.
-Third, all computer code (such as a statistical analysis in R) is executed in a virtual environment that guarantuees exact reproduction of results independent of the host operating system, R version, and installed package versions.
-Finally, dynamic document generation (also known as the literate programming paradigm) interweaves human-readable code and automatically produced results (such as point estimates, p values, or confidence intervals) to eliminate inconsistency errors such as those arising from careless copy and paste actions.
+Third, all computer code (such as a statistical analyses in R) is executed in a virtual environment that guarantuees exact reproduction of results independent of the host operating system, the locally installed R version, and installed package versions.
+Finally, dynamic document generation (also known as the literate programming paradigm) interweaves human-readable code and computed results (such as point estimates, p values, or confidence intervals) to eliminate inconsistency errors such as those arising from copy\&paste errors.
 
@@
 The translation of computational results into a human-readable summary, for example into a technical report, a presentation, or a journal manuscript, is time-consuming and error-prone.
-Typical errors result from copy\&paste mistakes, erroneous rounding, or missed updates of the manuscript when the computer code has changed.
-In order to create not only fully reproducible results but also fully reproducible reports, one needs to resort to the Literate Programming paradigm [@knuth1984literate], in which human-readable language and computer code are intermingled to create dynamic documents whose order follows the logic of thought rather then the order of the computer.
+Typical errors result from copy\&paste mistakes, erroneous rounding, or missed updates of the manuscript when the associated computer code and computed results have changed.
+In order to create not only fully reproducible results but also fully reproducible reports, we resort to the literate programming paradigm [@knuth1984literate], in which human-readable language and computer code are mixed to create dynamic documents whose order follows the logic of thought rather then the order of the computer.
 R Markdown is a simple markup language to create dynamic documents with embedded chunks of R code that can be exported to standard formats such as documents (docx, rtf, epub), presentations (ppt, html) or websites (html) using the `knitr` package [@xie2015; @xie2019].
@@
 
-# Version Management
+# Version Control
 
@@
 Each commit has a short, unique identifier (a hash code) and a human-readable description (commit message).
-Going back to one state is as easy as traversing the history of all commits and switching the repository to a given previous state; it is possible to visually track changes between different versions (similar to the track changes function offered in Microsoft Word).
+Going back to one state is as easy as traversing the history of all commits and switching the repository to a given previous state; it is possible to visually compare changes between different versions.
 The collection of all snapshots is called a "repository," which ideally tracks your entire R project.
@@
 # once script or data were changed, take a new snapshot
-git commit -a -m "complete data colections" # add (-a) and commit all changes
+git commit -a -m "complete data colections" # add and commit all changes
 ```
@@
 To keep track of all changes on your local computer, you only need to use `git add` and `git commit` or `git commit -a` to add and commit at the same time.
+Adding a file means to save its changes on the next commit.
 These commands need to be executed in the terminal, which you can access from within RStudio (`Shift + Alt + R`).
@@
 # -- type this in a terminal --
-git log # inspect all changes
-git checkout 77db06f78e # revert local directory to previous version with hash code '77db06f78e'
+
+# inspect all changes
+git log
+# revert local directory to previous version with hash '77db06f78e'
+git checkout 77db06f78e 
 ```
@@
 Alternatively, institutions can host an equally feature-rich open-source service called [GitLab](https://gitlab.com), avoiding the reliance on commercial service providers.
-Just sharing Git repositories on GitHub with the public is always free, private repositories (only visible to persons you invite) are [free for researchers](https://help.github.com/en/articles/applying-for-an-educator-or-researcher-discount) or have limited features.
+At the time of writing, sharing repositories on GitHub with the public is free, private repositories (only visible to persons you invite) are [free for researchers](https://help.github.com/en/articles/applying-for-an-educator-or-researcher-discount) or have limited features.
 After creating a user account, one can create a new repository and GitHub provides information on how to upload your repository from the terminal, e.g., for our repository (here with user name 'aaronpeikert' and repository name 'reproducible-research'):
@@
 
-[^1]: To comment on our paper, please add an issue to the GitHub repository of our paper: <https://github.com/aaronpeikert/reproducible-research/issues>.
+[^1]: To comment on our paper, please add an issue to the GitHub repository of our paper: [https://github.com/aaronpeikert/reproducible-research/issues](https://github.com/aaronpeikert/reproducible-research/issues).
 [^doi]: [https://guides.github.com/activities/citable-code/](https://guides.github.com/activities/citable-code/)
@@
 Even when you have obtained a given version of a project with the aim to reproduce reported results, and you can confirm that this version is unchanged, you may not know exactly how to reproduce the results because it may be unclear which files should be executed in which order.
-This is particularly the case when complex preprocessing pipelines are part of the computation and/or external dependencies (for example, a program needs to be called outside the R environment) are present.
+This is particularly the case when complex preprocessing pipelines are part of the computation or there are dependencies on external programs.
 Handling such dependencies is easy with Make because it allows you to manage dependencies by creating (computational) recipes to create or recreate files.

### 2020-05-19 manuscript.Rmd
further revisions

@@
 However, for our purposes, it is sufficient to master a valuable minimal subset of functions to ensure the reproducibility of scientific analyses.
+We recommend using RStudio, an integrated development environment (IDE) for R, which provides simplified access to essential features of some of the tools.
 
@@
 If there is a dependency missing, and there is no target to make it, Make stops with an error message.
-It is a convention to have the first target named `all`, which depends on everything.
+It is a convention to have the first target named `all`, which creates the entire project.
 Subsequently, the command `make` without any argument automatically creates everything possible in the project.
@@
 
-```{r buildpane-rstudio, fig.cap="Build pane in RStudio with access to Makefile target 'all'."}
+```{r buildpane-rstudio, eval=TRUE, echo=FALSE, fig.cap="Build pane in RStudio with access to Makefile target 'all'."}
 knitr::include_graphics("Images/build-pane.png")
@@
 
-Docker is a tool that allows encapsulation, sharing, and re-creation of a computational environment on most operating systems (Windows, macOS & Linux).
-Docker achieves these goals by setting up a virtual computer, on which it can execute a well-defined series of commands (e.g., installing software).
+Docker is a tool that allows encapsulation, sharing, and re-creation of a computational environment on most operating systems (Windows, macOS, & Linux).
+Docker achieves these goals by setting up a virtual computer, on which it can execute commands (e.g., installing software).
 It then saves the resulting state of the virtual computer in what is called an "image." This image can be started and execute commands on the virtual computer, e.g., running `Rscript` or `make`.
@@
 An image can be transferred and executed on any machine that has Docker installed.
-Irrespective of the machine that is executing the container, the computational environment is the same for the programs running inside the container.
-The most important advantage over traditional virtual machines is that containers are lightweight; that is, they start rapidly and do not need much storage space.
-Docker achieves this by reusing large parts of the host's operating system (the extend varies between Linux, macOS and Windows).
+Regardless of the machine that is executing the container, the computational environment is identical for the programs running inside the container.
+The most important advantage over traditional virtual machines is that containers are lightweight: they start rapidly, run with little overhead, and do not need much storage space.
+Docker achieves this by reusing large parts of the host's operating system (the extent varies between Linux, macOS and Windows).
 
-With the following example, we demonstrate the importance of storing computational environment.
+With the following example, we demonstrate the importance of documenting and storing the computational environment.
 Generally, with containers, we would like to safeguard against changes to the computational environment resulting in unexpected consequences, e.g., changes in the functionality or default options in packages or even in the R environment itself.
-While the R programming language is considered stable and much effort is put into backward compatibility, even basic functions like `sample()` (to randomly sample from a set) sometimes change their behaviour from one version to another.
-To ensure reproducibility in analyses based on a computer's pseudo random number generator (PRNG), it is good practice to rely on fixed PRNG seeds, which are numeric values that set the PRNG into a deterministic state, i.e., the sequence of random numbers reproduces exactly.
+While the R programming language is considered stable and much effort is put into backward compatibility, even basic functions like `read.csv()` (to load data) or `sample()` (to randomly sample from a set) sometimes change their behaviour from one version to another.
+For example, to ensure reproducibility of analyses based on a computer's pseudo random number generator (PRNG), it is good practice to rely on fixed PRNG seeds, which are numeric values that set the PRNG into a deterministic state, i.e., the sequence of random numbers reproduces exactly.
 Consider the following R code to randomly draw five numbers between 1 and 10:
@@
 
-The usual expectation is that this code delivers the same pseudo-random five numbers regardless of the operating system or R Version (due to `set.seed()`).
+The usual expectation is that this code delivers the same pseudo-random five numbers regardless of the operating system or R Version (because of `set.seed()`).
 Using Docker, we can start an image which contains the R Version 3.5.0, and execute the code there.
@@
 Only rarely does an analysis depend on base R only.
-Typically, a considerable number of packages is required that each may depend on multiple other packages, and the addition of each package will increase the likelihood of breaking reproducibility with every update to any of the packages or base R (the resulting frustration is sometimes referred to as _dependency hell_).
+Typically, a considerable number of packages is required that each may depend on multiple other packages.
+Each update of each package in this dependency hierarchy and updates to base R itself will increase the likelihood of breaking reproducibility (the resulting frustration is sometimes referred to as _dependency hell_).
 The whole endeavour of reproducibility is therefore at stake every time an update is rolled out.
-To ensure long-term reproducibility, Docker replicates the original computational environment of an analysis exactly.
+To ensure long-term reproducibility, our worflow replicates the original computational environment of an analysis exactly.
 Note, that we do not intend to advocate that software should not be updated; updates typically promote bugfixes and provide new functionality; our point is that full computational reproducibility is only achieved if the software versions used originally are precisely documented.
@@
 
-# execute a given command in a given container image; do not save
-# the state of the container; accept inputs from and return outputs to terminal
+# execute a command in a container image; do not save the state 
+# of the container; accept inputs from and return outputs to terminal
 docker run --rm -it <IMAGENAME> <COMMAND>
@@
 
-# start an interactive R session in the container named 'reproducible-research'
+# start an interactive R session in the 
+# container named 'reproducible-research'
 docker run --rm -it reproducible-research R
@@
 
+# set local variables for later use
 project := $(notdir $(CURDIR))
@@
 
+# determine if DOCKER=TRUE was given
+# if so, run everything in docker
+# if not, run everything locally
 ifeq ($(DOCKER),TRUE)
@@
 
+# default target is target 'manuscript.pdf'
 all: manuscript.pdf
 
+# build the docker container
 build: Dockerfile
-    docker build -t $(projekt) .
+    docker build -t $(project) .
 
+# build manuscript.pdf from Rmd file
+# run in docker if DOCKER=TRUE else locally
 manuscript.pdf: manuscript.Rmd reproducible-research.bib
@@
 Next, one needs to make sure that all files relevant to the analysis can be moved to another computer.
-To this end, it is recommendable that all files are within one folder (or enclosed subfolders within it) and all paths are relative to that folder because absolute paths are specific to a given computer.
+To this end, it is recommended that all files reside within a one folder (or enclosed subfolders within it) and all paths are relative to that folder because absolute paths are specific to a given computer.
 A robust solution to the problem of making sure that file access does not break across computing platforms are [RStudio projects](https://r4ds.had.co.nz/workflow-projects.html) and the `here` package [@here2017] to manage file access.
@@
 First, it takes care of the fact that path separator characters vary across operating systems (typically, slash or backslash).
-Second, anchor points of relative paths may differ depending on context.
+Second, it solves the issue that anchor points of relative paths may differ depending on context.
 For example, knitr interprets paths relative to the dynamic document, whereas R has a current working directory that may change over the course of an R session.
@@
 The folder where all the files reside that you need for analysis (code and data), is referred to as a 'project' (or sometimes as a 'research compendium').
-Working with projects is particularly convenient with RStudio, an integrated development environment (IDE) for R.
+Working with projects is particularly convenient with RStudio.
 It is useful to organize a data analysis project in a way that strictly segregates (raw) data and code by placing them in directories called `data` and `R` [see Section 4 in @marwickPackagingDataAnalytical2018]; there are also tools that automatize the standardized creation of folder structures such as `workflowr` [@workflowr].
@@
 The overarching goal of this paper was to provide a complete and easy-to-use workflow that allows confidence in the reproducibility of R-based data analyses.
-Analyses following our workflow can be reproduced with four commands (here shown for our paper):
+Analyses following our workflow can be reproduced with four commands (here shown for this manuscript):
 
@@
 
-That is, this reproduces a scientific report exactly without regard to the user, the operating system, the software, the timepoint, or interim changes to the involved files.
+The workflow enables reproduction a scientific report exactly without regard to the local operating system, locally installed software, time, or interim changes to the project files.
 To that end, the proposed workflow relies on tools that have been the foundation of reliable software development for years or even decades.
@@
 Each tool in the workflow reduces the chances of non-reproducibility.
-Dynamic reporting with R Markdown guarantees consistency between computational results and their reporting; Version control with Git ensures permanence and consistency across multiple versions of data and code; Dependency management with Make affirms defined entry-points and dependency resolution; containerization with Docker secures reproducibility of the full computational environment.
-We believe that this tightening of loose ends does not restrict researchers but enables them to operate on a solid basis to deliver sound and sustainable research.
+Dynamic reporting with R Markdown guarantees consistency between computational results and their reporting; Version control with Git ensures permanence and consistency across multiple versions of data and code; Dependency management with Make provides defined entry-points while mapping out dependencies between all components of a project; containerization with Docker guarantuees full computational reproducibility.
+We believe that the proposed combination of tools does not restrict researchers but enables them to operate on a solid basis to deliver transparent and sustainable research.
 
@@
 
-While our approach was designed to scale well with the complexity of a computationally intense project, we realize that this flexibility may not be straightforward to integrate into a researcher's everyday workflow.
-There are R packages that implement parts of our workflow and, thus, lower the threshold for adoption when the full flexibility provided by our workflow is not needed.
+While our approach was designed to scale well with the complexity of a computationally intense project, we realize that this flexibility may not be straightforward to integrate into researchers' everyday workflow.
+There are various R packages that implement parts of our workflow and, thus, lower the threshold for adoption when the full flexibility provided by our workflow is not needed.
 The use of R Markdown within a project, tracked with Git can be simplified with the [`workflowR`](https://github.com/jdblischak/workflowr) package [@workflowr].
@@
 The package [`reprex`](https://github.com/tidyverse/reprex) [**repr**oducible **ex**ample, @reprex] is also worth noting, but its scope is limited.
-A particularly noteworthy approach is provided by the [`worcs`](https://cran.r-project.org/web/packages/worcs) package [@worcs, @worcspaper], which is an R project template that creates a standardized file structure for code and data supporting version management with Git, package management with renv and dynamic document generation with R Markdown.
+A particularly noteworthy approach is provided by the [`worcs`](https://cran.r-project.org/web/packages/worcs) package [@worcs; @worcspaper], which is an R project template that creates a standardized file structure for code and data supporting version management with Git, package management with renv and dynamic document generation with R Markdown.
 We acknowledge that worcs is much easier to install and provides a one-click solution for the creation of reproducible projects.

### 2020-05-19 R/rmarkdown_fake.Rmd
update Fig 2 closes #39

--- 2019-11-12 R/rmarkdown.Rmd
@@
 ---
-title: "A simple R markdown example"
+title: "A simple R markdown example}"
 author: "Aaron Peikert & Andreas M. Brandmaier"
-date: '`r format(Sys.time(), "%B %d, %Y")`'
 output: pdf_document
-papersize: "a6"
-geometry:
-  - "landscape"
-  - margin=0.5cm
-header-includes:
-   - \usepackage[center,dvips,frame]{crop}
 ---
+
+### Silly Heading
 
@@
 ```
-
-# Silly Heading
 

### 2020-05-19 manuscript.Rmd

@@
 The package [`reprex`](https://github.com/tidyverse/reprex) [**repr**oducible **ex**ample, @reprex] is also worth noting, but its scope is limited.
-A particularly noteworthy approach is provided by the [`worcs`](https://cran.r-project.org/web/packages/worcs) package [@worcs; @worcspaper], which is an R project template that creates a standardized file structure for code and data supporting version management with Git, package management with renv and dynamic document generation with R Markdown.
+A particularly noteworthy approach is the [`worcs`](https://cran.r-project.org/web/packages/worcs) package [@worcs; @worcspaper], which is an R project template that creates a standardized file structure for code and data supporting version management with Git, package management with `renv` and dynamic document generation with R Markdown.
 We acknowledge that worcs is much easier to install and provides a one-click solution for the creation of reproducible projects.
@@
 We encourage authors to choose appropriate license forms that do not hinder others from freely downloading, using, and modifying the shared workflows and materials while, at the same time, ensuring recognition for the time and effort invested in creating the workflow in the first place.
-In our experience, the [Creative Commons license (CC-BY)](https://creativecommons.org/licenses/by/4.0/) is often appropriate for sharing texts, R Markdown files, generated figures, and other media, whereas scripts and any other computer code are often best shared under the [MIT license](https://opensource.org/licenses/MIT).
+In our experience, the [Creative Commons license (CC-BY)](https://creativecommons.org/licenses/by/4.0/) is often appropriate for sharing texts, R Markdown files, generated figures, and other media, whereas scripts and any other computer code are often best shared under the [MIT license](https://opensource.org/licenses/MIT) (or similar).
 Both licenses assure maximal freedom for future users while requiring the attribution of the original authors in derivative work.
@@
 The proposed workflow leverages various existing tools that are partly integrated into RStudio already.
-Parts of the proposed workflow have been integrated into stand-alone packages (such as worcs [@worcs], workflowr [@workflowr], holepunch  [@holepunch]), which are particularly recommendable to beginners but they are either incomplete in the sense of our proposal or rely on proprietary service providers.
-We hope that with increasing awareness of the challenges of computational reproducibility, the increased demand for unified and open solutions will lead to better integration of existing tools to allow reproducible workflows to become a standard in psychological research.
+Parts of the proposed workflow have been integrated into stand-alone packages [such as `worcs`, @worcs; `workflowr`, @workflowr; or `holepunch` @holepunch], which we particularly recommend to beginners;  but they are either incomplete in the sense of our proposal or rely on proprietary service providers.
+We hope that with increasing awareness of the challenges of computational reproducibility, the increased demand for unified and open solutions will lead to better integration of existing tools and services to allow reproducible workflows to become a standard in psychological research.
 

### 2020-05-19 R/rmarkdown.Rmd
updated rmarkdown example (Fig2/Fig3), possibly broken now

--- 2019-10-28 R/rmarkdown.Rmd
@@
 ---
-title: "A simple RMarkdown Example"
-author: "Aaron Peikert & Andreas M. Brandmaier"
-date: '`r format(Sys.time(), "%d %B, %Y")`'
+#title: "\\vspace{-1em}A simple R markdown example\\vspace{-1em}"
+#author: "Aaron Peikert & Andreas M. Brandmaier"
 output: pdf_document
 papersize: "a6"
+fontsize: 9pt
 geometry:
@@
 
-```{r setup}
+### Dynamic Document Demonstration
+
+```{r setup, echo=FALSE}
 library("knitr")
@@
 
-# Silly Heading
+This is a simple analysis of the `sleep` dataset (Student, 1908) taken from `help(t.test)`.
 
@@
 data("sleep")
-result <- t.test(extra ~ group, data = sleep)
+result <- t.test(extra ~ group, data = sleep, paired = TRUE)
 ```
 
-This is an example on students' sleep data taken from `help(t.test)`.
-
-`r apa_print.htest(result)$full_result`
-
-I can now assert that what I *believe* to be true
---- that there is a difference in means between the groups ---
-is `r ifelse(result$p.value > .025, "**not**", "")` supported by data.
+The difference in means of hours slept between the groups 
+was `r ifelse(result$p.value >= .05, "**not**", "")` 
+significantly different from zero (`r apa_print.htest(result)$full_result`).

### 2020-05-19 manuscript.Rmd
changed xref to Version Control

@@
 A possible downside is that in order to create an image from a Dockerfile, all software repositories need to be still available.
-Hence, to guarantee longterm reproducibility, it is best to archive the complete binary image at major points of the projects' progress, for example, on publication (ideally, using a release tag; see [Version Management] for details).
+Hence, to guarantee longterm reproducibility, it is best to archive the complete binary image at major points of the projects' progress, for example, on publication (ideally, using a release tag; see Section _Version Control_ for details).
 

### 2020-05-19 manuscript.Rmd
extended description of worcs in outlook

@@
 The proposed workflow leverages various existing tools that are partly integrated into RStudio already.
-Parts of the proposed workflow have been integrated into stand-alone packages [such as `worcs`, @worcs; `workflowr`, @workflowr; or `holepunch` @holepunch], which we particularly recommend to beginners;  but they are either incomplete in the sense of our proposal or rely on proprietary service providers.
+Parts of the proposed workflow have been integrated into stand-alone packages [such as `worcs`, @worcs; `workflowr`, @workflowr; or `holepunch` @holepunch], which we particularly recommend to beginners; in particular, `worcs` is a step-by-step procedure with best practices for Open Science from preregistration to publication.
+Still those approaches do either not guarantuee full computational reproducibility or rely on proprietary service providers.
 We hope that with increasing awareness of the challenges of computational reproducibility, the increased demand for unified and open solutions will lead to better integration of existing tools and services to allow reproducible workflows to become a standard in psychological research.

### 2020-05-19 manuscript.Rmd
removed a "particular"

@@
 The proposed workflow leverages various existing tools that are partly integrated into RStudio already.
-Parts of the proposed workflow have been integrated into stand-alone packages [such as `worcs`, @worcs; `workflowr`, @workflowr; or `holepunch` @holepunch], which we particularly recommend to beginners; in particular, `worcs` is a step-by-step procedure with best practices for Open Science from preregistration to publication.
+Parts of the proposed workflow have been integrated into stand-alone packages [such as `worcs`, @worcs; `workflowr`, @workflowr; or `holepunch` @holepunch], which we recommend to beginners; in particular, `worcs` is a step-by-step procedure with best practices for Open Science from preregistration to publication.
 Still those approaches do either not guarantuee full computational reproducibility or rely on proprietary service providers.

### 2020-05-19 manuscript.Rmd
rebuild with docker

--- 2020-05-19 manuscript.Rmd
@@
 
-if(!require("pacman"))install.packages("pacman")
+if(!requireNamespace("pacman"))install.packages("pacman")
 pacman::p_load("here", "tidyverse")
@@
 A possible downside is that in order to create an image from a Dockerfile, all software repositories need to be still available.
-Hence, to guarantee longterm reproducibility, it is best to archive the complete binary image at major points of the projects' progress, for example, on publication (ideally, using a release tag; see Section _Version Control_ for details).
+Hence, to guarantee longterm reproducibility, it is best to archive the complete binary image at major points of the projects' progress, for example, on publication (ideally, using a release tag; see [Version Management] for details).
 

### 2020-05-19 manuscript.Rmd
open port internally only #41

--- 2020-05-19 manuscript.Rmd
@@
 The second option is to start the container in the background and to interact with the container via the web browser and the RStudio server instance running in it.
-In order to do so, you need to supply a password to log into the RStudio server (`-e PASSWORD=<YOUR_PASS>`) and open a network service on a specified port (`-p 8787:8787`).
+In order to do so, you need to supply a password to log into the RStudio server (`-e PASSWORD=<YOUR_PASS>`) and open a network service on a specified port (`-p 127.0.0.1:8787:8787`).
 
 ```{bash, eval=FALSE}
-docker run -e PASSWORD=<YOUR_PASS> -p 8787:8787 image-name
+docker run -e PASSWORD=<YOUR_PASS> -p 127.0.0.1:787:8787 image-name
 ```

### 2020-05-19 manuscript.Rmd
fix cross ref (again) closes #43

--- 2020-05-19 manuscript.Rmd
@@
 
-Figure \ref{fig:schematic} gives an overview of how the four components of our workflow interact to ensure computational reproducibility.
+Figure \@ref(fig:schematic) gives an overview of how the four components of our workflow interact to ensure computational reproducibility.
 Before we describe the four components in more detail, we begin with a minimal description of the roles of each component.
@@
 Version control offers snapshots of your workflow at different time points identified by a unique identifier.
-The different frames in the z-axis of Figure  \ref{fig:schematic} illustrate the different versions of the project.
+The different frames in the z-axis of Figure  \@ref(fig:schematic) illustrate the different versions of the project.
 How different parts of a analysis and a report of it relate and in what order they need to be executed is documented using dependency management.
@@
 Of particular note are the `papaja` package [@papaja2018], which offers additional functions to enable American Psychological Association (APA) style document formatting, including a journal-style final typeset format, and the `stargazer` package [@stargazer2018], which provides journal-ready tables and reports of statistical models.
-Figure \ref{fig:source-rmarkdown} illustrates R Markdown syntax using the `papaja` package and Figure \ref{fig:rendered-result} shows the resulting rendered document.
+Figure \@ref(fig:source-rmarkdown) illustrates R Markdown syntax using the `papaja` package and Figure \@ref(fig:rendered-result) shows the resulting rendered document.
 
@@
 
-```{r rendered-result, eval = TRUE, echo = FALSE,  fig.cap="Rendered result of the source code shown in Figure \ref{fig:source-rmarkdown}."}
+```{r rendered-result, eval = TRUE, echo = FALSE,  fig.cap="Rendered result of the source code shown in Figure \\ref{fig:source-rmarkdown}."}
 knitr::include_graphics("R/rmarkdown.pdf")
@@
 RStudio also offers a graphical user interface for Git.
-For most basic operations, this interface is convenient and sufficient (see Figure \ref{fig:gitpane}).
+For most basic operations, this interface is convenient and sufficient (see Figure \@ref(fig:gitpane)).
 
@@
 
-`git push` or the green upward arrow in the Git pane (see Figure \ref{fig:gitpane}) uploads local updates. To download the remote Git repository on another computer, type into the terminal:
+`git push` or the green upward arrow in the Git pane (see Figure \@ref(fig:gitpane)) uploads local updates. To download the remote Git repository on another computer, type into the terminal:
 
@@
 Subsequently, the command `make` without any argument automatically creates everything possible in the project.
-The button `Build All` from within RStudio triggers this process (see Figure \ref{fig:buildpane-rstudio}).
+The button `Build All` from within RStudio triggers this process (see Figure \\@ref(fig:buildpane-rstudio)).
 
@@
 The `-it` flag tells Docker to run the command interactively, that is, to accept keyboard inputs and return outputs to the terminal.
-For example, this is the command to start an interactive R session inside a Docker image called `reproducible-research` (see Figure \ref{fig:docker-r-terminal} for a screenshot):
+For example, this is the command to start an interactive R session inside a Docker image called `reproducible-research` (see Figure \@ref(fig:docker-r-terminal) for a screenshot):
 
@@
 
-Figure \ref{fig:rstudio-in-docker} shows a screenshot of Rstudio running inside Docker accessed from a local web browser.
+Figure \@ref(fig:rstudio-in-docker) shows a screenshot of Rstudio running inside Docker accessed from a local web browser.
 

### 2020-05-19 manuscript.Rmd
fix small typo

--- 2020-05-19 manuscript.Rmd
@@
 
-manuscript.pdf: data/ìris_prepped.csv manuscript.Rmd
+manuscript.pdf: data/iris_prepped.csv manuscript.Rmd
   Rscript -e 'rmarkdown::render("manuscript.Rmd")'
@@
 The second option is to start the container in the background and to interact with the container via the web browser and the RStudio server instance running in it.
-In order to do so, you need to supply a password to log into the RStudio server (`-e PASSWORD=<YOUR_PASS>`) and open a network service on a specified port (`-p 127.0.0.1:8787:8787`).
+In order to do so, you need to supply a password to log into the RStudio server (`-e PASSWORD=<YOUR_PASS>`) and open a network service on a specified port (`-p 8787:8787`).
 
 ```{bash, eval=FALSE}
-docker run -e PASSWORD=<YOUR_PASS> -p 127.0.0.1:787:8787 image-name
+docker run -e PASSWORD=<YOUR_PASS> -p 8787:8787 image-name
 ```

### 2020-05-19 R/rmarkdown.Rmd

@@
 ---
-#title: "\\vspace{-1em}A simple R markdown example\\vspace{-1em}"
-#author: "Aaron Peikert & Andreas M. Brandmaier"
 output: pdf_document
-papersize: "a6"
 fontsize: 9pt
@@
   - margin=0.5cm
+  - paperwidth=5cm
+  - paperheight=10cm
 header-includes:

### 2020-05-19 manuscript.Rmd
fixes typos in port address

@@
 The second option is to start the container in the background and to interact with the container via the web browser and the RStudio server instance running in it.
-In order to do so, you need to supply a password to log into the RStudio server (`-e PASSWORD=<YOUR_PASS>`) and open a network service on a specified port (`-p 127.0.0.1:8787:8787`).
+In order to do so, you need to supply a password to log into the RStudio server (`-e PASSWORD=<YOUR_PASS>`) and open a local network service on a specified port (`-p 127.0.0.1:8787:8787`).
 
 ```{bash, eval=FALSE}
-docker run -e PASSWORD=<YOUR_PASS> -p 127.0.0.1:787:8787 image-name
+docker run -e PASSWORD=<YOUR_PASS> -p 127.0.0.1:8787:8787 image-name
 ```
 
-The address to connect to the RStudio server is your IP address (or `localhost` on Linux) in this scheme: `<IPADRESS>:8787`.
+The address to connect to the RStudio server is your IP address (or `localhost` on Linux) in this scheme: `<IPADDRESS>:8787`.
 This offers a fully functioning RStudio instance that runs in the image but is accessible through a local web browser.

### 2020-05-19 manuscript.Rmd
renamed xref
added figure caption fixed typo

@@
 title             : "A Reproducible Data Analysis Workflow with R Markdown, Git, Make, and Docker"
-shorttitle        : "Reproducible Data Analysis Workflow"
+shorttitle        : "Reproducible Workflow"
 
@@
 
-```{r source-rmarkdown, eval = TRUE, echo = FALSE,  fig.cap="Exemplary source of an R Markdown showing a combination of executable R code and English manuscript text."}
+```{r source-rmarkdown, eval = TRUE, echo = FALSE,  fig.cap="Exemplary source of an R Markdown showing a combination of executable R code and English manuscript text.
+Code is either given in separate chucks (grey background delimited by triple backticks) or inline (single backticks).
+"}
 knitr::include_graphics("Images/rmarkdown.png")
@@
 git add ./data/iris.csv ./R/analysis.R # track specific files
-git commit -m "add data & analysis" # take snapshot with comment
+git commit -m "added data and analysis" # take snapshot with comment
 # once script or data were changed, take a new snapshot
-git commit -a -m "complete data colections" # add and commit all changes
+git commit -a -m "completed data collection" # add and commit all changes
 ```
@@
 However, Docker needs unrestricted access rights to the system, which are rarely granted on high-performance computing clusters.
-For this case, Singularity provides a fully compatible alternative (see Section under [Linux]) that can be executed with limited access rights.
+For this case, Singularity provides a fully compatible alternative (see Section on Linux) that can be executed with limited access rights.
 

### 2020-05-20 manuscript.Rmd
fixed broken cross-reference

@@
 Subsequently, the command `make` without any argument automatically creates everything possible in the project.
-The button `Build All` from within RStudio triggers this process (see Figure \\@ref(fig:buildpane-rstudio)).
+The button `Build All` from within RStudio triggers this process (see Figure \@ref(fig:buildpane-rstudio)).
 
@@
 In the settings of Docker Desktop, you have to allow the sharing of your drive.
-Docker on Windows requires the path to be special; therefore, you need to hand-edit the `Makefile` and set `current_path` to the current directory and use `make all DOCKER=TRUE WINDOWS=TRUE`.
+Docker on Windows requires an unusual path specification (`C:Users\\aaron\\Documents\\reproducible-research` becomes `/c/Users/aaron/Documents/reproducible-research`).
+Therefore, you need to hand-edit the `Makefile` and set `current_path` to the project directory and use `make all DOCKER=TRUE WINDOWS=TRUE`.
 We hope that future releases of Docker for Windows will not require that workaround.

### 2020-05-20 manuscript.Rmd
revision

--- 2020-05-19 manuscript.Rmd
@@
     affiliation   : "1,3"
-    corresponding : yes    # Define only one corresponding author
+    corresponding : yes
     address       : "Max Planck Institute for Human Development, Lentzeallee 94, 14195 Berlin, Germany"

### 2020-05-20 R/rmarkdown.Rmd

@@
   - margin=0.5cm
-  - paperwidth=5cm
-  - paperheight=10cm
+  - paperwidth=6cm
+  - paperheight=12cm
 header-includes:

### 2020-05-20 manuscript.Rmd
fixes missing slash

--- 2020-05-20 manuscript.Rmd
@@
 In the settings of Docker Desktop, you have to allow the sharing of your drive.
-Docker on Windows requires the path to be special; therefore, you need to hand-edit the `Makefile` and set `current_path` to the current directory and use `make all DOCKER=TRUE WINDOWS=TRUE`.
+Docker on Windows requires an unusual path specification (e.g., `C:\\Users\\aaron\\Documents\\reproducible-research` becomes `/c/Users/aaron/Documents/reproducible-research`).
+Therefore, you currently need to hand-edit the `Makefile` and set `current_path` to the project directory and use `make all DOCKER=TRUE WINDOWS=TRUE`.
 We hope that future releases of Docker for Windows will not require that workaround.

### 2020-05-20 manuscript.Rmd
fixed typo

@@
 The arrows visualize dependencies, such as an analysis depending on the availability of a particular data file.
-Third, all computer code (such as a statistical analyses in R) is executed in a virtual environment that guarantuees exact reproduction of results independent of the host operating system, the locally installed R version, and installed package versions.
+Third, all computer code (such as a statistical analyses in R) is executed in a virtual environment that guarantees exact reproduction of results independent of the host operating system, the locally installed R version, and installed package versions.
 Finally, dynamic document generation (also known as the literate programming paradigm) interweaves human-readable code and computed results (such as point estimates, p values, or confidence intervals) to eliminate inconsistency errors such as those arising from copy\&paste errors.
@@
 Each tool in the workflow reduces the chances of non-reproducibility.
-Dynamic reporting with R Markdown guarantees consistency between computational results and their reporting; Version control with Git ensures permanence and consistency across multiple versions of data and code; Dependency management with Make provides defined entry-points while mapping out dependencies between all components of a project; containerization with Docker guarantuees full computational reproducibility.
+Dynamic reporting with R Markdown guarantees consistency between computational results and their reporting; Version control with Git ensures permanence and consistency across multiple versions of data and code; Dependency management with Make provides defined entry-points while mapping out dependencies between all components of a project; containerization with Docker guarantees full computational reproducibility.
 We believe that the proposed combination of tools does not restrict researchers but enables them to operate on a solid basis to deliver transparent and sustainable research.
@@
 Parts of the proposed workflow have been integrated into stand-alone packages [such as `worcs`, @worcs; `workflowr`, @workflowr; or `holepunch` @holepunch], which we recommend to beginners; in particular, `worcs` is a step-by-step procedure with best practices for Open Science from preregistration to publication.
-Still those approaches do either not guarantuee full computational reproducibility or rely on proprietary service providers.
+Still those approaches do either not guarantee full computational reproducibility or rely on proprietary service providers.
 We hope that with increasing awareness of the challenges of computational reproducibility, the increased demand for unified and open solutions will lead to better integration of existing tools and services to allow reproducible workflows to become a standard in psychological research.

### 2020-05-20 manuscript.Rmd
revised figure caption

@@
 
-```{r source-rmarkdown, eval = TRUE, echo = FALSE,  fig.cap="Exemplary source of an R Markdown showing a combination of executable R code and English manuscript text.
-Code is either given in separate chucks (grey background delimited by triple backticks) or inline (single backticks).
+```{r source-rmarkdown, eval = TRUE, echo = FALSE,  fig.cap="Exemplary excerpt of an R Markdown file showing a combination of executable R code, which will be dynamically rendered to content on document creation, and English manuscript text.
+Code is either given in separate chunks (shown in grey background delimited by triple backticks) or inline (single backticks).
+The resulting document is shown in Figure \\ref{fig:rendered-result}.
 "}

### 2020-05-21 manuscript.Rmd
added backticks
fix typo ì -> i

@@
 Assume the final product is a manuscript (`manuscript.pdf`).
-This manuscript is written in Rmarkdown (`manuscript.Rmd`) and includes dynamically generated plots from a raw data file (`data/iris.csv`) that needs to be preprocessed first using a separate script (`R/prepare_data.R`) into a prepared data file (`ìris_prepped.csv`).
+This manuscript is written in Rmarkdown (`manuscript.Rmd`) and includes dynamically generated plots from a raw data file (`data/iris.csv`) that needs to be preprocessed first using a separate script (`R/prepare_data.R`) into a prepared data file (`iris_prepped.csv`).
 A `Makefile` for these dependencies may look like this:
@@
 
-data/ìris_prepped.csv: R/prepare_data.R data/iris.csv
+data/iris_prepped.csv: R/prepare_data.R data/iris.csv
   Rscript -e 'source("R/prepare_data.R")'
@@
 If Make is called without an argument, the first target is built.
-To create `manuscript.pdf` (the second target in the file), the file `manuscript.Rmd` needs to be rendered, which depends on `data/ìris_prepped.csv`.
+To create `manuscript.pdf` (the second target in the file), the file `manuscript.Rmd` needs to be rendered, which depends on `data/iris_prepped.csv`.
 This dependency is itself a target (the third target in the file).
-To create `data/ìris_prepped.csv`, `R/prepare_data.R` and `data/iris.csv` are needed.
+To create `data/iris_prepped.csv`, `R/prepare_data.R` and `data/iris.csv` are needed.
 If you type `make manuscript.pdf`, Make first checks whether the dependencies do exist and, if not, creates them.
-Here, if `data/ìris_prepped.csv` does not exist, Make creates it by executing the third target.
+Here, if `data/iris_prepped.csv` does not exist, Make creates it by executing the third target.
 Also, if one of the dependencies of a target is newer than the target itself, Make updates everything that directly or indirectly depends on the target.
@@
 A robust solution to the problem of making sure that file access does not break across computing platforms are [RStudio projects](https://r4ds.had.co.nz/workflow-projects.html) and the `here` package [@here2017] to manage file access.
-The here package solves two common issues with relative paths.
+The `here` package solves two common issues with relative paths.
 First, it takes care of the fact that path separator characters vary across operating systems (typically, slash or backslash).

### 2020-05-21 manuscript.Rmd
fix double backslash no need to escape backslash in code

@@
 In the settings of Docker Desktop, you have to allow the sharing of your drive.
-Docker on Windows requires an unusual path specification (e.g., `C:\\Users\\aaron\\Documents\\reproducible-research` becomes `/c/Users/aaron/Documents/reproducible-research`).
+Docker on Windows requires an unusual path specification (e.g., `C:\Users\aaron\Documents\reproducible-research` becomes `/c/Users/aaron/Documents/reproducible-research`).
 Therefore, you currently need to hand-edit the `Makefile` and set `current_path` to the project directory and use `make all DOCKER=TRUE WINDOWS=TRUE`.

### 2020-05-21 manuscript.Rmd
mention screenshot of rstudio in docker earlier
bump release number

@@
 [^doi]: [https://guides.github.com/activities/citable-code/](https://guides.github.com/activities/citable-code/)
-[^sub]: We created an release for the submission: [https://github.com/aaronpeikert/reproducible-research/releases/tag/v0.1.1-submission](https://github.com/aaronpeikert/reproducible-research/releases/tag/v0.1.1-submission)
+[^sub]: We created an release for the submission: [https://github.com/aaronpeikert/reproducible-research/releases/tag/v0.2.0-submission](https://github.com/aaronpeikert/reproducible-research/releases/tag/v0.2.0-submission)
 
@@
 
+```{r rstudio-in-docker, eval = TRUE, echo = FALSE, fig.cap="RStudio running inside Docker"}
+knitr::include_graphics("Images/docker-rstudio.png")
+```
+
+Figure \@ref(fig:rstudio-in-docker) shows a screenshot of Rstudio running inside Docker accessed from a local web browser.
+
 By default, programs inside the container cannot access files on the local computer, thus requiring an explicit link to a local folder to enable access (and on macOS and Windows this also has to be allowed in the settings) :
@@
 
-```{r rstudio-in-docker, eval = TRUE, echo = FALSE, fig.cap="RStudio running inside Docker"}
-knitr::include_graphics("Images/docker-rstudio.png")
-```
-
-Figure \@ref(fig:rstudio-in-docker) shows a screenshot of Rstudio running inside Docker accessed from a local web browser.
+Figure \@ref(fig:rstudio-in-docker) shows a screenshot of the result.
 

### 2020-05-22 manuscript.Rmd
fixed installation guide for macOS

--- 2020-05-21 manuscript.Rmd
@@
 Assume the final product is a manuscript (`manuscript.pdf`).
-This manuscript is written in Rmarkdown (`manuscript.Rmd`) and includes dynamically generated plots from a raw data file (`data/iris.csv`) that needs to be preprocessed first using a separate script (`R/prepare_data.R`) into a prepared data file (`iris_prepped.csv`).
+This manuscript is written in Rmarkdown (`manuscript.Rmd`) and includes dynamically generated plots from a raw data file (`data/iris.csv`) that needs to be preprocessed first using a separate script (`R/prepare_data.R`) into a prepared data file (`ìris_prepped.csv`).
 A `Makefile` for these dependencies may look like this:
@@
 
-data/iris_prepped.csv: R/prepare_data.R data/iris.csv
+data/ìris_prepped.csv: R/prepare_data.R data/iris.csv
   Rscript -e 'source("R/prepare_data.R")'
@@
 If Make is called without an argument, the first target is built.
-To create `manuscript.pdf` (the second target in the file), the file `manuscript.Rmd` needs to be rendered, which depends on `data/iris_prepped.csv`.
+To create `manuscript.pdf` (the second target in the file), the file `manuscript.Rmd` needs to be rendered, which depends on `data/ìris_prepped.csv`.
 This dependency is itself a target (the third target in the file).
-To create `data/iris_prepped.csv`, `R/prepare_data.R` and `data/iris.csv` are needed.
+To create `data/ìris_prepped.csv`, `R/prepare_data.R` and `data/iris.csv` are needed.
 If you type `make manuscript.pdf`, Make first checks whether the dependencies do exist and, if not, creates them.
-Here, if `data/iris_prepped.csv` does not exist, Make creates it by executing the third target.
+Here, if `data/ìris_prepped.csv` does not exist, Make creates it by executing the third target.
 Also, if one of the dependencies of a target is newer than the target itself, Make updates everything that directly or indirectly depends on the target.
@@
 # install Docker via Homebrew
-brew cask docker
+brew cask install docker
 ```

### 2020-05-26 manuscript.Rmd
bumb release version number

--- 2020-05-21 manuscript.Rmd
@@
 [^doi]: [https://guides.github.com/activities/citable-code/](https://guides.github.com/activities/citable-code/)
-[^sub]: We created an release for the submission: [https://github.com/aaronpeikert/reproducible-research/releases/tag/v0.2.0-submission](https://github.com/aaronpeikert/reproducible-research/releases/tag/v0.2.0-submission)
+[^sub]: We created an release for the submission: [https://github.com/aaronpeikert/reproducible-research/releases/tag/v0.2.1-submission](https://github.com/aaronpeikert/reproducible-research/releases/tag/v0.2.1-submission)
 

### 2020-08-07 manuscript.Rmd
Fixed typo

@@
 Even when you lose access to the file on your computer, everything can be backed up on a remote Git server (like GitHub or GitLab).
-Further, one can reduce the liklihood of dead code accumulating (e.g., lines that have been commented out) because it is safe to simply remove unneeded code blocks and track their removal in Git.
+Further, one can reduce the likelihood of dead code accumulating (e.g., lines that have been commented out) because it is safe to simply remove unneeded code blocks and track their removal in Git.
 

### 2020-08-07 manuscript.Rmd
Fixed typo of iris_prepped.csv

@@
 Also, if one of the dependencies of a target is newer than the target itself, Make updates everything that directly or indirectly depends on the target.
-Here, if `data\iris.csv` is newer than `data\ìris_prepped.csv`, Make will attempt to recreate `data\ìris_prepped.csv` first and then `manuscript.pdf`.
+Here, if `data\iris.csv` is newer than `data\iris_prepped.csv`, Make will attempt to recreate `data\iris_prepped.csv` first and then `manuscript.pdf`.
 If there is a dependency missing, and there is no target to make it, Make stops with an error message.

### 2020-09-21 manuscript.Rmd
point windows home users to alternatives closes #63

@@
 Windows systems typically require the biggest efforts to install all necessary pieces of software.
-Note, that you must have either Windows Pro, Enterprise, Education, or Server installed, as Microsoft prevents the use of Docker on Windows Home.
+Note, that you must have either Windows Pro, Enterprise, Education, or Server installed, as Microsoft prevents the use of Docker on Windows Home[^home].
 There is a package manager for Windows called Chocolatey, which you can install from: [https://chocolatey.org/install](https://chocolatey.org/install).
@@
 We hope that future releases of Docker for Windows will not require that workaround.
+
+[^home]: See the section on [related approaches] for some alternatives to Docker in case you have to use Windows Home.
 
@@
 
-For other distributions, replace `apt` with the native package manager.
+For other distributions, replace `apt install` with your package manager's equivalent.
 You may need elevated rights for the installation; in this case, add `sudo` before the installation command.

### 2020-09-22 manuscript.Rmd
remove sloppy wording in definition closes #64
consistent slashes in make example

@@
 In this tutorial, we describe a workflow to ensure long-term and cross-platform reproducibility of data analyses in R [@baseR].
-Reproducibility is the ability to draw an identical conclusion from the same statistical analysis and the same data.
+Reproducibility is the ability to draw identical results from the same statistical analysis and the same data.
 For us,  statistical results are only reproducible if their generating, computational workflow is reported fully and transparently, and remains sustainably available, such that the workflow can be re-run by a different person or later in time, and that the results will be identical to the ones initially reported [@turingway2020; @herouxCompatibleReproducibilityTaxonomy2018; @claerboutElectronicDocumentsGive1992].
@@
 Also, if one of the dependencies of a target is newer than the target itself, Make updates everything that directly or indirectly depends on the target.
-Here, if `data\iris.csv` is newer than `data\iris_prepped.csv`, Make will attempt to recreate `data\iris_prepped.csv` first and then `manuscript.pdf`.
+Here, if `data/iris.csv` is newer than `data/iris_prepped.csv`, Make will attempt to recreate `data/iris_prepped.csv` first and then `manuscript.pdf`.
 If there is a dependency missing, and there is no target to make it, Make stops with an error message.

### 2020-09-24 manuscript.Rmd
mention zenondo directly closes #60

@@
 the resulting pdf of the manuscript or the docker image (see [Containerization]).
-From such a release a DOI can be created, making it easier to reference it (see this [GitHub Guide](https://guides.github.com/activities/citable-code/)[^doi]).
+From such a release [zenodo.org](https://zenodo.org) can create a DOI, making it easier to reference it (see this [GitHub Guide](https://guides.github.com/activities/citable-code/)[^doi]).
 

### 2020-09-24 manuscript.Rmd
better explain the Make example closes #65

@@
 Each recipe has a target (the name of the recipe) followed by a colon and a list of dependent targets or files.
+This is comparable to a cooking recipe where you have the resulting dish at the top, the required ingredients after that and at last the steps one has to follow to prepare the dish.
 If any of the dependencies have changed since the last time the target was built, the recipe's commands are executed to recreate the target file.
@@
 This manuscript is written in Rmarkdown (`manuscript.Rmd`) and includes dynamically generated plots from a raw data file (`data/iris.csv`) that needs to be preprocessed first using a separate script (`R/prepare_data.R`) into a prepared data file (`iris_prepped.csv`).
+You find a graphical representation of this example in Figure \@ref(fig:schematic).
 A `Makefile` for these dependencies may look like this:
@@
 If there is a dependency missing, and there is no target to make it, Make stops with an error message.
+This way Make ensures that all four dependencies of the manuscript (the raw data `data/iris.csv`, the data preparation script `R/prepare_data.R`, the prepared data `data/iris_prepped.csv` and the manuscript source file `manuscript.Rmd`) are correctly resolved.
 It is a convention to have the first target named `all`, which creates the entire project.

### 2020-09-24 manuscript.Rmd
be specific about CC-BY closes #61

@@
 We encourage authors to choose appropriate license forms that do not hinder others from freely downloading, using, and modifying the shared workflows and materials while, at the same time, ensuring recognition for the time and effort invested in creating the workflow in the first place.
-In our experience, the [Creative Commons license (CC-BY)](https://creativecommons.org/licenses/by/4.0/) is often appropriate for sharing texts, R Markdown files, generated figures, and other media, whereas scripts and any other computer code are often best shared under the [MIT license](https://opensource.org/licenses/MIT) (or similar).
+In our experience, the [Creative Commons - Attribution license (CC-BY)](https://creativecommons.org/licenses/by/4.0/) is often appropriate for sharing texts, R Markdown files, generated figures, and other media, whereas scripts and any other computer code are often best shared under the [MIT license](https://opensource.org/licenses/MIT) (or similar).
 Both licenses assure maximal freedom for future users while requiring the attribution of the original authors in derivative work.

### 2020-09-24 manuscript.Rmd
explain redoc workflow in more detail closes #67
fix error in the first build/make example closes #66

@@
 
-# obtain a local copy from the remote repository
+# (1) obtain a local copy of the remote repository
 git clone https://github.com/aaronpeikert/reproducible-research.git
-# enter the local copy
-cd iris
-# create all files from that project
+# (2) obtain a local copy of the remote repository
+cd reproducible-research
+# (3) run the analysis/data preparation etc. with the local R installation
 make all
@@
 Conversions between R Markdown and Word retain all changes and even support Word's track-changes feature.
+Hence, RMarkdown user can share a Word file with their collaborators, receive their changes in this file and transform it back to RMarkdown.
 

### 2020-09-25 manuscript.Rmd
thank Michèle Nuijten closes #68
also mention figshare as doi provider see #60

@@
    We are grateful to Julia Delius for her helpful assistance in language and style editing.
-We thank the reviewers and all contributors of our GitHub repository for helpful feedback on the paper.
+We thank the anonymous reviewer and Michèle Nuijten, and all contributors of our GitHub repository for helpful feedback on the paper.
 This paper is fully reproducible using the workflow described in this paper.
@@
 the resulting pdf of the manuscript or the docker image (see [Containerization]).
-From such a release [zenodo.org](https://zenodo.org) can create a DOI, making it easier to reference it (see this [GitHub Guide](https://guides.github.com/activities/citable-code/)[^doi]).
+From such a release [zenodo.org](https://zenodo.org) or [figshare.com](https://figshare.com) can create a DOI, making it easier to reference it (see this [GitHub Guide](https://guides.github.com/activities/citable-code/)[^doi]).
 

### 2020-09-30 manuscript.Rmd
minor modifications

@@
 In this tutorial, we describe a workflow to ensure long-term and cross-platform reproducibility of data analyses in R [@baseR].
-Reproducibility is the ability to draw identical results from the same statistical analysis and the same data.
-For us,  statistical results are only reproducible if their generating, computational workflow is reported fully and transparently, and remains sustainably available, such that the workflow can be re-run by a different person or later in time, and that the results will be identical to the ones initially reported [@turingway2020; @herouxCompatibleReproducibilityTaxonomy2018; @claerboutElectronicDocumentsGive1992].
+Reproducibility is the ability to obtain identical results from the same statistical analysis and the same data.
+For us, statistical results are only reproducible if their generating, computational workflow is reported completely and transparently, and remains permanently available, such that the workflow can be re-run by a different person or later in time, and that the results remain identical to those initially reported [@turingway2020; @herouxCompatibleReproducibilityTaxonomy2018; @claerboutElectronicDocumentsGive1992].
 The need to ensure reproducibility directly follows from commonly accepted rules of good scientific practice [such as the guidelines of the German Research Foundation; @dfg2019].
@@
 the resulting pdf of the manuscript or the docker image (see [Containerization]).
-From such a release [zenodo.org](https://zenodo.org) or [figshare.com](https://figshare.com) can create a DOI, making it easier to reference it (see this [GitHub Guide](https://guides.github.com/activities/citable-code/)[^doi]).
+From such a release, [zenodo.org](https://zenodo.org) or [figshare.com](https://figshare.com) can create a DOI, making it easier to reference and retrieve it (see the [GitHub Guide](https://guides.github.com/activities/citable-code/)[^doi]).
 
@@
 
-A `Makefile` obeys a simple scheme.
-It contains a list of recipes.
-Each recipe has a target (the name of the recipe) followed by a colon and a list of dependent targets or files.
-This is comparable to a cooking recipe where you have the resulting dish at the top, the required ingredients after that and at last the steps one has to follow to prepare the dish.
+Fundamentally, a `Makefile` is a list of recipes.
+Each recipe has a target (the name of the recipe) followed by a colon and a list of dependent targets or files and finally a list of system commands to create the target.
+This is similar to a cooking recipe where the name of the dish appears first, then the required ingredients and finally the steps to follow to prepare the dish.
 If any of the dependencies have changed since the last time the target was built, the recipe's commands are executed to recreate the target file.
@@
 Also, if one of the dependencies of a target is newer than the target itself, Make updates everything that directly or indirectly depends on the target.
-Here, if `data/iris.csv` is newer than `data/iris_prepped.csv`, Make will attempt to recreate `data/iris_prepped.csv` first and then `manuscript.pdf`.
+Here, if the original data (`data/iris.csv`) is newer than the preprocessed data (`data/iris_prepped.csv`) and thus was likely modified since the preprocessing was done the last time, Make will attempt to recreate `data/iris_prepped.csv` first before recreating `manuscript.pdf`.
 If there is a dependency missing, and there is no target to make it, Make stops with an error message.
@@
 
-[^home]: See the section on [related approaches] for some alternatives to Docker in case you have to use Windows Home.
+[^home]: See the section on [related approaches] for alternatives to Docker in case you cannot avoid Windows Home.
 
@@
 How can one best share a reproducible workflow?
-We believe that, ideally, a non-commercial public service provider should be found that guarantees permanent and reliable hosting of our reproducible workflows, such as the Open Science Framework [@fosterOpenScienceFramework2017].
-A provider mirroring and complementing the services offered by GitHub, Docker Hub, and MRAN would be desirable.
+We believe that, ideally, a non-commercial public service provider should be found that guarantees permanent and reliable hosting of reproducible workflows, such as the Open Science Framework [@fosterOpenScienceFramework2017].
+An independent provider mirroring and complementing the services offered by GitHub, Docker Hub, and MRAN would be desirable.
 Second, to ensure that other users are legally able to benefit from the shared materials, authors must choose an appropriate license format.
@@
 We encourage authors to choose appropriate license forms that do not hinder others from freely downloading, using, and modifying the shared workflows and materials while, at the same time, ensuring recognition for the time and effort invested in creating the workflow in the first place.
-In our experience, the [Creative Commons - Attribution license (CC-BY)](https://creativecommons.org/licenses/by/4.0/) is often appropriate for sharing texts, R Markdown files, generated figures, and other media, whereas scripts and any other computer code are often best shared under the [MIT license](https://opensource.org/licenses/MIT) (or similar).
+In our experience, the [Creative Commons - Attribution license (CC-BY)](https://creativecommons.org/licenses/by/4.0/) is often appropriate for sharing texts, R Markdown files, generated figures, and other media, whereas scripts and any other computer code are often best shared under the [MIT license](https://opensource.org/licenses/MIT) (or similar permissive licenses).
 Both licenses assure maximal freedom for future users while requiring the attribution of the original authors in derivative work.

### 2020-09-30 manuscript.Rmd
last change

@@
 That is, we have experienced the case that after writing up the first draft in R Markdown, we eventually had to generate a Word file that, from then on, was used as static file serving as a basis for multiple iterations among the co-authors.
-Retaining reproducibility in such situations then requires tedious manual synchronization of files across formats.
+Retaining reproducibility in such situations requires tedious manual synchronization of files across formats.
 This annoyance may be reduced with the [`redoc` package](https://noamross.github.io/redoc/articles/mixed-workflows-with-redoc.html) [@redoc], which enables a bidirectional synchronization between Word and R Markdown.
-Conversions between R Markdown and Word retain all changes and even support Word's track-changes feature.
-Hence, RMarkdown user can share a Word file with their collaborators, receive their changes in this file and transform it back to RMarkdown.
+Conversions between R Markdown and Word retain all changes and support Word's track-changes feature.
+Hence, RMarkdown users can share a Word file with their collaborators, receive their changes in this file and transform it back to RMarkdown.
 

### 2020-10-01 manuscript.Rmd
strike sentence thatrefers to old figure
fix another code comment refix #66

@@
 Version control offers snapshots of your workflow at different time points identified by a unique identifier.
-The different frames in the z-axis of Figure  \@ref(fig:schematic) illustrate the different versions of the project.
 How different parts of a analysis and a report of it relate and in what order they need to be executed is documented using dependency management.
@@
 git clone https://github.com/aaronpeikert/reproducible-research.git
-# (2) obtain a local copy of the remote repository
+# (2) enter the directory containing the project
 cd reproducible-research

### 2020-10-01 manuscript.Rmd
fix grammar

@@
 
-# BAD because of path that is specific to the computer
+# BAD because the path is specific to the computer/user
 iris <- read.csv("/home/aaron/reproducible-research/data/iris.csv")
-# GOOD because it is a relative path but slash depends on OS
+# GOOD because it is a relative path, but slash depends on OS
 iris <- read.csv("data/iris.csv")

### 2020-10-01 manuscript.Rmd
minor rewording

@@
 git clone https://github.com/aaronpeikert/workflow-showcase.git
-# (2) enter the directory containing the project
+# (2) enter the project directory 
 cd workflow-showcase

### 2020-10-02 manuscript.Rmd
swap order of Michèle and anonymous reviewer
containing the project -> project directory closes #72

@@
    We are grateful to Julia Delius for her helpful assistance in language and style editing.
-We thank the anonymous reviewer and Michèle Nuijten, and all contributors of our GitHub repository for helpful feedback on the paper.
+We thank Michèle Nuijten and the anonymous reviewer, and all contributors of our GitHub repository for helpful feedback on the paper.
 This paper is fully reproducible using the workflow described in this paper.
@@
 git clone https://github.com/aaronpeikert/reproducible-research.git
-# (2) enter the directory containing the project
+# (2) enter the project directory
 cd reproducible-research

### 2020-10-02 manuscript.Rmd
terminal -> comand line in code chunks closes #71

@@
 ```{bash, eval=FALSE}
-# -- type this in a terminal --
+# -- type this on the command line --
+
 git init # to initialize Git in the current directory
@@
 ```{bash, eval = FALSE}
-# -- type this in a terminal --
+# -- type this on the command line --
 
@@
 ```{bash, eval=FALSE}
-# -- type this in a terminal --
+# -- type this on the command line --
 
@@
 ```{bash, eval=FALSE}
-# -- type this in a terminal --
+# -- type this on the command line --
+
 git clone https://github.com/aaronpeikert/reproducible-research.git

### 2020-10-02 manuscript.Rmd
add link to coc in footnote again closes #62

@@
 A great resource to choose a license is [choosealicense.com](https://choosealicense.com), however, no resource, including our recommendation, replaces legal advice.
-To facilitate an inclusive environment, we recommend naming all contributors and including a [Code of Conduct](https://opensource.guide/code-of-conduct/) in your project.
+To facilitate an inclusive environment, we recommend naming all contributors and including a [Code of Conduct](https://opensource.guide/code-of-conduct/)[^coc] in your project.
+
+[^coc]: https://opensource.guide/code-of-conduct/
 

### 2020-10-02 manuscript.Rmd
minor revisions throughout

@@
    We are grateful to Julia Delius for her helpful assistance in language and style editing.
-We thank Michèle Nuijten and the anonymous reviewer, and all contributors of our GitHub repository for helpful feedback on the paper.
+We thank Michèle Nuijten and the anonymous reviewer, and all contributors to our GitHub repository for helpful feedback on the paper.
 This paper is fully reproducible using the workflow described in this paper.
@@
 The need to ensure reproducibility directly follows from commonly accepted rules of good scientific practice [such as the guidelines of the German Research Foundation; @dfg2019].
-Ensuring reproducibility is a prerequisite for replicability (the ability to reach a similar conclusion from the same analysis and _new_ data), and a means to increase the trustworthiness of empirical results [@epskamp2019rep].
+Ensuring reproducibility is a prerequisite for replicability (the ability to reach consistent conclusion from the same analysis and _new_ data), and a means to increase the trustworthiness of empirical results [@epskamp2019rep].
 Transparency and accessibility are central scientific values, and open, reproducible projects will increase the efficiency and veracity of knowledge accumulation [@nosek2012scientific].
@@
 We argue that only a workflow using all four concepts in unison can guarantee confidence in reproducing a scientific report [see @turingway2020 for similar arguments].
-Various implementations of these concepts exist, but we consider the following four best suited for analyses centered on the R environment [@r2019] but also allowing for external dependencies: R Markdown [@rmarkdown2019] for dynamic document generation, Git [@chaconProGit2014] for version control, Make [@feldmanMakeProgramMaintaining1979] for dependency management, and Docker [@merkelDockerLightweightLinux2014] for containerization.
-Each of these software solutions serves a valuable meta-scientific goal (reproducibility) and increases the researches' productivity.
+Various implementations of these concepts exist, but we consider the following four best suited for analyses centered on the R environment [@baseR] but also allowing for external dependencies: R Markdown [@rmarkdown2019] for dynamic document generation, Git [@chaconProGit2014] for version control, Make [@feldmanMakeProgramMaintaining1979] for dependency management, and Docker [@merkelDockerLightweightLinux2014] for containerization.
+Each of these software solutions serves a valuable meta-scientific goal (reproducibility) and increases the researchers' productivity.
 They all are highly flexible and powerful, so that mastering them fully, requires a considerable amount of practice.
@@
 Version control offers snapshots of your workflow at different time points identified by a unique identifier.
-How different parts of a analysis and a report of it relate and in what order they need to be executed is documented using dependency management.
-The arrows visualize dependencies, such as an analysis depending on the availability of a particular data file.
+How different parts of an analysis and a corresponding report relate to each other and in what order they need to be executed is documented using dependency management.
+The arrows in Figure  \@ref(fig:schematic) visualize dependencies, such as an analysis depending on the availability of a particular data file.
 Third, all computer code (such as a statistical analyses in R) is executed in a virtual environment that guarantees exact reproduction of results independent of the host operating system, the locally installed R version, and installed package versions.
@@
 
-The translation of computational results into a human-readable summary, for example into a technical report, a presentation, or a journal manuscript, is time-consuming and error-prone.
+The translation of computational results into a human-readable summary, for example into a technical report, a presentation, or a manuscript, is time-consuming and error-prone.
 Typical errors result from copy\&paste mistakes, erroneous rounding, or missed updates of the manuscript when the associated computer code and computed results have changed.
@@
 R Markdown is a simple markup language to create dynamic documents with embedded chunks of R code that can be exported to standard formats such as documents (docx, rtf, epub), presentations (ppt, html) or websites (html) using the `knitr` package [@xie2015; @xie2019].
-Several packages extend the functionality of knitr.
+Several packages extend the functionality of `knitr`.
 Of particular note are the `papaja` package [@papaja2018], which offers additional functions to enable American Psychological Association (APA) style document formatting, including a journal-style final typeset format, and the `stargazer` package [@stargazer2018], which provides journal-ready tables and reports of statistical models.
@@
 
-Even when you have obtained a given version of a project with the aim to reproduce reported results, and you can confirm that this version is unchanged, you may not know exactly how to reproduce the results because it may be unclear which files should be executed in which order.
+Even when you have obtained a given version of a project with the aim to reproduce reported results, and you can confirm that this version is unchanged, you may not know exactly how to reproduce the results because it may be unclear which scripts or commands must be executed in which order.
 This is particularly the case when complex preprocessing pipelines are part of the computation or there are dependencies on external programs.
@@
 Fundamentally, a `Makefile` is a list of recipes.
-Each recipe has a target (the name of the recipe) followed by a colon and a list of dependent targets or files and finally a list of system commands to create the target.
+Each recipe has a target (the name of the recipe) followed by a colon, a list of dependent targets or files, and finally a list of system commands to create the target.
 This is similar to a cooking recipe where the name of the dish appears first, then the required ingredients and finally the steps to follow to prepare the dish.
@@
 If you type `make manuscript.pdf`, Make first checks whether the dependencies do exist and, if not, creates them.
-Here, if `data/iris_prepped.csv` does not exist, Make creates it by executing the third target.
+Here, if `data/iris_prepped.csv` does not exist, Make creates it by executing the third target (running the preprocessing script `R/prepare_data.R`).
 Also, if one of the dependencies of a target is newer than the target itself, Make updates everything that directly or indirectly depends on the target.
@@
 
-If you have followed our workflow as presented thus far, a fellow researcher is only three commands away from fully reproducing your analysis.
-They would simply have to type the following commands in the terminal:
+If you have followed our workflow as presented thus far, you are (almost) only three commands away from fully reproducing our paper. You simply have to type the following commands on the command line:
 
@@
 
-If you execute the above on your system, there is a good chance that you cannot reproduce our manuscript and the `make all` command results in an error.
-Successful reproducibility relies on the crucial assumption that your computational environments are identically or sufficiently compatible, e.g., all software dependencies are installed (R and all additional R Packages) and no updates or other changes to the computational environment break or alter your analysis.
-As we will shortly see, ensuring full computational reproducibility requires one further level of documentation, that is, documentation of the computational environment.
+However, if you execute the above on your system, there is a good chance that you cannot reproduce our manuscript and the `make all` command results in an error.
+Successful reproducibility relies on the crucial assumption that your computational environment is identical to ours or sufficiently compatible, e.g., all software dependencies are installed (R and all additional R Packages) and no updates or other changes to the computational environment break or alter the original analysis.
+As we will shortly see, ensuring full computational reproducibility requires one further level of documentation, that is, documentation and reproduction of the computational environment.
 
@@
 The most important advantage over traditional virtual machines is that containers are lightweight: they start rapidly, run with little overhead, and do not need much storage space.
-Docker achieves this by reusing large parts of the host's operating system (the extent varies between Linux, macOS and Windows).
+Docker achieves this by reusing large parts of the host's operating system.
 
-With the following example, we demonstrate the importance of documenting and storing the computational environment.
+With the following example, we demonstrate the importance of documenting and (re-)storing the computational environment.
 Generally, with containers, we would like to safeguard against changes to the computational environment resulting in unexpected consequences, e.g., changes in the functionality or default options in packages or even in the R environment itself.
 While the R programming language is considered stable and much effort is put into backward compatibility, even basic functions like `read.csv()` (to load data) or `sample()` (to randomly sample from a set) sometimes change their behaviour from one version to another.
-For example, to ensure reproducibility of analyses based on a computer's pseudo random number generator (PRNG), it is good practice to rely on fixed PRNG seeds, which are numeric values that set the PRNG into a deterministic state, i.e., the sequence of random numbers reproduces exactly.
+For example, to ensure reproducibility of analyses based on a computer's pseudo-random number generator (PRNG), it is good practice to rely on fixed PRNG seeds, which are numeric values that set the PRNG into a deterministic state, i.e., the sequence of pseudo-random numbers reproduces exactly.
 Consider the following R code to randomly draw five numbers between 1 and 10:
@@
 As we see from these R outputs, the latter of the estimated confidence intervals does include zero while the former does not.
-Please note that one could discuss deeper issues about null hypothesis significance testing here, but with this example, we would simply like to stress that computational reproducibility in the strict sense requires capturing the full computational environment.
+Please note that one could discuss deeper issues with null hypothesis significance testing here, but with this example, we would simply like to stress that computational reproducibility in the strict sense requires capturing the full computational environment.
 
@@
 Those `RUN` statements can install dependencies that are not an R package, e.g., other programming languages like python or Matlab.
-The `WORKDIR` statement is not strictly necessary but saves time spent on writing the working directory.
+The `WORKDIR` statement is not strictly necessary but simplifies commands.
 The command `docker build -t image-name` creates an image named `image-name` from the Dockerfile in the project.
-A way to identify the dependencies automatically and generate a docker image out of them is provided in the [`liftR`](https://cran.r-project.org/web/packages/liftr/vignettes/liftr-intro.html) package [@liftr].
+A way to identify the dependencies automatically and generate a docker image from them is provided in the [`liftR`](https://cran.r-project.org/web/packages/liftr/vignettes/liftr-intro.html) package [@liftr].
 
@@
 In the following section, we share what we consider the easiest way to install those packages across common operating systems.
-However, installation processes may be subject to change, and we advise that readers also consult the documentations of the packages or see our collection of links to tutorials and installation instructions on our GitHub repository.
+However, installation processes may be subject to change, and we advise readers to also consult the documentations of the packages or see our collection of links to tutorials and installation instructions on our GitHub repository.
 
@@
 Next, one needs to make sure that all files relevant to the analysis can be moved to another computer.
-To this end, it is recommended that all files reside within a one folder (or enclosed subfolders within it) and all paths are relative to that folder because absolute paths are specific to a given computer.
+To this end, it is recommended that all files reside within one folder (or enclosed subfolders within it) and all paths are relative to that folder because absolute paths are specific to a given computer.
 A robust solution to the problem of making sure that file access does not break across computing platforms are [RStudio projects](https://r4ds.had.co.nz/workflow-projects.html) and the `here` package [@here2017] to manage file access.
@@
 ```{r}
+# -- R code --
+
 # create a dummy data.frame with two columns
@@
 
-The overarching goal of this paper was to provide a complete and easy-to-use workflow that allows confidence in the reproducibility of R-based data analyses.
+The overarching goal of this paper was to provide a complete workflow that allows confidence in the reproducibility of R-based data analyses.
 Analyses following our workflow can be reproduced with four commands (here shown for this manuscript):
@@
 
-The workflow enables reproduction a scientific report exactly without regard to the local operating system, locally installed software, time, or interim changes to the project files.
+The workflow enables the reproduction of a scientific report exactly without regard to the local operating system, locally installed software, time, or interim changes to the project files.
 To that end, the proposed workflow relies on tools that have been the foundation of reliable software development for years or even decades.
@@
 Each tool in the workflow reduces the chances of non-reproducibility.
-Dynamic reporting with R Markdown guarantees consistency between computational results and their reporting; Version control with Git ensures permanence and consistency across multiple versions of data and code; Dependency management with Make provides defined entry-points while mapping out dependencies between all components of a project; containerization with Docker guarantees full computational reproducibility.
-We believe that the proposed combination of tools does not restrict researchers but enables them to operate on a solid basis to deliver transparent and sustainable research.
+Dynamic reporting with R Markdown guarantees consistency between computational results and their reporting; version control with Git ensures permanence and consistency across multiple versions of data and code; dependency management with Make provides defined entry-points while mapping out dependencies between all components of a project; containerization with Docker guarantees full computational reproducibility.
+We believe that the proposed combination of tools does not limit researchers but enables them to operate on a solid basis to deliver transparent and sustainable research.
 
@@
 The use of R Markdown within a project, tracked with Git can be simplified with the [`workflowR`](https://github.com/jdblischak/workflowr) package [@workflowr].
-The [`drake`](https://github.com/ropensci/drake) package [@drake] is directly inspired by Make and takes an R centric approach, making it especially suited for projects only involving R, but can also handle external dependencies.
+The [`drake`](https://github.com/ropensci/drake) package [@drake] is directly inspired by Make and takes an R-centric approach, making it especially suited for projects only involving R, but it can also handle external dependencies.
 The [liftR ](https://cran.r-project.org/web/packages/liftr/vignettes/liftr-intro.html) package [@liftr] and the [`holepunch`](https://karthik.github.io/holepunch/) package [@holepunch] automatize the use of Docker.
@@
 There are several alternatives to Docker that manage dependencies on R packages.
-[`renv`](https://rstudio.github.io/renv/articles/renv.html) [@renv] is a way to freeze package version via local copies of packages in the project, but it does not guarantee a given base R version or system dependencies.
+[`renv`](https://rstudio.github.io/renv/articles/renv.html) [@renv] is a way to freeze package version via local copies of packages in the project, but it does not guarantee a given base R version or system dependencies beyond R.
 Similar approaches are taken by [`jetpack`](https://github.com/ankane/jetpack) [@jetpack], [`miniCRAN`](https://cran.r-project.org/web/packages/miniCRAN/vignettes/miniCRAN-introduction.html) [@miniCRAN] and [`checkpoint`](https://cran.r-project.org/web/packages/checkpoint/vignettes/checkpoint.html) [@checkpoint].

### 2020-10-02 abstract.Rmd
minor revisions throughout

@@
 The workflow leverages established tools and practices from software engineering.
-It combines the benefits of various open-source software tools including R Markdown, Git, Make, and Docker, whose interplay ensures seamless integration of version management, dynamic report generation conforming to various journal styles and full cross-platform and long-term computational reproducibility.
+It combines the benefits of various open-source software tools including R Markdown, Git, Make, and Docker, whose interplay ensures seamless integration of version management, dynamic report generation conforming to various journal styles, and full cross-platform and long-term computational reproducibility.
 The workflow ensures meeting the primary goals that 1) the reporting of statistical results is consistent with the actual statistical results (dynamic report generation), 2) the analysis exactly reproduces at a later point in time even if the computing platform or software is changed (computational reproducibility), and 3) changes at any time (during development and post-publication) are tracked, tagged, and documented while earlier versions of both data and code remain accessible.

### 2020-10-02 manuscript.Rmd
add link to latest release

@@
 GitHub allows you to archive and label a specific version of your repository in the form of a release.
-A release tags a particular commit with an arbitrary label, e.g., as "submission,"[^sub] "preprint," or "published," and archives also "binary" products of your code, e.g.
+A release tags a particular commit with an arbitrary label, e.g., as "submission,"[^sub] "preprint," or "published,"[^pub] and archives also "binary" products of your code, e.g.
 the resulting pdf of the manuscript or the docker image (see [Containerization]).
@@
 [^doi]: [https://guides.github.com/activities/citable-code/](https://guides.github.com/activities/citable-code/)
-[^sub]: We created an release for the submission: [https://github.com/aaronpeikert/reproducible-research/releases/tag/v0.2.1-submission](https://github.com/aaronpeikert/reproducible-research/releases/tag/v0.2.1-submission)
+[^sub]: We created an release for the submission: [https://github.com/aaronpeikert/reproducible-research/releases/tag/v0.1.1-submission](https://github.com/aaronpeikert/reproducible-research/releases/tag/v0.2.1-submission)
+[^pub]: We created an release for the publication: [https://github.com/aaronpeikert/reproducible-research/releases/latest](https://github.com/aaronpeikert/reproducible-research/releases/latest)
 

### 2021-03-19 manuscript.Rmd
update Dockerfile example in manuscript

@@
 # install additional R packages from github: papaja and wordcountaddin
+# the package version if fixed by hash (user/package@hash)
 RUN installGithub.r\
-  crsh/papaja benmarwick/wordcountaddin
+  crsh/papaja@b6cd70f benmarwick/wordcountaddin@fdf70d9
 # set the working directory inside the container

### 2021-03-23 manuscript.Rmd
update links to other packages

@@
 There are various R packages that implement parts of our workflow and, thus, lower the threshold for adoption when the full flexibility provided by our workflow is not needed.
-The use of R Markdown within a project, tracked with Git can be simplified with the [`workflowR`](https://github.com/jdblischak/workflowr) package [@workflowr].
-The [`drake`](https://github.com/ropensci/drake) package [@drake] is directly inspired by Make and takes an R-centric approach, making it especially suited for projects only involving R, but it can also handle external dependencies.
+The use of R Markdown within a project, tracked with Git can be simplified with the [`workflowR`](https://jdblischak.github.io/workflowr/articles/wflow-01-getting-started.html) package [@workflowr].
+The [`drake`](https://books.ropensci.org/drake/) package [@drake] is directly inspired by Make and takes an R-centric approach, making it especially suited for projects only involving R, but it can also handle external dependencies.
 The [liftR ](https://cran.r-project.org/web/packages/liftr/vignettes/liftr-intro.html) package [@liftr] and the [`holepunch`](https://karthik.github.io/holepunch/) package [@holepunch] automatize the use of Docker.

### 2022-03-31 README.Rmd
include nutshell figure after abstract
use relative path for image
add clean rebuild option in makefile
add note in case of unavailible services

@@
 ```{r abstract, child = 'abstract.Rmd'}
+```
+
+```{r, echo=FALSE}
+knitr::include_graphics("Images/nutshell.svg")
 ```
@@
 
+**Note: The build step may fail since services the build relies on are no longer available.
+In this case, you can download a ready build image from the [GitHub Releases](https://github.com/aaronpeikert/reproducible-research/releases/latest) and use `docker load -i reproducible-research.tar.gz` to load it.
+Continue with `make all DOCKER=TRUE`**
+
 **Note: Windows user need to manually edit the `Makefile` and set current_path to the current directory and use `make all DOCKER=TRUE WINDOWS=TRUE`.
@@
 ```{bash, eval=FALSE}
-git pull && make build && make -B DOCKER=TRUE
+git pull && make rebuild && make -B DOCKER=TRUE
 ```
